---
title: "analyze_vertical_growth"
author: "Carmen"
date: "November 9, 2017"
output: 
    html_document:
        toc: TRUE
---

This code analyzes how shrub removal affected seedling vertical growth.

```{r, include = F}
library(tidyverse)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
```

# Load data

```{r}
setwd("~/Shrubs-Seedlings/code/Shrub_experiment/analyze/")
load(file = "../../../compiled_data/shrub_removal_cleaned.Rdata")
```

# BACI ANALYSIS FOR PINES for 2018 GROWTH

# Restructure
```{r}
dfg2 <- gather(df, key = "pre_post", value = "relative_growth", c("growth_rel_2018", "pre_removal_growth_rel"))
dfg2 <- dfg2 %>% 
  mutate(pre_post = as.factor(paste(pre_post)))
```

```{r}
ggplot(dfg2 %>% filter(species == "PIPO"))+
  geom_boxplot(aes(x = interaction(shrubs, pre_post), y = relative_growth))
```

# Define pipo-only df
```{r}
dfpipo <- dfg2 %>% filter(species == "PIPO")
```


# Mixed effects linear model with island random effect

```{r}
PIPO_lm <- lme(relative_growth ~ shrubs*pre_post, 
               random = ~ 1|island, 
               data = dfpipo, 
               method = "ML")
anova(PIPO_lm)
```

## Check for  height effect

```{r}
PIPO_lm_ht <- lme(relative_growth ~ shrubs*pre_post + height_2016, 
                  random = ~ 1|island, 
                  data = dfpipo, 
                  method = "ML")
anova(PIPO_lm, PIPO_lm_ht)
```

I DO need to include height

## Check for height interaction
```{r}
PIPO_lm_int <- lme(relative_growth ~ shrubs*pre_post + shrubs*height_2016, 
                   random = ~ 1|island, 
                   data = dfpipo, method = "ML")
anova(PIPO_lm_ht, PIPO_lm_int)
```

I DO NOT need to include a shrub*height interaction


## Check for diameter effect
```{r}
PIPO_lm_dia <- lme(relative_growth ~ shrubs*pre_post +height_2016 +dia17_mm_Aug, 
                   random = ~ 1|island, 
                   data = dfpipo, 
                   method = "ML")
anova(PIPO_lm_ht, PIPO_lm_dia)
```

I SHOULD include diameter


## Check for shrub species effect
```{r}
dfpipo$shrub_species <- as.factor(dfpipo$shrub_species)
summary(dfpipo$shrub_species)
PIPO_lm_shspp <- lme(relative_growth ~ shrubs*pre_post + height_2016 + dia17_mm_Aug + shrub_species,
                      random = ~ 1|island,
                      data = dfpipo, method = "ML")
anova(PIPO_lm_dia, PIPO_lm_shspp)

PIPO_lm_shspp <- lme(relative_growth ~ shrubs*pre_post + height_2016 + dia17_mm_Aug + shrubs*shrub_species,
                      random = ~ 1|island,
                      data = dfpipo, method = "ML")
anova(PIPO_lm_dia, PIPO_lm_shspp)
```

No shrub species effect


## Establish best model
```{r}
pipo_lm_best <- lme(relative_growth ~ shrubs*pre_post +height_2016 +dia17_mm_Aug, 
                   random = ~ 1|island, 
                   data = dfpipo, 
                   method = "ML")
```


## Test for shrub effect in model with interaction - the true test for BACI
```{r}
PIPO_lm_noshrub <- lme(relative_growth ~ shrubs + pre_post + height_2016 + dia17_mm_Aug, 
                       random = ~ 1|island, 
                       data = dfpipo, 
                       method = "ML")
anova(PIPO_lm_noshrub, PIPO_lm_dia)
```

NO significant effect - but almost

## Visual
```{r}
plot(predictorEffect("shrubs", PIPO_lm_dia))
plot(predictorEffect("pre_post", PIPO_lm_dia))
drop1(PIPO_lm_dia, test = "Chisq")
```

### Make a prettier graph in ggplot

```{r}
effects_dfpipo <- as.data.frame(predictorEffect("pre_post", PIPO_lm_dia))
effects_dfpipo <- effects_dfpipo %>% 
  mutate(pre_post = ifelse(pre_post == "pre_removal_growth_rel", "before", "after"))
effects_dfpipo <- effects_dfpipo %>% 
  mutate(pre_post = as.factor(pre_post))
```

#### Reorder levels
```{r}
levels(effects_dfpipo$pre_post)
effects_dfpipo$pre_post = factor(effects_dfpipo$pre_post,levels(effects_dfpipo$pre_post)[c(2,1)])
levels(effects_dfpipo$pre_post)
```

#### Graph with reverse log transformed
```{r}
ggplot(effects_dfpipo, aes(x = pre_post, y = exp(fit), group = shrubs, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = exp(lower), ymax = exp(upper)), width = .2)+
  xlab("timing relative to shrub removal")+
  theme_bw()+
  theme(legend.position = c(0.8, 0.8))
```


### Save graph
```{r}
ggsave(file = "../../../results/figures/ShrubExperiment/PineVertBACI.png", width = 6, height = 5, dpi = 400)
```

# BACI ANALYSIS FOR FIRS for 2018 GROWTH


# Define ABCO-only df
```{r}
dfABCO <- dfg2 %>% filter(species == "ABCO")
```

# Mixed effects linear model with island random effect

```{r}
ABCO_lm <- lme(relative_growth ~ shrubs*pre_post, 
               random = ~ 1|island, 
               data = dfABCO, 
               method = "ML")
anova(ABCO_lm)
```

## Check for  height effect
```{r}
ABCO_lm_ht <- lme(relative_growth ~ shrubs*pre_post + height_2016, 
                  random = ~ 1|island, 
                  data = dfABCO, 
                  method = "ML")
anova(ABCO_lm, ABCO_lm_ht)
```

YES

## Check for height interaction
```{r}
ABCO_lm_int <- lme(relative_growth ~ shrubs*pre_post + shrubs*height_2016, 
                   random = ~ 1|island, 
                   data = dfABCO, 
                   method = "ML")
anova(ABCO_lm_ht, ABCO_lm_int)
```

NO

## Check for diameter effect
```{r}
ABCO_lm_dia <- lme(relative_growth ~ shrubs*pre_post +height_2016 +dia17_mm_Aug, 
                   random = ~ 1|island, 
                   data = dfABCO, 
                   method = "ML")
anova(ABCO_lm_ht, ABCO_lm_dia)
```

NO

## Check for shrub species effect
```{r}
dfABCO$shrub_species <- as.factor(dfABCO$shrub_species)
summary(dfABCO$shrub_species)
ABCO_lm_shspp <- lme(relative_growth ~ shrubs*pre_post + height_2016 + shrub_species,
                      random = ~ 1|island,
                      data = dfABCO, method = "ML")
anova(ABCO_lm_ht, ABCO_lm_shspp)

ABCO_lm_shspp_int <- lme(relative_growth ~ shrubs*pre_post + height_2016 + shrubs*shrub_species,
                      random = ~ 1|island,
                      data = dfABCO, method = "ML")
anova(ABCO_lm_shspp_int, ABCO_lm_shspp)
```

YES, but no interaction

## Establish best model
```{r}
ABCO_lm_best <- lme(relative_growth ~ shrubs*pre_post + height_2016 + shrub_species, 
                   random = ~ 1|island, 
                   data = dfABCO, 
                   method = "ML")
summary(ABCO_lm_best)
```

## Test for shrub effect in model with interaction - the true test for BACI
```{r}
ABCO_lm_noshrub <- lme(relative_growth ~ shrubs + pre_post + height_2016 + shrub_species, 
                       random = ~ 1|island, 
                       data = dfABCO, 
                       method = "ML")
anova(ABCO_lm_noshrub, ABCO_lm_best)
```

NO significant effect

## Visual
```{r}
plot(predictorEffect("shrubs", ABCO_lm_best))
plot(predictorEffect("shrub_species", ABCO_lm_best))
drop1(ABCO_lm_best, test = "Chisq")
```

### Make a prettier graph in ggplot

```{r}
effects_dfg2 <- as.data.frame(predictorEffect("pre_post", ABCO_lm_best))
effects_dfg2 <- effects_dfg2 %>% 
  mutate(pre_post = ifelse(pre_post == "pre_removal_growth_rel", "before", "after"))
effects_dfg2 <- effects_dfg2 %>% 
  mutate(pre_post = as.factor(pre_post))
```

#### Reorder levels
```{r}
levels(effects_dfg2$pre_post)
effects_dfg2$pre_post = factor(effects_dfg2$pre_post,levels(effects_dfg2$pre_post)[c(2,1)])
levels(effects_dfg2$pre_post)
```

#### Graph with reverse log transformed
```{r}
ggplot(effects_dfg2, aes(x = pre_post, y = exp(fit), group = shrubs, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = exp(lower), ymax = exp(upper)), width = .2)+
  xlab("timing relative to shrub removal")+
  theme_bw()+
  theme(legend.position = c(0.8, 0.8))
```

### Save graph
```{r}
ggsave(file = "../../../results/figures/ShrubExperiment/FirVertBACI.png", width = 6, height = 5, dpi = 400)
```
