---
title: "analyze_volume_2019"
author: "Carmen"
date: "April 23, 2019"
output: 
    html_document:
      toc: TRUE
---


```{r, include = F}
library(tidyverse)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
```

# Load data

```{r}
setwd("~/Shrubs-Seedlings/code/Shrub_experiment/analyze/")
load(file = "../../../compiled_data/shrub_removal_cleaned.Rdata")
```

```{r}
df %>% arrange(SdlgID) %>% select(1:5)
```

# Calculate volume

Note: all heights represent the height at the END of the growing season

Assuming parabolic cone shape (https://www.grc.nasa.gov/www/k-12/airplane/volume.html)
```{r}
df <- df %>% 
  mutate(vol_Aug17_cm2 = (2*pi*((dia17_mm_Aug/10)^2)*(ht_cm_2017))/15) %>% 
  mutate(vol_Aug18_cm2 = (2*pi*((dia18_mm/10)^2)*(height_2018))/15) 

hist(log(df$vol_Aug17_cm2))
hist(log(df$vol_Aug18_cm2))

```

# Calculate relative growth rate using volume
```{r}
df <- df %>% 
  mutate(growth_rel = (vol_Aug18_cm2 - vol_Aug17_cm2)/vol_Aug17_cm2)
hist(df$growth_rel, breaks = 50)
hist(log(df$growth_rel), breaks = 50)

```

NOTE: There are 0 values and negative values for growth. This makes it hard to ln transform the response variable, which is what I did for the fire footprints analysis.

# Mixed effects linear model with island random effect

```{r}
dfpipo <- df %>% filter(species == "PIPO")
pipo_lm <- lme( growth_rel ~ shrubs, random = ~ 1|island, 
                data = dfpipo, method = "ML")
summary(pipo_lm)
anova(pipo_lm)
```

## Check whether to include height
```{r}
pipo_lm_ht <- lme( growth_rel ~ shrubs + ht_cm_2017, random = ~ 1|island, 
                   data = dfpipo, method = "ML")
anova(pipo_lm, pipo_lm_ht)
```

No.

## Check for interaction between shrubs and height
```{r}
pipo_lm_int <- lme( growth_rel ~ shrubs*ht_cm_2017, random = ~ 1|island, 
                    data = dfpipo, method = "ML")
anova(pipo_lm, pipo_lm_int)
```

No.

## Check for diameter effect
```{r}
pipo_lm_dia <- lme( growth_rel ~ shrubs + dia17_mm_Aug,random = ~ 1|island,  
                    data = dfpipo, method = "ML")
anova(pipo_lm, pipo_lm_dia)
```

No. 

```{r}
pipo_lm_dia <- lme( growth_rel ~ shrubs * dia17_mm_Aug,random = ~ 1|island,  
                    data = dfpipo, method = "ML")
anova(pipo_lm, pipo_lm_dia)
```

No. 

## Test for shrub effect in best model
```{r}
pipo_lm_noshrub <- lme( growth_rel ~ 1, random = ~ 1|island, 
                        data = dfpipo, method = "ML")
anova(pipo_lm_noshrub, pipo_lm)
AIC(pipo_lm_noshrub)
AIC(pipo_lm)
```

Yes, there is a shrub effect

### Residuals
```{r}
dfpipo <- df %>% filter(species == "PIPO")
E <- resid(pipo_lm, type = "normalized")
dfpipo$E <- E
pred <- predict(pipo_lm)
dfpipo$pred <- pred
plot(pipo_lm)
plot(dfpipo$ht_cm_2017, E)
ggplot(dfpipo, aes(x = log(dfpipo$vol_Aug17_cm2), y = E))+
  geom_point()+
  geom_smooth()
```

### LOOK AT OUTLIER
```{r}
dfpipo %>% filter(E>6) %>% select(SdlgID, dia17_mm_Aug, ht_cm_2017, dia18_mm, height_2018, vol_Aug17_cm2, vol_Aug18_cm2, growth_rel, pred, E)
```

Seems like it's real data, not an error, so I'm keeping the outlier in.


## Visualize shrub effect
```{r}
plot(predictorEffect("shrubs", pipo_lm))
```

#### Graph with ggplot
```{r}
effects_df <- as.data.frame(predictorEffect("shrubs", pipo_lm))
```

```{r}
ggplot(effects_df, aes(x = shrubs, y = fit))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .2)+
  ggtitle("PIPO shrub removal effects")+
  ylim(c(0,.8))
```

### Graph raw data by shrub island 
```{r}
df_island <- dfpipo %>% 
  group_by(island, compartment, shrubs) %>% 
  summarize(mean_growth_rel = mean(growth_rel), n = n()) %>% 
  ungroup() %>% 
  group_by(compartment, island) 

df_island_diff <- spread(df_island %>% select(-n), key = shrubs, value = mean_growth_rel) %>% 
  arrange(island) %>% 
  mutate(higher = ifelse(removed > retained, "rem", "ret")) %>% 
  select(island, compartment, higher)

df_island_diff

df_island <- full_join(df_island, df_island_diff) %>% 
  group_by(island, compartment) %>% 
  mutate(n_island = sum(n))

ggplot(df_island, aes(x = shrubs, y = mean_growth_rel))+
  geom_point(aes(col = higher, alpha = n_island), size = 2)+
  geom_line(aes(group = interaction(island, compartment), alpha = n_island, col = higher), size = 1)+
  theme_bw()+
  ylab("Mean plot volumetric relative growth rate")+
  scale_color_discrete(name = "Plot with higher growth", labels = c("Removed", "Retained"))+
  scale_alpha_continuous(name = "n", range = c(.2, 1))

ggsave(file = "../../../results/figures/ShrubExperiment/PipoVol.png", width = 3, height = 5, dpi = 400)

```

## Count how many plot pairs had increases vs. decreases
```{r}
df_island %>% 
  select(island, compartment, higher) %>% 
  distinct() %>% 
  group_by(higher) %>% 
  summarize(n())

```

# REPEAT FOR FIRS

# Mixed effects linear model with island random effect

```{r}
ABCO_lm <- lme( growth_rel ~ shrubs, random = ~ 1|island, data = df %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm)
```

## Check whether to include height
```{r}
ABCO_lm_ht <- lme(growth_rel ~ shrubs+ht_cm_2017, random = ~ 1|island, data = df %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm, ABCO_lm_ht)
```

YES

## Check for interaction between shrubs and height
```{r}
ABCO_lm_int <- lme( growth_rel ~ shrubs*ht_cm_2017, random = ~ 1|island, data = df %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm_ht, ABCO_lm_int)
```

NO

## Check for diameter effect
```{r}
ABCO_lm_dia <- lme( growth_rel ~ shrubs+ ht_cm_2017 + dia17_mm_Aug,random = ~ 1|island,  data = df %>% filter(species == "ABCO" & !is.na(shrubs)), method = "ML")
summary(ABCO_lm_dia)
anova(ABCO_lm_dia)
anova(ABCO_lm_ht, ABCO_lm_dia)
```

YES

## Test for shrub effect in best model 
```{r}
ABCO_lm_noshrub <- lme( growth_rel ~ ht_cm_2017 + dia17_mm_Aug, random = ~ 1|island, data = df %>% filter(species == "ABCO" & !is.na(shrubs)), method = "ML")
anova(ABCO_lm_noshrub, ABCO_lm_dia)
AIC(ABCO_lm_noshrub)
AIC(ABCO_lm_dia)
```

YES, there is a significant effect.

### Residuals
```{r}
dfABCO <- df %>% filter(species == "ABCO")
E <- resid(ABCO_lm, type = "normalized")
dfABCO$E <- E
pred <- predict(ABCO_lm)
dfABCO$pred <- pred
plot(ABCO_lm)
plot(dfABCO$ht_cm_2017, E)
ggplot(dfABCO, aes(x = vol_Aug17_cm2, y = E))+
  geom_point()+
  geom_smooth()
```

### lOOK AT OUTLIER
```{r}
dfABCO %>% filter(vol_Aug17_cm2>1000) %>% select(SdlgID, dia17_mm_Aug, ht_cm_2017, dia18_mm, height_2018, vol_Aug17_cm2, vol_Aug18_cm2, growth_rel, pred, E)
```


## Visual
```{r}
plot(predictorEffect("shrubs", ABCO_lm_dia))
plot(predictorEffect("dia17_mm_Aug", ABCO_lm_dia))
plot(predictorEffect("ht_cm_2017", ABCO_lm_dia))

```

#### Graph with ggplot
```{r}
effects_df <- as.data.frame(predictorEffect("shrubs", ABCO_lm_dia))
```

```{r}
ggplot(effects_df, aes(x = shrubs, y = fit, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .2)+
  ggtitle("ABCO shrub removal effects")+
  ylim(c(0,.8))
```

### Graph raw data by shrub island 
```{r}
df_island <- dfABCO %>% 
  group_by(island, compartment, shrubs) %>% 
  summarize(mean_growth_rel = mean(growth_rel), n = n()) %>% 
  ungroup() %>% 
  group_by(compartment, island) 

df_island_diff <- spread(df_island %>% select(-n), key = shrubs, value = mean_growth_rel) %>% 
  arrange(island) %>% 
  mutate(higher = ifelse(removed > retained, "rem", "ret")) %>% 
  select(island, compartment, higher)

df_island_diff

df_island <- full_join(df_island, df_island_diff) %>% 
  group_by(island, compartment) %>% 
  mutate(n_island = sum(n))

ggplot(df_island %>% filter(!is.na(higher)), aes(x = shrubs, y = mean_growth_rel))+
  geom_point(aes(col = higher, alpha = n_island), size = 2)+
  geom_line(aes(group = interaction(island, compartment), alpha = n_island, col = higher), size = 1)+
  theme_bw()+
  ylab("Mean plot volumetric relative growth rate")+
  scale_color_discrete(name = "Plot with higher growth", labels = c("Removed", "Retained"))+
  scale_alpha_continuous(name = "n", range = c(.2, 1))

ggsave(file = "../../../results/figures/ShrubExperiment/PipoVol.png", width = 3, height = 5, dpi = 400)

```

## Count how many plot pairs had increases vs. decreases
```{r}
df_island %>% 
  select(island, compartment, higher) %>% 
  distinct() %>% 
  group_by(higher) %>% 
  summarize(n())

```

