---
title: "analyze_vertical_growth"
author: "Carmen"
date: "November 9, 2017"
output: 
    html_document:
        toc: TRUE
---

This code analyzes how shrub removal affected seedling vertical growth.

```{r, include = F}
library(tidyverse)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
```

# Load data

```{r}
load(file = "../../../compiled_data/shrub_removal_cleaned.Rdata")
```

# Plot
```{r}
ggplot(df %>% filter(species == "PIPO"))+
  geom_boxplot(aes(x = shrubs, y = post_removal_growth_rel))
ggplot(df %>% filter(species == "PIPO"))+
  geom_boxplot(aes(x = shrubs, y = pre_removal_growth_rel))
```


# Mixed effects linear model with island random effect

```{r}
pipo_lm <- lme(post_removal_growth_rel ~ shrubs, random = ~ 1|island, data = df %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm)
```

## Check for interaction between shrubs and height
```{r}
pipo_lm_int <- lme(post_removal_growth_rel ~ shrubs*height_2016, random = ~ 1|island, data = df %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm, pipo_lm_int)
AIC(pipo_lm_int)
AIC(pipo_lm)
```

There is definitely an interaction that I should include

## Check for diameter effect
```{r}
pipo_lm_dia <- lme(post_removal_growth_rel ~ shrubs*height_2016 + dia17_mm_Aug,random = ~ 1|island,  data = df %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm_int, pipo_lm_dia)
AIC(pipo_lm_dia)
AIC(pipo_lm_int)
```

I should definitely include diameter

## Test for shrub effect in model with interaction and dia
```{r}
pipo_lm_noshrub <- lme(post_removal_growth_rel ~ height_2016 + dia17_mm_Aug, random = ~ 1|island, data = df %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm_noshrub, pipo_lm_dia)
AIC(pipo_lm_noshrub)
AIC(pipo_lm_dia)
```

Shrub removal definitely has a negative relationship with PIPO growth

## Visual
```{r}
plot(predictorEffect("shrubs", pipo_lm_dia))
```

# REPEAT WITH BACI ANALYSIS STRUCTURE

# Restructure
```{r}
dfg <- gather(df, key = "pre_post", value = "relative_growth", c("post_removal_growth_rel", "pre_removal_growth_rel"))
dfg <- dfg %>% 
  mutate(pre_post = as.factor(paste(pre_post)))
```

```{r}
ggplot(dfg %>% filter(species == "PIPO"))+
  geom_boxplot(aes(x = interaction(shrubs, pre_post), y = relative_growth))
```

# Mixed effects linear model with island random effect

```{r}
pipo_lm <- lme(relative_growth ~ shrubs*pre_post, random = ~ 1|island, data = dfg %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm)
```

## Check for  height

```{r}
ggplot(dfg %>% filter(species == "PIPO"))+
  geom_point(aes(x = height_2016, y = relative_growth))
```

```{r}
pipo_lm_int <- lme(relative_growth ~ shrubs*pre_post + height_2016, random = ~ 1|island, data = dfg %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm, pipo_lm_int)
AIC(pipo_lm_int)
AIC(pipo_lm)
```

I do NOT need to include height

## Check for diameter effect
```{r}
pipo_lm_dia <- lme(relative_growth ~ shrubs*pre_post +  dia17_mm_Aug, random = ~ 1|island, data = dfg %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm, pipo_lm_dia)
AIC(pipo_lm_dia)
AIC(pipo_lm_int)
```

I should NOT include diameter

## Test for shrub effect in model with interaction - the true test for BACI
```{r}
pipo_lm_noshrub <- lme(relative_growth ~ pre_post + shrubs, random = ~ 1|island, data = dfg %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(pipo_lm_noshrub, pipo_lm)
AIC(pipo_lm_noshrub)
AIC(pipo_lm)
```

NO significant effect

## Visual
```{r}
plot(predictorEffect("shrubs", pipo_lm))
plot(predictorEffect("pre_post", pipo_lm))
drop1(pipo_lm_dia, test = "Chisq")
```

### Make a prettier graph in ggplot

```{r}
effects_dfg <- as.data.frame(predictorEffect("pre_post", pipo_lm))
effects_dfg <- effects_dfg %>% 
  mutate(pre_post = ifelse(pre_post == "post_removal_growth_rel", "after", "before"))
effects_dfg <- effects_dfg %>% 
  mutate(pre_post = as.factor(pre_post))
```

#### Reorder levels
```{r}
levels(effects_dfg$pre_post)
effects_dfg$pre_post = factor(effects_dfg$pre_post,levels(effects_dfg$pre_post)[c(2,1)])
levels(effects_dfg$pre_post)
```

#### Graph with reverse log transformed
```{r}
ggplot(effects_dfg, aes(x = pre_post, y = exp(fit), group = shrubs, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = exp(lower), ymax = exp(upper)), width = .2)
```

# REPEAT BACI ANALYSIS FOR FIRS

```{r}
ggplot(dfg %>% filter(species == "ABCO"))+
  geom_boxplot(aes(x = interaction(shrubs, pre_post), y = relative_growth))
```

# Mixed effects linear model with island random effect

```{r}
ABCO_lm <- lme(relative_growth ~ shrubs*pre_post, random = ~ 1|island, data = dfg %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm)
```

## Check for  height effect
```{r}
ggplot(dfg %>% filter(species == "ABCO"))+
  geom_point(aes(x = height_2016, y = relative_growth))
```

```{r}
ABCO_lm_ht <- lme(relative_growth ~ shrubs*pre_post + height_2016, random = ~ 1|island, data = dfg %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm, ABCO_lm_ht)
AIC(ABCO_lm_ht)
AIC(ABCO_lm)
```

I DO need to include height

## Check for height interaction
```{r}
ABCO_lm_int <- lme(relative_growth ~ shrubs*pre_post + shrubs*height_2016, random = ~ 1|island, data = dfg %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm_ht, ABCO_lm_int)
AIC(ABCO_lm_int)
AIC(ABCO_lm_ht)
```

I DO need to include a shrub*height interaction


## Check for diameter effect
```{r}
ABCO_lm_dia <- lme(relative_growth ~ shrubs*pre_post +height_2016 +dia17_mm_Aug, random = ~ 1|island, data = dfg %>% filter(species == "ABCO" & !is.na(shrubs)), method = "ML")
anova(ABCO_lm_ht, ABCO_lm_dia)
AIC(ABCO_lm_ht)
AIC(ABCO_lm_int)
```

I should NOT include diameter

## Test for shrub effect in model with interaction - the true test for BACI
```{r}
ABCO_lm_noshrub <- lme(relative_growth ~ pre_post + shrubs+height_2016, random = ~ 1|island, data = dfg %>% filter(species == "ABCO" & !is.na(shrubs)), method = "ML")
AIC(ABCO_lm_noshrub)
AIC(ABCO_lm_ht)
anova(ABCO_lm_noshrub, ABCO_lm_ht)
```

NO significant effect

## Visual
```{r}
plot(predictorEffect("shrubs", ABCO_lm_ht))
plot(predictorEffect("pre_post", ABCO_lm_ht))
drop1(ABCO_lm_ht, test = "Chisq")
```

### Make a prettier graph in ggplot

```{r}
effects_dfg <- as.data.frame(predictorEffect("pre_post", ABCO_lm_ht))
effects_dfg <- effects_dfg %>% 
  mutate(pre_post = ifelse(pre_post == "post_removal_growth_rel", "after", "before"))
effects_dfg <- effects_dfg %>% 
  mutate(pre_post = as.factor(pre_post))
```

#### Reorder levels
```{r}
levels(effects_dfg$pre_post)
effects_dfg$pre_post = factor(effects_dfg$pre_post,levels(effects_dfg$pre_post)[c(2,1)])
levels(effects_dfg$pre_post)
```

#### Graph with reverse log transformed
```{r}
ggplot(effects_dfg, aes(x = pre_post, y = exp(fit), group = shrubs, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = exp(lower), ymax = exp(upper)), width = .2)+
  ggtitle("ABCO shrub removal effects")+
  xlab("timing relative to shrub removal")
```

# REPEAT BACI ANALYSIS FOR FIRS for 2018 GROWTH

# Restructure
```{r}
dfg2 <- gather(df, key = "pre_post", value = "relative_growth", c("growth_rel_2018", "pre_removal_growth_rel"))
dfg2 <- dfg2 %>% 
  mutate(pre_post = as.factor(paste(pre_post)))
```

```{r}
ggplot(dfg2 %>% filter(species == "ABCO"))+
  geom_boxplot(aes(x = interaction(shrubs, pre_post), y = relative_growth))
```

# Mixed effects linear model with island random effect

```{r}
ABCO_lm <- lme(relative_growth ~ shrubs*pre_post, random = ~ 1|island, data = dfg2 %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm)
```

## Check for  height effect
```{r}
ggplot(dfg2 %>% filter(species == "ABCO"))+
  geom_point(aes(x = height_2016, y = relative_growth))
```

```{r}
ABCO_lm_ht <- lme(relative_growth ~ shrubs*pre_post + height_2016, random = ~ 1|island, data = dfg2 %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm, ABCO_lm_ht)
AIC(ABCO_lm_ht)
AIC(ABCO_lm)
```

I DO need to include height

## Check for height interaction
```{r}
ABCO_lm_int <- lme(relative_growth ~ shrubs*pre_post + shrubs*height_2016, random = ~ 1|island, data = dfg2 %>% filter(species == "ABCO"), method = "ML")
anova(ABCO_lm_ht, ABCO_lm_int)
AIC(ABCO_lm_int)
AIC(ABCO_lm_ht)
```

I DO need to include a shrub*height interaction


## Check for diameter effect
```{r}
ABCO_lm_dia <- lme(relative_growth ~ shrubs*pre_post +height_2016 +dia17_mm_Aug, random = ~ 1|island, data = dfg2 %>% filter(species == "ABCO" & !is.na(shrubs)), method = "ML")
anova(ABCO_lm_ht, ABCO_lm_dia)
AIC(ABCO_lm_ht)
AIC(ABCO_lm_dia)
```

I should NOT include diameter

## Test for shrub effect in model with interaction - the true test for BACI
```{r}
ABCO_lm_noshrub <- lme(relative_growth ~ pre_post + shrubs+height_2016, random = ~ 1|island, data = dfg2 %>% filter(species == "ABCO" & !is.na(shrubs)), method = "ML")
AIC(ABCO_lm_noshrub)
AIC(ABCO_lm_ht)
anova(ABCO_lm_noshrub, ABCO_lm_ht)
```

NO significant effect

## Visual
```{r}
plot(predictorEffect("shrubs", ABCO_lm_ht))
plot(predictorEffect("pre_post", ABCO_lm_ht))
drop1(ABCO_lm_ht, test = "Chisq")
```

### Make a prettier graph in ggplot

```{r}
effects_dfg2 <- as.data.frame(predictorEffect("pre_post", ABCO_lm_ht))
effects_dfg2 <- effects_dfg2 %>% 
  mutate(pre_post = ifelse(pre_post == "pre_removal_growth_rel", "before", "after"))
effects_dfg2 <- effects_dfg2 %>% 
  mutate(pre_post = as.factor(pre_post))
```

#### Reorder levels
```{r}
levels(effects_dfg2$pre_post)
effects_dfg2$pre_post = factor(effects_dfg2$pre_post,levels(effects_dfg2$pre_post)[c(2,1)])
levels(effects_dfg2$pre_post)
```

#### Graph with reverse log transformed
```{r}
ggplot(effects_dfg2, aes(x = pre_post, y = exp(fit), group = shrubs, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = exp(lower), ymax = exp(upper)), width = .2)+
  ggtitle("ABCO shrub removal effects")+
  xlab("timing relative to shrub removal")
```


# REPEAT BACI ANALYSIS FOR PINES for 2018 GROWTH

# Restructure
```{r}
dfg2 <- gather(df, key = "pre_post", value = "relative_growth", c("growth_rel_2018", "pre_removal_growth_rel"))
dfg2 <- dfg2 %>% 
  mutate(pre_post = as.factor(paste(pre_post)))
```

```{r}
ggplot(dfg2 %>% filter(species == "PIPO"))+
  geom_boxplot(aes(x = interaction(shrubs, pre_post), y = relative_growth))
```

# Mixed effects linear model with island random effect

```{r}
PIPO_lm <- lme(relative_growth ~ shrubs*pre_post, random = ~ 1|island, data = dfg2 %>% filter(species == "PIPO"), method = "ML")
anova(PIPO_lm)
```

## Check for  height effect
```{r}
ggplot(dfg2 %>% filter(species == "PIPO"))+
  geom_point(aes(x = height_2016, y = relative_growth))
```

```{r}
PIPO_lm_ht <- lme(relative_growth ~ shrubs*pre_post + height_2016, random = ~ 1|island, data = dfg2 %>% filter(species == "PIPO"), method = "ML")
anova(PIPO_lm, PIPO_lm_ht)
AIC(PIPO_lm_ht)
AIC(PIPO_lm)
```

I DO need to include height

## Check for height interaction
```{r}
PIPO_lm_int <- lme(relative_growth ~ shrubs*pre_post + shrubs*height_2016, random = ~ 1|island, data = dfg2 %>% filter(species == "PIPO"), method = "ML")
anova(PIPO_lm_ht, PIPO_lm_int)
AIC(PIPO_lm_int)
AIC(PIPO_lm_ht)
```

I DO NOT need to include a shrub*height interaction


## Check for diameter effect
```{r}
PIPO_lm_dia <- lme(relative_growth ~ shrubs*pre_post +height_2016 +dia17_mm_Aug, random = ~ 1|island, data = dfg2 %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
anova(PIPO_lm_ht, PIPO_lm_dia)
AIC(PIPO_lm_ht)
AIC(PIPO_lm_dia)
```

I SHOULD include diameter

## Test for shrub effect in model with interaction - the true test for BACI
```{r}
PIPO_lm_noshrub <- lme(relative_growth ~ pre_post + shrubs+height_2016 +dia17_mm_Aug, random = ~ 1|island, data = dfg2 %>% filter(species == "PIPO" & !is.na(shrubs)), method = "ML")
AIC(PIPO_lm_noshrub)
AIC(PIPO_lm_dia)
anova(PIPO_lm_noshrub, PIPO_lm_dia)
```

NO significant effect

## Visual
```{r}
plot(predictorEffect("shrubs", PIPO_lm_dia))
plot(predictorEffect("pre_post", PIPO_lm_dia))
drop1(PIPO_lm_dia, test = "Chisq")
```

### Make a prettier graph in ggplot

```{r}
effects_dfg2 <- as.data.frame(predictorEffect("pre_post", PIPO_lm_dia))
effects_dfg2 <- effects_dfg2 %>% 
  mutate(pre_post = ifelse(pre_post == "pre_removal_growth_rel", "before", "after"))
effects_dfg2 <- effects_dfg2 %>% 
  mutate(pre_post = as.factor(pre_post))
```

#### Reorder levels
```{r}
levels(effects_dfg2$pre_post)
effects_dfg2$pre_post = factor(effects_dfg2$pre_post,levels(effects_dfg2$pre_post)[c(2,1)])
levels(effects_dfg2$pre_post)
```

#### Graph with reverse log transformed
```{r}
ggplot(effects_dfg2, aes(x = pre_post, y = exp(fit), group = shrubs, col = shrubs))+
  geom_point()+
  geom_line()+
  ylab("predicted relative growth rate")+
  geom_errorbar(aes(ymin = exp(lower), ymax = exp(upper)), width = .2)+
  ggtitle("PIPO shrub removal effects")+
  xlab("timing relative to shrub removal")
```
