---
title: "Fir volumetric growth analysis - shrub experiment"
author: "Carmen"
date: "November 14, 2017"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(dplyr)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
```


# Load final data frames
```{r}
load(file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia_final.Rdata")
df_dia <- df
remove(df)
load(file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_vert_final.Rdata")
df_vert <- df
remove(df)
```

# Filter to only seedlings with workable data for both diameter and vertical growth
```{r}
df_vert$leader_damage <- as.character(df_vert$leader_damage)
df <- inner_join(df_dia, df_vert)
```

# Clean up extreme and negative values

## Take out outliers on both ends, defined as greater than two SD from the mean
```{r}
nrow(df)
hist(df$dia_growth_mm, breaks = 20)
mean_minSD <- mean(df$dia_growth_mm) - 1.5*sd(df$dia_growth_mm)
mean_minSD
mean_plusSD <- mean(df$dia_growth_mm) + 1.5*sd(df$dia_growth_mm)
mean_plusSD
df <- df %>% 
  filter(!(dia_growth_mm < mean_minSD |
             dia_growth_mm > mean_plusSD))
hist(df$dia_growth_mm)
nrow(df)
```

## Adjust  so everything's above 0
```{r}
hist(df$Aug_dia_ave_mm)
df <- df %>% 
  mutate(Aug_dia_ave_mm = Aug_dia_ave_mm - min(dia_growth_mm)) %>% 
  mutate(dia_growth_mm = (dia_growth_mm - min(dia_growth_mm))) 
hist(df$Aug_dia_ave_mm)  
hist(df$dia_growth_mm)
```

# Calculate relative growth rate

## Calculate pre volume
```{r}
df <- df %>% 
  mutate(vol_pre_cm2 = ((spring_dia_ave_mm/10)^2) * (ht_cm - gr_2017_cm) * (pi/12))
hist(df$vol_pre_cm2)
```

## Calculate post volume
```{r}
df <- df %>% 
  mutate(vol_post_cm2 = ((Aug_dia_ave_mm/10)^2) * ht_cm * (pi/12))
hist(df$vol_post_cm2)
```

## Check
```{r}
hist(log(df$vol_post_cm2))
ggplot(df)+
  geom_point(aes(x = vol_pre_cm2, y = vol_post_cm2))+
  scale_x_continuous(limits = c(0,50))+
  scale_y_continuous(limits = c(0,50))+
  geom_abline(slope = 1,intercept = 0)
ggplot(df %>% filter(species == "ABCO"))+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_pre_cm2)-log(vol_post_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)

ggplot(df %>% filter(species == "PIPO"))+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_pre_cm2)-log(vol_post_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)
ggplot(df %>% filter(species == "ABCO"))+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_pre_cm2)-log(vol_post_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)
ggplot(df %>% filter(species == "PIPO"))+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_pre_cm2)-log(vol_post_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)
```

NOTE: This plot looks very different if I set the outlier cutoff to 2xSD, making it so that all the values are adjusted more to the right. I chose to stick with the 1xSD cutoff for that reason. 

## Check
```{r}
df[14,] %>% select(spring_dia_ave_mm, ht_cm, vol_pre_cm2)
(df[14,"spring_dia_ave_mm"]/10)^2 * (df[14,"ht_cm"] -  df[14,"gr_2017_cm"]) * (pi/12)

df[12,] %>% select(Aug_dia_ave_mm, ht_cm, vol_post_cm2)
(df[12,"Aug_dia_ave_mm"]/10)^2 * df[12,"ht_cm"] * (pi/12)

```

## Calculate RGR
```{r}
df <- df %>% 
  mutate(RGR = (log(vol_post_cm2)- log(vol_pre_cm2)))
hist(df$RGR, breaks = 20)
hist(log(df$RGR), breaks = 20)
```

### Take a look

```{r}
ggplot(df)+
  geom_boxplot(aes(x = interaction(shrubs, species), y = RGR))
```


```{r}
ggplot(df)+
  geom_point(aes(x = vol_pre_cm2, y = RGR))+
  scale_x_continuous(limits = c(0,300))
ggplot(df)+
  geom_point(aes(x = vol_pre_cm2, y = log(RGR)))+
  scale_x_continuous(limits = c(0,300))
```


```{r}
ggplot(df)+
  geom_boxplot(aes(x = interaction(shrubs, species), y = log(RGR)))
```

```{r}
ggplot(df, aes(x = ht_cm, y = log(RGR)))+
  geom_point()+
  stat_smooth(method = "gam")
```

Log transformation looks better than no log transformation:

```{r}
ggplot(df, aes(x = ht_cm, y = RGR))+
  geom_point()+
  stat_smooth(method = "gam")
```



# Plot mean and SE RGR for both species
```{r}
df_summary <- df %>% 
  group_by(shrubs, species) %>% 
  summarise(mean_RGR = mean(RGR), sd= sd(RGR), n = n(), SE = sd(RGR)/sqrt(n()))
df_summary
figure_meanSE <- ggplot(df_summary, aes(x = interaction(shrubs, species), col = species))+
  geom_errorbar(aes(ymin = mean_RGR-SE, ymax = mean_RGR+SE), width = .1)+
  geom_point(aes(y = mean_RGR), size = 2)+
  theme_bw()+
  theme(legend.position = c(.8,.8),
        text = element_text(size=10)
        )+
  ylab("relative growth rate (RGR)")+
  xlab("")+
  scale_x_discrete(labels = c("shrubs\nremoved","control","shrubs\nremoved", "control"))+
  scale_colour_manual(
    name = "Species",
    labels = c("fir", "pine"),
    values = c("#1b3f4c","#a02313")
    )
  
    
figure_meanSE
```

Save this figure
```{r}
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
pdf("meansSE.pdf", width = 3, height = 3, useDingbats = F)
figure_meanSE
dev.off()
```


# Standardize continuous variables I plan to use in the model
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

## Analyze RGR in relation to shrub removal
```{r}
df_abco <- df %>% filter(species == "ABCO")
```


```{r}
df_abco <- df_abco %>% 
  mutate(ht_norm = normalize(ht_cm)) %>% 
  mutate(shrubarea_sqrt_norm = normalize(sqrt(shrubarea)) ) %>% 
  mutate(ht_log_norm = normalize(log(ht_cm)))
```

```{r}
ggplot(df_abco, aes(x = ht_cm, y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
ggplot(df_abco, aes(x = ht_norm, y = log(RGR)))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
ggplot(df_abco, aes(x = ht_log_norm, y = log(RGR)))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
```

The log relationship between ht and RGR means that we have to transform one of them


# Stats

THESE ARE SUPER WACKY! NEED TO FIX EVERYTHING

### Mixed effects with island effect
```{r}
MM1_abco <- lme(RGR ~ shrubs + ht_log_norm + shrub_species, random =~1|island, data = df_abco, method = "ML", weights = varExp(form=~ht_norm))
MM1_null_abco <- lme(RGR ~ ht_log_norm + shrub_species, random =~1|island, method = "ML",data = df_abco, weights = varExp(form=~ht_norm))
anova(MM1_abco, MM1_null_abco)
summary(MM1_abco)
```


```{r}
df_pipo <- df %>% filter(species == "PIPO")
```


```{r}
df_pipo <- df_pipo %>% 
  mutate(ht_norm = normalize(ht_cm)) %>% 
  mutate(shrubarea_sqrt_norm = normalize(sqrt(shrubarea)) ) %>% 
  mutate(ht_log_norm = normalize(log(ht_cm)))
```

The log relationship between ht and RGR means that we have to transform one of them


### Mixed effects with island effect - PIPO
```{r}
MM1_pipo <- lme(RGR ~ shrubs + ht_log_norm + shrub_species + shrubs:ht_log_norm, 
                data = df_pipo, 
                random =~ 1 | island, 
                weights = varExp(form=~ht_log_norm),
                method = "ML")
MM1_null_pipo <- lme(RGR ~ ht_log_norm + shrub_species, 
                     data = df_pipo,
                     weights = varExp(form=~ht_log_norm),
                     random =~ 1 | island, 
                method = "ML")
anova(MM1_pipo, MM1_null_pipo)
summary(MM1_pipo)
AIC(MM1_pipo)
AIC(MM1_null_pipo)
```

### Simple linear model - PIPO
### Mixed effects with island effect - PIPO
```{r}
MM2_pipo <- lm(RGR ~ shrubs + ht_log_norm + shrub_species + shrubs:ht_log_norm, 
                data = df_pipo)
                
MM2_null_pipo <- lm(RGR ~ ht_log_norm + shrub_species, data = df_pipo)

anova(MM2_pipo, MM2_null_pipo)
summary(MM1_pipo)
AIC(MM2_pipo)
AIC(MM2_null_pipo)
```




### Residuals
```{r}
E <- resid(MM1_pipo, type = "normalized")
pred <- predict(MM1_pipo)
plot(MM1_pipo)

plot(df[df$species == "PIPO","shrubs"], E)
plot(df[df$species == "PIPO","shrub_species"], E)
plot(df_pipo$ht_log_norm, E)
```

<!-- # Plot predictions -->

<!-- ### Create MyData -->
<!-- ```{r} -->
<!-- MyData_abco <- expand.grid(shrubs = c("retained", "removed"), ht_norm = c(seq(min(df$ht_norm), max(df$ht_norm), length = 20)), shrub_species = c("CEIN","CECO","ARPA"), shrubarea_sqrt_norm= seq(min(df$shrubarea_aqrt_norm, max(df$shrubarea_sqrt_norm)), length = 20), island = unique(df$island)) -->
<!-- ``` -->

<!-- #### Take out MyData_abco for islands without ABCO -->
<!-- ```{r} -->
<!-- MyData_abco <- filter(MyData_abco, island != "A") -->
<!-- ``` -->


<!-- ### Add pred -->
<!-- ```{r} -->
<!-- MyData_abco$predMM <- predict(MM1_abco, newdata = MyData_abco) -->
<!-- ``` -->

<!-- ## Overall effect of removal on growth -->
<!-- ```{r} -->
<!-- MyData_summary_abco <- MyData_abco %>%  -->
<!--   group_by(shrubs, shrub_species, island) %>%  -->
<!--   summarise(mean_predMM = mean(predMM)) -->
<!-- MyData_summary_abco -->
<!-- figure <- ggplot(MyData_summary_abco,aes(x = shrubs, y = mean_predMM)) + -->
<!--    geom_boxplot(data = df %>% filter(species == "ABCO"), aes(x = shrubs, y = RGR)) + -->
<!--   geom_point(aes(col = shrub_species, shape = shrub_species), size = 2)+ -->
<!--   geom_line(aes(col = shrub_species, linetype = shrub_species, group = interaction(shrub_species, island)))+ -->
<!--   theme_bw()+ -->
<!--   scale_color_manual(values=c("#1a9850", "#d73027", "pink"))+ -->
<!--   labs(y = "Predicted relative growth rate")+ -->
<!--   ggtitle("Shrub removal increased relative growth rate") -->

<!-- figure -->
<!-- setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/") -->
<!-- png("RGR.png") -->
<!-- figure -->
<!-- dev.off() -->
<!-- ``` -->

<!-- ### Plot effect of height -->
<!-- ```{r} -->
<!-- MyData_abco$ht_norm <- as.numeric(paste(MyData_abco$ht_norm)) -->
<!-- MyData_abco_height_summary <-  MyData_abco %>%  -->
<!--   group_by(shrub_species, shrubs, ht_norm) %>%  -->
<!--   summarise(predMM = mean(predMM)) -->
<!-- MyData_abco_height_summary -->
<!-- ggplot(MyData_abco_height_summary)+ -->
<!--   geom_line(aes(x = ht_norm, y = predMM, group = interaction(shrubs, shrub_species), col = shrub_species, linetype = shrubs))+ -->
<!--   theme_bw() -->
<!-- ``` -->

<!-- The taller the seedling, the smaller the predicted growth rate. I think there's a serious GLS issue here.  -->

<!-- # GLS analysis, no island effect -->

<!-- ## Which is better, varExp or VarPower? (test according to Zuur Ch. 4) -->
<!-- ```{r} -->
<!-- G1_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species+ ht_norm:shrubs,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "REML") -->
<!-- G2_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species+ ht_norm:shrubs,  -->
<!--                weights = varPower(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "REML") -->
<!-- AIC(G1_abco) -->
<!-- AIC(G2_abco) -->

<!-- E1 <- resid(G1_abco, type = "normalized") -->
<!-- plot(df[df$species == "ABCO","ht_cm"], E1) -->
<!-- E2 <- resid(G2_abco, type = "normalized") -->
<!-- plot(df[df$species == "ABCO","ht_cm"], E2) -->
<!-- ``` -->

<!-- VarExp is definitely better! -->

<!-- ## See what nonessential predictors are helpful -->

<!-- ### ht:removal interaction? -->
<!-- ```{r} -->
<!-- G1_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "ML") -->
<!-- G3_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species+ ht_norm:shrubs,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "ML") -->
<!-- AIC(G1_abco) -->
<!-- AIC(G3_abco) -->
<!-- anova(G1_abco, G3_abco) -->
<!-- ``` -->

<!-- Interaction between ht and removal does not help! -->

<!-- ### species:removal interaction? -->
<!-- ```{r} -->
<!-- G4_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species + shrub_species:shrubs,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "ML") -->
<!-- AIC(G1_abco) -->
<!-- AIC(G4_abco) -->
<!-- anova(G1_abco, G4_abco) -->
<!-- ``` -->

<!-- Interaction between shrub species and removal does not help! -->

<!-- ### shrubarea? -->

<!-- ```{r} -->
<!-- G5_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species + shrubarea_sqrt_norm,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "ML") -->
<!-- AIC(G1_abco) -->
<!-- AIC(G5_abco) -->
<!-- anova(G1_abco, G5_abco) -->
<!-- ``` -->

<!-- Shrub area does not help! -->

<!-- ## Compare to null model -->
<!-- ```{r} -->
<!-- G1_abco <- lme(log(RGR) ~ ht_norm + shrubs + shrub_species,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "ML") -->
<!-- G1_abco_null <- lme(log(RGR) ~ ht_norm  + shrub_species,  -->
<!--                weights = varExp(form=~ht_norm), -->
<!--                random =~ 1 | island,  -->
<!--                data = df %>% filter(species == "ABCO"), -->
<!--                method = "ML") -->
<!-- summary(G1_abco) -->
<!-- anova(G1_abco, G1_abco_null) -->
<!-- ``` -->


<!-- ## Plot -->

<!-- ### Add pred -->
<!-- ```{r} -->
<!-- MyData_abco$predMM_gls <- predict(G1_abco, newdata = MyData_abco) -->
<!-- ``` -->

<!-- ## Overall effect of removal on diameter growth -->
<!-- ```{r} -->
<!-- MyData_summary_abco <- MyData_abco %>%  -->
<!--   group_by(shrubs, shrub_species, island) %>%  -->
<!--   summarise(mean_predMM_gls = mean(predMM_gls)) -->
<!-- MyData_summary_abco -->
<!-- figure <- ggplot(MyData_summary_abco, -->
<!--                  aes(x = shrubs, y = mean_predMM_gls)) + -->
<!--    geom_boxplot(data = df %>% filter(species == "ABCO"),  -->
<!--                 aes(x = shrubs, y = RGR)) + -->
<!--   geom_point( -->
<!--     aes(col = shrub_species, shape = shrub_species), size = 2)+ -->
<!--   geom_line( -->
<!--     aes(col = shrub_species,  -->
<!--         linetype = shrub_species,  -->
<!--         group = interaction(shrub_species, island)))+ -->
<!--   theme_bw()+ -->
<!--   scale_color_manual(values=c("#1a9850", "#d73027", "pink"))+ -->
<!--   labs(y = "Predicted relative growth rate")+ -->
<!--   ggtitle("Shrub removal increased relative growth rate") -->

<!-- figure -->
<!-- setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/") -->
<!-- png("RGR.png") -->
<!-- figure -->
<!-- dev.off() -->
<!-- ``` -->

<!-- ## Residuals -->

<!-- ```{r} -->
<!-- E <- resid(G1_abco, type = "normalized") -->
<!-- pred <- predict(G1_abco) -->
<!-- plot(G1_abco) -->
<!-- plot(df[df$species == "ABCO","shrubs"], E) -->
<!-- plot(df[df$species == "ABCO","shrub_species"], E) -->
<!-- ``` -->


