---
title: "Pine volumetric growth analysis - shrub experiment"
author: "Carmen"
date: "November 15, 2017"
output: html_document
---
---
title: "Volumetric growth analysis - shrub experiment"
author: "Carmen"
date: "November 14, 2017"
output: html_document
---

```{r, include = F}
library(dplyr)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
```


# Load final data frames
```{r}
load(file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia_final.Rdata")
df_dia <- df
remove(df)
load(file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_vert_final.Rdata")
df_vert <- df
remove(df)
```

# Filter to only seedlings with workable data for both diameter and vertical growth
```{r}
df_vert$leader_damage <- as.character(df_vert$leader_damage)
df <- inner_join(df_dia, df_vert)
```

# Clean up extreme and negative values

## Take out outliers on both ends, defined as greater than two SD from the mean
```{r}
nrow(df)
hist(df$dia_growth_mm)
mean_minSD <- mean(df$dia_growth_mm) - 2*sd(df$dia_growth_mm)
mean_plusSD <- mean(df$dia_growth_mm) + 2*sd(df$dia_growth_mm)
df <- df %>% 
  filter(!(dia_growth_mm < mean_minSD |
             dia_growth_mm > mean_plusSD))
hist(df$dia_growth_mm)
nrow(df)
```

## Adjust  so everything's above 0
```{r}
df <- df %>% 
  mutate(dia_growth_mm = (dia_growth_mm - min(dia_growth_mm)))
```


# Calculate relative growth rate

## Calculate pre volume
```{r}
df <- df %>% 
  mutate(vol_pre_cm2 = ((spring_dia_ave_mm/10)^2) * (ht_cm - gr_2017_cm) * (pi/12))
hist(df$vol_pre_cm2)
```

## Calculate post volume
```{r}
df <- df %>% 
  mutate(vol_post_cm2 = ((Aug_dia_ave_mm/10)^2) * ht_cm * (pi/12))
hist(df$vol_post_cm2)
```

## Check
```{r}
df[14,] %>% select(spring_dia_ave_mm, ht_cm, vol_pre_cm2)
(df[14,"spring_dia_ave_mm"]/10)^2 * (df[14,"ht_cm"] -  df[14,"gr_2017_cm"]) * (pi/12)

df[12,] %>% select(Aug_dia_ave_mm, ht_cm, vol_post_cm2)
(df[12,"Aug_dia_ave_mm"]/10)^2 * df[12,"ht_cm"] * (pi/12)

```

## Calculate RGR
```{r}
df <- df %>% 
  mutate(RGR = (log(vol_post_cm2)- log(vol_pre_cm2)))
hist(df$RGR)
```

### Take a look
```{r}
ggplot(df)+
  geom_boxplot(aes(x = interaction(shrubs, species), y = RGR))
```

```{r}
ggplot(df, aes(x = ht_cm, y = RGR))+
  geom_point()+
  stat_smooth(method = "gam")
```


# Standardize continuous variables I plan to use in the model
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

```{r}
df <- df %>% 
  mutate(ht_norm = normalize(ht_cm)) %>% 
  mutate(shrubarea_sqrt_norm = normalize(sqrt(shrubarea)) )
```

## Take out CEIN
```{r}
df <- df %>% filter(shrub_species != "CEIN")
```


# Stats

## Analyze RGR in relation to shrub removal

### Mixed effects with island effect
```{r}
MM1_pipo <- lme(RGR ~ shrubs + ht_norm + shrub_species + ht_norm:shrubs, random =~1|island, data = df %>% filter(species == "PIPO"), method = "ML")
MM1_null_pipo <- lme(RGR ~ ht_norm + shrub_species + ht_norm:shrubs, random =~1|island, method = "ML",data = df %>%  filter(species == "PIPO"))
anova(MM1_pipo, MM1_null_pipo)
summary(MM1_pipo)
```

### Residuals
```{r}
E <- resid(MM1_pipo, type = "normalized")
pred <- predict(MM1_pipo)
plot(MM1_pipo)
plot(df[df$species == "PIPO","shrubs"], E)
plot(df[df$species == "PIPO","shrub_species"], E)
plot(df[df$species == "PIPO","ht_norm"], E)
```

# Plot predictions

### Create MyData
```{r}
MyData_pipo <- expand.grid(shrubs = c("retained", "removed"), ht_norm = c(seq(min(df$ht_norm), max(df$ht_norm), length = 20)), shrub_species = c("CECO","ARPA"), shrubarea_sqrt_norm= seq(min(df$shrubarea_aqrt_norm, max(df$shrubarea_sqrt_norm)), length = 20), island = unique(df$island))
```

#### Take out MyData_pipo for islands without PIPO
```{r}
MyData_pipo <- filter(MyData_pipo, island != "A")
```


### Add pred
```{r}
MyData_pipo$predMM <- predict(MM1_pipo, newdata = MyData_pipo)
```

## Overall effect of removal on diameter growth
```{r}
MyData_summary_pipo <- MyData_pipo %>% 
  group_by(shrubs, shrub_species, island) %>% 
  summarise(mean_predMM = mean(predMM))
MyData_summary_pipo
figure <- ggplot(MyData_summary_pipo,aes(x = shrubs, y = mean_predMM)) +
   geom_boxplot(data = df %>% filter(species == "PIPO"), aes(x = shrubs, y = RGR)) +
  geom_point(aes(col = shrub_species, shape = shrub_species), size = 2)+
  geom_line(aes(col = shrub_species, linetype = shrub_species, group = interaction(shrub_species, island)))+
  theme_bw()+
  scale_color_manual(values=c("#1a9850", "#d73027", "pink"))+
  labs(y = "Predicted relative growth rate")+
  ggtitle("Shrub removal increased relative growth rate")
 
figure
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
png("RGR.png")
figure
dev.off()
```

### Plot effect of height
```{r}
MyData_pipo$ht_norm <- as.numeric(paste(MyData_pipo$ht_norm))
MyData_pipo_height_summary <-  MyData_pipo %>% 
  group_by(shrub_species, shrubs, ht_norm) %>% 
  summarise(predMM = mean(predMM))
MyData_pipo_height_summary
ggplot(MyData_pipo_height_summary)+
  geom_line(aes(x = ht_norm, y = predMM, group = interaction(shrubs, shrub_species), col = shrub_species, linetype = shrubs))+
  theme_bw()
```

The taller the seedling, the smaller the predicted growth rate. I think there's a serious GLS issue here. 

# GLS analysis, no island effect

## Which is better, varExp or VarPower? (test according to Zuur Ch. 4)
```{r}
G1_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species+ ht_norm:shrubs, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "REML")
G2_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species+ ht_norm:shrubs, 
               weights = varPower(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "REML")
AIC(G1_pipo)
AIC(G2_pipo)

E1 <- resid(G1_pipo, type = "normalized")
plot(df[df$species == "PIPO","ht_cm"], E1)
E2 <- resid(G2_pipo, type = "normalized")
plot(df[df$species == "PIPO","ht_cm"], E2)
```

VarExp is definitely better!

## See what nonessential predictors are helpful

### ht:removal interaction?
```{r}
G1_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "ML")
G3_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species+ ht_norm:shrubs, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "ML")
AIC(G1_pipo)
AIC(G3_pipo)
anova(G1_pipo, G3_pipo)
```

Interaction between ht and removal does not help!

### species:removal interaction?
```{r}
G4_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species + shrub_species:shrubs, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "ML")
AIC(G1_pipo)
AIC(G4_pipo)
anova(G1_pipo, G4_pipo)
```

Interaction between shrub species and removal helps if CEIN is included, but otherwise does not. 

### shrubarea?

```{r}
G5_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species +  shrubarea_sqrt_norm, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "ML")
AIC(G1_pipo)
AIC(G5_pipo)
anova(G1_pipo, G5_pipo)
```

Shrub area maybe helps.

## Compare to null model
```{r}
G1_pipo <- lme(RGR ~ ht_norm + shrubs + shrub_species, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "ML")
G1_pipo_null <- lme(RGR ~ ht_norm  + shrub_species, 
               weights = varExp(form=~ht_norm),
               random =~ 1 | island, 
               data = df %>% filter(species == "PIPO"),
               method = "ML")
summary(G1_pipo)
anova(G1_pipo, G1_pipo_null)
```


## Plot

### Add pred
```{r}
MyData_pipo$predMM_gls <- predict(G1_pipo, newdata = MyData_pipo)
```

## Overall effect of removal on diameter growth
```{r}
MyData_summary_pipo <- MyData_pipo %>% 
  group_by(shrubs, shrub_species, island) %>% 
  summarise(mean_predMM_gls = mean(predMM_gls))
MyData_summary_pipo
figure <- ggplot(MyData_summary_pipo,aes(x = shrubs, y = mean_predMM_gls)) +
  geom_point(aes(col = shrub_species, shape = shrub_species), size = 2)+
  geom_line(aes(col = shrub_species, linetype = shrub_species, group = interaction(shrub_species, island)))+
  theme_bw()+
  scale_color_manual(values=c("#1a9850", "#d73027", "pink"))+
  labs(y = "Predicted relative growth rate")+
  ggtitle("Shrub removal increased relative growth rate")
 
figure
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
png("RGR.png")
figure
dev.off()
```

# The effect is reverse for CEIN so I took that species out

## Residuals
```{r}
E <- resid(G1_pipo, type = "normalized")
pred <- predict(G1_pipo)
plot(G1_pipo)
plot(df[df$species == "PIPO","shrubs"], E)
plot(df[df$species == "PIPO","shrub_species"], E)
```


# Boxplots
```{r}
ggplot(df %>% filter(species == "PIPO" & ht_cm > median(df$ht_cm)))+
  geom_boxplot(aes(x = interaction(shrubs, shrub_species), y = RGR))+
  geom_point(aes(x = interaction(shrubs, shrub_species), y = mean(RGR)), col = "red")
```

# Check for outliers with high RGR!