---
title: "Shrub experiment seedling data - cleaning"
author: "Carmen"
date: "November 7, 2017"
output: 
    html_document:
        toc: TRUE
---

This code compiles and cleans seedling diameter, height, and vertical growth data. It then combines these data with treatment, shrub species, and compartment data. Later code makes pretty graphs and analyses with the cleaned and combined data.

```{r, include = F}
library(dplyr)
library(readxl)
library(ggplot2)
```


# Pull in data
```{r}
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/Compiled/")
dfsp_plots <- read_excel("Spring_seedling_data_compiled_20171107.xlsx", sheet = 1)
dfsp <- read_excel("Spring_seedling_data_compiled_20171107.xlsx", sheet = "dia")
dfAu <- read_excel("Aug_seedling_data_compiled_110717.xlsx", sheet = "data")
```

# Make all island names uppercase
```{r}
dfsp <- dfsp %>% mutate(island = toupper(island))
dfAu <- dfAu %>% mutate(island = toupper(island))
```

# Overall cleaning not specific to vert or dia
 
## Combine spring and August data
```{r}
dfsp
nrow(dfsp)
dfsp %>% group_by(compartment, island) %>% summarise(n())
dfAu
nrow(dfAu)
str(dfAu)
dfAu %>% group_by(compartment, island) %>% summarise(n())
df <- full_join(dfsp, dfAu, by = c("compartment", "island", "plot", "seedling tag #"))
nrow(df)
str(df)
```

## Rename columns
```{r}
df <- df %>% 
  rename(spring_dia_1_mm = `dm at base 1 (mm)`) %>% 
  rename(spring_dia_2_mm = `dm at base 2 (mm)`) %>% 
  rename(Aug_dia_1_mm = bas_dia_mm_1) %>% 
  rename(Aug_dia_2_mm = bas_dia_mm_2) %>% 
  rename(growing = `growing yet.`) %>% 
  rename(sdlg = `seedling tag #`) %>% 
  rename(notes.spring = notes.x) %>% 
  rename(notes.Aug = notes.y) %>% 
  rename(leader_damage = `leader broke or died`) %>% 
  select(-`DIA/ht check`)
df
```

## Remove unnecessary columns

### Column for diameter at 10 cm (very partial)
```{r}
df <- df %>% 
  select(-`dm at 10 (1)`, -`dm at 10 (2)`)
```
 
### Growing column (very partial)
```{r}
df <- df %>% 
  mutate(growing = toupper(growing)) %>% 
  mutate(growing = ifelse(growing == "Y", "YES", growing))
```
 
### Duplicate species columns 

#### Check that species match
```{r}
df %>% filter(species.x != species.y) %>% select(compartment, island, plot, sdlg, species.x, species.y, notes.Aug, notes.spring)
```

For 180-G-1, ABCO is correct based on the note from August 2017. For 180-C-2-109, PIPO is correct based on note from Aug 2018. Therefore, the column species.y has correct species

#### Remove species.x column since species.y is more accurate
```{r}
df <- df %>% select(-species.x) %>% rename(species = species.y)
```
 
### For the column for height before 2017 growth, there are only 11 seedlings with data. Check those.
```{r}
df_check_ht <- df %>% 
  filter(!is.na(`height (before 2017 growth)`)) %>% 
  filter(is.na(notes.Aug) | notes.Aug != "thrown out") %>% 
  mutate(check_ht = `height (before 2017 growth)`+gr_2017_cm - ht_cm) %>% 
  select(sdlg, check_ht, gr_2017_cm, ht_cm, `height (before 2017 growth)`, notes.spring, notes.Aug)
df_check_ht
df <- df %>% 
  select(-`height (before 2017 growth)`)
```

These don't all look good. However, there's nothing I can do but delete the column for pre-2017 height. 

## Correct typos 

### Diameter typo
```{r}
df <- df %>% 
  mutate(spring_dia_2_mm = ifelse(sdlg == 21 & spring_dia_2_mm == 28, 2.8, spring_dia_2_mm))
filter(df, sdlg == 21)
```

## Take out seedlings need to be thrown out of the study entirely

### Before throwing out, save a version of the data set to check against 2018 data for 2018 data completeness.
```{r}
save(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_unedited.Rdata")
```


### Seedlings with a comment to throw them out
```{r}
npre <- nrow(df)
npre
df %>% filter(is.na(Aug_dia_1_mm)) %>% select(notes.spring, notes.Aug)
df <- df %>% 
  filter(!is.na(Aug_dia_1_mm)) 
npost <- nrow(df)
```

#### This removed `r npre-npost` seedlings, bringing the total to `r npost`.

### Throw out dead seedlings, except the one with the note that it died after completing 2017 growth
```{r}
npre <- nrow(df)
npre
nrow(df)
df %>% filter(notes.Aug == "3 yrs of growth died after 2017 growth completed")
df <- df %>% filter(dead == "A" | notes.Aug == "3 yrs of growth died after 2017 growth completed")
npost <- nrow(df)
npost
```

#### This removed `r npre-npost` trees, bringing the total to `r npost`

# Combine with plot data to assign each row a date and removed/retained based on plot data

### Clean up plot data
```{r}
dfsp_plots <- dfsp_plots %>% 
  rename(date_sdlg = `date - seedlings`, time_sdlg = `time - seedling meas`, source_sdlg = `source - seedlings`)
dfsp_plots_small <- dfsp_plots %>% 
  dplyr::select(compartment, island, plot, shrub_species, shrubs, date_sdlg, time_sdlg, source_sdlg)
```

### Join plot data with seedling data
```{r}
df <- full_join(df, dfsp_plots_small, by = c("compartment", "island", "plot"))
```

### Check 
```{r}
df %>% filter(is.na(shrubs)) %>% select(compartment, island, plot) %>% distinct()
```

# Diameter growth

## Calculate average diameter growth
```{r}
df_dia <- df %>%
  rowwise() %>% 
  mutate(spring_dia_ave_mm = mean(c(spring_dia_1_mm, spring_dia_2_mm), na.rm = T) ) %>% 
  mutate(Aug_dia_ave_mm = mean(c(Aug_dia_1_mm, Aug_dia_2_mm), na.rm = T))
```

## Remove seedling with missing Spring 2017 diameter data
```{r}
npre <- nrow(df_dia)
npre
df_dia %>% 
  filter(is.na(spring_dia_1_mm)) %>% 
  select(compartment, island, plot, sdlg, notes.spring, notes.Aug, spring_dia_1_mm, spring_dia_2_mm, species, gr_2017_cm, Aug_dia_1_mm, dead)
df_dia <- df_dia %>% filter(!(is.na(spring_dia_1_mm)))
npost <- nrow(df_dia)
npost
```

I looked at the original data sheets for all 3 of the seedlings missing spring data, and found nothing.

### This removed `r npre-npost` seedlings, bringing the total to `r npost`.

## Calculate diameter growth and relative diameter growth
```{r}
df_dia <- df_dia %>% 

  mutate(dia_growth_mm = Aug_dia_ave_mm - spring_dia_ave_mm) %>% 
  mutate(dia_growth_rel = dia_growth_mm/(spring_dia_ave_mm))
```

## Take out seedlings with clear errors in diameter measurements
```{r}
npre <- nrow(df_dia)
npre
df_dia %>% filter(dia_growth_rel>1)
df_dia <- df_dia %>% 
  filter(dia_growth_rel<1) 
npost <- nrow(df_dia)
```

### This removed `r npre-npost` seedlings, bringing the total to `r npost`.

## Plot diameter growth in relation to treatment, shrub species, and tree species
```{r}
dia_growth_boxplots <- ggplot(data = df_dia)+
  geom_boxplot(aes(x = interaction(shrubs, species, shrub_species), y = dia_growth_rel, col = shrubs))
dia_growth_boxplots
df_dia %>% 
  group_by(species, shrub_species, shrubs) %>% 
  summarise(mean(dia_growth_rel), median(dia_growth_rel))
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
png("dia_growth_boxplots.jpeg", width = 1500, height = 450, res = 72)
dia_growth_boxplots
dev.off()
```

```{r}
ggplot(data = df_dia)+
  geom_boxplot(aes(x = interaction(shrubs, species, shrub_species), y = dia_growth_rel, col = shrubs))
figure <- ggplot(data = df_dia)+
  geom_boxplot(aes(x = interaction(shrubs, species), y = dia_growth_rel, col = shrubs))+
  labs(y = "2017 relative diameter growth")+
  ggtitle("Relative diameter growth was higherin removal plots")
figure
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
png("dia_growth_boxplots_simple.png")
figure
dev.off()
```

# Vertical growth

## Calculate relative vertical growth for each year
```{r}
df_vert <- df %>% 
  mutate(gr_2017_rel = gr_2017_cm/(ht_cm-gr_2017_cm)) %>% 
  mutate(gr_2016_rel = gr_2016_cm/(ht_cm - gr_2017_cm - gr_2016_cm)) %>% 
  mutate(gr_2015_rel = gr_2015_cm/(ht_cm - gr_2017_cm - gr_2016_cm - gr_2015_cm)) 
```

## Take out seedlings with broken leaders in 2017 or 2016, unless it died after growth was complete

The way I leave in the seadling that died after growth was complete is by not listing its `lead_damage` record below, which was `20172016,2015`

```{r}
npre <- nrow(df_vert)
npre
summary(as.factor(df_vert$leader_damage))
df_vert %>% filter(!is.na(leader_damage) & leader_damage != 0)
df_vert <- df_vert %>% 
  filter(!leader_damage %in% c("2016", "2017", "2016,2017", "2017 and 2016", "2017,2015", "2017,2016", "gall in 2016 growth"))
npost <- nrow(df_vert)
npost
```

### This removed `r npre-npost` trees, bringing the total to `r npost`

## Calculate vertical 2017 growth/2016 growth ratio variable and compare it between treatments

### Check distribution of unaltered relative growth by year
```{r}
hist(df_vert$gr_2017_rel)
hist(log(df_vert$gr_2017_rel))
hist(df_vert$gr_2017_rel)
hist(sqrt(df_vert$gr_2017_rel))
```

### Create gr_17_over_16 variable
```{r}
df_vert <- df_vert %>% 
  mutate(gr_17_over_16 = gr_2017_rel/gr_2016_rel)
```

### Check distribution
```{r}
hist(df_vert$gr_17_over_16)
hist(log(df_vert$gr_17_over_16)) # don't use log because there are some zeros (plus sqrt transformation is better)
hist(sqrt(df_vert$gr_17_over_16))
```

## Check the NA value of `gr_17_over_16` and remove it
```{r}
npre <- nrow(df_vert)
na <- df_vert %>% filter(is.na(gr_17_over_16))
na
df_vert <- df_vert %>% filter(!is.na(gr_17_over_16))
npost <- nrow(df_vert)
```

### This removed `r npre-npost` trees for the reason: `r na$notes.Aug`, bringing the df_vert total to `r npost`

## Plot vertical growth by treatment type and tree species and save the figure
```{r}
ggplot(data = df_vert)+
  geom_boxplot(aes(x = interaction(shrubs, species, shrub_species), y = sqrt(gr_17_over_16), col = shrubs))
figure <- ggplot(data = df_vert)+
  geom_boxplot(aes(x = interaction(shrubs, species), y = sqrt(gr_17_over_16), col = shrubs))+
  labs(y = "2017 relative growth/2016 relative growth, square root transformed")+
  ggtitle("Change in vertical growth between 2017 and 2016\nwas lower in removal plots")
figure
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
png("vert_growth_boxplots.png")
figure
dev.off()
```

# Save data frames
```{r}
write.csv(df_vert, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_vert.csv", row.names = F)
save(df_vert, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_vert.Rdata")

write.csv(df_dia, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia.csv", row.names = F)
save(df_dia, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia.Rdata")

write.csv(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df.csv", row.names = F)
save(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df.Rdata")

```

