---
title: "Shrub experiment seedling data - cleaning"
author: "Carmen"
date: "November 7, 2017"
output: 
    html_document:
        toc: TRUE
---

This code compiles and cleans seedling diameter, height, and vertical growth data. It then combines these data with treatment, shrub species, and compartment data. Later code makes pretty graphs and analyses with the cleaned and combined data.

```{r, include = F}
library(dplyr)
library(readxl)
library(ggplot2)
```


# Pull in data
```{r}
dfsp_plots <- read_excel("~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/Compiled/Spring_seedling_data_compiled_20171107.xlsx", sheet = 1)
dfsp <- read_excel("~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/Compiled/Spring_seedling_data_compiled_20171107.xlsx", sheet = "dia")
dfAu <- read_excel("~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/Compiled/Seedling_vert_Aug_2017_compiled_110717.xlsx", sheet = "data")
```

# Make all island names uppercase
```{r}
dfsp <- dfsp %>% mutate(island = toupper(island))
dfAu <- dfAu %>% mutate(island = toupper(island))
```


# Combine spring and August data

## Check and correct for trees with two stems
```{r}
dfAu %>% 
  group_by(compartment, island, plot, `seedling tag #`) %>% 
  mutate(n = n()) %>% 
  filter(n>1)
```

```{r}
dfAu %>% filter(compartment == 180 & island == "G" & plot == 2 & `seedling tag #` == 118 & notes == "first stem")
dfAu <- dfAu %>% 
  mutate(`seedling tag #` = ifelse(compartment == 180 & island == "G" & plot == 2 & `seedling tag #` == 118 & notes == "first stem", "118A", `seedling tag #`)) %>% 
    mutate(`seedling tag #` = ifelse(compartment == 180 & island == "G" & plot == 2 & `seedling tag #` == 118 & notes == "second stem", "118B", `seedling tag #`))
dfAu %>% filter(compartment == 180 & island == "G" & plot == 2 )
```

```{r}
new_row <- dfsp %>%  
  filter(compartment == 180 & island == "G" & plot == 2 & `seedling tag #` == 118) %>% 
  mutate(`seedling tag #`=c("118A"))
new_row
dfsp <- dfsp %>% 
  mutate(`seedling tag #` = ifelse(compartment == 180 & island == "G" & plot == 2 & `seedling tag #` == 118, "118B", `seedling tag #`))
dfsp <- bind_rows(dfsp, new_row)
dfsp %>% 
  filter(compartment == 180 & island == "G" & plot == 2 )
```

## Combine data frame
```{r}
dfsp
nrow(dfsp)
dfsp %>% group_by(compartment, island) %>% summarise(n())
dfAu
nrow(dfAu)
str(dfAu)
dfAu %>% group_by(compartment, island) %>% summarise(n())
df <- full_join(dfsp, dfAu, by = c("compartment", "island", "plot", "seedling tag #"))
nrow(df)
str(df)
```

## CHeck the two stemmers
```{r}
df %>% 
  filter(compartment == 180 & island == "G" & plot == 2 )
```


# Overall cleaning not specific to vert or dia

## Rename columns
```{r}
df <- df %>% 
  rename(spring_dia_1_mm = `dm at base 1 (mm)`) %>% 
  rename(spring_dia_2_mm = `dm at base 2 (mm)`) %>% 
  rename(Aug_dia_1_mm = bas_dia_mm_1) %>% 
  rename(Aug_dia_2_mm = bas_dia_mm_2) %>% 
  rename(growing = `growing yet.`) %>% 
  rename(sdlg = `seedling tag #`) %>% 
  rename(notes.spring = notes.x) %>% 
  rename(notes.Aug = notes.y) %>% 
  rename(leader_damage = `leader broke or died`) %>% 
  select(-`DIA/ht check`)
df
```

## Remove unnecessary columns

### Column for diameter at 10 cm (very partial)
```{r}
df <- df %>% 
  select(-`dm at 10 (1)`, -`dm at 10 (2)`)
```
 
### Growing column (very partial)
```{r}
df <- df %>% 
  mutate(growing = toupper(growing)) %>% 
  mutate(growing = ifelse(growing == "Y", "YES", growing))
```
 
### Duplicate species columns 

#### Check that species match
```{r}
df %>% filter(species.x != species.y) %>% select(compartment, island, plot, sdlg, species.x, species.y, notes.Aug, notes.spring)
```

For 180-G-1, ABCO is correct based on the note from August 2017. For 180-C-2-109, PIPO is correct based on note from Aug 2018. Therefore, the column species.y has correct species

#### Remove species.x column since species.y is more accurate
```{r}
df <- df %>% select(-species.x) %>% rename(species = species.y)
```
 
### For the column for height before 2017 growth, there are only 11 seedlings with data. Check those.
```{r}
df_check_ht <- df %>% 
  filter(!is.na(`height (before 2017 growth)`)) %>% 
  filter(is.na(notes.Aug) | notes.Aug != "thrown out") %>% 
  mutate(check_ht = `height (before 2017 growth)`+gr_2017_cm - ht_cm) %>% 
  select(sdlg, check_ht, gr_2017_cm, ht_cm, `height (before 2017 growth)`, notes.spring, notes.Aug)
df_check_ht
df <- df %>% 
  select(-`height (before 2017 growth)`)
```

These don't all look good. However, there's nothing I can do but delete the column for pre-2017 height. 

## Correct typos 

### Diameter typo
```{r}
df <- df %>% 
  mutate(spring_dia_2_mm = ifelse(sdlg == 21 & spring_dia_2_mm == 28, 2.8, spring_dia_2_mm))
filter(df, sdlg == 21)
```

## Take out seedlings need to be thrown out of the study entirely

### Before throwing out, save a version of the data set to check against 2018 data for 2018 data completeness.
```{r}
save(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_unedited.Rdata")
```

### Seedlings with a comment to throw them out
```{r}
npre <- nrow(df)
npre
df %>% filter(is.na(Aug_dia_1_mm)) %>% select(notes.spring, notes.Aug)
df <- df %>% 
  filter(!is.na(Aug_dia_1_mm)) 
npost <- nrow(df)
```

#### This removed `r npre-npost` seedlings, bringing the total to `r npost`.

### Throw out dead seedlings, except the one with the note that it died after completing 2017 growth
```{r}
npre <- nrow(df)
npre
nrow(df)
df %>% filter(notes.Aug == "3 yrs of growth died after 2017 growth completed")
df <- df %>% filter(dead == "A" | notes.Aug == "3 yrs of growth died after 2017 growth completed")
npost <- nrow(df)
npost
```

#### This removed `r npre-npost` trees, bringing the total to `r npost`

# Combine with plot data to assign each row a date and removed/retained based on plot data

### Clean up plot data
```{r}
dfsp_plots <- dfsp_plots %>% 
  rename(date_sdlg = `date - seedlings`, time_sdlg = `time - seedling meas`, source_sdlg = `source - seedlings`)
dfsp_plots_small <- dfsp_plots %>% 
  dplyr::select(compartment, island, plot, shrub_species, shrubs, date_sdlg, time_sdlg, source_sdlg)
```

### Join plot data with seedling data
```{r}
df <- full_join(df, dfsp_plots_small, by = c("compartment", "island", "plot"))
```

### Check 
```{r}
df %>% filter(is.na(shrubs)) %>% select(compartment, island, plot) %>% distinct()
```

# Diameter growth

## Calculate average diameter growth
```{r}
df_dia <- df %>%
  rowwise() %>% 
  mutate(spring_dia_ave_mm = mean(c(spring_dia_1_mm, spring_dia_2_mm), na.rm = T) ) %>% 
  mutate(Aug_dia_ave_mm = mean(c(Aug_dia_1_mm, Aug_dia_2_mm), na.rm = T))
```

## Remove seedling with missing Spring 2017 diameter data
```{r}
npre <- nrow(df_dia)
npre
df_dia %>% 
  filter(is.na(spring_dia_1_mm)) %>% 
  select(compartment, island, plot, sdlg, notes.spring, notes.Aug, spring_dia_1_mm, spring_dia_2_mm, species, gr_2017_cm, Aug_dia_1_mm, dead)
df_dia <- df_dia %>% filter(!(is.na(spring_dia_1_mm)))
npost <- nrow(df_dia)
npost
```

I looked at the original data sheets for all 3 of the seedlings missing spring data, and found nothing.

### This removed `r npre-npost` seedlings, bringing the total to `r npost`.

## Calculate diameter growth and relative diameter growth
```{r}
df_dia <- df_dia %>% 
  mutate(dia_growth_mm = Aug_dia_ave_mm - spring_dia_ave_mm) %>% 
  mutate(dia_growth_rel = dia_growth_mm/(spring_dia_ave_mm))
```

## Take out seedlings with clear errors in diameter measurements
```{r}
npre <- nrow(df_dia)
npre
df_dia %>% filter(dia_growth_rel>1)
df_dia <- df_dia %>% 
  filter(dia_growth_rel<1) 
npost <- nrow(df_dia)
```

### This removed `r npre-npost` seedlings, bringing the total to `r npost`.

# Establish df_Vert

```{r}
df_vert <- df
```


# Save data frames
```{r}
write.csv(df_vert, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_vert.csv", row.names = F)
save(df_vert, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_vert.Rdata")

write.csv(df_dia, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia.csv", row.names = F)
save(df_dia, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia.Rdata")

write.csv(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df.csv", row.names = F)
save(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df.Rdata")

```

