---
title: "CLean diameter data for both species"
author: "Carmen"
date: "April 11, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(ggplot2)
library(tidyverse)
require(nlme)
library(kableExtra)
library(knitr)
library(lme4)
```


# Load data
```{r}
load(file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings2.Rdata")
```

# General Clean


## Take out trees that aren't PIPO or ABCO
```{r}
df %>% group_by(Species) %>% 
  summarise(n())
df <- df %>%
  filter(Species %in% c("ABCO", "PIPO"))
```


## Convert numeric columns to numeric where needed
```{r}
df <- df %>% 
  mutate(BasDia2017.mm.1 = as.numeric(paste(BasDia2017.mm.1))) %>% 
  mutate(BasDia2017.mm.2 = as.numeric(paste(BasDia2017.mm.2))) 
```

## Take out trees with no second dia measurement
```{r}
df <- df %>% filter(!(BasDia2017.mm.2== 0 & BasDia2017.mm.1 == 0))
```

## Restructure so there's only one row per seedling 
```{r}
df_dia <- df %>% 
  select(-VertGrowth, -VertGrowth_Rel, -Year) %>% 
  distinct()
remove(df)
```

## A seedling was defined as an outlier using analysis of residuals after the fact
It had suspiciously lower basal diameter in 2017 compared to 2016, and I take it out here
```{r}
df_dia <- df_dia %>% filter(!(Fire == "CLVD" & Sdlg == "11"))
```

## Lump quercus with "other" shrub genus since it only has 3 observations
```{r}
df_dia <- df_dia %>% 
  ungroup() %>% 
  mutate(ShrG1 = ifelse(ShrG1 == "Quercus", "Other", as.character(ShrG1)))
df_dia$ShrG1 <- as.factor(df_dia$ShrG1)
```

## Take out repeat row
```{r}
nrow(df_dia)
df_dia <- df_dia %>% 
  filter(!(Sdlg == "117" & LAI == 0))
nrow(df_dia)
```

### Delete trees whose measurements are obviously errors
```{r}
df_dia <- df_dia %>% 
  filter(!Sdlg == 27)
```


### Correct tree where units were clearly recorded in cm instead of mm
```{r}
df_dia %>% 
  filter(Sdlg == 290) %>% 
  select(BasDia2016.cm, BasDia2017.mm.1, BasDia2017.mm.2)
df_dia <- df_dia %>% 
  mutate(BasDia2017.mm.1 = ifelse(Sdlg == 290, BasDia2017.mm.1*10, BasDia2017.mm.1)) %>% 
  mutate(BasDia2017.mm.2 = ifelse(Sdlg == 290, BasDia2017.mm.2*10, BasDia2017.mm.2))
```


# Calculate diameter growth

## Calculate diameter average in 2017
```{r}
df_dia <- tidyr::replace_na(df_dia, replace = list(BasDia2017.mm.2 = 0, BasDia2017.mm.1 = 0))
df_dia <- df_dia %>% 
  mutate(BasDia2017.mm.ave = (BasDia2017.mm.1+BasDia2017.mm.2)/2)
```

## Adjust 2016 diameters to mm and correct for calipers error
```{r}
df_dia <- df_dia %>% 
  mutate(BasDia2016.mm = BasDia2016.cm*10-1.1)
```

## Calculate growth
```{r}
df_dia <- df_dia %>% 
  mutate(DiaGrowth.mm = ifelse(BasDia2017.mm.1 >0, BasDia2017.mm.ave - BasDia2016.mm, NA))
df_dia %>% 
  filter(BasDia2017.mm.ave >0 & !is.na(DiaGrowth.mm)) %>% 
  summarise(mean(DiaGrowth.mm))
```

## Calculate relative diameter growth
```{r}
df_dia <- df_dia %>% 
  mutate(DiaGrowth.mm_rel = DiaGrowth.mm/(BasDia2016.cm*10))
hist(df_dia$DiaGrowth.mm_rel)
```

# Deal with netative values

## Count how many there are 
```{r}
df_dia %>% 
  filter(DiaGrowth.mm<0) %>% 
  summarize(n())
  nrow(df_dia)
df_dia %>% 
  group_by(Species) %>% 
  filter(DiaGrowth.mm<0) %>% 
  summarize(n())
  nrow(df_dia)
```

<!-- ## Take out outliers on both ends, defined as greater than two SD from the mean -->

<!-- ### Define bounds of + or - 2 SD -->
<!-- ```{r} -->
<!-- mean(df_dia$DiaGrowth.mm_rel) - 2*sd(df_dia$DiaGrowth.mm_rel) -->
<!-- mean(df_dia$DiaGrowth.mm_rel) + 2*sd(df_dia$DiaGrowth.mm_rel) -->
<!-- ``` -->

<!-- ### Look at the trees outside those bounds -->
<!-- ```{r} -->
<!-- df_dia %>%  -->
<!--    filter(DiaGrowth.mm_rel < (mean(df_dia$DiaGrowth.mm_rel) - 2*sd(df_dia$DiaGrowth.mm_rel)) | -->
<!--               DiaGrowth.mm_rel > mean(df_dia$DiaGrowth.mm_rel) + 2*sd(df_dia$DiaGrowth.mm_rel)) %>%  -->
<!--    select(Sdlg, Ht_cm, Species, DiaGrowth.mm, DiaGrowth.mm_rel, BasDia2016.mm, BasDia2017.mm.1, BasDia2017.mm.2, notes, DataSheet2017) %>%  -->
<!--   arrange(DiaGrowth.mm) -->
<!-- ``` -->

<!-- ### Filter them out -->
<!-- ```{r} -->
<!-- df_dia <- df_dia %>%  -->
<!--   filter(!(DiaGrowth.mm_rel < (mean(df_dia$DiaGrowth.mm_rel) - 2*sd(df_dia$DiaGrowth.mm_rel)) | -->
<!--              DiaGrowth.mm_rel > mean(df_dia$DiaGrowth.mm_rel) + 2*sd(df_dia$DiaGrowth.mm_rel))) -->
<!-- ``` -->

<!-- ### Look at the new data distribution -->
<!-- ```{r} -->
<!-- hist(df_dia$DiaGrowth.mm_rel) -->
<!-- df_dia %>%  -->
<!--   filter(DiaGrowth.mm_rel < 0) %>%  -->
<!--   summarise(n()) -->
<!-- 39/nrow(df_dia) -->
<!-- ``` -->


<!-- ## Adjust so that no values are below 0  -->
<!-- ```{r} -->
<!-- df_dia <- df_dia %>%  -->
<!--   mutate(DiaGrowth.mm_rel = (DiaGrowth.mm_rel - min(df_dia$DiaGrowth.mm_rel))) %>%  -->
<!--   mutate(DiaGrowth.mm = (DiaGrowth.mm - min(df_dia$DiaGrowth.mm)))  -->
<!-- ``` -->


# Save data
```{r}
save(df_dia, file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_dia.Rdata")
```

