---
title: "Fir volume cleaning"
author: "Carmen"
date: "March 25, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(ggplot2)
library(dplyr)
library(kableExtra)
library(knitr)
library(tidyr)
```

# Data wrangling

## Load final data frame
```{r}
load("~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_dia.Rdata")
load("~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_vert.Rdata")
df_vert <- df
remove(df)
```

## Delete rows for vert growth before 2016
```{r}
df_vert <- df_vert %>% 
  filter(Year != 2015)
```

## Restructure vert data so there's one row per seedling
```{r}
df_vert <- df_vert %>% select(-VertGrowth_Rel)
df_vert <- spread(df_vert, key = Year, value = VertGrowth) %>% 
  rename(VertGr_2016 = `2016`) %>% 
  rename(VertGr_2017 = `2017`)
```

### Remove duplicate seedling
```{r}
sort(df_vert$Sdlg)
nrow(df_vert)
length(unique(df_vert$Sdlg))
df_vert[df_vert %>% select(Sdlg) %>% duplicated(),]
df_vert <- df_vert %>% 
  filter(!(Sdlg == "117" & LAI == 0))


sort(df_dia$Sdlg)
nrow(df_dia)
length(unique(df_dia$Sdlg))
dups <- df_dia[df_dia %>% select(Sdlg) %>% duplicated(),]
df_dia %>% filter(Sdlg %in% dups$Sdlg)
df_dia <- df_dia %>% filter(!(Sdlg %in% dups$Sdlg & file_lat == "AMR - Aerial - 2.csv"))
```


```{r}
nrow(df_vert)
df_vert <- df_vert %>% 
  filter(!(Sdlg == "117" & LAI == 0))
nrow(df_vert)
```


## Filter to only seedlings with workable data for both diameter and vertical growth
```{r}
df_vert <- df_vert %>% select(Sdlg, VertGr_2016, VertGr_2017)
df <- inner_join(df_dia, df_vert)
```

## Rename columns to distinguish between seedling ht and shrub ht
```{r}
df <- df %>% 
  rename(tree_Ht_cm = Ht_cm, tree_Ht_norm = Ht_norm)
```

## Take out PSME so there's only ABCO
```{r}
df <- df %>% 
  filter(Species == "ABCO")
```

## Clean up extreme and negative values

<!-- ## Take out outliers on both ends, defined as greater than two SD from the mean -->
<!-- ```{r} -->
<!-- nrow(df) -->
<!-- mean_minSD <- mean(df$DiaGrowth.mm) - 2*sd(df$DiaGrowth.mm) -->
<!-- mean_minSD -->
<!-- mean_plusSD <- mean(df$DiaGrowth.mm) + 2*sd(df$DiaGrowth.mm) -->
<!-- mean_plusSD -->
<!-- df <- df %>%  -->
<!--   filter(!(DiaGrowth.mm < mean_minSD | -->
<!--              DiaGrowth.mm > mean_plusSD)) -->
<!-- hist(df$DiaGrowth.mm) -->
<!-- nrow(df) -->
<!-- ``` -->


<!-- ## Adjust diameter growth so that all values are > 0 -->
<!-- ```{r} -->
<!-- min(df$BasDia2017.mm.ave) -->
<!-- hist(df$BasDia2017.mm.ave) -->
<!-- df <- df %>%  -->
<!--   mutate(BasDia2017.mm.ave = (BasDia2017.mm.ave - min(df$DiaGrowth.mm))) -->
<!-- min(df$BasDia2017.mm.ave) -->
<!-- hist(df$BasDia2017.mm.ave) -->
<!-- df <- df %>%  -->
<!--   mutate(DiaGrowth.mm = (DiaGrowth.mm - min(DiaGrowth.mm))) -->
<!-- ``` -->

## Check for seedlings with more than one row
```{r}
nrow(df)
length(unique(df$Sdlg))
df %>% filter(duplicated(Sdlg))
df <- df %>% 
  distinct()
```

# Calculate relative growth rate

## Replace NA vertical growth with 0 
```{r}
df <- df %>%
  mutate(VertGr_2017 = ifelse(is.na(VertGr_2017), 0, VertGr_2017)) %>% 
   mutate(VertGr_2016 = ifelse(is.na(VertGr_2016), 0, VertGr_2016))
```

## Calculate pre volume assuming parabolic cone shape (https://www.grc.nasa.gov/www/k-12/airplane/volume.html)
```{r}
df <- df %>% 
  mutate(vol_pre_cm2 = (2*pi*((BasDia2016.mm/10)^2)*(tree_Ht_cm - VertGr_2017  - VertGr_2016))/15)
hist(df$vol_pre_cm2)
```

## Calculate post volume
```{r}
df <- df %>% 
  mutate(vol_post_cm2 = (2*pi*((BasDia2017.mm.ave/10)^2)*tree_Ht_cm)/15)

hist(df$vol_post_cm2)
```

### Compare to pre
```{r}
plot(log(df$vol_post_cm2), log(df$vol_pre_cm2))
```

## Calculate volume growth 
```{r}
df <- df %>% 
  mutate(vol_growth_cm2 = vol_post_cm2 - vol_pre_cm2)
```


## Calculate RGR, using a demonimator of 2 since there were 2 growing seasons between initial and final measurements
```{r}
df <- df %>% 
  mutate(RGR = (log(vol_post_cm2)- log(vol_pre_cm2))/2)
hist(df$RGR)
```

# Take a look

```{r}
ggplot(df, aes(x = tree_Ht_cm, y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
ggplot(df, aes(x = log(tree_Ht_cm), y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
```

# Save
```{r}
save(df, file = "~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_vol.Rdata")
```

