---
title: Fir volume growth analysis -fire footprints
author: Carmen
date: November 19, 2017
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(ggplot2)
library(dplyr)
require(nlme)
library(kableExtra)
library(knitr)
library(lme4)
library(tidyr)
library(randomForest)
library(VSURF)
```

# Data wrangling

## Load final data frame
```{r}
load("~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_dia.Rdata")
load("~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_vert.Rdata")
df_vert <- df
remove(df)
```

## Delete rows for vert growth before 2016
```{r}
df_vert <- df_vert %>% 
  filter(Year != 2015)
```

## Restructure vert data so there's one row per seedling
```{r}
df_vert <- df_vert %>% select(-VertGrowth_Rel)
df_vert <- spread(df_vert, key = Year, value = VertGrowth) %>% 
  rename(VertGr_2016 = `2016`) %>% 
  rename(VertGr_2017 = `2017`)
```

### Remove duplicate seedling
```{r}
nrow(df_vert)
df_vert <- df_vert %>% 
  filter(!(Sdlg == "117" & LAI == 0))
nrow(df_vert)
```


## Filter to only seedlings with workable data for both diameter and vertical growth
```{r}
df_vert <- df_vert %>% select(Sdlg, VertGr_2016, VertGr_2017)
df <- inner_join(df_dia, df_vert)
```

## Rename columns to distinguish between seedling ht and shrub ht
```{r}
df <- df %>% 
  rename(tree_Ht_cm = Ht_cm, tree_Ht_norm = Ht_norm)
```

## Take out PSME so there's only ABCO
```{r}
df <- df %>% 
  filter(Species == "ABCO")
```

## Clean up extreme and negative values

```{r}
nrow(df)
hist(df$DiaGrowth_rel, breaks = 10)
hist(df$DiaGrowth_rel)
nrow(df)
```


## Take out outliers on both ends, defined as greater than two SD from the mean
```{r}
nrow(df)
hist(df$DiaGrowth, breaks = 20)
mean_minSD <- mean(df$DiaGrowth) - 2*sd(df$DiaGrowth)
mean_minSD
mean_plusSD <- mean(df$DiaGrowth) + 2*sd(df$DiaGrowth)
mean_plusSD
df <- df %>% 
  filter(!(DiaGrowth < mean_minSD |
             DiaGrowth > mean_plusSD))
hist(df$DiaGrowth)
nrow(df)
```


## Adjust diameter growth so that all values are > 0
```{r}
hist(df$DiaGrowth)
min(df$BasDia2017.mm.ave)
hist(df$BasDia2017.mm.ave)
df <- df %>% 
  mutate(BasDia2017.mm.ave = (BasDia2017.mm.ave - min(df$DiaGrowth)))
min(df$BasDia2017.mm.ave)
hist(df$BasDia2017.mm.ave)
df <- df %>% 
  mutate(DiaGrowth = (DiaGrowth - min(DiaGrowth)))
```

## Check for seedlings with more than one row
```{r}
nrow(df)
length(unique(df$Sdlg))
df %>% filter(duplicated(Sdlg))
df <- df %>% 
  distinct()
```

# Calculate relative growth rate

## Replace NA vertical growth with 0 
```{r}
df <- df %>%
  mutate(VertGr_2017 = ifelse(is.na(VertGr_2017), 0, VertGr_2017)) %>% 
   mutate(VertGr_2016 = ifelse(is.na(VertGr_2016), 0, VertGr_2016))
```

## Calculate pre volume
```{r}
df <- df %>% 
  mutate(vol_pre_cm2 = ((BasDia2016.mm/10)^2) * (tree_Ht_cm - VertGr_2017  - VertGr_2016) * (pi/12))
hist(df$vol_pre_cm2)
```

## Calculate post volume
```{r}
df <- df %>% 
  mutate(vol_post_cm2 = ((BasDia2017.mm.ave/10)^2) * tree_Ht_cm * (pi/12))
hist(df$vol_post_cm2)
```

## Calculate RGR
```{r}
df <- df %>% 
  mutate(RGR = (log(vol_post_cm2)- log(vol_pre_cm2))/2)
hist(df$RGR)
```

# Take a look

```{r}
ggplot(df, aes(x = tree_Ht_cm, y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
ggplot(df, aes(x = log(tree_Ht_cm), y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
```

# Normalize independent variables

```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

```{r}
df <- df %>% 
  mutate(tree_Ht_norm = normalize(tree_Ht_cm)) %>% 
  mutate(shrubarea2_sqrt_norm = normalize(sqrt(shrubarea2)) ) %>% 
  mutate(tree_Ht_log_norm = normalize(log(tree_Ht_cm)))
```

## Visualize pre and post volumes on a log scale

```{r}
ggplot(df)+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)
```

```{r}
ggplot(df)+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_post_cm2)-log(vol_pre_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))
```

## Plot relative growth rate vs. volume to check for a possible relationship

### least squares
```{r}
ggplot(df, aes(x = log(vol_pre_cm2), y = RGR))+
  geom_point()+
  geom_smooth(method = 'lm')
```

### loess
```{r}
ggplot(df, aes(x = log(vol_pre_cm2), y = RGR))+
  geom_point()+
  geom_smooth()
```

This looks like I should maybe include volume in the model of growth rate or adjust how I analyze growth rates

## Plot relative growth rate vs. ht to check for a relationship

### least squares
```{r}
ggplot(df, aes(x = tree_Ht_log_norm, y = RGR))+
  geom_point()+
  geom_smooth(method = 'lm')
```

### loess
```{r}
ggplot(df, aes(x = tree_Ht_log_norm, y = RGR))+
  geom_point()+
  geom_smooth()
```

There appears to be less of a relationship between height and growth than between volume and growth

# Mixed effects with random fire effect
```{r}
MM1 <- lme(RGR ~Cov1.2_norm, random =~1|Fire, data = df, method = "ML")

MM1_null <- lme(RGR ~ 1, random =~1|Fire, method = "ML",data = df)
anova(MM1, MM1_null)
summary(MM1)
```

```{r}
decdf <- read.csv("~/Shrubs-Seedlings-master_Dec2018/code/FireFootprints_analysis/df_firvol_Dec2018.csv")
setdiff(as.numeric(df$Sdlg), decdf$Sdlg)
setdiff(decdf$Sdlg, as.numeric(df$Sdlg))
```


# Try using VSURF
```{r}
xdf <- df %>% 
  select(Fire, 
         shrubarea2, 
         shrubarea3,
         Cov1, 
         Cov1.2, 
         Cov1.3, 
         Ht1, 
         Ht1.2, 
         Ht1.3,
         ShrubSpp01, 
         ShrubSpp02, 
         ShrubSpp03, 
         IAG, 
         shrubarea2_sqrt_norm, 
         Cov1.2_norm, 
         Cov1.3_norm,
         ShrG1, 
         BasDia2016.cm, 
         vol_pre_cm2, 
         tree_Ht_log_norm)

names(xdf)
vsurf <- VSURF(x = xdf, y = df$RGR, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
```

## Try using all variables, including those with missing values

### Remove columns that are not explanatory variables or are totally redundant (eg same as another var but diff units)
```{r}
df_vars <- df %>% 
  select(-Sdlg, -Status, -notes, -Light_File, -Map, -DataSheet2017, -Date, -resid, -DiaGrowth, -DiaGrowth_rel, -BasDia2016.cm, -VertGr_2016, -VertGr_2017)
```

```{r}
vsurf2 <- VSURF(RGR ~ ., data = df_vars, na.action = na.omit)
```

The order thing is annoying. Better to reorder variables before running vsurf than to have to figure out the otder afterwards. Since RGR is the 57th variable out of 58, I probably don't need to worry aboutt he order issue.

### See what variables came out
```{r}
vsurf2$varselect.pred
df_vars[vsurf2$varselect.pred]

```

### Run a random forest with just those variables and look at their relative importances
```{r}
fit <- randomForest(formula = GRG)
```


# Build a random forest to get a sense of what variables are important
```{r}
fit <- randomForest(y = df$RGR, x = df_vars[vsurf2$varselect.pred], 
                    importance = TRUE,
                    scale = TRUE,
                    na.action = na.omit)
fit
varImpPlot(fit)
```


# Then identify exactly what variables to use in an AIC table based on the random forest above

## Find all possible variable combinations

### Define variables of interest
```{r}
v1 <- c("shrubarea2_sqrt_norm", "Cov1.2_norm" , "Cov1.3_norm" , "IAG", "ShrG1", "shrubarea2_sqrt_norm:tree_Ht_log_norm", "Cov1.2_norm:tree_Ht_log_norm", "BasDia2017.mm.ave", "log(vol_pre_cm2)", "tree_Ht_log_norm")
combs <- do.call(c, lapply(seq_along(v1), function(i) combn(v1, i, FUN = list)))
```

### Remove biologically redundant variable combinations
```{r}
list.bad <- list()
for(i in 1:length(combs)){
  if("Cov1.2_norm" %in% combs[i][[1]] & "Cov1.3_norm" %in% combs[i][[1]])
    {
    list.bad <- c(list.bad, i)
  }
  if("tree_Ht_log_norm" %in% combs[i][[1]] & "log(vol_pre_cm2)" %in% combs[i][[1]])
    {
    list.bad <- c(list.bad, i)
    }
  if("shrubarea2_sqrt_norm" %in% combs[i][[1]] & ("Cov1.2_norm" %in% combs[i][[1]] | "Cov1.3_norm" %in% combs[i][[1]]     )){
    list.bad <- c(list.bad, i)
    } 
}
list.bad <- unlist(list.bad)
combs.bad <- combs[list.bad]
combs <- setdiff(combs, combs.bad)
length(combs)
```

## Table
```{r}
table <- data.frame(Model = character(), Formula = character(), AIC = numeric())
for(i in 1:length(combs)){
  vars <- combs[i]
  vars <- paste(vars[[1]], collapse="+")
  f <- formula(paste("(RGR ~ ", vars,")"))
  M <- lme(f,data = df, random = ~ 1| Fire)
  ModelName <- paste("M", i, sep="_")
  add <- data.frame(c(ModelName),c(vars), c(AIC(M)))
  table <- rbind(table, add)
}
colnames(table) <- c("Model","Variables", "AIC")
table_final <- kable(table %>% arrange(AIC), format = "markdown") 
head(table_final, n = 10)
```

# Likelihood ratio test using the top two models
```{r}
model1 <- lme(RGR ~ log(vol_pre_cm2),data = df, random = ~ 1| Fire)
model2 <- lme(RGR ~ Cov1.3_norm+BasDia2016.cm+tree_Ht_log_norm,data = df, random = ~ 1| Fire)
summary(model1)
summary(model2)
nullmodel <- lme(RGR ~ 1, random = ~1|Fire, data= df)
summary(nullmodel)
```

# Likelihood ratio test using variables from VSURF
```{r}
modelvsurf <- lme(RGR ~ Cov1.3+BasDia2016.cm + ShrubSpp02 + Ht1.3 + vol_pre_cm2,data = df, random = ~ 1| Fire, method = "ML")
modelvsurfnull <- lme(RGR ~ BasDia2016.cm + vol_pre_cm2, data = df, random = ~ 1| Fire, method = "ML")
summary(modelvsurf)
anova(modelvsurf, modelvsurfnull)
```


### Residuals
```{r}
E <- resid(model1, type = "normalized")
pred <- predict(model1)
plot(model1)
plot(df$shrubarea2_sqrt_norm, E)
plot(df$tree_Ht_log_norm, E)
plot(log(df$vol_pre_cm2), E)
```

Do the residuals look good?

# Plot

## Make fake data
```{r}
MyData <-  rbind(
  expand.grid( 
    Fire = unique(df$Fire),
    tree_Ht_cm = mean(df$tree_Ht_cm),
    Cov1.2 = seq(min(df$Cov1.2), max(df$Cov1.2), length = 20)
    ) 
)
```

## normalize MyData
```{r}
MyData$tree_Ht_norm <- (MyData$tree_Ht_cm - mean(df$tree_Ht_cm))/sd(df$tree_Ht_cm)
MyData$tree_Ht_log_norm <- (log(MyData$tree_Ht_cm) - mean(log(df$tree_Ht_cm)))/sd(log(df$tree_Ht_cm))
MyData$Cov1.2_norm <- (MyData$Cov1.2 - mean(df$Cov1.2))/sd(df$Cov1.2)
```

```{r}
MM1
MyData$pred <- predict(MM1, newdata = MyData)
```

### Summary plot
```{r}
MyData_summary_simple <- MyData %>% 
  group_by(Cov1.2) %>%
  mutate(pred_simple = mean(pred)) %>% 
  ungroup()

brown <- "#5b3d07"
figure_summary <- ggplot(MyData_summary_simple)+
  labs(title = "Fir seedling volumetric growth")+
  ylab("relative growth rate (RGR)")+
  xlab("shrub cover 0-2 m from seedling")+
  geom_line(aes(y = pred_simple, x =Cov1.2), col = "#1b3f4c")+
  geom_point(data=df, aes(x = Cov1.2, y = RGR), col = "#1b3f4c")+
  theme_bw()+
  theme(plot.background = element_rect(fill = '#e9ebe8', colour = '#e9ebe8'),
        panel.background = element_rect(fill="#e9ebe8"),
        text = element_text(size=20, colour = brown),
        axis.text.x = element_text(colour=brown), 
        axis.text.y = element_text(colour=brown),
        panel.border = element_rect(colour = brown)
        )+
  scale_y_continuous(limits = c(0,1.12))
figure_summary
```

# Save figure
```{r}
setwd("C:/Users/Carmen/Documents/Shrubs-Seedlings/results/figures")
pdf("fir_RGR.pdf", width = 5.5, height = 7.25, pointsize = 30,useDingbats = F)
figure_summary
dev.off()
```

# Calculate pseudo r squared
```{r}
meanGrth <- mean(df$RGR)
sse <- (df$RGR - predict(MM1))^2
sst <- (df$RGR - meanGrth)^2
R2 <- 1 - (sum(sse)/sum(sst))
R2
```

# Save final data
```{r}
save(df, file = "~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_vol.Rdata")
```
