---
title: "Growth Mortality Initial Tree Ring Analysis"
author: "Carmen"
date: "December 11, 2018"
output: 
  html_document:
    toc: TRUE
---

# Sources: 

-  https://rpubs.com/andybunn/r4tree-rings
-  Speer, J. H. (2010). Fundamentals of tree-ring research. Tucson: University of Arizona Press.
-  to ask an expert questions, go to itrdbfor.org

# General notes:

-  **Standardization** = **Detrending** = removing age-related growth trends and other long-term variability that can be considered noise
    - often, a climatologist's signal is an ecologist's noise
    - produces a series mean of 1
    - Technique #1: **negative exponential curve**
        - determined by the geometry of how much narrower a ring gets when you add the same volume of wood to a larger cylinder
        - deterministic = same eqn for every data set
        - most common method
        - `ModNegEx` method in `detrend` in `dplR` tries to fit a nonlinear model, and if that doesn't work it fits a linear model
    - Technique #2: **cubic smoothing spline**
        - a 40 year one removes 50% of variance, removes 99% of variance at 126 years, and removes only 1% of variance at 12.67 years
        - more flexible than negative exponential curve
    - Technique #3: **regional curve standardization** 
        - advantage: splines often remove long-term trends that are not age-related, like the Little Ice Age
        - based on the biological age of each ring
            - therefore you have to age each core
        - regional curve = average growth for that stand, which is removed from each core
        - must have at least 5 trees for any section of the curve and at least 40 trees at densest part of the curve 
    - Technique #4: **signal-free method**
        - don't know anything about this other than that Andy Bunn says there's been a lot of interest in including it in `dplR` but it's not there yet
    - once a `rwl` data set is detrended, it becomes a `rwi` (ring-width index) file
        - at which point every series has a mean of 1
    - use `detrend` in `dplR` to detrend
    - sometimes people detrend different series in the same data set using different methods, which you could do using `i.detrend.series`
        
        
```{r, include = F}
require(dplR)
require(tidyverse)
```

# Example with built-in data
```{r}
data(ca533)
```

## This is a special kind of data frame specific to dendro analysis
```{r}
class(ca533)
```

In order to create a rwl data frame, you use the function `read.rwl`

## Time (years) is an attribute of this class
```{r}
head(time(ca533))
```

# Rwl data frames can easily be plotted in a dendro way
```{r}
plot(ca533, plot.type = "spag")
```

# And described with descriptive stats
```{r}
rwl.report(ca533)
summary(ca533)
```

**Note: Bunn thinks that mean sensitivity is a terrible statistic that should not be used

## which you can then plot
```{r}
ggplot(summary(ca533))+
  geom_boxplot(aes(y = ar1))+
  geom_jitter(aes(y = ar1, x = 0), width = .01)
```

# Detrend
```{r}
ca533.rwi <- detrend(rwl = ca533, method = "ModNegExp")
class(ca533.rwi)
```

## This results in each column (series) having a mean of 1
```{r}
summary(colMeans(ca533.rwi, na.rm = T))
```

## You can also detrend a single series at a time
```{r}
CAM011.rwi <- detrend.series(y = ca533[, "CAM011"], verbose = TRUE)
```

This shows the performance of multiple detrending options

# Crossdating a sample data set
```{r}
data(co021)
dat <- co021
```

