---
title: "simulations_plot_ShrubSpecies"
author: "Carmen"
date: "3/6/2020"
output: html_document
---

# If you're running this code on its own, you have to define data_date and n_reps at the top

```{r}
strt <- Sys.time()
```

```{r, include = F}
require(tidyverse)
require(gridExtra)
require(egg)
require(grid)
```

```{r}
data_date <<- "2020-12-17"
n_reps <- 1000
```


```{r}
namemin <- paste("~/Ch3_Simulation_Results/Simulation15_",  data_date, "_", n_reps, "_MIN.Rdata", sep = "")

namemedian <- paste("~/Ch3_Simulation_Results/Simulation16_", data_date, "_", n_reps, "_MEDIAN.Rdata", sep = "")

namemax <- paste("~/Ch3_Simulation_Results/Simulation17_", data_date, "_", n_reps, "_MAX.Rdata", sep = "")
```

```{r}
load(namemin)
min <- dfsimallreps
remove(dfsimallreps)

load(namemedian)
median <- dfsimallreps
remove(dfsimallreps)

load(namemax)
max <- dfsimallreps
remove(dfsimallreps)
```

```{r}
min <- min %>% 
  mutate(sim_spp = "MIN")
median <- median %>% 
  mutate(sim_spp = "MEDIAN")
max <- max %>% 
  mutate(sim_spp = "MAX")
```

```{r}
nrow_total <- nrow(min) + nrow(median) + nrow(max)
``` 

```{r}
df <- full_join(min, median)
remove(min, median)
df <- full_join(df, max)
remove(max)
```

```{r}
nrow_total == nrow(df)
```

```{r}
df %>% 
  filter(!is.finite(pred_exp) | is.na(pred_exp)) %>% 
  group_by(emerged, Years, sim_spp) %>% 
  count()
```

<!-- ```{r} -->
<!-- df <- df %>%  -->
<!--  mutate(pred_exp = ifelse(!is.finite(pred_exp) | !is.na(pred_exp), 999, pred_exp)) -->
<!-- ``` -->


## Summarize

### Mean
```{r}
summary_emerge <- df %>% 
  group_by(Years, Species, rep, sim_spp) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species, sim_spp) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, Species, sim_spp) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```

### Test
```{r}
df %>% 
  filter(Years == 40) %>% 
  group_by(sim_spp, Species) %>% 
  summarize(mean())
```


## Filter out the flat parts - take out years if they differ by <0.001 (0.1%)
```{r}
max_year <- summary_emerge %>% 
  mutate(mean = round(mean, 3)) %>% 
  group_by(Species, sim_spp) %>% 
  mutate(maxmean = max(mean)) %>% 
  filter(mean == maxmean) %>% 
  ungroup() %>%
  group_by(Species, sim_spp) %>% 
  filter(Years == min(Years)) %>%
  ungroup() %>% 
  filter(Years ==max(Years)) %>% 
  dplyr::select(Years) %>% 
  unlist()

max_year
```

```{r}
summary_emerge <- summary_emerge %>% 
  filter(Years < unique(max_year) + 1)
```

# Save summarized data
```{r}
filename <- paste("../../../../results/data/summary_emerge_prop_gradient_", data_date,"_", n_reps, ".Rdata", sep = "")
save(summary_emerge, file = filename)
```

