---
title: "simulations_plot_ShrubSpecies"
author: "Carmen"
date: "3/6/2020"
output: 
  html_document:
    toc: TRUE
---

# If you're running this code on its own, you have to define data_date and n_reps at the top

```{r}
strt <- Sys.time()
```

```{r, include = F}
require(tidyverse)
require(gridExtra)
require(egg)
require(grid)
```

```{r}
data_date <<- "2020-11-02"
n_reps <<- "20"
```

```{r}
namecein <- paste("~/Ch3_Simulation_Results/Simulation5_", data_date, "_", n_reps, "_CEIN.Rdata", sep = "")

namearpa <- paste("~/Ch3_Simulation_Results/Simulation4_", data_date, "_", n_reps, "_ARPA.Rdata", sep = "")

namececo <- paste("~/Ch3_Simulation_Results/Simulation3_", data_date, "_", n_reps, "_CECO.Rdata", sep = "")
```

```{r}
load(namearpa)
arpa <- dfsimallreps
remove(dfsimallreps)

load(namececo)
ceco <- dfsimallreps
remove(dfsimallreps)

load(namecein)
cein <- dfsimallreps
remove(dfsimallreps)
```

```{r}
arpa <- arpa %>% 
  mutate(sim_spp = "ARPA")
ceco <- ceco %>% 
  mutate(sim_spp = "CECO")
cein <- cein %>% 
  mutate(sim_spp = "CEIN")
```

```{r}
nrow_total <- nrow(arpa) + nrow(ceco) + nrow(cein)
``` 

```{r}
df <- full_join(arpa, ceco)
remove(arpa, ceco)
df <- full_join(df, cein)
remove(cein)
```

This should be TRUE:
```{r}
nrow_total == nrow(df)
```

```{r}
df %>% 
  filter(!is.finite(pred_exp) | is.na(pred_exp)) %>% 
  group_by(emerged, Years) %>% 
  count()
```

<!-- ```{r} -->
<!-- df <- df %>%  -->
<!--  mutate(pred_exp = ifelse(!is.finite(pred_exp) | !is.na(pred_exp), 999, pred_exp)) -->
<!-- ``` -->


## Summarize

### Mean
```{r}
summary_emerge <- df %>% 
  group_by(Years, Species, rep, sim_spp) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species, sim_spp) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, Species, sim_spp) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```


## Filter out the flat parts - take out years if they differ by <0.001 (0.1%)
```{r}
max_year <- summary_emerge %>% 
  mutate(mean = round(mean, 3)) %>% 
  group_by(Species, sim_spp) %>% 
  mutate(maxmean = max(mean)) %>% 
  filter(mean == maxmean) %>% 
  ungroup() %>%
  group_by(Species, sim_spp) %>% 
  filter(Years == min(Years)) %>%
  ungroup() %>% 
  filter(Years ==max(Years)) %>% 
  dplyr::select(Years) %>% 
  unlist()

max_year
```

```{r}
summary_emerge <- summary_emerge %>% 
  filter(Years < unique(max_year) + 1)
```

# Save summarized data
```{r}
filename <- paste("../../../../results/data/summary_emerge_prop_ShrubSpp_", data_date,"_", n_reps, ".Rdata", sep = "")
save(summary_emerge, file = filename)
```

