---
title: 'Analysis 5, Seedling Growth by genus : Pinus'
author: "Carmen"
date: "March 3, 2017"
output: html_document
---

- Main goal: see if it makes sense to add pila to existing pipo seedling growth model

```{r, include=F}
setwd("~/Shrubs-Seedlings/Rdata")
load(file="master_data.Rdata")
library(nlme)
```

## Look for the model with the best AIC

```{r}
dfPinus <- df[df$Species %in% c("PIPO","PILA") & !df$Fire %in% c("PLKN","WRTS"),]
varPower <- varPower(form=~Ht.cm)

table <- data.frame(model=character(),
                 AIC=numeric(), stringsAsFactors=FALSE)

M1F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M1F.pinus)
table[1,] <- c("M1F.pinus", AIC(M1F.pinus))


M2F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire  + Fire*Ht.cm +IAG + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M2F.pinus)
table[2,] <- c("M2F.pinus", AIC(M2F.pinus))

M3F.pinus <- gls(LastYearGrth.cm ~  Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire  + Fire*Ht.cm +IAG ,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M3F.pinus)
table[3,] <- c("M3F.pinus", AIC(M3F.pinus))

M4F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.3 + BasDia.cm + Ht1.3 + Fire  + Fire*Ht.cm +Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M4F.pinus)
table[4,] <- c("M4F.pinus", AIC(M4F.pinus))

M5F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + Ht1.3 + Fire  + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M5F.pinus)
table[5,] <- c("M5F.pinus", AIC(M5F.pinus))

M6F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.2 + Fire + Fire*Ht.cm + Ht.cm*Ht1.2,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M6F.pinus) 
table[6,] <- c("M6F.pinus", AIC(M6F.pinus))

M7F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M7F.pinus)
table[7,] <- c("M7F.pinus", AIC(M7F.pinus))

table
```

## Try to refine further by using different cover segments and height segments

```{r}
M8F.pinus <- gls(LastYearGrth.cm ~ Ht.cm +  BasDia.cm + Cov1.3 + Ht1.2 + Fire + Fire*Ht.cm + Ht.cm*Ht1.2,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M8F.pinus) 
table <- rbind(table, c("M8F.pinus", AIC(M8F.pinus)))

M9F.pinus = update(M8F.pinus, . ~ . - Cov1.3 + Cov1.2)
AIC(M9F.pinus)
table <- rbind(table, c("M9F.pinus", AIC(M9F.pinus)))

M10F.pinus = update(M8F.pinus, . ~ . - Cov1.3 + Cov1)
AIC(M10F.pinus)
table <- rbind(table, c("M10F.pinus", AIC(M10F.pinus)))

M11F.pinus = update(M8F.pinus, . ~ . - Cov1.3 + Cov1.2 - Ht1.2 + Ht1.3 - Ht.cm*Ht1.2 + Ht.cm*Ht1.3)
AIC(M10F.pinus)
table <- rbind(table, c("M11F.pinus", AIC(M11F.pinus)))

M12F.pinus = update(M1F.pinus, . ~ . - BasDia.cm)
AIC(M12F.pinus)
table <- rbind(table, c("M12F.pinus", AIC(M12F.pinus)))
anova(M12F.pinus, M1F.pinus)

M13F.pinus =  update(M1F.pinus, . ~ . + ImmedAboveHt.cm)
AIC(M13F.pinus)
table <- rbind(table, c("M13F.pinus", AIC(M13F.pinus)))

M13F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M13F.pinus)

table
```


## Show these models graphically 

```{r}
pred <- as.data.frame(predict(M13F.pinus))
pred$Cov <- dfPinus$Cov1.2
pred$Ht1.2 <- dfPinus$Ht1.2
pred$Ht.cm <- dfPinus$Ht.cm
pred$Fire <- dfPinus$Fire
pred$Grth <- dfPinus$LastYearGrth.cm
pred$ShrG1 <- dfPinus$ShrG1
pred$Species <- dfPinus$Species
#pred <- arrange(pred, predict(M3.lme))
library(ggplot2)
ggplot(pred)+
  geom_point(aes(y=predict(M13F.pinus),x=Ht.cm,col=pred$Ht1.2, shape=pred$Fire))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted values, Model 6F:Pinus")
ggplot(pred)+
  geom_point(aes(y=Grth,x=Ht.cm,col=pred$Ht1.2, shape=pred$Fire))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Measured values")
pila <- pred[pred$Species=="PILA",]
ggplot(pred)+
  geom_point(aes(y=Grth,x=Ht.cm,col=pred$Ht1.2, shape=pred$Fire))+
  geom_point(data=pila, aes(y=pila[1],x=pila$Ht.cm, size=2, shape=pila$Fire, col=pila$Ht1.2))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Measured values, with larger symbols for PILA than PIPO")

E.pinus <- resid(M13F.pinus, type="normalized")
coplot(E.pinus ~ Ht.cm | Fire, data=dfPinus, ylab="Normalized residuals", col=dfPinus$Fire)
```

- The predicted values and the residuals look good
- Are the different slopes caused by different shrub heights or covers in different patches (ie covariance between fire and shrub heights or covers or shrub species?)
```{r}
ggplot(pred)+
  geom_point(aes(y=predict(M1F.pinus),x=Ht.cm,col=pred$ShrG1, shape=pred$Fire))+
  labs(title = "Relationship between shrub species and fire, Model 6F:Pinus")
```

- It appears there is a relationship but the shrub species in the flatter-sloped fires are also in other fires
- Potential conclusions from comparing abco and pinus results
    - basal diameter matters for ABCO but not for Pinus
    - shrub cover almost matters for Pinus but not for ABCO


## Now double check that I'm using the best GLS 

```{r}
library(nlme)
# None
M1Fa.pinus <-  gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3 ,
                 data= dfPinus, method = "ML")
AIC(M1Fa.pinus)
table <- rbind(table, c("M1Fa.pinus", AIC(M1Fa.pinus)))
      
table[2,] <- c("M2F.pinus", AIC(M2F.pinus))
  
M1Fb.pinus <- gls(LastYearGrth.cm ~Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3 ,
                weights = varExp(form=~ Ht.cm),
                 data= dfPinus, method = "ML")
AIC(M1Fb.pinus)
table <- rbind(table, c("M1Fb.pinus", AIC(M1Fb.pinus)))

M1Fc.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3 ,
                weights = varConstPower(form=~Ht.cm),
                 data= dfPinus, method = "ML")
AIC(M1Fc.pinus)
table <- rbind(table, c("M1Fc.pinus", AIC(M1Fc.pinus)))

table
```

## Try the same model but with DIFN!

```{r}
dfPinus.D <- dfPinus[!is.na(dfPinus$DIFN),]
plot(dfPinus.D$DIFN, dfPinus.D$LastYearGrth.cm)
tableDIFN <- data.frame(model=character(),
                 AIC=numeric(), stringsAsFactors=FALSE)

M1D.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + DIFN + Fire +Fire * Ht.cm + Ht.cm*DIFN,
                weights = varPower,
                 data= dfPinus.D, method = "ML")
AIC(M1D.pinus)
tableDIFN[1,] <- c("M1D.pinus", AIC(M1D.pinus))

M2D.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3,
                weights = varPower,
                 data= dfPinus.D, method = "ML")
AIC(M2D.pinus)
anova(M1D.pinus, M2D.pinus)

M3D.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + DIFN + Fire +Fire * Ht.cm,
                weights = varPower,
                 data= dfPinus.D, method = "ML")
AIC(M3D.pinus)

E1D.pinus <- resid(M1D.pinus, type="normalized")
coplot(E1D.pinus ~ Ht.cm | Fire, data=dfPinus.D, ylab="Normalized residuals", main="DIFN model residuals", col=dfPinus.D$Fire)


M2D.pinus <- lme(LastYearGrth.cm ~ Ht.cm + Ht1 + BasDia.cm,
                random=~1+Ht1|Fire,
                weights = varPower,
                 data= dfPinus.D,
                method = "ML")
AIC(M2D.pinus)

M4D.pinus <- gls(LastYearGrth.cm ~ Ht.cm + + DIFN + Fire + Fire*Ht.cm,
                weights = varPower,
                 data= dfPinus.D, method = "ML")
AIC(M4D.pinus)

M5D.pinus <- gls(LastYearGrth.cm ~ Ht.cm + DIFN + Ht1.2 + Fire + Fire*Ht.cm,
                weights = varPower,
                 data= dfPinus.D, method = "ML")
AIC(M5D.pinus)
tableDIFN <- rbind(tableDIFN, c("M5D.pinus", AIC(M5D.pinus)))

M6D.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Ht1.2 + Fire + Fire*Ht.cm,
                weights = varPower,
                 data= dfPinus.D, method = "ML")
AIC(M6D.pinus)
tableDIFN <- rbind(tableDIFN, c("M6D.pinus", AIC(M6D.pinus)))

tableDIFN

```

## Main Findings

1. The best model to describe Pinus seedling growth with the data I have is 
```{r}
M1F.pinus <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M1F.pinus)

M1F.pinus.2 <- gls(LastYearGrth.cm ~ Ht.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "ML")
AIC(M1F.pinus.2)
```

  - Must rerun it with REML for parameter estimation - important!
  - For the purposes of SORTIE, having basal diameter in there is not that helpful. Plus it is only marginally significant in the model. Also, the effect of shrub cover is tiny so I'll delete that too. So the SORTIE model should be 

```{r}
SORTIE_pinus <- gls(LastYearGrth.cm ~ Ht.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPinus, method = "REML")
summary(SORTIE_pinus)
```

  - So this makes the final equation: 
  
  Height growth in cm = -.39 + 0.19*Ht.cm - .006*shrub ht 
  
2. DIFN is not useful for Pinus
3. Shrub cover is not super important but shrub height is
4. Shrub genus does not appear to be important

## Graphically display predictions for a set of fake values

```{r, include=F}
library(dplyr)
```

```{r}
Ht.cm <- c(rep(seq(0,300,5),4))
length(Ht.cm)
fakedata <- as.data.frame(Ht.cm)
fakedata$Ht1.3 <- c(rep(c(5,50,100,150),61))
fakedata$Fire <- as.factor(c(rep("AMCR",244)))
predicted <- predict(SORTIE_pinus, fakedata)

pred <- as.data.frame(predicted)
pred$Shrub_height <- fakedata$Ht1.3
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

dfCLVD <- subset(dfPinus, dfPinus$Fire=="CLVD" & !is.na(dfPinus$LastYearGrth.cm))
dfAMCR<- subset(dfPinus, dfPinus$Fire=="AMCR" & !is.na(dfPinus$LastYearGrth.cm))
dfFRDS <- subset(dfPinus, dfPinus$Fire=="FRDS" & !is.na(dfPinus$LastYearGrth.cm))
dfSTAR <- subset(dfPinus, dfPinus$Fire=="STAR" & !is.na(dfPinus$LastYearGrth.cm))

plot_AMCR <- ggplot(pred[pred$Shrub_height<51&pred$Ht.cm <= max(dfAMCR$Ht.cm),])+
  geom_line(aes(y= predicted ,x=Ht.cm,group=Shrub_height,col=Shrub_height))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted seedling growth,\nAmerican Complex Fire")+
  ylim(0,55)+
  xlim(0,300)+
  geom_point(data=dfAMCR,aes(y=LastYearGrth.cm,x=Ht.cm,col=Ht1.3))+
  xlab("Seedling height")+
  ylab("Seedling growth")
  
plot_AMCR

fakedata$Fire <- as.factor(c(rep("CLVD",244)))
predicted <- predict(SORTIE_pinus, fakedata)
pred <- as.data.frame(predicted)
pred$Shrub_height <- fakedata$Ht1.3
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire
pred$Grth <- fakedata$LastYearGrth.cm

predicted_clvd <- predicted

CLVD <- ggplot(pred[pred$Ht.cm<=max(dfCLVD$Ht.cm),])+
  geom_line(aes(y= predicted ,x=Ht.cm,group=Shrub_height,col=Shrub_height))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted seedling growth,\nCleveland Fire")+
  ylim(0,55)+
  xlim(0,300)+
  geom_point(data=dfCLVD,aes(y=LastYearGrth.cm,x=Ht.cm,col=Ht1.3))+
  xlab("Seedling height")+
  ylab("Seedling growth")


fakedata$Fire <- as.factor(c(rep("FRDS",244)))
predicted <- predict(SORTIE_pinus, fakedata)
pred <- as.data.frame(predicted)
pred$Shrub_height <- fakedata$Ht1.3
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire
pred$Grth <- fakedata$LastYearGrth.cm

predicted_freds <- predicted

FRDS <- ggplot(pred[pred$Ht.cm<=max(dfFRDS$Ht.cm),])+
  geom_line(aes(y= predicted ,x=Ht.cm,group=Shrub_height,col=Shrub_height))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted seedling growth,\nFreds Fire")+
  ylim(0,55)+
  xlim(0,300)+
  geom_point(data=dfFRDS,aes(y=LastYearGrth.cm,x=Ht.cm,col=Ht1.3))

FRDS

fakedata$Fire <- as.factor(c(rep("STAR",244)))
predicted <- predict(SORTIE_pinus, fakedata)
pred <- as.data.frame(predicted)
pred$Shrub_height <- fakedata$Ht1.3
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire
pred$Grth <- fakedata$LastYearGrth.cm

STAR <- ggplot(pred[pred$Shrub_height>51&pred$Ht.cm <= max(dfSTAR$Ht.cm),])+
  geom_line(aes(y= predicted,x=Ht.cm,group=Shrub_height,col=Shrub_height))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted seedling growth,\nStar Fire")+
  ylim(0,55)+
  xlim(0,300)+
  geom_point(data=dfSTAR,aes(y=LastYearGrth.cm,x=Ht.cm,col=Ht1.3))

STAR
```

```{r, include=F}
require(gridExtra)
```

```{r}
grid.arrange(STAR, FRDS,plot_AMCR,CLVD,ncol=2,nrow=2)

```

## How do these results inform my initial SORTIE inputs?

- How do I deal with non-continuous shrub cover given that this model only takes shrub height into consideration?
```{r}
hist(dfPinus$Cov1)
nrow(dfPinus[dfPinus$Cov1<100,])
```
- There are only 10 pine observations that have seedlings with less than 100 cm shrub cover across 4 1-m transects
- Perhaps for the seedlings with less than 100 cm shrub cover, I could use a separate growth function that is not dependent upon shrubs
    - I could derive this function from Welch data, maybe
    
### Add variable for whether the seedling is immediately overtopped by shrubs
```{r}

dfPinus <- tbl_df(dfPinus)
dfPinus <- dfPinus%>%
  mutate(Topped = ifelse(dfPinus$ImmedAboveHt.cm>0,1,0))

top_AMCR <- ggplot(data=dfPinus[dfPinus$Fire=="AMCR",], aes(y=LastYearGrth.cm, x=Topped))+
  geom_point(aes(colour = Fire))
top_STAR <- ggplot(data=dfPinus[dfPinus$Fire=="STAR",], aes(y=LastYearGrth.cm, x=Topped))+
  geom_point(aes(colour = Fire))
top_CLVD <- ggplot(data=dfPinus[dfPinus$Fire=="CLVD",], aes(y=LastYearGrth.cm, x=Topped))+
  geom_point(aes(colour = Fire))
top_FRDS <- ggplot(data=dfPinus[dfPinus$Fire=="FRDS",], aes(y=LastYearGrth.cm, x=Topped))+
  geom_point(aes(colour = Fire))
grid.arrange(top_STAR, top_FRDS,top_AMCR,top_CLVD,ncol=2,nrow=2)

box_AMCR <- ggplot(data=dfPinus[dfPinus$Fire=="AMCR",],aes(factor(Topped),LastYearGrth.cm))+
         geom_boxplot()
box_STAR <- ggplot(data=dfPinus[dfPinus$Fire=="STAR",],aes(factor(Topped),LastYearGrth.cm))+
         geom_boxplot()
box_CLVD <- ggplot(data=dfPinus[dfPinus$Fire=="CLVD",],aes(factor(Topped),LastYearGrth.cm))+
         geom_boxplot()
box_FRDS <- ggplot(data=dfPinus[dfPinus$Fire=="FRDS",],aes(factor(Topped),LastYearGrth.cm))+
         geom_boxplot()
grid.arrange(box_AMCR, box_FRDS, box_CLVD, box_STAR, ncol=2,nrow=2)

```

- Look at whether there's a pattern in the residuals for seedlings with different cover attributes that can inform me on where the shrub cover threshold is below which shrub height doesn't matter for seedling growth
- Also look at residuals for PILA compared to PIPO

- First, check that residuals are in the same order as the dfPinus data and predictions
```{r}
head(dfPinus$LastYearGrth.cm-residuals(SORTIE_pinus))
head(predict(SORTIE_pinus))
```
- Good. Now check out the residuals
```{r}
dfPinus$Predicted <- predict(SORTIE_pinus)
dfPinus$Residuals <- resid(SORTIE_pinus, type="normalized")
```


```{r}
ggplot(dfPinus)+
  geom_point(aes(y=Residuals,x=Cov1))+
  geom_point(data=dfPinus[dfPinus$Cov1.2<300,],aes(y=Residuals,x=Cov1.2),color="red",size=2)

mean(dfPinus[dfPinus$Topped==0,]$Residuals)
mean(dfPinus$Residuals)


ggplot()+
  geom_point(aes(y=residuals(SORTIE_pinus),x=names(residuals(SORTIE_pinus))))+
  geom_point(aes(y=residuals(SORTIE_pinus)[dfPinus$Topped==0],x=names(residuals(SORTIE_pinus)[dfPinus$Topped==0])),color="red",size=2)

length(residuals(SORTIE_pinus)[residuals(SORTIE_pinus)>0])
length(residuals(SORTIE_pinus)[residuals(SORTIE_pinus)<0])
length(residuals(SORTIE_pinus)[dfPinus$Cov1<300&residuals(SORTIE_pinus)<0])
length(residuals(SORTIE_pinus)[dfPinus$Cov1<300&residuals(SORTIE_pinus)>0])

```

