---
title: "Analysis 3 - Seedling Growth by Species: PIPO"
author: "Carmen"
date: "March 1, 2017"
output: html_document
---

- This is a based on some of what I learned in Analysis 2
    - for example, in terms of using VarPower
- Note: delete seedlings from PLKN and WRTS because there are only two PIPO in each
```{r, include=F}
setwd("~/Shrubs-Seedlings/Rdata")
load(file="master_data.Rdata")
library(nlme)
```

## Look for the model with the best AIC

```{r}
dfPIPO <- df[df$Species=="PIPO" & !df$Fire %in% c("PLKN","WRTS"),]
varPower <- varPower(form=~Ht.cm)

table <- data.frame(model=character(),
                 AIC=numeric(), stringsAsFactors=FALSE)

M1F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M1F.pipo)
AIC(M1F.pipo)
table[1,] <- c("M1F.pipo", AIC(M1F.pipo))

M2F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire  + Fire*Ht.cm +IAG + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M2F.pipo)
AIC(M2F.pipo)
table[2,] <- c("M2F.pipo", AIC(M2F.pipo))

M3F.pipo <- gls(LastYearGrth.cm ~  Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire  + Fire*Ht.cm +IAG ,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M3F.pipo)
AIC(M3F.pipo)
table[3,] <- c("M3F.pipo", AIC(M3F.pipo))

M4F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.3 + BasDia.cm + Ht1.3 + Fire  + Fire*Ht.cm +Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M4F.pipo)
AIC(M4F.pipo)
table[4,] <- c("M4F.pipo", AIC(M4F.pipo))

M5F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + Ht1.3 + Fire  + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M5F.pipo)
AIC(M5F.pipo)
table[5,] <- c("M5F.pipo", AIC(M5F.pipo))

M6F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.2 + Fire + Fire*Ht.cm + Ht.cm*Ht1.2,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M6F.pipo)
AIC(M6F.pipo) 
table[6,] <- c("M6F.pipo", AIC(M6F.pipo))

M7F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPIPO, method = "ML")
anova(M7F.pipo)
AIC(M7F.pipo)
table[7,] <- c("M7F.pipo", AIC(M7F.pipo))
```

## Try to refine further by using different cover segments and height segments

```{r}
M8F.pipo <- gls(LastYearGrth.cm ~ Ht.cm +  BasDia.cm + Cov1.3 + Ht1.2 + Fire + Fire*Ht.cm + Ht.cm*Ht1.2,
                weights = varPower,
                 data= dfPIPO, method = "ML")
AIC(M8F.pipo) 
anova(M8F.pipo)
table <- rbind(table, c("M8F.pipo", AIC(M8F.pipo)))

M9F.pipo = update(M8F.pipo, . ~ . - Cov1.3 + Cov1.2)
AIC(M9F.pipo)
table <- rbind(table, c("M9F.pipo", AIC(M9F.pipo)))

M10F.pipo = update(M8F.pipo, . ~ . - Cov1.3 + Cov1)
AIC(M10F.pipo)
table <- rbind(table, c("M10F.pipo", AIC(M10F.pipo)))

M11F.pipo = update(M8F.pipo, . ~ . - Cov1.3 + Cov1.2 - Ht1.2 + Ht1.3 - Ht.cm*Ht1.2 + Ht.cm*Ht1.3)
AIC(M10F.pipo)
table <- rbind(table, c("M11F.pipo", AIC(M11F.pipo)))

M12F.pipo = update(M1F.pipo, . ~ . - BasDia.cm)
AIC(M12F.pipo)
table <- rbind(table, c("M12F.pipo", AIC(M12F.pipo)))

M13F.pipo =  update(M1F.pipo, . ~ . + ImmedAboveHt.cm)
AIC(M13F.pipo)
table <- rbind(table, c("M13F.pipo", AIC(M13F.pipo)))
```


## Show these models graphically 

```{r}
pred <- as.data.frame(predict(M1F.pipo))
pred$Cov <- dfPIPO$Cov1.2
pred$Ht1.2 <- dfPIPO$Ht1.2
pred$Ht.cm <- dfPIPO$Ht.cm
pred$Fire <- dfPIPO$Fire
pred$Grth <- dfPIPO$LastYearGrth.cm
pred$ShrG1 <- dfPIPO$ShrG1
#pred <- arrange(pred, predict(M3.lme))
library(ggplot2)
ggplot(pred)+
  geom_point(aes(y=predict(M1F.pipo),x=Ht.cm,col=pred$Ht1.2, shape=pred$Fire))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted values, Model 6F:PIPO")
ggplot(pred)+
  geom_point(aes(y=Grth,x=Ht.cm,col=pred$Ht1.2, shape=pred$Fire))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Measured values")
E6F.pipo <- resid(M1F.pipo, type="normalized")
coplot(E6F.pipo ~ Ht.cm | Fire, data=dfPIPO, ylab="Normalized residuals", col=dfPIPO$Fire)
```

- The predicted values and the residuals look good
- Are the different slopes caused by different shrub heights or covers in different patches (ie covariance between fire and shrub heights or covers or shrub species?)
```{r}
ggplot(pred)+
  geom_point(aes(y=predict(M1F.pipo),x=Ht.cm,col=pred$ShrG1, shape=pred$Fire))+
  labs(title = "Relationship between shrub species and fire, Model 6F:PIPO")
```

- It appears there is a relationship but the shrub species in the flatter-sloped fires are also in other fires
- Potential conclusions from comparing abco and pipo results
    - basal diameter matters for ABCO but not for PIPO
    - shrub cover almost matters for PIPO but not for ABCO



## Now double check that I'm using the best GLS 

```{r}
library(nlme)
# None
M1Fa.pipo <-  gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3 ,
                 data= dfPIPO, method = "ML")
AIC(M1Fa.pipo)
table <- rbind(table, c("M1Fa.pipo", AIC(M1Fa.pipo)))
      
table[2,] <- c("M2F.pipo", AIC(M2F.pipo))
  
M1Fb.pipo <- gls(LastYearGrth.cm ~Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3 ,
                weights = varExp(form=~ Ht.cm),
                 data= dfPIPO, method = "ML")
AIC(M1Fb.pipo)
table <- rbind(table, c("M1Fb.pipo", AIC(M1Fb.pipo)))

M1Fc.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3 ,
                weights = varConstPower(form=~Ht.cm),
                 data= dfPIPO, method = "ML")
AIC(M1Fc.pipo)
table <- rbind(table, c("M1Fc.pipo", AIC(M1Fc.pipo)))
```


```{r}
# M6F.pipo <- gls(LastYearGrth.cm ~ Ht.cm +  BasDia.cm + Ht1.2 + Fire + Fire*Ht.cm,
#                 weights = varPower,
#                  data= dfPIPO, method = "ML")
# anova(M6F.pipo)
# AIC(M6F.pipo)
# 
# M6R.pipo <- gls(LastYearGrth.cm ~ Ht.cm +  BasDia.cm + Ht1.2 + Fire + Fire*Ht.cm,
#                 weights = varPower,
#                 random = ~1|FirePatch,  
#                  data= dfPIPO, method = "ML")
# anova(M6F.pipo)
# AIC(M6F.pipo)
```
## Try the same model but with DIFN!

```{r}
dfPIPO.D <- dfPIPO[!is.na(dfPIPO$DIFN),]
plot(dfPIPO.D$DIFN, dfPIPO.D$LastYearGrth.cm)
tableDIFN <- data.frame(model=character(),
                 AIC=numeric(), stringsAsFactors=FALSE)

M1D.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + DIFN + Fire +Fire * Ht.cm + Ht.cm*DIFN,
                weights = varPower,
                 data= dfPIPO.D, method = "ML")
AIC(M1D.pipo)
anova(M1D.pipo)
tableDIFN[1,] <- c("M1D.pipo", AIC(M1D.pipo))

M2D.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire +Fire * Ht.cm + Ht.cm * Ht1.3,
                weights = varPower,
                 data= dfPIPO.D, method = "ML")
AIC(M2D.pipo)
anova(M2D.pipo)
anova(M1D.pipo, M2D.pipo)

M3D.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + DIFN + Fire +Fire * Ht.cm,
                weights = varPower,
                 data= dfPIPO.D, method = "ML")
AIC(M3D.pipo)

E1D.pipo <- resid(M1D.pipo, type="normalized")
coplot(E1D.pipo ~ Ht.cm | Fire, data=dfPIPO.D, ylab="Normalized residuals", main="DIFN model residuals", col=dfPIPO.D$Fire)


M2D.pipo <- lme(LastYearGrth.cm ~ Ht.cm + Ht1 + BasDia.cm,
                random=~1+Ht1|Fire,
                weights = varPower,
                 data= dfPIPO.D,
                method = "ML")
AIC(M2D.pipo)
anova(M2D.pipo)

M4D.pipo <- gls(LastYearGrth.cm ~ Ht.cm + + DIFN + Fire + Fire*Ht.cm,
                weights = varPower,
                 data= dfPIPO.D, method = "ML")
AIC(M4D.pipo)

M5D.pipo <- gls(LastYearGrth.cm ~ Ht.cm + DIFN + Ht1.2 + Fire + Fire*Ht.cm,
                weights = varPower,
                 data= dfPIPO.D, method = "ML")
AIC(M5D.pipo)
tableDIFN <- rbind(tableDIFN, c("M5D.pipo", AIC(M5D.pipo)))

M6D.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Ht1.2 + Fire + Fire*Ht.cm,
                weights = varPower,
                 data= dfPIPO.D, method = "ML")
AIC(M6D.pipo)
tableDIFN <- rbind(tableDIFN, c("M6D.pipo", AIC(M6D.pipo)))

```

## Main Findings

1. The best model to describe PIPO seedling growth with the data I have is (must rerun it with REML)
```{r}
M1F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Cov1.2 + BasDia.cm + Ht1.3 + Fire + Fire*Ht.cm + Ht.cm*Ht1.3,
                weights = varPower,
                 data= dfPIPO, method = "REML")
```
2. DIFN is not useful for PIPO
3. Shrub cover is not important but shrub height is
4. Shrub genus does not appear to be important
5. The effect of shrub height on seedling growth is small compared to the effect of seedling height and fire 

## Graphically display predictions for a set of fake values

```{r}
Ht.cm <- c(rep(seq(0,300,5),4))
length(Ht.cm)
fakedata <- as.data.frame(Ht.cm)
fakedata$BasDia.cm <- c(rep(2,244))
fakedata$Ht1.3 <- c(rep(c(10,80,150,220),61))
fakedata$Fire <- as.factor(c(rep("AMCR",244)))
fakedata$Cov1.2 <- c(rep(c(200,200,200,200,800,800,800,800),30),200,200,200,200)

predicted <- predict(M1F.pipo, fakedata)

pred <- as.data.frame(predicted)
pred$Ht1.3 <- fakedata$Ht1.3
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire
pred$Grth <- fakedata$LastYearGrth.cm

AMCR <- ggplot(pred)+
  geom_point(aes(y= predicted ,x=Ht.cm,col=Ht1.3))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted seedling growth ~ \nHt, shrub height (Ht1.3) \nand shrub cover, American Complex \nFire, for best PIPO model")+
  ylim(0,55)

fakedata$Fire <- as.factor(c(rep("CLVD",244)))
predicted2 <- predict(M1F.pipo, fakedata)

pred2 <- as.data.frame(predicted2)
pred2$Ht1.3 <- fakedata$Ht1.3
pred2$Ht.cm <- fakedata$Ht.cm
pred2$Fire <- fakedata$Fire
pred2$Grth <- fakedata$LastYearGrth.cm

CLVD <- ggplot(pred2)+
  geom_point(aes(y= predicted2 ,x=Ht.cm,col=Ht1.3))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted seedling growth ~ \nHt, shrub height (Ht1.3) \nand shrub cover, Cleveland \nFire, for best PIPO model")+
  ylim(0,55)

require(gridExtra)
grid.arrange(AMCR, CLVD,ncol=2)

```

### What happens if you run a model for just the small seedlings? - haven't gotten there yet 
```{r}
# dfPIPO.s <- subset(dfPIPO, dfPIPO$Ht.cm < 70)
# nrow(dfPIPO.s)
# M1s.pipo <- lme(LastYearGrth.cm ~  Ht.cm + Ht1.3 + Fire + Fire*Ht.cm,
#                  data= dfPIPO.s)
# anova(M1s.pipo)
```

