---
title: "Clean and combine - Welch data"
author: "Carmen"
date: "March 9, 2017"
output: html_document
---

### This is my first go at the Welch data

```{r,include=F}
library(RODBC)
conn <- odbcDriverConnect("Driver={Microsoft Access Driver (*.mdb, *.accdb)};DBQ=C:/Users/Carmen/Dropbox (Stephens Lab)/SORTIE/Data from Hugh/RegenDatabase_1_13_16_updatedDates.accdb")
library(dplyr)
library(ggplot2)
```

## Import data from Access

```{r}
plot <- sqlFetch(conn, "Plot_data")
regen <- sqlFetch(conn, "tree_regen")
dist <- sqlFetch(conn, "nearest_regen")
seed_trees <- sqlFetch(conn, "seed_tree")
trees <- sqlFetch(conn, "surviving_trees")
```

*can I use the num_dead column to get at mortality? - well, only 5 plots have dead recorded*

## Combine regen and plot data
```{r}
plot <- tbl_df(plot)
regenselect <- regen %>% select(ID,Regen_Plot,OldPLOTID,Count_total,`0yr`,Ct_1yr,Ct_2yr,Ct_3yr,Ct_4yr,Ct_5yr,`6yr`,`7yr`,`8yr`,`9yr`,`10yr`,`11+yr`,Type,Species)
regenselect[is.na(regenselect)] <- 0
regenselect <- tbl_df(regenselect)
plot <- plot %>% select(ID,Regen_Plot,Fire,SHRUB,FIRE_SEV,`Year of Fire`,Date)
plot[is.na(plot)] <- 0
plot$ObsYr <- substr(plot$Date,1,4)
plot$ObsYr <- as.numeric(plot$ObsYr)
plot$Years <- plot$ObsYr-plot$`Year of Fire`
regenplot <- left_join(regenselect,plot,by="Regen_Plot") %>%
  filter(FIRE_SEV>3) %>%
  filter(SHRUB>20)
```

## Gather age columns and rename ages
```{r}
library(tidyr)
regentidy <- gather(regenplot, "age","count",5:16)
```

- rename ages
```{r}
ages <- unique(regentidy$age)
for(i in 0:length(ages)){
  regentidy$age[regentidy$age==ages[(i+1)]]<- i
}
regentidy$count[is.na(regentidy$count)]<- 0
regentidy$age <- as.numeric(regentidy$age)
regentidy$count <- as.numeric(regentidy$count)
regentidy <- regentidy[regentidy$count!=0,]
```

## Plot by age, for pines and ABCO

```{r}

regen_pinus <- regentidy[regentidy$Species %in% c('PIPO','PILA','PINUS','PIJE'),]

regen_abco <- regentidy[regentidy$Species %in% c('ABCO','ABIES'),]

ggplot(data=regen_pinus,aes(age,fill=as.factor(Years)))+
  geom_bar(position="dodge")+
  labs(title="Plots with pine seedlings, all plots in high sev with shrub cover>20%")+
  scale_x_continuous(breaks=c(0:11))+
  guides(fill=guide_legend(title="Years since fire"))

ggplot(data=regen_pinus,aes(age,count))+
  stat_summary(fun.y=sum,geom="bar",aes(fill=as.factor(Years)),position="dodge")+
    scale_x_continuous(breaks=c(0:11))+
  labs(title="Pine seedling totals, all plots in high sev with shrub cover>20%")+
   guides(fill=guide_legend(title="Years since fire"))

ggplot(data=regen_abco,aes(age,fill=as.factor(Years)))+
  geom_bar(position="dodge")+
  labs(title="Plots with ABCO seedlings, all plots in high sev with shrub cover>20%")+
  scale_x_continuous(breaks=c(0:11))+
  guides(fill=guide_legend(title="Years since fire"))
```

## By species, only plots with >50% shrub cover

```{r}
ggplot(data=regen_abco,aes(age,count))+
  stat_summary(fun.y=sum,geom="bar",aes(fill=as.factor(Years)),position="dodge")+
    scale_x_continuous(breaks=c(0:11))+
  labs(title="ABCO seedling totals, all plots in high sev with  shrub cover>50%")+
   guides(fill=guide_legend(title="Years since fire"))

ggplot(data=regen_pinus[regen_pinus$SHRUB>50,],aes(age,fill=as.factor(Years)))+
  geom_bar(position="dodge")+
  labs(title="Plots with pine seedlings, all plots in high sev with shrub cover>50%")+
  scale_x_continuous(breaks=c(0:11))+
  guides(fill=guide_legend(title="Years since fire"))

ggplot(data=regen_pinus[regen_pinus$SHRUB>50,],aes(age,count))+
  stat_summary(fun.y=sum,geom="bar",aes(fill=as.factor(Years)),position="dodge")+
    scale_x_continuous(breaks=c(0:11))+
  labs(title="Pine seedling totals, all plots in high sev with shrub cover>50%")+
   guides(fill=guide_legend(title="Years since fire"))


ggplot(data=regen_abco[regen_abco$SHRUB>50,],aes(age,fill=as.factor(Years)))+
  geom_bar(position="dodge")+
  labs(title="Plots with ABCO seedlings, all plots in high sev with shrub cover>50%")+
  scale_x_continuous(breaks=c(0:11))+
  guides(fill=guide_legend(title="Years since fire"))

ggplot(data=regen_abco[regen_abco$SHRUB>50,],aes(age,count))+
  stat_summary(fun.y=sum,geom="bar",aes(fill=as.factor(Years)),position="dodge")+
    scale_x_continuous(breaks=c(0:11))+
  labs(title="ABCO seedling totals, all plots in high sev withshrub cover>50%")+
   guides(fill=guide_legend(title="Years since fire"))


```


## Look at zero y/o seedlings to understand seedling establishment

```{r}
nrow(regen_abco[regen_abco$age==0,])
summary(as.factor(regen_abco[regen_abco$age==0,]$Years))
```

- There are only 5 plots with an 0-yr abco seedlings! So that's interesting but doesn't allow me to model abco establishment
- That's how few abco established out of how many seedlings? Very few!
```{r}
sum(regen_abco$count)
5/1002
```

## What about 1-yr-olds?
```{r}
sum(regen_abco$count)
5/1002
ggplot(data=regen_abco[regen_abco$age<2,],aes(Years))+
  geom_bar()+
  labs(title="Plots with 0 or 1 year-old fir seedlings, all plots in high sev with shrub cover>20%")
```


```{r}
nrow(regen_pinus[regen_pinus$age==0,])
sum(regen_pinus$count)
32/1104
ggplot(data=regen_pinus[regen_pinus$age==0,],aes(Years))+
  geom_bar()+
  labs(title="Plots with 0 year-old pine seedlings, all plots in high sev with shrub cover>20%")
nrow(regen_abco[regen_abco$age<2,])
summary(regen_abco[regen_abco$age<2,]$Fire)
summary(regen_abco[regen_abco$age<2,]$SHRUB)


```

-1-yr-old pines:

```{r}
nrow(regen_pinus[regen_pinus$age<2,])
ggplot(data=regen_pinus[regen_pinus$age<2,],aes(Years))+
  geom_bar()+
  labs(title="Plots with 0 or 1 year-old pine seedlings, all plots in high sev with shrub cover>20%")
summary(regen_pinus[regen_pinus$age<2,]$Fire)
summary(regen_pinus[regen_pinus$age<2,]$SHRUB)

```

## Main findings:

- Establishment of 0-yr-old seedlings in these plots 5-12 years old appears to be almost negligible
- This introduces a basic design question I need to address in SORTIE: will I model seedling establishment, and if so how?
- I could take a closer look at my American Fire data in high severity areas
- There seems to be a peak in seedling establishment 2-4 years after fire - either that or mortality is high (or the seedlings were aged poorly)
- I could potentially model establishment using my small sample size 