---
title: "diameter growth analysis - shrub removal"
author: "Carmen"
date: "November 9, 2017"
output: 
    html_document:
        toc: TRUE
---

This code analyzes how shrub removal affected seedling radial growth.

```{r, include = F}
library(dplyr)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
```

# Load data
```{r}
df <- read.csv(file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia.csv")
dfshrub <- read.csv(file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/dfshrub.csv")
```

# Add shrub data
```{r}
df <- full_join(df, dfshrub)
summary(df$shrubarea)
```

# Look at sample sizes
```{r}
df %>% group_by(species, shrub_species, shrubs) %>% 
  summarise(n())
```

# Check for missing values
```{r}
summary(df$dia_growth_rel)
summary(df$species)
summary(df$shrub_species)
summary(df$ht_cm)
```

# Look at height distributions by shrub species
```{r}
ggplot(df)+
  geom_boxplot(aes(x = interaction(shrub_species, species), y = ht_))
```

# How does variance of RGR compare for different heights?
```{r}
sd(unlist(df[df$ht_cm < median(df$ht_cm),"RGR"]))
sd(unlist(df[df$ht_cm > median(df$ht_cm),"RGR"]))
```

Taller seedlings have nearly half the variance of shorter seedlings. This hints we'll need a GLS model. 

# How many of the measurements are below 0?
```{r}
df %>% 
  summarise(n_neg = sum(dia_growth_rel<0), n_pos = sum(dia_growth_rel>=0), n_total = n()) 
```

# Stats

## Distribution of diameter growth
```{r}
hist(df$dia_growth_rel, breaks = 30)
```

Looks pretty normal.

## Mixed effects with island effect
```{r}
MM1 <- lme(dia_growth_rel ~ shrubs + ht_cm +  species + shrub_species + shrubarea + shrubs:species + shrubs:shrub_species + shrubs:ht_cm, random =~1|island, data = df, method = "ML")
MM1_noremoval <- lme(dia_growth_rel ~ ht_cm + species + shrub_species + shrubarea, random =~1|island, method = "ML",data = df)
anova(MM1, MM1_noremoval)
```

### Residuals
```{r}
E <- residuals(MM1)
pred <- predict(MM1)
plot(MM1, which=c(1))
plot(df$shrubs, E)
plot(df$species, E)
plot(df$shrub_species, E)
```

# Plot predictions

### Create MyData
```{r}
MyData <- expand.grid(shrubs = c("retained", "removed"), ht_cm = c(seq(10,300, length = 20)), species = c("ABCO", "PIPO"), shrub_species = c("CEIN","CECO","ARPA"), shrubarea = seq(33000,120000, length = 20))
```

### Add pred
```{r}
MyData$predMM <- predict(MM1, level = 0, newdata = MyData)
```

## Overall effect of removal on diameter growth
```{r}
MyData_summary <- MyData %>% 
  group_by(shrubs, shrub_species, species) %>% 
  summarise(mean_predMM = mean(predMM))
figure <- ggplot(MyData_summary,aes(x = shrubs, y = mean_predMM)) +
  geom_point(aes(col = species, shape = shrub_species), size = 2)+
  geom_line(aes(col = species, linetype = shrub_species, group = interaction(shrub_species, species)))+
  theme_bw()+
  scale_color_manual(values=c("#1a9850", "#d73027"))+
  labs(y = "Predicted relative diameter growth")+
  ggtitle("Shrub removal increased diameter growth,\nespecially for whitethorn")
setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/results/figures/")
png("dia_growth.png")
figure
dev.off()
```

### Plot effect of height
```{r}
MyData$ht_cm <- as.numeric(paste(MyData$ht_cm))
MyData_height_summary <-  MyData %>% 
  group_by(species, shrubs, ht_cm) %>% 
  summarise(predMM = mean(predMM))
  
ggplot(MyData_height_summary)+
  geom_line(aes(x = ht_cm, y = predMM, group = interaction(shrubs, species), col = species, linetype = shrubs))+
  scale_color_manual(values = c('#d73027','#66bd63','#f46d43','#a6d96a','#fdae61','#1a9850'))+
  theme_bw()+
  ggtitle("The effect of removal was stronger for larger seedlings")+
  labs(y = "Predicted relative growth rate")+
  labs(x = "Seedling height in 2017 (cm)")
```

The taller the seedling, the smaller the effect of removal. Makes sense.

### Effect of pre-treatment shrub cover 
```{r}
MyData_shrubarea_summary <- MyData %>% 
  group_by(species, shrubs, shrubarea) %>% 
  summarise(predMM = mean(predMM))
ggplot(MyData_shrubarea_summary)+
  geom_line(aes(x = shrubarea, y = predMM, group = interaction(shrubs, species), linetype = shrubs, col = species))
```

The greater the pre-treatment shrub cover, the greater the predicted diameter growth. 


# Save final data frame
```{r}
save(df, file = "~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/JOINED_DATA/df_dia_final.Rdata")
```