---
title: "prelim_analysis"
author: "Carmen"
date: "November 18, 2017"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
```


# Load data
```{r}
df <- read_excel("~/../Dropbox (Stephens Lab)/SORTIE/Growth_mortality/data/details/compiled/archives/Mort_details_compiled_Nov08_2017.xlsx", sheet = 2)
```

# Delete seedlings with missing data or pathogen damage
```{r}
summary(as.factor(df$Practice))
df <- df %>% 
  filter(is.na(Practice) | Practice != "1") %>% 
  filter(is.na(PATH_DAMAGE) | PATH_DAMAGE != "1")
```

# Clean

## Fix typos
```{r}
df <- df %>% 
  mutate(DEAD_ALIVE = ifelse(DEAD_ALIVE == "alive", "ALIVE", DEAD_ALIVE)) %>% 
  mutate(DEAD_ALIVE = ifelse(DEAD_ALIVE == "dead", "DEAD", DEAD_ALIVE)) %>% 
  mutate(SPECIES = ifelse(SPECIES == "abco", "ABCO", SPECIES)) %>% 
  mutate(SPECIES = ifelse(SPECIES == "pipo", "PIPO", SPECIES))
```

```{r}
df %>% filter(is.na(as.numeric(df$LAST_YR_GR_cm)))
df %>% filter(is.na(as.numeric(df$MINUS_1_GR_cm)))
df %>% filter(is.na(as.numeric(df$MINUS_2_GR_cm)))
```

## Make numeric
```{r}
df <- df %>% 
  mutate(LAST_YR_GR_cm = as.numeric(LAST_YR_GR_cm)) %>% 
  mutate(MINUS_1_GR_cm = as.numeric(MINUS_1_GR_cm)) %>% 
  mutate(MINUS_2_GR_cm = as.numeric(MINUS_2_GR_cm))
```

```{r}
df <- df %>% 
  mutate(DEAD = ifelse(DEAD_ALIVE == "DEAD", 1, 0))
```



# calculate relative growth rate
```{r}
df <- df %>% 
  mutate(LAST_YR_GR_rel = LAST_YR_GR_cm/(HEIGHT - LAST_YR_GR_cm)) %>% 
  mutate(MINUS_1_GR_rel = MINUS_1_GR_cm/(HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm)) %>% 
  mutate(MINUS_2_GR_rel = MINUS_2_GR_cm/(HEIGHT - LAST_YR_GR_cm - MINUS_2_GR_cm - MINUS_1_GR_cm)) %>% 
  select(LAST_YR_GR_rel, MINUS_1_GR_rel, MINUS_2_GR_rel, everything())
```


# Restructure
```{r}
df <- gather(df, "Year","VertGrowth",1:3)
```

# Remove obvious errors
```{r}
df <- df %>% filter(VertGrowth<60)
```



# Calculate average growth rate for past 3 years
```{r}
df <- df %>% 
  group_by(SEEDLING, SPECIES, DEAD, HEIGHT) %>% 
  summarise(vert_ave = mean(VertGrowth))
```


# Plot seedling growth rate against whether it's dead
```{r}
ggplot(df)+
  geom_boxplot(aes(x = interaction(SPECIES, DEAD), y = vert_ave))
```

# GLM for abco


```{r}
dfa <- df %>% filter(SPECIES == "ABCO")
dfp <- df %>% filter(SPECIES == "PIPO")
```

```{r}
dfa <- dfa %>% filter(vert_ave>0)
```

```{r}
dfa %>% group_by(DEAD) %>% 
  summarise(mean(vert_ave))
```

```{r}
dfp <- dfp %>% filter(vert_ave<.8)
```


```{r}
GLM_abco <- glm(DEAD ~ log(vert_ave)*HEIGHT, family = binomial(link = "logit"), data = dfa)
anova(GLM_abco, test = "Chi")
GLM_a_null <- glm(DEAD ~ 1, family = binomial(link = "logit"), data = dfa)
AIC(GLM_a_null)
AIC(GLM_abco)
anova(GLM_a_null, GLM_abco, test = "Chi")
```


# GLM for pipo
```{r}
GLM_pipo <- glm(DEAD ~ log(vert_ave)+HEIGHT, family = binomial(link = "logit"), data = dfp)
anova(GLM_pipo, test = "Chi")
drop1(GLM_pipo, test = "Chisq")
GLM_p_null <- glm(DEAD ~ 1, family = binomial(link = "logit"), data = dfp)
AIC(GLM_p_null)
AIC(GLM_pipo)
anova(GLM_p_null, GLM_pipo, test = "Chi")
```


## plot effects
```{r}
plot(predictorEffect("vert_ave", GLM_pipo))

```


Significant for pines but not firs

# Plot
```{r}
MyData <- data.frame(vert_ave = seq(min(dfp$vert_ave), max(dfp$vert_ave), length = 100))
```


```{r}
 MyData$pred <- predict(GLM_pipo, newdata = MyData, type = "response")
figure <-ggplot(MyData)+
  geom_line(aes(x = vert_ave, y = pred), size = 2, col = "#a02313")+
  theme_bw()+
  xlab("Mean relative vertical growth")+
  ylab("Predicted probability of mortality")+
  ggtitle("Pine mortality in relation to vertical growth")+
   theme(
    plot.background = element_rect(fill = '#e9ebe8', colour = '#e9ebe8'),
        panel.background = element_rect(fill="#e9ebe8"),
        text = element_text(size=20),
        axis.title = element_text(colour = "#a02313"),
        plot.title = element_text(colour = "#a02313")
    )
figure
```

```{r}
setwd("~/../Dropbox (Stephens Lab)/SORTIE/Growth_mortality/results/figures/")
pdf("pine_vert_mort.pdf", width = 7.5, height = 4.5, pointsize = 30,useDingbats = F)
figure
dev.off()
```

# NEXT STEPS: RE-ANALYZE EVERTHING USING THE PROPER FORMULA FOR RELATIVE GROWTH RATE