---
title: "Diameter ~ Height"
author: "Carmen"
date: "November 1, 2019"
output: 
  html_document:
    toc: TRUE
---

# This code creates linear models for diameter in relation to height for ABCO and PIPO based on dendro data

```{r}
require(tidyverse)
require(readxl)
require(effects)
require(nlme)
require(MuMIn)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/Simulations/")
load(file = "../../compiled_data/growth_mortality/dendro_all_vars.Rdata")
load(file = "../../compiled_data/growth_mortality/df_vert.Rdata")
dfvert <- df
remove(df)
```

# Clean and wrangle data

## Check for missing data
```{r}
join <- tbl_df(join)
join %>% 
  filter(is.na(DEAD))
```

```{r}
dfvert %>% filter(is.na(rel))
```


These missing data are because of pathogen damage, etc. so they should be filtered out
```{r}
join <- join %>% 
  filter(!is.na(DEAD))
```

## Remove unnecesary variables
```{r}
join <- join %>% 
  dplyr::select(-BAS_DIA_1_mm, -BAS_DIA_2_mm, -BARK, -PERC_NEEDLES, -SUPPRESSED, -`BLACK DOTS`, -MECH_DAMAGE, -PATH_DAMAGE, -exclude_reason, -mateless, -`dia/ht`, -NEEDLE_COLOR, -`Growing?`, -BUDS, -TIME)
```

<!-- ## Filter to just live seedlings for now -->
<!-- ```{r} -->
<!-- join <- join %>%  -->
<!--   filter(DEAD==0) -->
<!-- ``` -->

## Restructure to make one row per tree and one column for year
```{r}
df <- gather(join, key = "year", value = "growth", 5:17) %>% 
  tbl_df() %>% 
  dplyr::select(tree, series, year, growth, everything()) %>% 
  arrange(tree) %>% 
  filter(!is.na(growth))
df
join %>% filter(tree ==1) %>% dplyr::select(1:20)
```

```{r}
df <- df %>% 
  group_by(tree, year) %>% 
  mutate(mean_growth_dia = mean(growth)*2) %>% 
  dplyr::select(-growth,-radius, -series) %>% 
  distinct() %>% 
  dplyr::select(tree, year, mean_growth_dia, everything()) %>% 
  rename(BAS_DIA_AVE_mm = BAS_DIA_AVE)
df
```

## Take out growth years before 2011
```{r}
df <- df %>% 
  filter(year>2010)
```

# _

# Combine with `dfvert` to get info on when dead trees died
```{r}
dfvert <- dfvert %>% 
  dplyr::select(SEEDLING, Notes, BAS_DIA_AVE, ht1, ht2, ht3, ht4, gr1, gr2, gr3, death_year, gr_rel_ave) %>% 
  rename(tree = SEEDLING)
```

```{r}
df <- left_join(df, dfvert)
```

# NOTE: I am not fixing the death year because it doesn't matter to me. I'm not including calenar year as a fixed effect.

## Calculate relative vertical growth for each year
```{r}
df %>% 
   dplyr::select(tree, DEAD_ALIVE, year, year, HEIGHT, LAST_YR_GR_cm, gr1, gr2, gr3, ht1, ht2, ht3, ht4) %>% 
  filter(DEAD_ALIVE == "ALIVE")
```

```{r}
df <- df %>% 
  mutate(relgrvert = case_when(
    year == 2017 ~ gr1,
    year == 2016 ~ gr2,
    year == 2015 ~ gr3,
    TRUE ~ 0
  ))
```

### Check
```{r}
df %>% 
  filter(is.na(relgrvert)) %>% 
  dplyr::select(tree, DEAD_ALIVE, year, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm, MINUS_2_GR_cm, gr1, gr2, gr3, ht1, ht2, ht3, relgrvert) %>%
  arrange(tree)
```

```{r}
df <- df %>% 
  filter(tree != 119) %>% 
  filter(!(year == 2015 & is.na(MINUS_2_GR_cm)))
```

```{r}
df %>% 
  filter(is.na(relgrvert)) %>% 
  dplyr::select(tree, DEAD_ALIVE, year, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm, MINUS_2_GR_cm, gr1, gr2, gr3, ht1, ht2, ht3, relgrvert) %>%
  arrange(tree)
```

```{r}
df <- df %>% 
  mutate(relgrvert = case_when(
    is.na(relgrvert) & year == 2016 ~ MINUS_1_GR_cm/(HEIGHT-LAST_YR_GR_cm-MINUS_1_GR_cm),
    is.na(relgrvert) & year == 2015 ~ MINUS_2_GR_cm/(HEIGHT-LAST_YR_GR_cm-MINUS_1_GR_cm - MINUS_2_GR_cm),
    TRUE ~ relgrvert
  ))
```

Check
```{r}
df %>% 
  filter(tree %in% c(1, 214) & year == 2016) %>% 
  dplyr::select(tree, DEAD_ALIVE, year, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm, MINUS_2_GR_cm, gr1, gr2, gr3, ht1, ht2, ht3, relgrvert) %>%
  arrange(tree) %>% 
  mutate(CHECK = MINUS_1_GR_cm/(HEIGHT-LAST_YR_GR_cm-MINUS_1_GR_cm))
```

```{r}
df %>% 
  filter(tree %in% c(1, 214) & year == 2015) %>% 
  dplyr::select(tree, DEAD_ALIVE, year, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm, MINUS_2_GR_cm, gr1, gr2, gr3, ht1, ht2, ht3, relgrvert) %>%
  arrange(tree) %>% 
  mutate(CHECK = MINUS_2_GR_cm/(HEIGHT-LAST_YR_GR_cm-MINUS_1_GR_cm-MINUS_2_GR_cm))
```
 

```{r}
df %>% 
   dplyr::select(tree, DEAD_ALIVE, year, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm, MINUS_2_GR_cm, gr1, gr2, gr3, ht1, ht2, ht3, relgrvert) %>%
  arrange(tree)
```


## Restructure height so there's yearly height and yearly height growth
```{r}
df <- df %>% 
  mutate(pre_growth_height = case_when(
    year == 2017 ~ HEIGHT-LAST_YR_GR_cm,
    year == 2016 ~ HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm,
    year == 2015 ~ HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm - MINUS_2_GR_cm,
    TRUE~0
  )) %>% 
   mutate(ht_growth = case_when(
    year == 2017 ~ LAST_YR_GR_cm,
    year == 2016 ~ MINUS_1_GR_cm,
    year == 2015 ~ MINUS_2_GR_cm,
    TRUE~0
  )) 
```

### Check
```{r}
df %>% 
   dplyr::select(tree, DEAD_ALIVE, year, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm, MINUS_2_GR_cm,gr1, gr2, gr3, ht1, ht2, ht3,ht4, relgrvert, pre_growth_height, ht_growth) %>%
  arrange(tree)
```

## Add a column for yearly diameter by subtracting growth from diameter for each year
```{r}
df <- df %>% 
  filter(year >2010) %>% 
  ungroup() %>% 
    mutate(cumulative_growth = ifelse(year == 2016, mean_growth_dia, 0)) %>% 
    group_by(tree) %>% 
  mutate(cumulative_growth = ifelse(year == 2015, mean_growth_dia + max(cumulative_growth), cumulative_growth)) %>% 
  mutate(cumulative_growth = ifelse(year == 2014, mean_growth_dia + max(cumulative_growth), cumulative_growth)) 
```

```{r}
df <- df %>% 
  mutate(pre_growth_diameter = BAS_DIA_AVE_mm-cumulative_growth) 
```

### Check
```{r}
df %>% 
  dplyr::select(tree, year, mean_growth_dia, cumulative_growth, pre_growth_diameter, BAS_DIA_AVE_mm, everything())  
df
```

## Calculate relative diameter growth 
```{r}
df <- df %>% 
  mutate(relgrdia = (mean_growth_dia)/pre_growth_diameter)
```
 
## Throw out the two observations where pre-growth diameter is below 0
```{r}
df <- df %>% 
  filter(pre_growth_diameter>0)
```

# Throw out tree/year combos without vertical growth
```{r}
df <- df %>% 
  filter(ht_growth>0)
```

# Take out more columns
```{r}
df <- df %>% 
  dplyr::select(-Practice, `got it`, -`Tops cut?`)
```

# Take out tree with missing height data
```{r}
df <- df %>% 
  filter(tree != 119)
```

# _
# _

# Model dia growth in relation to previous dia and height, both species together
```{r}
df$SPECIES <- as.factor(paste(df$SPECIES))
class(df$SPECIES)
M1 <- lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height + SPECIES + pre_growth_height*ht_growth, data = df, random = ~ 1|tree)
summary(M1)
hist(log(df$relgrdia))
hist(df$relgrdia)
ggplot(df, aes(x = HEIGHT, y = log(relgrdia), col = SPECIES))+
  geom_point()+
  theme_minimal()+
  geom_smooth(method = "lm")

```

# Do AIC model selection
```{r}
AIC(M1)
AIC(lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height + SPECIES + pre_growth_height, data = df, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height + SPECIES + ht_growth, data = df, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + SPECIES + ht_growth, data = df, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter*ht_growth + pre_growth_height + SPECIES , data = df, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter + ht_growth*pre_growth_height + SPECIES , data = df, random = ~ 1|tree))
```

```{r}
M2 <- lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + SPECIES + ht_growth, data = df, random = ~ 1|tree)
```


# Pseudo r2
```{r}
r.squaredGLMM(M2)
```

# Plot predictor effects
```{r}
plot(effects::predictorEffect("pre_growth_height", M2))
plot(effects::predictorEffect("pre_growth_diameter", M2))
plot(predictorEffect("ht_growth", M2))
```

# Plot range of dia and ht in these data

## diameter
```{r}
ggplot(df)+
  geom_freqpoly(aes(BAS_DIA_AVE_mm/10,col = SPECIES, stat(density)))+
  theme_minimal()+
  xlab("Diameter (cm)")+
  theme(text = element_text(size = 20))
```

## height

```{r}
ggplot(df)+
  geom_freqpoly(aes(HEIGHT, col = SPECIES, stat(density)))+
  theme_minimal()+
  xlab("Height (cm)")+
  theme(text = element_text(size = 20))
```

## Ratio
```{r}
ggplot(df)+
  geom_freqpoly(aes((BAS_DIA_AVE_mm/10)/HEIGHT, col = SPECIES, stat(density)))+
  theme_minimal()+
  xlab("Diameter:Height (cm:cm)")+
  theme(text = element_text(size = 20))
```

# _
# _
# Model dia in relation to ht and competition, species separate

## ABCO
```{r}
abcodf <- df %>% filter(SPECIES == "ABCO")
lmeabco <- lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, random = ~ 1|tree, data = abcodf, method = "ML")
summary(lmeabco)
drop1(lmeabco, test = "Chisq")
r.squaredGLMM(lmeabco)
```

### Try the model that worked best for PIPO
```{r}
LMPIPO_final_abco <- lm(relgrdia ~ pre_growth_diameter*relgrvert + pre_growth_height,data = abcodf)
summary(LMPIPO_final_abco)
plot(predictorEffect("relgrvert", LMPIPO_final_abco))
```

# That worked pretty well!

### Do AIC model selection
```{r}
AIC(lmeabco)
AIC(lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height*ht_growth,  random = ~ 1|tree, data = abcodf))
AIC(lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height + ht_growth, data = abcodf, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, data = abcodf, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter*ht_growth + pre_growth_height ,data = abcodf, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter + ht_growth*pre_growth_height, data = abcodf, random = ~ 1|tree))
```

```{r}
lmeabco <- lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, data = abcodf, random = ~ 1|tree, method = "ML")
summary(lmeabco)
drop1(lmeabco, test = "Chisq")
r.squaredGLMM(lmeabco)
```

### Plot predictor effects
```{r}
plot(predictorEffect("pre_growth_diameter", lmeabco))
plot(predictorEffect("pre_growth_height", lmeabco))
plot(predictorEffect("ht_growth", lmeabco))
```


## PIPO
```{r}
pipodf <- df %>% filter(SPECIES == "PIPO")
lmepipo <- lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, random = ~ 1|tree, data = pipodf, method = "ML")
summary(lmepipo)
drop1(lmepipo, test = "Chisq")
r.squaredGLMM(lmepipo)
```

```{r}
hist(pipodf$relgrvert)
```


### Model with random effect
```{r}
lmepipo <- lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, data = pipodf, random = ~ 1|tree, method = "ML")
summary(lmepipo)
drop1(lmepipo, test = "Chisq")
r.squaredGLMM(lmepipo)
```

### Do AIC model selection

#### no log of dia growth
```{r}
AIC(lme(relgrdia ~ pre_growth_diameter*pre_growth_height*relgrvert,  random = ~ 1|tree, data = pipodf))
AIC(lme(relgrdia ~ pre_growth_diameter*pre_growth_height + relgrvert, data = pipodf, random = ~ 1|tree))
AIC(lme(relgrdia ~ pre_growth_diameter + pre_growth_height + relgrvert, data = pipodf, random = ~ 1|tree))
AIC(lme(relgrdia ~ pre_growth_diameter*relgrvert + pre_growth_height ,data = pipodf, random = ~ 1|tree))
AIC(lme(relgrdia ~ pre_growth_diameter + relgrvert*pre_growth_height, data = pipodf, random = ~ 1|tree))
```

##### 1. Look at best model
```{r}
lmepipo <- lme(relgrdia ~ pre_growth_diameter + pre_growth_height + relgrvert, data = pipodf, random = ~ 1|tree, method = "ML")
summary(lmepipo)
drop1(lmepipo, test = "Chisq")
r.squaredGLMM(lmepipo)
```

#### log of dia growth
```{r}
AIC(lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height*ht_growth,  random = ~ 1|tree, data = pipodf))
AIC(lme(log(relgrdia) ~ pre_growth_diameter*pre_growth_height + ht_growth, data = pipodf, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, data = pipodf, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter*ht_growth + pre_growth_height ,data = pipodf, random = ~ 1|tree))
AIC(lme(log(relgrdia) ~ pre_growth_diameter + ht_growth*pre_growth_height, data = pipodf, random = ~ 1|tree))
```

##### 2. Look at best model
```{r}
lmepipo2 <- lme(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + ht_growth, data = pipodf, random = ~ 1|tree, method = "ML")
summary(lmepipo)
drop1(lmepipo, test = "Chisq")
r.squaredGLMM(lmepipo)
```


### Model without random effect

#### Do AIC model selection with log transformation
```{r}
AIC(lm(log(relgrdia) ~ pre_growth_diameter*pre_growth_height*relgrvert,  data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter*pre_growth_height + relgrvert, data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + relgrvert, data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter*relgrvert + pre_growth_height ,data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter*log(relgrvert) + pre_growth_height ,data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter + relgrvert*pre_growth_height, data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter + pre_growth_height + log(relgrvert), data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter + pre_growth_height * log(relgrvert), data = pipodf))
AIC(lm(log(relgrdia) ~ pre_growth_diameter + pre_growth_height * relgrvert, data = pipodf))
```


##### 3. Look at best model WITH transformation
```{r}
lmpipo <- lm(log(relgrdia) ~ pre_growth_diameter*relgrvert + pre_growth_height, data = pipodf)
summary(lmpipo)
AIC(lmpipo)
drop1(lmpipo, test = "F")
plot(lmpipo)
plot(predictorEffect("relgrvert", lmpipo))
plot(predictorEffect("pre_growth_diameter", lmpipo))
plot(predictorEffect("pre_growth_height", lmpipo))
```

#### Do AIC model selection WITHOUT log transformation
```{r}
AIC(lmpipo)
AIC(lm(relgrdia ~ pre_growth_diameter*pre_growth_height*relgrvert,  data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter*pre_growth_height +relgrvert, data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter + pre_growth_height + relgrvert, data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter*relgrvert + pre_growth_height ,data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter + relgrvert*pre_growth_height, data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter + pre_growth_height + log(relgrvert), data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter + pre_growth_height * log(relgrvert), data = pipodf))
AIC(lm(relgrdia ~ pre_growth_diameter + pre_growth_height * relgrvert, data = pipodf))
```

##### 4. Look at best model without transformation
```{r}
lmpipo <- lm(relgrdia ~ pre_growth_diameter*relgrvert + pre_growth_height,data = pipodf)
summary(lmpipo)
AIC(lmpipo)
drop1(lmpipo, test = "F")
plot(predictorEffect("relgrvert", lmpipo))
plot(predictorEffect("pre_growth_diameter", lmpipo))
plot(predictorEffect("pre_growth_height", lmpipo))
```

##### Same structure, lme
```{r}
lmepipo <- lm(relgrdia ~ pre_growth_diameter*relgrvert + pre_growth_height,data = pipodf)
summary(lmpipo)
AIC(lmpipo)

```


# BEST PIPO MODEL OVERALL
```{r}
LMPIPO_final <- lm(relgrdia ~ pre_growth_diameter*relgrvert + pre_growth_height + year,data = pipodf)
summary(LMPIPO_final)
```



### Plot predictor effects
```{r}
plot(predictorEffect("pre_growth_diameter", LMPIPO_final))
plot(predictorEffect("pre_growth_height", LMPIPO_final))
plot(predictorEffect("relgrvert", LMPIPO_final))
```


# Save
```{r}
save(LMPIPO_final_abco, file ="../../results/coefficients/LM_dia_abco_dendro.Rdata")
save(LMPIPO_final, file ="../../results/coefficients/LM_dia_pipo_dendro.Rdata")
```

