---
title: "simulations_plot"
author: "Carmen"
date: "2/13/2020"
output: 
  html_document:
    toc: TRUE
---

```{r}
name <- "~/Simulation1_2020-03-27_1000_empirical_welch_empiricalcoef_empiricalgrowth_empiricalindex.Rdata"
load(name)
strt <- Sys.time()
```

```{r}
setwd("~/Shrubs-Seedlings/code/Simulations/")
```

```{r, include = F}
require(tidyverse)
require(gridExtra)
```

# Take out trees with non-finite or NA pred because they had already emerged and things got wacky
```{r}
dfsimallreps %>% 
  filter(!is.finite(pred_exp) | is.na(pred_exp)) %>% 
  group_by(emerged) %>% 
  count()
```

```{r}
dfsimallreps <- dfsimallreps %>% 
  filter(is.finite(pred_exp) & !is.na(pred_exp))
```

# Summarize

## Focal tree height by year for trees that haven't yet emerged
```{r}
dfsimallreps_summary <- dfsimallreps %>% 
  ungroup() %>%
  mutate(rep = as.factor(paste(rep))) %>% 
  group_by(rep, Years, Species) %>% 
  filter(emerged==0) %>% 
  summarize(mean_ht_years = mean(Ht_cm1), mean_dia_years = mean(dia.cm)) 
```

```{r}
ht <- ggplot(dfsimallreps_summary, aes(x = as.factor(Years), y = mean_ht_years, fill = Species, col = Species))+
  geom_boxplot(alpha = .2, outlier.shape = NA)+
  geom_smooth(aes(x = as.factor(Years), y = mean_ht_years, group = Species, col = Species), size = 1)+
  xlab("Years since fire")+
  ylab("Average tree ht (cm) by simulation")+
  theme_bw()+
  scale_color_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  scale_fill_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  theme(text = element_text(size = 16), 
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 8))+
  theme_minimal()
ht
```

## Diameter growth patterns 
```{r}
dia <- ggplot(dfsimallreps_summary, aes(x = as.factor(Years), y = mean_dia_years, fill = Species, col = Species))+
  geom_boxplot(alpha = .2, outlier.shape = NA)+
  geom_smooth(aes(x = as.factor(Years), y = mean_dia_years, group = Species, col = Species), size = 1)+
  xlab("Years since fire")+
  ylab("Average tree dia (cm) by simulation")+
  theme_bw()+
  scale_color_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  scale_fill_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  theme(text = element_text(size = 16), 
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 8))+
  theme_minimal()
dia
```

### Diameter::Height growth patterns
```{r}
ratio <- ggplot(dfsimallreps_summary, aes(x = as.factor(Years), y = mean_dia_years/mean_ht_years, fill = Species, col = Species))+
  geom_boxplot(alpha = .2, outlier.shape = NA)+
  geom_smooth(aes(x = as.factor(Years), y = mean_dia_years/mean_ht_years, group = Species, col = Species), size = 1)+
  xlab("Years since fire")+
  ylab("Average tree dia:ht ratio by simulation")+
  theme_bw()+
  scale_color_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  scale_fill_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  theme(text = element_text(size = 16), 
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 8))+
  theme_minimal()
ratio
```

## Counts by year
```{r}
dfsimallreps_summary <- dfsimallreps %>% 
  ungroup() %>% 
  group_by(rep, Years, Species) %>% 
  summarize(count = n()) %>% 
  mutate(count = as.numeric(count))
```

```{r}
title <- paste("Results for", max(dfsimallreps$Years), "years across", max(dfsimallreps$rep), "iterations", sep = " ")

ggplot(dfsimallreps_summary)+
  geom_smooth(aes(x = as.factor(Years), y = count, fill = Species, col = Species), size = 1)+
  geom_boxplot(aes(x = as.factor(Years), y = count, fill = Species, col = Species), alpha = .2, outlier.shape = NA)+
  ggtitle(title)+
  xlab("Years since fire")+
  ylab("Tree count")+
  theme_bw()+
  scale_color_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  scale_fill_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  theme(text = element_text(size = 16), 
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 8))

```

## Shrub height by year
```{r}
dfsimallreps_summary <- dfsimallreps %>% 
  filter(!is.na(Ht_cm1)) %>% 
  ungroup() %>% 
  group_by(rep, Years, Ht1.3, Cov1.3, ShrubSpp03) %>% 
  summarize(mean_shrub_ht = mean(Ht1.3), mean_shrub_cov = mean(Cov1.3), mean_ht = mean(Ht_cm1))
```

```{r}
ggplot(dfsimallreps_summary, aes(x = Years, y = mean_shrub_ht, group = ShrubSpp03, col = ShrubSpp03))+
  #geom_line(aes(group = Sdlg), alpha = .2)+
  geom_smooth()+
  ggtitle(paste("Results for", max(dfsimallreps$rep), "simulations", sep = " "))+
  xlab("Years since fire")+
  ylab("Shrub height")+
  theme_minimal()
```

## Shrub cover by year
```{r}
ggplot(dfsimallreps_summary, aes(x = Years, y = mean_shrub_cov, group = ShrubSpp03, col = ShrubSpp03))+
  #geom_line(aes(group = Sdlg), alpha = .2)+
  geom_smooth()+
  ggtitle(paste("Results for", max(dfsimallreps$rep), "simulations", sep = " "))+
  xlab("Years since fire")+
  ylab("Shrub cover")+
  theme_bw()
```

## Focal tree height by shrub species and year
```{r}
ggplot(dfsimallreps_summary, aes(x = Years, y = mean_ht, group = ShrubSpp03, col = ShrubSpp03))+
  #geom_line(aes(group = interaction(rep, ID_withinrep)), alpha = .05)+
  geom_smooth()+
  ggtitle(paste("Results for", max(dfsimallreps$rep), "simulations", sep = " "))+
  xlab("Years since fire")+
  ylab("Focal tree height")+
  ylim(0,350)+
  theme_minimal()
```

## Growth by climate year
```{r}
ggplot(dfsimallreps %>% filter(Species == "PIPO") %>% filter(emerged == 0))+
  geom_boxplot(aes(x = as.factor(Year), y = pred_exp), outlier.shape = NA)+
  theme_minimal()+
  ylim(0,0.5)+
  xlab("")+
  ylab("Predicted relative growth rate")
```

```{r}
dfsimallreps %>% 
   filter(Species == "PIPO") %>% filter(emerged == 0) %>% 
  group_by(Year) %>% 
  summarize(n(), mean(pred_exp))
```

# Plot how many trees emerge above the shrub canopy each year

## How many trees are neither emerged nor dead at the end? Should be none.
```{r}
dfsimallreps %>% 
  filter(Years == max(Years)) %>% 
  group_by(emerged, Species) %>% 
  count()
```

## Plot emergents per year 

### Summarize
```{r}
summary_emerge <- dfsimallreps %>% 
  group_by(Years, Species, rep) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  mutate(Species = fct_relevel(Species, "PIPO", "ABCO"))
```

### Counts

```{r}
# plot_emerge_count <- ggplot(summary_emerge)+
#   theme_minimal()+
#   geom_boxplot(aes(x = as.factor(Years), y = total_emergents, fill = Species), alpha = .8)+
#   ylab("Number of emerged trees")+
#   xlab("Years since fire")+
#   ggtitle(paste("Results for", max(dfsimallreps$rep), "simulations", sep = " "))
```

### Proportions

```{r}
plot_emerge_prop <- ggplot(summary_emerge)+
  theme_minimal()+
  geom_boxplot(aes(x = as.factor(Years), y = prop_emerged, fill = Species), alpha = .8, outlier.shape = NA)+
  ylab("Proportion of initial juvenile trees that have emerged")+
  xlab("Years since fire")+
  #ggtitle(paste("Results for", max(dfsimallreps$rep), "simulations", sep = " "))+
  theme(legend.position = c(0.8, 0.2))+
  scale_x_discrete(breaks =c("10","20","30", "40"))+
  theme(text = element_text(size = 13), 
        legend.title = element_blank(),
        legend.text=element_text(size=13))+
  xlab(element_blank())+
  scale_fill_discrete(labels = c( "Ponderosa pine", "White fir"))

plot_emerge_prop
```

```{r}
#grid.arrange(plot_emerge_count, plot_emerge_prop, ncol = 1)
```

## Check to make sure there are in fact trees that emerged before the simulation started
```{r}
dfsimallreps %>% 
  filter(rep %in% c(1:10) & Years==9) %>% 
  dplyr::select(Species, Sdlg, Ht1.3, Ht_cm1) %>% 
  mutate(emerge = ifelse(Ht_cm1*.75>Ht1.3, 1, 0)) %>% 
  group_by(Species, emerge) %>% 
  count()
```

## Plot number and proportion of original seedlings left at the end

```{r}
title <- paste("Results for", max(dfsimallreps$rep), "simulations at\n", max(dfsimallreps$Years),"years after fire", sep = " ")
```

```{r}
# summary_count <- dfsimallreps %>% 
#   group_by(rep, Species) %>% 
#   filter(Years == max(Years)) %>% 
#   summarize(emerged = sum(emerged)) 
# 
# boxplot_trees <- ggplot(summary_count)+
#   geom_boxplot(aes(x = Species, y = emerged))+
#   theme_minimal()+
#   ylab("Number of emerged trees")+
#   ggtitle(title)
```

```{r}
summary_prop <- dfsimallreps %>% 
  ungroup() %>% 
  group_by(Years, Species, rep) %>% 
  mutate(total= n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  group_by(rep, initial_n, Species) %>% 
  filter(Years== max(Years)) %>% 
  summarize(emerged = sum(emerged)) %>% 
  mutate(prop_emerged = emerged/initial_n) 
```

```{r}
boxplot_prop <- ggplot(summary_prop)+
  geom_boxplot(aes(x = Species, y = prop_emerged))+
  theme_minimal()+
  ylab("Proportion of initial trees")+
  xlab(element_blank())+
  scale_x_discrete(labels = c("White fir", "Ponderosa pine"))+
  theme(text = element_text(size = 14))
boxplot_prop
```


# What's different about the iterations where 100% of trees emerge?
```{r}
hundoP <- summary_prop %>% 
  ungroup() %>% 
  filter(prop_emerged==1) %>% 
  dplyr::select(rep) %>% 
  unlist()
hundoP
```

## What do the hundoP reps look like?
```{r}
ggplot(dfsimallreps %>% filter(rep %in% hundoP))+
  geom_smooth(aes(x = Years, y = Ht_cm1, group = rep))+
  geom_smooth(aes(x = Years, y = Ht1.3, group = rep), col = "red")+
  theme_minimal()
```

```{r, echo = F}
print("The whole thing took")
Sys.time()-strt
```

# Are ending proportions significantly different between pine and fir?
```{r}
ggplot(summary_prop)+
  geom_histogram(aes(x = prop_emerged, fill = Species), bins =20)+
  theme_minimal()
glm <- glm(prop_emerged ~ Species, family = binomial, data = summary_prop, weights = initial_n)
summary(glm)
```

# Sensitivity to error 

## mortality
```{r}
dfsimallreps %>% 
  ungroup() %>% 
  group_by(Years, Species, rep) %>% 
  mutate(total= n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  group_by(rep, initial_n, Species, coef_gr_mort_pipo) %>% 
  filter(Years== max(Years)) %>% 
  summarize(emerged = sum(emerged)) %>% 
  mutate(prop_emerged = emerged/initial_n) %>% 
  filter(Species == "PIPO") %>% 
  ggplot(aes(x = coef_gr_mort_pipo, y = prop_emerged))+
  geom_point()+
  geom_smooth()+
  theme_minimal()+
  ggtitle("PIPO")

dfsimallreps %>% 
  ungroup() %>% 
  group_by(Years, Species, rep) %>% 
  mutate(total= n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  group_by(rep, initial_n, Species, coef_gr_mort_abco) %>% 
  filter(Years== max(Years)) %>% 
  summarize(emerged = sum(emerged)) %>% 
  mutate(prop_emerged = emerged/initial_n) %>% 
  filter(Species == "ABCO") %>% 
  ggplot(aes(x = coef_gr_mort_abco, y = prop_emerged))+
  geom_point()+
  geom_smooth()+
  theme_minimal()+
  ggtitle("ABCO")
```

## Growth
```{r}
dfsimallreps %>% 
  ungroup() %>% 
  group_by(Years, Species, rep) %>% 
  mutate(total= n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  group_by(rep, initial_n, Species, coef_gr_shrubarea) %>% 
  filter(Years== max(Years)) %>% 
  summarize(emerged = sum(emerged)) %>% 
  mutate(prop_emerged = emerged/initial_n) %>% 
  filter(Species == "PIPO") %>% 
  ggplot(aes(x = coef_gr_shrubarea, y = prop_emerged))+
  geom_point()+
  geom_smooth()+
  theme_minimal()+
  ggtitle("PIPO")
```


# Long-term goals

1.  Use Kristen's data or Hugh's data for initial conditions
2.  Improve dispersal kernel based on Kristen/Hugh's data 
4.  Include residual surviving trees and their seed dispersal
5.  Include seed dispersal of post-fire regen once it reaches reproductive age 
6.  Add customization of patch size and shape
8.  Change sapling growth equations once they emerge from the shrub canopy
9.  Update display of shrub competition after simulation years
7.  apply to King, American River Complex, rest of the fires I measured
5.  what does shrub competition mean for new recruitment?

