---
title: "simulations_plot_ShrubSpecies COEFFICIENT"
author: "Carmen"
date: "3/6/2020"
output: html_document
---


```{r}
strt <- Sys.time()
setwd("~/Shrubs-Seedlings/code/Simulations/")
```

```{r, include = F}
require(tidyverse)
require(egg)
```

```{r}
date <- "2020-04-08"

namearpa <- paste("~/Simulation8_", date, "_1000_empirical_ARPAcoef_empiricalgrowth.Rdata", sep = "")

namecein <- paste("~/Simulation7_", date, "_1000_empirical_CEINcoef_empiricalgrowth.Rdata", sep = "")

namececo <- paste("~/Simulation6_", date, "_1000_empirical_CECOcoef_empiricalgrowth.Rdata", sep = "")
```

```{r}
load(namecein)
cein <- dfsimallreps
remove(dfsimallreps)
```

```{r}
max_year <- cein %>% 
    group_by(Years, emerged) %>% 
    summarize(sum_emerged = sum(emerged)) %>% 
    filter(emerged ==0) %>% 
    ungroup() %>% 
    summarize(max_year = max(Years)) %>% 
    unlist()
max_year
cein <- cein %>% 
  filter(Years < max_year+1)
```



```{r}
load(namearpa)
arpa <- dfsimallreps
remove(dfsimallreps)

load(namececo)
ceco <- dfsimallreps
remove(dfsimallreps)
```

```{r}
arpa <- arpa %>% 
  mutate(simspp = ifelse(Species == "PIPO", "ARPA", "ABCO")) %>% 
  mutate(source = "arpa")
ceco <- ceco %>% 
  mutate(simspp = ifelse(Species == "PIPO", "CECO", "ABCO")) %>% 
  mutate(source = "ceco")
cein <- cein %>% 
  mutate(simspp = ifelse(Species == "PIPO", "CEIN", "ABCO")) %>% 
  mutate(source = "cein")
nrow_total <-  nrow(cein) + nrow(ceco) + nrow(arpa)
```

```{r}
df <- full_join(arpa, ceco)
remove(arpa, ceco)
df <- full_join(df, cein)
remove(cein)
```

# Check df nrows
```{r}
nrow(df) == nrow_total
```


# Take out trees with non-finite or NA pred because they had already emerged and things got wacky
```{r}
df %>% 
  filter(!is.finite(pred_exp) | is.na(pred_exp)) %>% 
  group_by(emerged) %>% 
  count()
```


```{r}
df <- df %>% 
 mutate(pred_exp = ifelse(!is.finite(pred_exp) | !is.na(pred_exp), 999, pred_exp))
```

# Check that abco all looks the same

```{r}
summary_emerge <- df %>% 
  group_by(Years, rep, simspp, source, Species) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, simspp, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, simspp, source) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```

```{r}
summary_emerge %>% 
  group_by(simspp) %>% 
  mutate(distance_from_50 = mean-.5) %>% 
  filter(distance_from_50>0) %>% 
  filter(distance_from_50 == min(distance_from_50))
```


```{r}
ggplot(summary_emerge %>% filter(simspp == "ABCO"))+
  theme_minimal()+
  geom_line(aes(x = Years, y = mean, col = source, group = source), size = 1)+
  geom_ribbon(aes(x = Years, ymin = mean-sd, ymax = mean+sd, fill = source, group = source), alpha = .5)+
  xlim(8,38)+
  ylab("Proportion of initial juvenile trees that have emerged")+
  xlab("Years since fire")+
  #ggtitle(paste("Results for", max(dfsimallreps$rep), "simulations", sep = " "))+
  theme(legend.position = c(0.7, 0.3))
```

# Summarize
```{r}
summary_emerge <- df %>% 
  group_by(Years, rep, simspp, source, Species) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, simspp, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, simspp, Species) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```

## Filter out the flat parts - take out years if they differ by <0.001 (0.1%)
```{r}
max_year <- summary_emerge %>% 
  mutate(mean = round(mean, 3)) %>% 
  group_by(Species, simspp) %>% 
  mutate(maxmean = max(mean)) %>% 
  filter(mean == maxmean) %>% 
  ungroup() %>%
  group_by(Species, simspp) %>% 
  filter(Years == min(Years)) %>%
  ungroup() %>% 
  filter(Years ==max(Years)) %>% 
  select(Years) %>% 
  unlist()

max_year
```

```{r}
summary_emerge <- summary_emerge %>% 
  filter(Years < max_year + 1)
```


## Save summary
```{r}
save(summary_emerge, file = "../../results/data/summary_emerge_coef.Rdata")
```
