---
title: "Growth_mort FOR MANY RUNS AT ONCE"
author: "Carmen"
date: "October 31, 2019"
output: 
  html_document:
    toc: TRUE
---

# This script runs simulations without being spatially explicit. Thus, the functions `shrubclump` and `ratify_r` were removed.

```{r, include = F}
setwd("~/Shrubs-Seedlings/code/Simulations/") # change this if needed

pkgs <- c("tidyverse", "sf", "knitr", "mgcv", "gridExtra", "doParallel")
lapply(pkgs, require, character.only = T)
strt <- Sys.time()
```

#### Iteration function
```{r}
print(Sys.time())
source("functions/iterate.R")
```

# Define stuff

## Define iterations
```{r}
iterations <- 10
```

## Define length of simulation
```{r}
years_max <- 50
```

## Define fire
```{r}
fire <- "AMRC"
```

## Sensitivity methods
```{r}
climate_method <- "historic" #historic, random, uniform_2015, uniform_2016, uniform_2017
conifer_species_method <- "empirical" #random, empirical
shrub_method <- "welch" #empirical, welch, CECO, ARPA, CHSE, LIDE, OTHER
shrub_coefficient <- "empirical"  #emprical, CECO, ARPA, CHSE, LIDE, Other
shrub_heightgrowth <- "empirical"  #emprical, CECO, ARPA, CEIN
shrub_initial_index <- "empirical" # empirical, CECO, ARPA, CEIN
n_seedlings <- 200
```

## Define filter function
```{r}
filter_years <- function(){
  max_year <- dfsimallreps %>% 
    group_by(Years, emerged) %>% 
    summarize(sum_emerged = sum(emerged)) %>% 
    filter(emerged ==0) %>% 
    ungroup() %>% 
    summarize(max_year = max(Years)) %>% 
    unlist()

dfsimallreps <<- dfsimallreps %>% 
  filter(Years < max_year+1)
}
```


# Iterate for base model
```{r, warning=FALSE}
time <- Sys.time()

dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient, shrub_heightgrowth, shrub_initial_index)


Sys.time()-time
```

```{r, echo = F}
if(nrow(dfsimallreps)==iterations){
  print("ALERT! Something went wrong.")
}
```

# Filter out years in which nothing happens
```{r}
filter_years()
```
 
# Save results
```{r}
filename <-  paste("~/Simulation1_",  Sys.Date(), "_", iterations, "_", conifer_species_method, "_", shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth_", shrub_initial_index, "index.Rdata", sep = "")

save(dfsimallreps, file =filename)
```

# Repeat for establishment-neutral model
```{r}
conifer_species_method <- "random" #random, empirical
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient, shrub_heightgrowth, shrub_initial_index)

filename <-  paste("~/Simulation2_",  Sys.Date(), "_", iterations, "_", conifer_species_method, "_", shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth_", shrub_initial_index, "index.Rdata", sep = "")

save(dfsimallreps, file =filename)
```


# Repeat for single shrub species models

```{r}
conifer_species_method <- "empirical" #random, empirical
```


## CECO
```{r}
remove(dfsimallreps)
shrub_method <- "CECO" #empirical, welch, CECO, ARPA, CHSE, LIDE, OTHER
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)
Sys.time()-time

filter_years()

filename <-  paste("~/Simulation3_",  Sys.Date(), "_", iterations, "_",  shrub_method, ".Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## ARPA
```{r}
remove(dfsimallreps)
shrub_method <- "ARPA" #empirical, welch, CECO, ARPA, CHSE, LIDE, OTHER
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation4_",  Sys.Date(), "_", iterations, "_",  shrub_method, ".Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## CEIN 
```{r}
remove(dfsimallreps)
shrub_method <- "CEIN" #empirical, welch, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)
print(paste("That took"))
Sys.time()-time
filter_years()
filename <-  paste("~/Simulation5_",  Sys.Date(), "_", iterations, "_",  shrub_method, ".Rdata", sep = "")
save(dfsimallreps, file =filename)
```


# Repeat for single shrub COEFFICIENT models

```{r}
shrub_method <- "welch" #empirical, welch, CECO, ARPA, CHSE, LIDE, OTHER
```

## CECO
```{r}
remove(dfsimallreps)
shrub_coefficient <- "CECO"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation6_",  Sys.Date(), "_", iterations, "_",  shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## CEIN
```{r}
remove(dfsimallreps)
shrub_coefficient <- "CEIN"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation7_",  Sys.Date(), "_", iterations, "_",  shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## ARPA
```{r}
remove(dfsimallreps)
shrub_coefficient <- "ARPA"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation8_",  Sys.Date(), "_", iterations, "_",  shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```


# Repeat for single shrub GROWTH models

```{r}
shrub_coefficient <- "empirical" #empirical, welch, CECO, ARPA, CHSE, LIDE, OTHER
```

## CECO
```{r}
remove(dfsimallreps)
shrub_heightgrowth <- "CECO"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation9_",  Sys.Date(), "_", iterations, "_",  shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## CEIN
```{r}
remove(dfsimallreps)
shrub_heightgrowth <- "CEIN"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation10_",  Sys.Date(), "_", iterations, "_",  shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## ARPA

```{r}
remove(dfsimallreps)
shrub_heightgrowth <- "ARPA"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()
dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient,  shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation11_",  Sys.Date(), "_", iterations, "_",  shrub_method, "_", shrub_coefficient, "coef_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

# Repeat for single shrub INITIAL INDEX models

```{r}
shrub_heightgrowth <- "empirical" #empirical, welch, CECO, ARPA, CHSE, LIDE, OTHER
```


## CECO
```{r}
remove(dfsimallreps)
shrub_initial_index <- "CECO"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()

dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient, shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation12_",  Sys.Date(), "_", iterations, "_", shrub_method, "_", shrub_coefficient, "coef_", shrub_initial_index, "index_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## CEIN
```{r}
remove(dfsimallreps)
years_max <- 70
shrub_initial_index <- "CEIN"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()

dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient, shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation13_",  Sys.Date(), "_", iterations,  "_", shrub_method, "_", shrub_coefficient, "coef_", shrub_initial_index, "index_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```

## ARPA
```{r}
remove(dfsimallreps)
years_max <- 50
shrub_initial_index <- "ARPA"  #emprical, CECO, ARPA, CEIN
time <- Sys.time()

dfsimallreps <- iterate(iterations, fire, years_max, climate_method, conifer_species_method, shrub_method, n_seedlings, shrub_coefficient, shrub_heightgrowth, shrub_initial_index)

Sys.time()-time
filter_years()
filename <-  paste("~/Simulation14_",  Sys.Date(), "_", iterations, "_", shrub_method, "_", shrub_coefficient, "coef_", shrub_initial_index, "index_", shrub_heightgrowth, "growth.Rdata", sep = "")
save(dfsimallreps, file =filename)
```
