---
title: "Growth_mort"
author: "Carmen"
date: "October 31, 2019"
output: pdf_document
---

# This script runs simulations with the most up-to-date functions

```{r, include = F}
setwd("~/Shrubs-Seedlings/code/Simulations/")
require(raster)
require(tidyverse)
require(tmap)
require(sf)
require(sp)
require(tmaptools)
require(knitr)
strt <- Sys.time()
```

# Define parameters
```{r}
fire <- "AMRC"
year <- "2016"
years <- 20
iterations <- 10
max_shrub_ht_cm <- 250
max_shrub_ht_years <- 15
n_seedlings <- 100
length_m <- 40
height_m <- 40
lambda <- 4
shrub_clumpiness <- 7
```

# Load functions

#### Create shrub patch
```{r}
source("functions/shrubclump.R")
```

#### Initialization function
```{r}
source("functions/initialize.R")
```

#### Height growth functions
```{r}
source("functions/abcogrowth.R")
source("functions/pipogrowth.R")
```

#### Diameter growth functions
```{r}
source("functions/abcodia.R")
source("functions/pipodia.R")
```

#### Mortality functions
```{r}
source("functions/abcomort.R")
source("functions/pipomort.R")
```

#### Shrub growth functions
```{r}
source("functions/abco_shrubgrowth.R")
source("functions/pipo_shrubgrowth.R")
```

#### Simulation function
```{r}
source("functions/sim.R")
```

#### Iteration function
```{r}
source("functions/iterate.R")
```

# Initialize

## Create clumped shrub pattern
```{r}
shrubclump()
plot(r)
```

## Randomly select seedlings from data and place them on the shrub patch
```{r}
initialize()
```

## Plot patch with seedlings
```{r}
max_shrub <- max(r@data@attributes[[1]]$sqrt_shrubarea3)
r@data@attributes[[1]]$shrub_rel <- r@data@attributes[[1]]$sqrt_shrubarea3/max_shrub
```

```{r}
pts.sf.pipo.graph <- pts.sf.pipo %>% 
  rename("Pine height" = Ht_cm1)
pts.sf.abco.graph <- pts.sf.abco %>% 
  rename("Fir height" = Ht_cm1)
```

```{r}
tm_shape(p)+
  tm_borders(col = "black", lwd= 5)+
tm_shape(r)+
  tm_raster(col = "sqrt_shrubarea3", title = "Shrub competition index", palette = "Greys", alpha = .5)+
tm_shape(r)+
  tm_raster(col = "ShrubSpp03", alpha = .2, title = "Shrub species", palette = "Set1")+
  tm_layout(asp=1:1, legend.outside = T)+
tm_shape(pts.sf.pipo.graph)+
  tm_symbols(size = "Pine height", col = "darkgreen", size.max = 500, border.col = "white", border.lwd = .05)+
tm_shape(pts.sf.abco.graph)+
  tm_symbols(size = "Fir height", col = "darkblue", size.max = 500, border.col = "white", border.lwd = .05)
```

# Simulate across years
```{r, warning=FALSE}
suppressMessages(sim(years))
```

## Tests
```{r}
ggplot(dfsimall %>% filter(Species == "ABCO"))+
  geom_line(aes(x = Years, y = Ht1.3, col = ShrubSpp03, group = Sdlg))+
  theme_bw()
```

```{r}
ggplot(dfsimall %>% filter(Species == "PIPO"))+
  geom_line(aes(x = Years, y = Ht1.3, col = ShrubSpp03, group = Sdlg))+
  theme_bw()
```

```{r}
pts.sf.pipo.graph <- pts.sf.pipo %>% 
  rename("Pine height (cm)" = Ht_cm1)
pts.sf.abco.graph <- pts.sf.abco %>% 
  rename("Fir height (cm)" = Ht_cm1)
```

```{r}
tm_shape(p)+
  tm_borders(col = "black", lwd= 5)+
tm_shape(r)+
  tm_raster(col = "sqrt_shrubarea3", title = "Shrub competition index", palette = "Greys", alpha = .5)+
tm_shape(r)+
  tm_raster(col = "ShrubSpp03", alpha = .7, title = "Shrub species", palette = "Pastel1")+
  tm_layout(asp=1:1, legend.outside = T, legend.title.size = 4, legend.text.size = 2)+
tm_shape(pts.sf.pipo.graph)+
  tm_symbols(size = "Pine height (cm)", col = "#DCC960", size.max = 300, border.lwd = 3)+
tm_shape(pts.sf.abco.graph)+
  tm_symbols(size = "Fir height (cm)", col = "#899DA4", size.max = 300, border.lwd = 3)
```

# Iterate 
```{r, warning=FALSE}
iterate(iterations)
dfsimallreps %>% 
  group_by(rep) %>% 
  summarize(mean(Ht_cm1))
```

# Summarize

## Height by year
```{r}
dfsimallreps_summary <- dfsimallreps %>% 
  ungroup() %>%
  mutate(rep = as.factor(paste(rep))) %>% 
  group_by(rep, Years, Species) %>% 
  mutate(mean_ht_years = mean(Ht_cm1))
dfsimallreps_summary %>% dplyr::select(rep, Years, mean_ht_years) %>% summary()
```

```{r}
ggplot(dfsimallreps_summary, aes(x = as.factor(Years), y = mean_ht_years, fill = Species, col = Species))+
  geom_boxplot(alpha = .2, outlier.alpha = .02)+
  geom_smooth(aes(x = as.factor(Years), y = mean_ht_years, group = Species, col = Species), size = 1)+
  ggtitle("Results for 1000 simulations")+
  xlab("Years since fire")+
  ylab("Average tree ht (cm) by simulation")+
  theme_bw()+
  scale_color_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  scale_fill_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  theme(text = element_text(size = 16), 
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 8))
ggsave(file = "../../results/figures/sim_1000_hts.png", width = 6, height =4, dpi = 400)
```

## Counts by year
```{r}
dfsimallreps_summary <- dfsimallreps %>% 
  ungroup() %>% 
  group_by(rep, Years, Species) %>% 
  mutate(count = n()) %>% 
  mutate(count = as.numeric(count))
```

```{r}
ggplot(dfsimallreps_summary, aes(x = as.factor(Years), y = count, fill = Species, col = Species))+
  geom_boxplot(alpha = .2, outlier.alpha = .02)+
  geom_smooth(aes(x = as.factor(Years), y = count, fill = Species, col = Species), size = 1)+
  ggtitle("Results for 1000 simulations")+
  xlab("Years since fire")+
  ylab("Tree count")+
  theme_bw()+
  scale_color_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  scale_fill_manual(values = c("#899DA4", "#9A8822"), labels = c("Fir", "Pine"))+
  theme(text = element_text(size = 16), 
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 8))
ggsave(file = "../../results/figures/sim_1000_count.png", width = 6, height = 4, dpi = 400)
```

## Shrub height by year
```{r}
dfsimallreps_summary <- dfsimallreps %>% 
  ungroup() %>% 
  group_by(rep, Years, Ht1.3, ShrubSpp03) %>% 
  mutate(mean_shrub_ht = mean(Ht1.3))
```

```{r}
ggplot(dfsimallreps_summary, aes(x = Years, y = mean_shrub_ht, group = ShrubSpp03, col = ShrubSpp03))+
  geom_line(aes(group = Sdlg), alpha = .2)+
  geom_smooth()+
  ggtitle("Results for 1000 simulations")+
  xlab("Years since fire")+
  ylab("Shrub height")+
  theme_bw()
```

```{r}
print(Sys.time() - strt)
```


# Next steps to improve the model

1.  Use Kristen's data or Hugh's data for initial conditions
2.  Improve dispersal kernel based on Kristen/Hugh's data 
3.  Improve shrub growth based on data - DONE
4.  Include residual surviving trees and their seed dispersal
5.  Include seed dispersal of post-fire regen once it reaches reproductive age 
6.  Add customization of patch size and shape
7.  Add customization of whether the weather conditions reflect those of 2015, 2016, or 2017
8.  Change sapling growth equations once they emerge from the shrub canopy


For next week:
- Improve shrub growth based on data
- display dominant shrub species
- make the shrub grid dependent upon surrounding cells so it's not so checkerboard
- Update display of shrub competition after simulation years
- what does shrub competition mean for new recruitment?
- "emergent year" = when 50% of trees are above shrub canopy
- maybe submit to American Naturalist
- Global Change Biology - mixing up the years
- no overstory reproduction for now 
- apply to King, American River Complex, rest of the fires I measured
- switch diameter equation to be from dendro work


