---
title: "simulations_plot"
author: "Carmen"
date: "2/13/2020"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
require(tidyverse)
require(gridExtra)
```

```{r}
name <- "Simulation1_2020-10-30_1000_empirical_welch_empiricalcoef_empiricalgrowth_empiricalindex"
df_filename <- paste("~/Ch3_Simulation_Results/", name, ".Rdata", sep = "")
load(df_filename)
strt <- Sys.time()
```

```{r}
setwd("~/../Shrubs-Seedlings/code/Simulations/")
```

# Find emerged trees at 8 years (first simulation year)
```{r}
dfsimallreps %>% filter(Years == 8) %>% filter(emerged ==1) %>% dplyr::select(Sdlg, Species, ShrubSpp03, Ht1.3, Ht_cm1, emerged) %>% 
  distinct()
```

These trees are either the one ABCO that had emerged at data collection OR trees under CEIN, because the Welch CEIN heights I drew from included some very short CEIN

# Find out what is going on with non-finite or NA pred 
```{r}
dfsimallreps %>% 
  filter(!is.finite(pred_exp) | is.na(pred_exp)) %>% 
  group_by(Years, ShrubSpp03) %>% 
  arrange(Years) %>% 
  count() 
```

## Make sure # reps, seedlings per rep, and minimum year matches what I expect
```{r}
length(unique(dfsimallreps$rep))
```

```{r}
min(dfsimallreps$Years)
```

```{r}
dfsimallreps %>% 
  group_by(rep) %>% 
  filter(Years == min(Years)) %>% 
  count() %>% 
  filter(n != 100)
```

# Plot how many trees emerge above the shrub canopy each year

## How many trees are neither emerged nor dead at the end? Should be none.
```{r}
dfsimallreps %>% 
  filter(Years == max(Years)) %>% 
  group_by(emerged, Species) %>% 
  count() %>% 
  filter(emerged==0)
```

# Calculate mean and SD starting # of each species
```{r}
df8 <- dfsimallreps %>% filter(Years == min(Years))
```

```{r}
df8 %>% 
  group_by(rep, Species) %>% 
  count() %>% 
  group_by(Species) %>% 
  summarize(mean(n), sd(n), min(n), max(n))
```

# Summarize emergents per year per rep
```{r}
summary_emerge <- dfsimallreps %>% 
  group_by(Years, Species, rep) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  mutate(Species = fct_relevel(Species, "PIPO", "ABCO"))
```

# Calculate % emergence by the end
```{r}
summary_emerge %>% 
  group_by(Species) %>% 
  filter(Years == max(Years)) %>% 
  summarize(mean_prop = mean(prop_emerged), sd = sd(prop_emerged))
```

# Plot emergents per year 

## Counts

```{r}
plot_emerge_count <- ggplot(summary_emerge)+
  theme_minimal()+
  geom_boxplot(aes(x = as.factor(Years), y = total_emergents, fill = Species), alpha = .8, outlier.shape = NA)+
  ylab("Number of emerged trees")+
  xlab("Years since fire")+
  scale_x_discrete(breaks =c("10","20","30", "40"))+
  theme_bw()+
    theme(text = element_text(size = 11), 
        legend.title = element_blank(),
        legend.text=element_text(size=11),
        legend.position = "none")
plot_emerge_count
```

## Proportions

```{r}
plot_emerge_prop <- ggplot(summary_emerge)+
  geom_boxplot(aes(x = as.factor(Years), y = prop_emerged, fill = Species), alpha = .8, outlier.shape = NA)+
  ylab("Proportion of initial trees that have emerged")+
  xlab("Years since fire")+
  scale_x_discrete(breaks =c("10","20","30", "40"))+
  theme_bw()+
  theme(text = element_text(size = 11), 
        legend.title = element_blank(),
        legend.text=element_text(size=11),
        legend.position = c(0.7, 0.2))+
  scale_fill_discrete(labels = c( "Ponderosa pine", "White fir"))

plot_emerge_prop
```


# Plot Q50 

## Summarize with means
```{r}
summary_emerge <- dfsimallreps %>% 
  group_by(Years, Species, rep) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, Species) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```

## Calculate bounds
```{r}
Q50 <- summary_emerge %>% 
  mutate(dist50 = mean-.5) %>% 
  filter(dist50>0) %>%
  ungroup() %>% 
  group_by(Species) %>% 
  filter(dist50 == min(dist50)) %>% 
  mutate(bound = "mean")

Q50_ub <- summary_emerge %>% 
  mutate(lowerdist50 = mean-sd-.5) %>% 
  filter(lowerdist50>0) %>%
  ungroup() %>% 
  group_by(Species) %>% 
  filter(lowerdist50 == min(lowerdist50)) %>% 
  mutate(bound = "upper")

Q50_lb <- summary_emerge %>% 
  mutate(upperdist50 = mean+sd-.5) %>% 
  filter(upperdist50>0) %>%
  ungroup() %>% 
  group_by(Species) %>% 
  filter(upperdist50 == min(upperdist50)) %>% 
  mutate(bound = "lower")
```

## Reformat table
```{r}
Q50 <- full_join(Q50, Q50_lb) 
Q50 <- full_join(Q50, Q50_ub) 

Q50 <- Q50 %>% 
  dplyr::select(Years, Species, bound) 

Q50 <- pivot_wider(Q50, names_from = bound, values_from = Years) %>% 
  dplyr::select(Species, lower, mean, upper)

Q50
```


# Make sure the starting ratios of the results match Welch ratios

## Load Welch ratios
```{r}
load("../../results/coefficients/welch_ratios.Rdata")
```

## By conifer species
```{r}
dfsimallreps %>% 
  filter(Years == min(Years)) %>% 
  group_by(Species) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(prop = n/sum(n)) 
welch_ratios %>% 
  group_by(Species) %>% 
  summarize(n = sum(trees)) %>% 
  ungroup() %>% 
  mutate(prop = n/sum(n)) 
```

## Average across reps
```{r}
init_counts_by_rep <-  dfsimallreps %>% 
  filter(Years == min(Years)) %>% 
  group_by(Species, rep) %>% 
  count() %>% 
  group_by(Species) %>% 
  summarize(mean(n))
init_counts_by_rep
```

## By conifer/shrub combo
```{r}
dfsimallreps %>% 
  filter(Years == min(Years)) %>% 
  group_by(Species, ShrubSpp03) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(prop = n/sum(n)) %>% 
  arrange(ShrubSpp03)
welch_ratios
```

# Density - Plot on a per-area basis

Each plot in Welch's data was 60 m^2. 

## Load Welch data on # plots
```{r}
load("../../results/coefficients/n_plots_welch.Rdata")
```

## Calculate total plot area and density in the welch data
```{r}
welch_area_m2 <-  60*n_plots
welch_area_ha <- welch_area_m2/10000
welch_area_ha
```

```{r}
welch_density <- welch_ratios %>% 
  group_by(Species) %>% 
  summarize(trees_welch = sum(trees)) %>% 
  mutate(density_m2 = trees_welch/welch_area_m2) %>% 
  mutate(density_ha = density_m2*10000)
welch_density
```

Welch data has 144 ABCO and 101 PIPO per ha.


## Calculate initial area equivalent for my results

### At the first year of the simulation, there were 56.5 ABCO and 39 PIPO
```{r}
calc_area <- dfsimallreps %>% 
  filter(Years == min(Years)) %>% 
  group_by(rep, Species) %>% 
  count() %>% 
  group_by(Species) %>% 
  summarize(mean_trees_sim = mean(n)) 
calc_area
```

### So how many ha is the simulated area?
```{r}
area <- full_join(calc_area, welch_density) %>% 
  mutate(area_sim_ha = mean_trees_sim/density_ha) %>% 
  summarize(area_sim_ha = mean(area_sim_ha))
area_sim_ha <- area %>%  unlist()
area_sim_ha
```

### Add column for initial area to summary_emerge

```{r}
summary_emerge <- dfsimallreps %>% 
  group_by(Years, Species, rep) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  mutate(Species = fct_relevel(Species, "PIPO", "ABCO")) %>% 
  mutate(density_ha = total_emergents/area_sim_ha)
summary_emerge
```

```{r}
plot_emerge_density <- ggplot(summary_emerge)+
  theme_minimal()+
  geom_boxplot(aes(x = as.factor(Years), y = density_ha, fill = Species), alpha = .8, outlier.shape = NA)+
  ylab("Density of emerged trees per ha")+
  xlab("Years since fire")+
  scale_x_discrete(breaks =c("10","20","30", "40"))+
  theme_bw()+
    theme(text = element_text(size = 11), 
        legend.title = element_blank(),
        legend.text=element_text(size=11),
        legend.position = "none")
plot_emerge_density
filename <- paste("../../results/figures/Simulations/base_density_boxplot", name, ".png", sep = "")
ggsave(file = filename, width = 5, height =4, dpi = 400)
```

# Calculate average density at end of the simulation
```{r}
summary_emerge %>% 
  group_by(rep, Species) %>% 
  filter(prop_emerged==max(prop_emerged)) %>% 
  arrange(rep) %>% 
  filter(Years == min(Years)) %>% 
  group_by(Species) %>% 
  summarize(mean_ha = mean(density_ha), min_ha = min(density_ha), max_ha = max(density_ha), sd_ha = sd(density_ha))
```


# Plot the way I plotted the species ones

## Summarize
```{r}
summary_emerge_prop <- dfsimallreps %>% 
  group_by(Years, rep, Species) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, Species) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```

### Filter out the flat parts - take out years if they differ by <0.001 (0.1%)
```{r}
max_year <- summary_emerge_prop %>% 
  mutate(mean = round(mean, 3)) %>% 
  group_by(Species) %>% 
  mutate(maxmean = max(mean)) %>% 
  filter(mean == maxmean) %>% 
  ungroup() %>%
  group_by(Species) %>% 
  filter(Years == min(Years)) %>%
  ungroup() %>% 
  filter(Years ==max(Years)) %>% 
  dplyr::select(Years) %>% 
  unlist()

max_year
```

```{r}
summary_emerge_prop <- summary_emerge_prop %>% 
  filter(Years < max_year + 2)
```

## Save prop summary
```{r}
save(summary_emerge_prop, file = "../../results/data/summary_emerge_prop_base_Oct30.Rdata")
```

## Plot 
```{r}
ggplot(summary_emerge_prop)+
  theme_bw()+
  geom_line(aes(x = Years, y = mean, col = Species, group = Species, linetype = Species), size = 1)+
  geom_ribbon(aes(x = Years, ymin = mean-sd, ymax = mean+sd, fill = Species, group = Species), alpha = .3)+
  ylab("Proportion of initial trees that have emerged")+
  xlab("Years since fire")+
  theme(legend.position = c(0.75, 0.3))+
   scale_fill_manual(values = c( "#899DA4", "#9A8822"), 
                                labels = c("White fir", "Ponderosa pine"))+

  scale_color_manual(values = c("#899DA4", "#9A8822"), 
                                   labels = c("White fir", "Ponderosa pine"))+

  scale_linetype_discrete(     labels = c("White fir", "Ponderosa pine"))+

  
  theme(text = element_text(size = 11), 
        
        legend.title = element_blank(),
        legend.spacing=unit(0, "cm"),
        legend.text=element_text(size=10),
        legend.margin = margin(0, 0, 0, 0)
        )+
  xlab(element_blank())+
  xlim(8,max_year)+
  ylim(c(-0.05,1))
filename <- paste("../../results/figures/Simulations/base_prop_line_Oct30", name, ".png", sep = "")
ggsave(file = filename, width = 5, height =4, dpi = 400)
```


## Plot Density

```{r}
summary_emerge_density <- dfsimallreps %>% 
  group_by(Years, Species, rep) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  mutate(density_ha = total_emergents/area_sim_ha) %>% 
  group_by(Years, Species) %>% 
  summarize(mean = mean(density_ha), sd = sd(density_ha))
```

```{r}
summary_emerge_density <- summary_emerge_density %>% 
  filter(Years < max_year + 2)
```

## Save density summary
```{r}
save(summary_emerge_density, file = "../../results/data/summary_emerge_density_base_Oct30.Rdata")
```

```{r}
ggplot(summary_emerge_density)+
  theme_bw()+
  geom_line(aes(x = Years, y = mean, col = Species, group = Species, linetype = Species), size = 1)+
  geom_ribbon(aes(x = Years, ymin = mean-sd, ymax = mean+sd, fill = Species, group = Species), alpha = .3)+
  
  xlab("Years since fire")+
  theme(legend.position = c(0.75, 0.3))+
    ylab("Density of emerged trees per ha")+
  scale_fill_manual(values = c( "#899DA4", "#9A8822"), 
                                labels = c("White fir", "Ponderosa pine"))+

  scale_color_manual(values = c("#899DA4", "#9A8822"), 
                                   labels = c("White fir", "Ponderosa pine"))+

  scale_linetype_discrete(     labels = c("White fir", "Ponderosa pine"))+

  
  theme(text = element_text(size = 11), 
        
        legend.position = "none"
        )

filename <- paste("../../results/figures/Simulations/base_density_line_Oct30", name, ".png", sep = "")
ggsave(file = filename, width = 5, height =4, dpi = 400)
```

# Calculate mean and SD years to end of simulation
```{r}
summary_emerge %>% 
  group_by(rep, Species) %>% 
  filter(total_emergents == max(total_emergents)) %>% 
  filter(Years==min(Years)) %>% 
  group_by(Species) %>% 
  summarize(mean = mean(Years), min = min(Years), max = max(Years), sd = sd(Years))
```
