---
title: "simulations_plot_ShrubSpecies"
author: "Carmen"
date: "3/6/2020"
output: html_document
---

```{r}
strt <- Sys.time()
setwd("~/../Shrubs-Seedlings/code/Simulations/")
```

```{r, include = F}
require(tidyverse)
require(gridExtra)
require(egg)
require(grid)
```


```{r}
date <- "2020-10-27"

namecein <- paste("~/Ch3_Simulation_Results/Simulation5_", date, "_100_CEIN.Rdata", sep = "")

namearpa <- paste("~/Ch3_Simulation_Results/Simulation4_", date, "_100_ARPA.Rdata", sep = "")

namececo <- paste("~/Ch3_Simulation_Results/Simulation3_", date, "_1000_CECO.Rdata", sep = "")
```

```{r}
load(namearpa)
arpa <- dfsimallreps
remove(dfsimallreps)

load(namececo)
ceco <- dfsimallreps
remove(dfsimallreps)

load(namecein)
cein <- dfsimallreps
remove(dfsimallreps)
```

```{r}
arpa <- arpa %>% 
  mutate(sim_spp = "ARPA")
ceco <- ceco %>% 
  mutate(sim_spp = "CECO")
cein <- cein %>% 
  mutate(sim_spp = "CEIN")
```

```{r}
nrow_total <- nrow(arpa) + nrow(ceco) + nrow(cein)
``` 

```{r}
df <- full_join(arpa, ceco)
remove(arpa, ceco)
df <- full_join(df, cein)
remove(cein)
```

```{r}
nrow_total == nrow(df)
```

# Take out trees with non-finite or NA pred because they had already emerged and things got wacky
```{r}
df %>% 
  filter(!is.finite(pred_exp) | is.na(pred_exp)) %>% 
  group_by(emerged) %>% 
  count()
```

```{r}
df <- df %>% 
 mutate(pred_exp = ifelse(!is.finite(pred_exp) | !is.na(pred_exp), 999, pred_exp))
```


## Summarize

### Mean
```{r}
summary_emerge <- df %>% 
  group_by(Years, Species, rep, sim_spp) %>% 
  summarize(total_emergents = sum(emerged), total = n()) %>% 
  ungroup() %>% 
  group_by(rep, Species, sim_spp) %>% 
  mutate(initial_n = max(total)) %>% 
  ungroup() %>% 
  mutate(prop_emerged = total_emergents/initial_n) %>% 
  group_by(Years, Species, sim_spp) %>% 
  summarize(mean = mean(prop_emerged), sd = sd(prop_emerged))
```


## Filter out the flat parts - take out years if they differ by <0.001 (0.1%)
```{r}
max_year <- summary_emerge %>% 
  mutate(mean = round(mean, 3)) %>% 
  group_by(Species, sim_spp) %>% 
  mutate(maxmean = max(mean)) %>% 
  filter(mean == maxmean) %>% 
  ungroup() %>%
  group_by(Species, sim_spp) %>% 
  filter(Years == min(Years)) %>%
  ungroup() %>% 
  filter(Years ==max(Years)) %>% 
  dplyr::select(Years) %>% 
  unlist()

max_year
```

```{r}
summary_emerge <- summary_emerge %>% 
  filter(Years < max_year + 1)
```

# Save summarized data
```{r}
save(summary_emerge, file = "../../results/data/summary_emerge_prop_ShrubSpp.Rdata")
```

