---
title: "Calc_Welch_ratios"
output: 
  html_document:
    toc: TRUE
---


```{r, include = F}
require(tidyverse)
require(lubridate)
require(readxl)
require(gridExtra)
```

```{r}
load(file = "~/../Shrubs-Seedlings/compiled_data/welch_5_8_yrs.Rdata")
```

# Summarize cover of dominant shrub by regen plot
```{r}
welch_5.8years_summary <- welch_subset_5.8years %>% 
  ungroup() %>% 
  group_by(Regen_Plot) %>% 
  filter(Cover == max(Cover)) %>% 
  mutate(shrubarea = Cover*modal_ht_cm) %>% 
  group_by(Regen_Plot) %>% 
  filter(shrubarea == max(shrubarea)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species %in% c("ARPA6", "ARVI"), "ARPA/ARVI", Shrub_species)) %>% 
  group_by(Shrub_species) %>% 
  mutate(mean_cov = mean(Cover)) %>% 
  ungroup() %>% 
  mutate(Shrub_species = fct_reorder(Shrub_species, desc(mean_cov))) %>% 
  select(Regen_Plot, Shrub_species, Cover) %>% 
  distinct()
```

## Check
```{r}
nrow(welch_5.8years_summary) == length(unique(welch_5.8years_summary$Regen_Plot))
```

### Make sure there's only one row per plot/Species combo
```{r}
welch_5.8years_summary %>% 
  group_by(Regen_Plot) %>% 
  count() %>% 
  arrange(desc(n))
```

#### Fix the one plot with a duplicate
```{r}
welch_5.8years_summary %>% 
  filter(Regen_Plot == "CUB1300199")
```


```{r}
welch_subset_5.8years %>% 
  filter(Regen_Plot == "CUB1300199") %>% 
  select(Regen_Plot, Shrub_species, Cover, Count_total, Species)
```

```{r}
welch_5.8years_summary <- welch_5.8years_summary %>% 
  filter(!(Regen_Plot == "CUB1300199" & Shrub_species == "QUVA"))
```


## Plot for supplement

```{r}
welch_5.8years_summary <- welch_5.8years_summary %>% 
  group_by(Shrub_species) %>% 
  mutate(mean_cov = mean(Cover))
```


```{r}
ggplot(welch_5.8years_summary)+
  geom_boxplot(aes(x = Shrub_species, y = Cover))+
  theme_minimal()+
  geom_point(aes(x = Shrub_species, y = mean_cov), col = "red")+
  xlab("Shrub species")
ggsave(file = "../../../../results/figures/Simulations/welch_cein_ceco_ht_compare.png", width = 8)
```



# Look at overall combinations of shrub and tree species across Welch data at 5-8 years post-fire

```{r}
welch_summary_combos <- welch_subset_5.8years %>% 
  ungroup() %>% 
  group_by(Regen_Plot) %>% 
  filter(Cover == max(Cover)) %>% 
  mutate(shrubarea = Cover*modal_ht_cm) %>% 
  group_by(Regen_Plot) %>% 
  filter(shrubarea == max(shrubarea)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species %in% c("ARPA6", "ARVI"), "ARPA/ARVI", Shrub_species)) %>% 
  ungroup()  %>% 
  select(Regen_Plot, Shrub_species, Cover, Species, Count_total) %>% 
  arrange(Regen_Plot)
welch_summary_combos
```

# Remove plots with unapplicable shrub species

## Plots dominated by shrub species that weren't even in the Other category in my Fire Footprint data, which included:

 ARNE   ARPA   ARVI   CECO   CEIN   CEPR   CHFO   CHSE   CONU   LIDE   none  other 
  PREM   QUKE   QUVA willow   NA's 
```{r}
sort(unique(welch_summary_combos$Shrub_species))
```

### Look at what's getting removed
```{r}
welch_summary_combos %>% 
  filter(Shrub_species %in% c("SYMO", "RIBES"))
```

### Remove it
```{r}
welch_5.8years_summary <- welch_5.8years_summary %>% 
  filter(!Shrub_species %in% c("SYMO", "RIBES"))
```

```{r}
welch_summary_combos <- welch_summary_combos %>% 
  filter(!Shrub_species %in% c("SYMO", "RIBES"))
```


## Take out CHFO because it's not in my American River fire

CHFO isn't in the top 4 species ny # trees in the Welch data, and it would be difficult to initialize without knowing who establishes under it

THIS IS THE STEP THAT DRASTICALLY CHANGES THE RESULTS BECAUSE THERE'S SO MUCH PIPO UNDER CHFO

### Look at what's getting removed
```{r}
welch_summary_combos %>% 
  filter(Species == "PIPO" & Shrub_species == "CHFO")
```

```{r}
welch_summary_combos %>% 
  filter(Species == "ABCO" & Shrub_species == "CHFO")
```


What to do about CHSE? There are some problems:
1. There are no PIPO/CHSE combos in my data
2. There is TONS of ABCO under CHSE according to Welch data
3. CHSE was not one of the species with a coefficient in Chapter 2
4. Why did I leave CHSE separate in the simulations instead of lumping it under "other"? The only reason is that it's separate for shrub growth models from Welch data. 
5. So what I need to do is lump it with "other" for the shrub gams and then it can be lumped with "other" for everything.
6. Then I can leave in ALL trees under CHSE. Since there are only 4 PIPO, it won't change the results much.
6. Lumping shouldn't change the GAMs much, either, because the GAM shapes are similar for CHSE and for Other

### Remove it

```{r}
welch_5.8years_summary <- welch_5.8years_summary %>% 
  filter(Shrub_species != "CHFO")
```

```{r}
welch_summary_combos <- welch_summary_combos %>% 
  filter(Shrub_species != "CHFO")
```

# Save the welch subsets
```{r}
save(welch_subset_5.8years, file = "~/../Shrubs-Seedlings/compiled_data/welch_5_8_yrs_noCHFO.Rdata")
save(welch_5.8years_summary, file = "~/../Shrubs-Seedlings/compiled_data/welch_shrub_5_8_yrs_noCHFO.Rdata")
```

# Count total regen plots, with or without seedlings present, excluding those dominated by CHFO
```{r}
n_plots <- length(unique(welch_5.8years_summary$Regen_Plot))
n_plots
save(n_plots, file = "~/../Shrubs-Seedlings/results/coefficients/n_plots_welch.Rdata")
```


# Calculate ratios
```{r}
welch_ratios <- welch_summary_combos %>% 
  group_by(Shrub_species, Species) %>% 
  summarize(plots = n(), trees = sum(na.omit(Count_total))) %>% 
  filter(Species %in% c("ABCO", "PIPO"))
welch_ratios
```

## Reformat so it's compatible with my shrub data

### Which shrub species are most common?
```{r}
welch_ratios %>% 
  summarize(sum(plots), trees = sum(trees)) %>% 
  arrange(desc(trees))
```

### Which tree species is most common?
```{r}
welch_ratios %>% 
  group_by(Species) %>% 
  summarize(sum(trees))
```

```{r}
main_spp <- c("ARPA", "CECO", "CEIN", "Other")

welch_ratios <- welch_ratios %>% 
  ungroup() %>% 
  mutate(Shrub_species = case_when(
    Shrub_species == "CEIN3" ~ "CEIN",
    Shrub_species == "ARPA/ARVI" ~ "ARPA",
    !Shrub_species %in% main_spp ~ "Other",
    TRUE ~ as.character(Shrub_species)
  )) %>% 
  group_by(Shrub_species, Species) %>% 
  summarize(trees = sum(trees)) %>% 
  ungroup() %>% 
  mutate(prop = trees/sum(trees))

welch_ratios
  ```

```{r}
welch_ratios %>% 
  group_by(Species) %>% 
  summarize(sum(prop))
```


```{r}
save(welch_ratios, file = "../../../../results/coefficients/welch_ratios.Rdata")
```


# EXTRA STUFF THAT'S NOT NECESSARY 

## Summarize MY AMRC data

```{r}
load(file="../../../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
me <- df
remove(df)
```

```{r}
me_AMRC <- me %>%
  dplyr::select(Sdlg, Species, Cov1.3, Ht1.3, ShrubSpp03, shrubarea3, BasDia2016.cm, Ht2016.cm_spring, heatload, incidrad, Slope.Deg, Elevation, Fire, Years, Year) %>%
  filter(Fire == "AMRC") %>% 
  filter(Year==2016) %>%
  mutate(Cov_prop = Cov1.3/1200) %>%
  distinct() %>%
  droplevels()
```


### Count 

#### Tree species by shrub species
```{r}
me_AMRC %>% 
  filter(Species %in% c("PIPO", "ABCO")) %>% 
  group_by(Species, ShrubSpp03) %>% 
  summarize(sum = n()) 
```

#### Total # trees in my focal spp combinations, I guess?
```{r}
me_AMRC %>% 
  filter(Species %in% c("PIPO", "ABCO")) %>% 
  group_by(Species, ShrubSpp03) %>% 
  summarize(sum = n()) %>% 
  filter(ShrubSpp03 %in% c("ARPA", "CECO", "CEIN")) %>% 
  ungroup() %>% 
  summarize(sum(sum))
```

### Plot species composition in my AMRC data
```{r}
me_shrspp_AMRC <- me_AMRC %>% 
  filter(Species %in% c("PIPO", "ABCO")) %>% 
  group_by(Species, ShrubSpp03) %>% 
  summarize(sum = n()) %>% 
  ungroup() %>% 
  group_by(Species) %>% 
  mutate(prop = sum/sum(sum)) %>% 
  #filter(ShrubSpp03 %in% c("ARPA", "CECO", "CHSE")) %>% 
  ggplot()+
  geom_col(aes(x = Species, y = prop, fill = ShrubSpp03), position = "dodge")+
  theme_minimal()+
  ylim(c(0, 0.7))+
  ggtitle("My AMRC data")
me_shrspp_AMRC
```

## Compare shrub species composition in Welch AMRC data to my AMRC data

### Filter welch to American River Complex
```{r}
welch_AMRC <- welch_subset_5.8years %>% 
  filter(Fire == "AMERICAN RIVER")
```

### Look at the shrubs
```{r}
welch_AMRC %>% 
  group_by(Shrub_species) %>% 
  count()
```

### Restructure data so that there's only one row per tree species per regen plot

Choose the dominant shrub species by choosing highest shrub cover, then highest shrubarea
```{r}
welch_shrspp <- welch_AMRC %>% 
  ungroup() %>% 
  group_by(Regen_Plot) %>% 
  filter(Cover == max(Cover)) %>% 
  mutate(shrubarea = Cover*modal_ht_cm) %>% 
  group_by(Regen_Plot) %>% 
  filter(shrubarea == max(shrubarea)) %>% 
  dplyr::select(Regen_Plot, Shrub_species, Species, Cover, modal_ht_cm, Fire, Count_total)
welch_shrspp
```

### Make sure there's only one row per plot/Species combo
```{r}
welch_shrspp %>% 
  group_by(Regen_Plot, Species) %>% 
  filter(n()>1)
```

### See total trees in each species and shrub combo
```{r}
welch_shrspp <- welch_shrspp %>% 
  filter(Shrub_species %in% c("CECO", "ARPA6", "CHSE11", "QUVA", "CEIN3")) %>% 
  mutate(Shrub_species = ifelse(!Shrub_species %in% c("CECO", "ARPA6", "CEIN3", "CHFO"), "Other", Shrub_species)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species == "ARPA6", "ARPA", Shrub_species)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species == "CEIN3", "CEIN", Shrub_species)) %>% 
  filter(Species %in% c("PIPO", "ABCO")) %>% 
  group_by(Species, Shrub_species) %>% 
  summarize(plots = n(), trees = sum(Count_total)) %>% 
  ungroup() %>% 
  mutate(proportion = trees/sum(trees))
welch_shrspp
```

### Proportion of fir vs. pine?
```{r}
welch_shrspp %>% 
  ungroup() %>% 
  group_by(Species) %>% 
  summarize(total = sum(trees)) %>% 
  ungroup() %>% 
  mutate(prop = round(total/sum(total), 2))
```


## Compare shrub height between Welch AMRC data and my AMRC data
```{r}
welch_shr_ht <- ggplot(welch_AMRC %>% filter(Shrub_species %in% c("ARPA6", "CECO", "CHSE11")))+
  geom_boxplot(aes(x = Shrub_species, y = modal_ht_cm, fill = Shrub_species))+
  theme_minimal()+
  ylim(15, 125)+
  ggtitle("Welch data at 5 years\nSHRUB HEIGHT")+
  theme(legend.position = "none") 
welch_shr_ht
```

```{r}
me_shr_ht <- ggplot(me_AMRC %>% filter(ShrubSpp03 %in% c("ARPA", "CECO", "CHSE")))+
  geom_boxplot(aes(x = ShrubSpp03, y = Ht1.3, fill = ShrubSpp03))+
  theme_minimal()+
  ylim(15, 125)+
  ggtitle("My data at 8 years\nSHRUB HEIGHT")+
  theme(legend.position = "none") 
me_shr_ht
```

```{r}
grid.arrange(welch_shr_ht, me_shr_ht, ncol = 2)
```
