---
title: "Clean_Welch_regen_densities"
author: "Carmen"
date: "3/26/2020"
output: html_document
---


```{r, include = F}
setwd("~/../Shrubs-Seedlings/code/Shrub_growth_time/")
require(tidyverse)
require(lubridate)
require(readxl)
```

# Load Welch data
```{r}
plot <- read_excel("../../data/Data_for_shrub_growth/Welch_data/plot_data_RegenDatabase_1_13_16_updatedDates.xlsx") 
seed_tree <- read_excel("../../data/Data_for_shrub_growth/Welch_data/seed_tree_RegenDatabase_1_13_16_updatedDates.xlsx") 
shrub <- read_excel("../../data/Data_for_shrub_growth/Welch_data/shrub_regen_RegenDatabase_1_13_16_updatedDates.xlsx") 
tree <- read_excel("~/Shrubs-Seedlings/data/Data_for_shrub_growth/Welch_data/tree_regen_RegenDatabase_1_13_16_updatedDates.xlsx") 
```

# Are there plots without any tree regen?
```{r}
no_regen_plots <- anti_join(plot, tree, by = "Regen_Plot") %>% 
  filter(Fire == "AMERICAN RIVER") %>% 
  select(Regen_Plot, FIRE_SEV) %>% 
  filter(FIRE_SEV %in% c(4,5))
tree %>% filter(Regen_Plot == "AMR1300209")
```

There are 7 plots in the American River Complex Fire that don't have any regen. 4 of them are in high severity.

```{r}
seed_tree %>% 
  filter(Regen_Plot %in% no_regen_plots$Regen_Plot)
```

The plots that have no tree regen in AMRC and are fire sev 4 or 5 have seed trees < 20 m. This means that the 9 plots with trees in the AMRC from the Welch data are the 9 I analyze. 

# Take out unnecessary columns

## Plot
```{r}
plot <- plot %>% 
  dplyr::select(ID, Regen_Plot, Fire, Date, `Year of Fire`, FIRE_SEV)
```

## shrub
```{r}
shrub <- shrub %>% 
  dplyr::select(-comments, Cover, -ID, -modal_ht_ft, -OldPLOTID)
```

# Join shrub data and plot data
```{r}
df <- left_join(shrub, plot, by = "Regen_Plot")
```

# Rename shrub column
```{r}
head(df)
df <- df %>% 
   rename(Shrub_species = Species) 
head(df)
```

# Filter plots by fire severity
```{r}
df <- df %>% 
  filter(FIRE_SEV > 3)
```

# Reshape seed tree table to join with df
```{r}
seed_tree <- seed_tree %>% 
  group_by(Regen_Plot) %>% 
  filter(Dist_m == min(Dist_m)) %>% 
  dplyr::select(Regen_Plot, Dist_m) %>% 
  rename(seed_tree_dist_m = Dist_m) %>% 
  distinct()
```


# Join seed tree table with df
```{r}
df <- left_join(df, seed_tree)
```


# If seed tree distance is NA, replace it with 999 based on March 27 email with Derek Young
```{r}
df <- df %>% 
  mutate(seed_tree_dist_m = ifelse(is.na(seed_tree_dist_m), 999, seed_tree_dist_m))
```

# Filter df to plots with >20 m distance to seed tree
```{r}
distdf <- df %>% 
  filter(Fire == "AMERICAN RIVER") %>% 
  dplyr::select(Regen_Plot, seed_tree_dist_m) %>% 
  distinct() %>% 
  mutate(far = ifelse(seed_tree_dist_m>20, 1, 0)) 

distdf %>% 
  group_by(far) %>% 
  count()


df <- df %>% 
  filter(seed_tree_dist_m>20)
```

# Filter to shrub species of interest
```{r}
summary(as.factor(df$Shrub_species))
df <- df %>% 
  filter(Shrub_species %in% c("ARNE", "ARPA", "ARPA6", "ARVI", "CECO", "CEIN","CEIN3", "CEPR", "CHFO", "CHSE11", "PREM", "QUVA", "RIBES", "RICE", "RIRO", "SYMO"))
summary(as.factor(df$Shrub_species))
```

# Fix species name typos
```{r}
df <- df %>% 
  mutate(Shrub_species = 
           case_when(Shrub_species == "ARPA" ~ "ARPA6",
                     Shrub_species == "CEIN" ~ "CEIN3",
                    Shrub_species == "RICE" ~ "RIBES",
                     Shrub_species == "RIRO" ~ "RIBES",
                     TRUE ~ Shrub_species))
```

# Check for NA modal shrub ht
```{r}
df %>% filter(is.na(modal_ht_cm))
```

# Look at column names
```{r}
names(df)
```

# Filter to American River Complex Fire
```{r}
df <- df %>% filter(Fire == "AMERICAN RIVER")
```

# Count plots
```{r}
df %>% 
  group_by(Regen_Plot) %>% 
  summarize(sum())
  
```

# List plots with CEIN 
```{r}
df %>% 
  group_by(Regen_Plot) %>% 
  filter(Shrub_species == "CEIN3")
```


# Compare to plot table
```{r}
plot %>% 
  filter(Fire == "AMERICAN RIVER") %>% 
  group_by(FIRE_SEV, seed_tree_dist_m) %>% 
  count()
```

