---
title: "Count_Welch_Plots_5-8yrs"
output: 
  html_document:
    toc: TRUE
---

To count plots, I can only include plots that meet the following requirements:
- no 

```{r, include = F}
setwd("~/../Shrubs-Seedlings/code/Simulations/parameter_dev/Davis_initialization/")
require(tidyverse)
require(lubridate)
require(readxl)
```

```{r}
load(file = "~/../Shrubs-Seedlings/compiled_data/welch_5_8_yrs.Rdata")
welch_5.8yrs <- welch_subset_5.8years
```

```{r}
welch_5.8yrs %>% 
  group_by(Regen_Plot) %>% 
  summarize(sum_regen = sum(Count_total)) %>% 
  arrange(sum_regen)
```

# Remove plots with dominant shrub species that aren't in my AMRC data, even in the "other" category of my shrub data

The shrub species that get to stay are: ARPA   CECO   CEIN   CEPR   CHFO   CHSE   LIDE PREM   QUVA 


## Find the dominant shrub species of each plot
```{r}
welch_5.8yrs_shrubdom <- welch_5.8yrs %>% 
  ungroup() %>% 
  group_by(Regen_Plot) %>% 
  filter(Cover == max(Cover)) %>% 
  mutate(shrubarea = Cover*modal_ht_cm) %>% 
  group_by(Regen_Plot) %>% 
  filter(shrubarea == max(shrubarea)) %>% 
  dplyr::select(Regen_Plot, Shrub_species) %>% 
  distinct()
summary(as.factor(welch_5.8yrs_shrubdom$Shrub_species))
```

## Make a list of all the plots that have dominant shrub species outside of the ones in my AMRC data
```{r}
no_go_regen_plots <- welch_5.8yrs_shrubdom %>% 
  filter(Shrub_species %in% c("ARNE", "RIBES", "SYMO"))
```

## Remove those plots from the welch data
```{r}
welch_5.8yrs <- welch_5.8yrs %>% 
  filter(!Regen_Plot %in% no_go_regen_plots$Regen_Plot)
```

# Count total regen plots
```{r}
n_plots <- length(unique(welch_5.8yrs$Regen_Plot))
n_plots
```

# Take out repeated plots and plots without PIPO or ABCO seedlings in order to tabulate regen
```{r}
welch_5.8yrs <- welch_5.8yrs %>% 
  mutate(Species = ifelse(is.na(Species), "0", Species)) %>% 
  filter(Species %in% c("0", "PIPO", "ABCO")) %>% 
  dplyr::select(years_since_fire, Regen_Plot, Fire, Count_total, Species) %>% 
  distinct()
```

# Count ABCO vs. PIPO
```{r}
welch_5.8yrs %>% 
  group_by(Species) %>% 
  summarize(sum(Count_total))
```


# Plot densities of regen histogram
```{r}
welch_5.8yrs %>% 
  mutate(Species = ifelse(is.na(Species), "0", Species)) %>% 
  filter(Species %in% c("0", "PIPO", "ABCO")) %>% 
  group_by(Regen_Plot) %>% 
  summarize(sum_regen = sum(Count_total)) %>% 
  ggplot()+
  geom_histogram(aes(sum_regen))+
  theme_minimal()
```

# Check plot seedling densities against Access database (this checked out) 
```{r}
welch_5.8yrs %>% 
  mutate(Species = ifelse(is.na(Species), "0", Species)) %>% 
  filter(Species %in% c("0", "PIPO", "ABCO")) %>% 
  group_by(Regen_Plot) %>% 
  summarize(sum_regen = sum(Count_total)) %>% 
  arrange(desc(sum_regen))
```

# Calculate density of fir and pine per ha using plot size

Each plot was 60 m2 according to Welch Ecosphere paper

Total area of all plots = 
```{r}
total_area_ha <- n_plots*60/10000
total_area_ha
```
ha.

In this area, there were a total of 
```{r}
welch_5.8yrs %>% 
  mutate(Species = ifelse(is.na(Species), "0", Species)) %>% 
  filter(Species %in% c("0", "PIPO", "ABCO")) %>% 
  group_by(Species) %>% 
  summarize(sum_regen = sum(Count_total)) 
```
tree per species.

This equals the following densities:

ABCO: 
```{r}
472/total_area_ha
```
130 trees per ha


PIPO:
```{r}
406/total_area_ha
```
111.8 trees per ha
