---
title: "Fill_in_CEIN"
output: 
  html_document:
    toc: TRUE
---


```{r, include = F}
setwd("~/../Shrubs-Seedlings/code/Simulations/parameter_dev/Davis_initialization/")
require(tidyverse)
require(lubridate)
require(readxl)
require(gridExtra)
```

# Load data

## Welch data
```{r}
tree <- read_excel("~/../Shrubs-Seedlings/data/Data_for_shrub_growth/Welch_data/tree_regen_RegenDatabase_1_13_16_updatedDates.xlsx") 
load(file = "~/../Shrubs-Seedlings/compiled_data/shrub_growth/Welch_df.Rdata")
```



```{r}
welch_5.8years_summary <- welch_subset_5.8years %>% 
  ungroup() %>% 
  group_by(Regen_Plot) %>% 
  filter(Cover == max(Cover)) %>% 
  mutate(shrubarea = Cover*modal_ht_cm) %>% 
  group_by(Regen_Plot) %>% 
  filter(shrubarea == max(shrubarea)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species %in% c("ARPA6", "ARVI"), "ARPA/ARVI", Shrub_species)) %>% 
  group_by(Shrub_species) %>% 
  mutate(mean_cov = mean(Cover)) %>% 
  ungroup() %>% 
  mutate(Shrub_species = fct_reorder(Shrub_species, desc(mean_cov))) 

ggplot(welch_5.8years_summary)+
  geom_boxplot(aes(x = Shrub_species, y = Cover))+
  theme_minimal()+
  geom_point(aes(x = Shrub_species, y = mean_cov), col = "red")+
  xlab("Shrub species")
ggsave(file = "../../../../results/figures/Simulations/welch_cein_ceco_ht_compare.png", width = 8)

```


# Clean Welch tree data 

## Remove unnecessary columns
```{r}
tree <- tree %>% 
  dplyr::select(ID, Regen_Plot, Count_total, Type, Species, num_dead, seed_veg_plant, tallest_age, tallest_ht_cm, tallest_lastyr_cm, Remarks, number_sprouts)
```

## Combine with previously cleaned Welch data with info on shrubs 
```{r}
df <- left_join(df, tree, by = "Regen_Plot") 
```

```{r}
welch_all_fires <- df
```

```{r}
remove(df)
```


## Remove Zaca fire
```{r}
welch_all_fires <- welch_all_fires %>% 
  filter(Fire != "ZACA")
```


# Load my data

```{r}
load(file="~/../Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings_vert.Rdata")
me <- df
remove(df)
```

# Try to fill in missing CEIN data in my surveys

## See what initial HEIGHTS of CEIN look like compared to other shrub species at years 5-8 after fire, according to Welch data
```{r}
welch_all_fires %>% 
  filter(years_since_fire %in% c(5:8)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species %in% c("ARPA6", "ARVI"), "ARPA/ARVI", Shrub_species)) %>% 
  group_by(Shrub_species) %>% 
  mutate(mean_ht = mean(modal_ht_cm)) %>% 
  ungroup() %>% 
  mutate(Shrub_species = fct_reorder(Shrub_species, desc(mean_ht))) %>%
ggplot()+
  geom_boxplot(aes(x = Shrub_species, y = modal_ht_cm))+
  theme_minimal()+
  geom_point(aes(x = Shrub_species, y = mean_ht), col = "red")
```


# Save CEIN height data for use in simulations, limiting to fires that are 5-8 years old
```{r}
welch_CEIN_hts <- welch_all_fires %>% 
  filter(years_since_fire %in% c(5:8)) %>% 
  filter(Shrub_species == "CEIN3") %>% 
  mutate(shrubarea3 = sqrt(Cover*modal_ht_cm)) %>% 
  dplyr::select(years_since_fire, Regen_Plot, Cover, modal_ht_cm, Fire, Shrub_species) %>% 
  distinct()

welch_CEIN_hts
```

```{r}
save(welch_CEIN_hts, file = "../../../../data/welch_CEIN_hts.Rdata")
```

  
## Repeat for CEIN COVER, limiting it to plots where CEIN is dominant and fires that are 5-8 years old
```{r}
welch_subset_5.8years <- welch_all_fires %>% 
  filter(years_since_fire %in% c(5:8)) 
```


```{r}
welch_5.8years_summary <- welch_subset_5.8years %>% 
  ungroup() %>% 
  group_by(Regen_Plot) %>% 
  filter(Cover == max(Cover)) %>% 
  mutate(shrubarea = Cover*modal_ht_cm) %>% 
  group_by(Regen_Plot) %>% 
  filter(shrubarea == max(shrubarea)) %>% 
  mutate(Shrub_species = ifelse(Shrub_species %in% c("ARPA6", "ARVI"), "ARPA/ARVI", Shrub_species)) %>% 
  group_by(Shrub_species) %>% 
  mutate(mean_cov = mean(Cover)) %>% 
  ungroup() %>% 
  mutate(Shrub_species = fct_reorder(Shrub_species, desc(mean_cov))) 

ggplot(welch_5.8years_summary)+
  geom_boxplot(aes(x = Shrub_species, y = Cover))+
  theme_minimal()+
  geom_point(aes(x = Shrub_species, y = mean_cov), col = "red")+
  xlab("Shrub species")
ggsave(file = "../../../../results/figures/Simulations/welch_cein_ceco_ht_compare.png", width = 8)
```

