---
title: "add_site_class"
author: "Carmen"
date: "June 27, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
require(readxl)
require(dplyr)
require(tidyr)
require(ggplot2)
```

# Read in raster
```{r}
setwd("~/../Documents/Shrubs-Seedlings/code/FireFootprints_analysis/")
load("../../compiled_data/fire_footprints/master_seedlings.Rdata")
```

# Create functions

## Create site class variable
```{r}
df$siteclass <- 0
```

## by seedling
```{r}
scs <- function(seedling, site_class){
  df <<- df %>% 
    mutate(siteclass = ifelse(Sdlg == seedling, site_class, siteclass))
}
```

## by set of seedlings
```{r}
scss <- function(seedlings, site_class){
  df <<- df %>% 
    mutate(siteclass = ifelse(Sdlg %in% seedlings, site_class, siteclass))
}
```


## by patch
```{r}
scp <- function(patch, site_class){
  df <<- df %>% 
    mutate(siteclass = ifelse(FirePatch == patch, site_class, siteclass))
}
```

# Apply function by comparing points to site class layer "C:/Users/Carmen/Dropbox (Stephens Lab)/SORTIE/FireFootprints_2017/data/Site_class/Wilson_site_class_1_7_250m.tif" manually in QGIS file "Users\Carmen\Dropbox (Stephens Lab)\Chapter_2\Figure components\Seedling_locations.qgs"
```{r}
scp("AMRC-NW", 4)
scp("AMRC-NE", 4)
scp("AMRC-3", 3)
scp("AMRC-2", 4)
scss(c(85, 86, 87, 81, 83), 3)
scp("STAR-NW", 6)
scss(c(170:174), 4)
scp("STAR-mid", 4)
scp("STAR-S", 4)
scp("CLVD-N", 3)
scp("CLVD-SE", 4)
scp("FRDS-N", 4)
scp("FRDS-JB", 4)
scp("FRDS-mid", 4)
scp("FRDS-SE", 2)
scss(c(26, 27), 4)
scp("WRTS-1", 3)
scs(122, 6)
```

```{r}
df %>% 
  filter(siteclass == 0) %>% 
  dplyr::select(FirePatch) %>% 
  distinct()
```


```{r}
df %>% 
  dplyr::select(Sdlg, FirePatch, siteclass) %>% 
  filter(Sdlg %in% c(85, 86, 89, 227, 265, 161, 172, 176, 187,12, 130))
```

# Compare to site classes at Blodgett
```{r}

```

```{r}
summary(as.factor(df$siteclass))
```

# Save
```{r}
save(df, file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings1.Rdata")
```

