---
title: 'Analysis 6: Pinus with shrubarea'
author: "Carmen"
date: "March 9, 2017"
output: html_document
---
```{r, include=F}
setwd("~/Shrubs-Seedlings/Rdata")
load(file="master_data.Rdata")
library(nlme)
df <- subset(df, df$Species != "CADE")
dfPinus <- df[df$Species%in%c("PIPO","PILA") & !df$Fire %in% c("PLKN","WRTS"),]
varPower <- varPower(form=~Ht.cm)
library(plot3D)
```

- Best model without using shrubarea, to compare this analysis' results to:
```{r}
M1F.pipo <- gls(LastYearGrth.cm ~ Ht.cm + Fire + Ht.cm*Fire + Ht1.3 + Ht1.3*Ht.cm,
                weights = varPower,
                data= dfPinus, method = "ML")
AIC(M1F.pipo)

```

## Try using shrub volume

```{r}
dfPinus$shrubvol <- dfPinus$Ht1.2*(dfPinus$Cov1.2^2)

M.interact <- gls(LastYearGrth.cm ~ Ht.cm + Fire + Ht.cm*Fire + shrubvol*Ht.cm + shrubvol,
                weights = varPower,
                data= dfPinus, method = "ML")
AIC(M.interact)

E <- resid(M.interact, type="normalized")
coplot(E ~ Ht.cm | Fire, data=dfPinus, ylab="Normalized residuals", col=dfPinus$Fire)

plot(dfPinus$Cov1.2, resid(M.interact))
hist(resid(M.interact))

dfPinus$shrubvol <- dfPinus$Ht1*(dfPinus$Cov1^2)
M.interact <- gls(LastYearGrth.cm ~ Ht.cm + Fire + Ht.cm*Fire + shrubvol*Ht.cm + shrubvol,
                weights = varPower,
                data= dfPinus, method = "ML")
AIC(M.interact)
dfPinus$shrubvol <- dfPinus$Ht1.3*(dfPinus$Cov1.3^2)
M.interact <- gls(LastYearGrth.cm ~ Ht.cm + Fire + Ht.cm*Fire + shrubvol*Ht.cm + shrubvol,
                weights = varPower,
                data= dfPinus, method = "ML")
AIC(M.interact)
```

Results for shrub volume:

- AIC for 1-3 meters, cover squared: 587.6655
- AIC for 1-2 meters, cover squared: 587.6485
- AIC for 1 meter, cover squared: 590.1969

So it seems that 1-2 m makes the most sense

## Try with shrub area

```{r}
dfPinus$shrubarea <- dfPinus$Ht1.2*(dfPinus$Cov1.2)
M.interact2 <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm+shrubarea+Fire,
                  weights = varPower,
                  data= dfPinus, method = "ML")
AIC(M.interact2)
E <- resid(M.interact2, type="normalized")
coplot(E ~ Ht.cm | Fire, data=dfPinus, ylab="Normalized residuals", col=dfPinus$Fire)

plot(dfPinus$Cov1.2, resid(M.interact2))
hist(resid(M.interact2))

```

Results using shrub area:
- AIC for 1-3 meters: 587.8378
- AIC for 1-2 meters: 587.8073
- AIC for 1 meter:  591.0605

### Check out residuals in more detail
```{r}
E <- resid(M.interact2, type="normalized")
coplot(E ~ shrubarea | Fire, data=dfPinus, ylab="Normalized residuals", col=dfPinus$Fire)
coplot(E ~ Cov1.2 | Fire, data=dfPinus, ylab="Normalized residuals", col=dfPinus$Fire)
coplot(E ~ Ht1.2 | Fire, data=dfPinus, ylab="Normalized residuals", col=dfPinus$Fire)


plot(dfPinus$Cov1.2, E)
plot(dfPinus$Ht1.2, E)
plot(dfPinus$shrubarea, E)
hist(E)
plot(predict(M.interact2),E)
```

## Re-evaluate with REML
```{r}
M.interact2 <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm +shrubarea+Fire,
                   weights = varPower,
                   data= dfPinus, method = "REML")
```

## Plot predictions

```{r, echo=F}
library(ggplot2)
Ht.cm <- c(rep(seq(0,300,5),4))
fakedata <- as.data.frame(Ht.cm)
fakedata$shrubarea <- c(rep(c( 44375,88750,133125,177500 ),61))
fakedata$Fire <- as.factor(rep("AMCR",244))
predicted <- predict(M.interact2, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

dfAMCR <- subset(dfPinus, dfPinus$Fire=="AMCR")
dfCLVD <- subset(dfPinus, dfPinus$Fire=="CLVD")
dfSTAR <- subset(dfPinus, dfPinus$Fire=="STAR")
dfFRDS <- subset(dfPinus, dfPinus$Fire=="FRDS")

plotAMCR <- ggplot(pred[pred$Ht.cm<max(dfAMCR$Ht.cm) & pred$shrubarea<=max(dfAMCR$shrubarea) & pred$shrubarea>=min(dfAMCR$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfPinus$shrubarea)))+
  labs(title = "Pine seedling growth ~ \nHt, shrub area, AMRC fire")+
  ylim(0,55)+
  xlim(0,300)+
  xlab("Seedling height (cm)")+
  ylab("Seedling growth (cm)")+
  geom_point(data=dfAMCR, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

Ht.cm <- c(rep(seq(0,300,5),4))
fakedata <- as.data.frame(Ht.cm)
fakedata$shrubarea <- c(rep(c( 44375,88750,133125,177500 ),61))
fakedata$Fire <- as.factor(rep("FRDS",244))
predicted <- predict(M.interact2, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotFRDS <- ggplot(pred[pred$Ht.cm<max(dfFRDS$Ht.cm) & pred$shrubarea<=max(dfFRDS$shrubarea) & pred$shrubarea>=min(dfFRDS$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfPinus$shrubarea)))+
  labs(title = "Pine seedling growth ~ \nHt, shrub area, Freds Fire")+
  ylim(0,55)+
  xlim(0,300)+
  xlab("Seedling height (cm)") +
  ylab("Seedling growth (cm)")+
  geom_point(data=dfFRDS, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

Ht.cm <- c(rep(seq(0,300,5),4))
fakedata <- as.data.frame(Ht.cm)
fakedata$shrubarea <- c(rep(c( 44375,88750,133125,177500 ),61))
fakedata$Fire <- as.factor(rep("STAR",244))
predicted <- predict(M.interact2, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotSTAR <- ggplot(pred[pred$Ht.cm<max(dfSTAR$Ht.cm) & pred$shrubarea<=max(dfSTAR$shrubarea) & pred$shrubarea>=min(dfSTAR$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red", limits=c(0,max(dfPinus$shrubarea)))+
  labs(title = "Pine seedling growth ~ \nHt, shrub area, Star Fire")+
  ylim(0,55)+
  xlim(0,300)+
  xlab("Seedling height (cm)") +
  ylab("Seedling growth (cm)")+
  geom_point(data=dfSTAR, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

fakedata$Fire <- as.factor(rep("CLVD",244))
predicted <- predict(M.interact2, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotCLVD <- ggplot(pred[pred$Ht.cm<max(dfCLVD$Ht.cm) & pred$shrubarea<=max(dfCLVD$shrubarea) & pred$shrubarea>=min(dfCLVD$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfPinus$shrubarea)))+
  labs(title = "Pine seedling growth ~ \nHt, shrub area, CLVD Fire")+
  ylim(0,55)+
  xlim(0,300)+
  xlab("Seedling height (cm)") +
  ylab("Seedling growth (cm)")+
  geom_point(data=dfCLVD, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

require(gridExtra)
grid.arrange(plotAMCR, plotFRDS,plotSTAR,plotCLVD,nrow=2,ncol=2 )

```

## Plot in 3D
```{r}
plot3d(dfPinus[,c("Ht.cm","LastYearGrth.cm","shrubarea")], pch=as.integer(dfPinus$LastYearGrth.cm), , type="s", size=2)

scatterplot3d(dfPinus[,c("Ht.cm","LastYearGrth.cm","shrubarea")])

plot3D <- function(data, model, Ht_range, shrub_range, Fire, theta, species){
  data<- data[data$Fire==Fire,]
  xy <- expand.grid(Ht.cm=Ht_range, shrubarea=shrub_range, Fire = Fire)
  predmatrix <- matrix (nrow = 60, ncol = 60,  data = predict(model, newdata = data.frame(xy),interval = "prediction"))
  scatter3D(z=data$LastYearGrth.cm, x=data$Ht.cm, y=data$shrubarea, pch=18, cex=1, cex.axis=.5,theta = theta, phi = 20, ticktype = "detailed",
            xlab = "Height of juvenile conifer (cm)", ylab = "shrub cover*height (cm*cm)", zlab = "Conifer vertical growth (cm)",  
            zlim=c(0,max(data[data$Fire==Fire,"LastYearGrth.cm"])+10),
            xlim=c(0,max(data[data$Fire==Fire,"Ht.cm"])+10),
           surf = list(z=predmatrix, x=Ht_range,y=shrub_range,
                     facets = NA),
            main=paste(species, "seedling growth,",Fire,"Fire",sep=" "))
   }

plot3D(data=dfPinus, model=M.interact2, Ht_range=seq(0,max(dfPinus$Ht.cm[dfPinus$Fire=="FRDS"]),length.out=60),shrub_range=seq(0,max(dfPinus$shrubarea[dfPinus$Fire=="FRDS"]),length.out=60), Fire = "FRDS", theta=-150, species = "PIPO")

plot3D(data=dfPinus, model=M.interact2, Ht_range=seq(min(dfPinus$Ht.cm[dfPinus$Fire=="STAR"]),max(dfPinus$Ht.cm[dfPinus$Fire=="STAR"]),length.out=60),shrub_range=seq(min(dfPinus$shrubarea[dfPinus$Fire=="STAR"]),max(dfPinus$shrubarea[dfPinus$Fire=="STAR"]),length.out=60), Fire = "STAR", theta=-150, species = "PIPO")


plot3D(data=dfPinus, model=M.interact2, Ht_range=seq(0,max(dfPinus$Ht.cm[dfPinus$Fire=="AMCR"]),length.out=60),shrub_range=seq(0,max(dfPinus$shrubarea[dfPinus$Fire=="AMCR"]),length.out=60), Fire = "AMCR", theta=-150, species = "PIPO")

plot3D(data=dfPinus, model=M.interact2, Ht_range=seq(min(dfPinus$Ht.cm[dfPinus$Fire=="CLVD"]),max(dfPinus$Ht.cm[dfPinus$Fire=="CLVD"]),length.out=60),shrub_range=seq(min(dfPinus$shrubarea[dfPinus$Fire=="CLVD"]),max(dfPinus$shrubarea[dfPinus$Fire=="CLVD"]),length.out=60), Fire = "CLVD", theta=-150, species = "PIPO")


```


## Check AIC with similar models

```{r}

M.shrubarea <- gls(LastYearGrth.cm ~ Ht.cm + Fire + Ht.cm*Fire + shrubarea*Ht.cm + shrubarea,
                   weights = varPower,
                   data= dfPinus, method = "ML")
AIC(M.shrubarea)

AIC(update(M.shrubarea, . ~ . - Fire))
AIC(update(M.shrubarea, . ~ . - Fire-shrubarea))
AIC(update(M.shrubarea, . ~ . - shrubarea))
AIC(update(M.shrubarea, . ~ . - Fire-Fire*Ht.cm))
anova(M.shrubarea, (update(M.shrubarea, . ~ . - -shrubarea)))
AIC(update(M.shrubarea, . ~ . + BasDia.cm))
```

- hmm taking out main effects lowers AIC and doesn't have much effect on the model, likely because the intercept is already 0. But it seemes very non-standard to take out main effect so I won't 

## Are individual variables significant?

```{r}
summary(M.shrubarea)
```

- height*shrubarea interaction is significant
- for interactions with fire, must use ANOVA

```{r}
anova(M.shrubarea, (update(M.shrubarea, . ~ . - Fire*Ht.cm)))
```

- Yes, the fire interaction with Ht.cm is significant

## Calculate pseudo R^2

-proportion of variance explained by the model relative to
that explained by the simple mean of the data

```{r}
pred <- as.data.frame(predict(M.shrubarea))
pred$Ht.cm <- dfPinus$Ht.cm
pred$Fire <- dfPinus$Fire
pred$Grth <- dfPinus$LastYearGrth.cm
pred$Shrubarea <- dfPinus$shrubarea
#pred <- arrange(pred, predict(M3.lme))
library(ggplot2)
predicted <- ggplot(pred)+
  geom_point(aes(y=predict(M.shrubarea),x=Ht.cm,col=pred$Shrubarea, shape=pred$Fire))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Predicted values, Best Pinus model")

meas <- ggplot(dfPinus)+
  geom_point(aes(y=LastYearGrth.cm,x=Ht.cm,col=shrubarea, shape=Fire))+
  scale_color_gradient(low="blue",high="red")+
  labs(title = "Measured Pinus growth")

grid.arrange(predicted, meas,nrow=1,ncol=2)
meanGrth <- mean(dfPinus$LastYearGrth.cm)
sse <- (dfPinus$LastYearGrth.cm - predict(M.shrubarea))^2
sst <- (dfPinus$LastYearGrth.cm - meanGrth)^2
R2 <- 1 - (sum(sse)/sum(sst))
R2
```

## Try similar methods with DIFN

#### First run a model that's the same as the best model above but without 
```{r}
M.DIFN <- gls(LastYearGrth.cm ~ Ht.cm + Fire + Ht.cm*Fire + shrubarea*Ht.cm + shrubarea + DIFN.53,
                   weights = varPower,
                   data= dfPinus[!is.na(dfPinus$DIFN.53),], method = "ML")
AIC(M.DIFN)
```

Plot residuals
``` {r}
E <- resid(M.DIFN, type="normalized")
coplot(E ~ Ht.cm | Fire, data=dfPinus[!is.na(dfPinus$DIFN.53),], ylab="Normalized residuals", col=dfPinus[!is.na(dfPinus$DIFN.53),]$Fire)


```

Test AIC of similar models with and without DIFN.all and DIFN.53
Below, baseline is best model above without any DIFN. Positive numbers mean that the model is better than the model without DIFN in it.
```{r}
baseline <- update(M.DIFN, . ~ . - DIFN.53)
AIC(baseline) - AIC(M.DIFN)
AIC(baseline) - AIC(update(M.DIFN, . ~ . - DIFN.53 + DIFN.all))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - DIFN.53 + DIFN.all + DIFN.all*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - shrubarea-shrubarea*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - shrubarea-shrubarea*Ht.cm+DIFN.53*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - shrubarea -shrubarea*Ht.cm + DIFN.53*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . + DIFN.53*Ht.cm))
```

