---
title: "Analysis 4 - Seedling Growth by Species: ABCO"
author: "Carmen Tubbesing"
date: "3/1/2017"
output: html_document
---
- This is a based on some of what I learned in Analysis 2
    - for example, in terms of using VarPower
- Note: delete seedlings from PLKN and FRDS because there are not many ABCO in them
```{r, include=F}
setwd("~/Shrubs-Seedlings/Rdata")
load(file="master_data.Rdata")
library(nlme)
library(plot3D)
```

- Define function for calculating pseudo R^2
```{r}
pseudoR2_ABCO <- function(model, data){
  meanGrth <- mean(dfABCO$LastYearGrth.cm)
  sse <- (data$LastYearGrth.cm - predict(model))^2
  sst <- (data$LastYearGrth.cm - meanGrth)^2
  R2 <- 1 - (sum(sse)/sum(sst))
  return(R2)
}

```

#### Define function for 3D plotting
```{r}

plot3D <- function(data, model, Ht_range, shrub_range, Fire){
  xy <- expand.grid(Ht.cm=Ht_range, shrubarea=shrub_range, Fire = Fire)
  predmatrix <- matrix (nrow = 60, ncol = 60,  data = predict(model, newdata = data.frame(xy),interval = "prediction"))
  scatter3D(z=data$LastYearGrth.cm, x=data$Ht.cm, y=data$shrubarea, pch=18, cex=2, theta = 45, phi = 20, ticktype = "detailed",
            xlab = "Ht", ylab = "shrubarea", zlab = "growth",  
            surf = list(z=predmatrix, x=Ht.cm,y=shrubarea,
                        facets = NA),
            main="fakedata")
}

```


```{r}
dfABCO <- df[df$Species=="ABCO" & !df$Fire %in% c("PLKN","FRDS"),]
```

- Whittling it down to the top 4 fires for ABCO leaves out only three ABCO:
```{r}
nrow(df[df$Species=="ABCO",])
nrow(dfABCO)
```

## Use GLS function varPower
```{r}
varPower <- varPower(form=~Ht.cm)
```

## Start out with the best model for PIPO and test similar models

```{r}

dfABCO$shrubarea <- dfABCO$Ht1.2*(dfABCO$Cov1.2)
ABCO.M1 <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm+shrubarea+Fire,
                  weights = varPower,
                  data= dfABCO, method = "ML")
pseudoR2_ABCO(ABCO.M1, dfABCO)

AIC(ABCO.M1)
summary(ABCO.M1)
```

## Test importance of including shrubarea

```{r}
anova(ABCO.M1, update(ABCO.M1, . ~ . - shrubarea-shrubarea*Ht))
summary(ABCO.M1)
pseudoR2_ABCO(update(ABCO.M1, . ~ . - shrubarea-shrubarea*Ht), dfABCO)

table <- data.frame(model=character(), AIC=numeric(), stringsAsFactors=FALSE)
```

## Try adding basal diameter

```{r}
ABCO.M2 <- update(ABCO.M1, .~. +BasDia.cm)
AIC(ABCO.M2)
table[1,] <- c("ABCO.M1", AIC(ABCO.M1))
table <- rbind(table, c("ABCO.M2", AIC(ABCO.M2)))
table
anova(ABCO.M1,ABCO.M2)
summary(ABCO.M2)

ABCO.M3 <- update(ABCO.M1, .~. + BasDia.cm +BasDia.cm*Ht.cm)
AIC(ABCO.M3)
```
- So, adding basal diameter is ok but not necessary, and interaction between basal diameter and seedling height does not help

## How do the residuals looks?
```{r}
E <- resid(ABCO.M1, type="normalized")
coplot(E ~ Ht.cm | Fire, data=dfABCO, ylab="Normalized residuals", col=dfABCO$Fire)
```

## Try using shrubarea with ABCO

- The models with and without shrubarea are almost identical in AIC and logLik

## Plot with predict(fakedata)

- First, fit the model again using REML
```{r}
ABCO.M1 <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm+shrubarea+Fire,
                  weights = varPower,
                  data= dfABCO, method = "REML")
pseudoR2_ABCO(ABCO.M1, dfABCO)

```

## Try out 3D plotting function
```{r}
data=dfABCO
model = ABCO.M1
Ht_range = seq(0,150,length.out=60)
shrub_range = seq(44375,177500,length.out=60)
Fire = "AMCR"

plot3D <- function(data, model, Ht_range, shrub_range, Fire, theta){
  xy <- expand.grid(Ht.cm=Ht_range, shrubarea=shrub_range, Fire = Fire)
  predmatrix <- matrix (nrow = 60, ncol = 60,  data = predict(model, newdata = data.frame(xy),interval = "prediction"))
  scatter3D(z=data$LastYearGrth.cm, x=data$Ht.cm, y=data$shrubarea, pch=18, cex=2, theta = theta, phi = 20, ticktype = "detailed",
            xlab = "Ht", ylab = "shrubarea", zlab = "growth",  
            surf = list(z=predmatrix, x=Ht_range,y=shrub_range,
                        facets = NA),
            main="fakedata")
}

plot3D(data=dfABCO, model=ABCO.M1, Ht_range = seq(0,150,length.out=60),
shrub_range = seq(44375,177500,length.out=60),
Fire = "AMCR", theta=60)
```

```{r}

```



```{r}
library(ggplot2)
Ht.cm <- c(rep(seq(0,150,length.out=61),4))
fakedata <- as.data.frame(Ht.cm)
fakedata$shrubarea <- c(rep(c( 44375,88750,133125,177500 ),61))
fakedata$Fire <- as.factor(rep("AMCR",244))
predicted <- predict(ABCO.M1, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

dfAMCR <- subset(dfABCO, dfABCO$Fire=="AMCR")
dfCLVD <- subset(dfABCO, dfABCO$Fire=="CLVD")
dfSTAR <- subset(dfABCO, dfABCO$Fire=="STAR")
dfFRDS <- subset(dfABCO, dfABCO$Fire=="WRTS")

plotAMCR <- ggplot(pred[pred$Ht.cm<max(dfAMCR$Ht.cm) & pred$shrubarea<=max(dfAMCR$shrubarea) & pred$shrubarea>=min(dfAMCR$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfABCO$shrubarea)))+
  labs(title = "ABCO seedling growth ~ \nHt, shrub area, Am. Riv. Complex fire")+
  ylim(0,25)+
  xlim(0,125)+
  xlab("Seedling height (cm)")+
  ylab("Seedling growth (cm)")+
  geom_point(data=dfAMCR, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

plotAMCR

fakedata$Fire <- as.factor(rep("WRTS",244))
predicted <- predict(ABCO.M1, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotWRTS <- ggplot(pred[pred$Ht.cm<max(dfFRDS$Ht.cm) & pred$shrubarea<=max(dfFRDS$shrubarea) & pred$shrubarea>=min(dfFRDS$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfABCO$shrubarea)))+
  labs(title = "ABCO seedling growth ~ \nHt, shrub area, Wrights Fire")+
  ylim(0,25)+
  xlim(0,225)+
  xlab("Seedling height (cm)") +
  ylab("Seedling growth (cm)")+
  geom_point(data=dfFRDS, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

plotWRTS

fakedata$Fire <- as.factor(rep("STAR",244))
predicted <- predict(ABCO.M1, fakedata)
pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotSTAR <- ggplot(pred[pred$Ht.cm<max(dfSTAR$Ht.cm) & pred$shrubarea<=max(dfSTAR$shrubarea) & pred$shrubarea>=min(dfSTAR$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red", limits=c(0,max(dfABCO$shrubarea)))+
  labs(title = "ABCO seedling growth ~ \nHt, shrub area, Star Fire")+
  ylim(0,25)+
  xlim(0,150)+
  xlab("Seedling height (cm)") +
  ylab("Seedling growth (cm)")+
  geom_point(data=dfSTAR, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))
plotSTAR

fakedata$Fire <- as.factor(rep("CLVD",244))
predicted <- predict(ABCO.M1, fakedata)
pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotCLVD <- ggplot(pred[pred$Ht.cm<max(dfCLVD$Ht.cm) & pred$shrubarea<=max(dfCLVD$shrubarea) & pred$shrubarea>=min(dfCLVD$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfABCO$shrubarea)))+
  labs(title = "ABCO seedling growth ~ \nHt, shrub area, Cleveland Fire")+
  ylim(0,25)+
  xlim(0,150)+
  xlab("Seedling height (cm)") +
  ylab("Seedling growth (cm)")+
  geom_point(data=dfCLVD, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

plotCLVD
require(gridExtra)
all.fires <- grid.arrange(plotAMCR, plotWRTS,plotSTAR,plotCLVD,nrow=2,ncol=2 )
all.fires
```

## What happens if you analyze AMCR data alone and remove fire effect?

- Note: I removed an outlier from the AMCR data, which switched the effect of shrubarea from positive to negative at high values of growth and seedling ht

```{r}
ABCO.M1 <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm + shrubarea + Fire,
                  weights = varPower,
                  data= dfABCO, method = "REML")
AIC(ABCO.M1)
AIC(update(ABCO.M1, . ~ . - Fire - Fire*Ht.cm))

ABCO.AMCR <- gls(LastYearGrth.cm ~ Ht.cm + shrubarea*Ht.cm + shrubarea,
                  weights = varPower,
                  data= dfAMCR[dfAMCR$Ht.cm<100,], method = "REML")

predicted <- predict(ABCO.AMCR, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire

plotAMCR.2 <-ggplot(pred[pred$Ht.cm<max(dfAMCR$Ht.cm) & pred$shrubarea<=max(dfAMCR$shrubarea) & pred$shrubarea>=min(dfAMCR$shrubarea),])+
  geom_line(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfABCO$shrubarea)))+
  labs(title = "ABCO seedling growth ~ \nHt, shrub area, Am. Riv. Complex only model")+
  ylim(0,25)+
  xlim(0,125)+
  xlab("Seedling height (cm)")+
  ylab("Seedling growth (cm)")+
  geom_point(data=dfAMCR, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))
plotAMCR.2

AMCR.compare <- grid.arrange(plotAMCR,plotAMCR.2,nrow=2,ncol=2)
```

- Answer: It's important.

## Graph effect of change in basal diameter

```{r}
ABCO.M2 <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm + shrubarea + Fire + BasDia.cm,
                  weights = varPower,
                  data= dfABCO, method = "REML")
Ht.cm <- c(rep(seq(0,150,length.out=61),4))
fakedata <- as.data.frame(Ht.cm)
fakedata$shrubarea <- c(rep(c( 44375,88750,133125,177500 ),61))
fakedata$Fire <- as.factor(rep("AMCR",244))
fakedata$BasDia.cm <- c(rep(1,122),rep(3,122))
predicted <- predict(ABCO.M2, fakedata)

pred <- as.data.frame(predicted)
pred$shrubarea <- fakedata$shrubarea
pred$Ht.cm <- fakedata$Ht.cm
pred$Fire <- fakedata$Fire
pred$BasDia.cm <- fakedata$BasDia.cm

plotAMCR <- ggplot(pred[pred$Ht.cm<max(dfAMCR$Ht.cm) & pred$shrubarea<=max(dfAMCR$shrubarea) & pred$shrubarea>=min(dfAMCR$shrubarea),])+
  geom_point(aes(y= predicted ,group=shrubarea,x=Ht.cm,col=shrubarea,size=BasDia.cm))+
  scale_color_gradient(low="blue",high="red",limits=c(0,max(dfABCO$shrubarea)))+
  labs(title = "ABCO seedling growth ~ \nHt, shrub area, Am. Riv. Complex fire")+
  ylim(0,25)+
  xlim(0,125)+
  xlab("Seedling height (cm)")+
  ylab("Seedling growth (cm)")+
  geom_point(data=dfAMCR, aes(x=Ht.cm,y=LastYearGrth.cm,col=shrubarea))

plotAMCR
hist(dfAMCR$BasDia.cm)
```

- There is a negative relationship between basal diameter and vertical seedling growth - how will I incorporate that into SORTIE given that I was hoping to consistently model diameter allometrically with height? Perhaps this summer when I remeasure diameter I can relate basal diameter growth with height growth

## Try including DIFN

```{r}
baseline <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm + shrubarea + Fire,
                  weights = varPower,
                  data= dfABCO[!is.na(dfABCO$DIFN.53),], method = "ML")
M.DIFN <- gls(LastYearGrth.cm ~ Ht.cm + Ht.cm*Fire + shrubarea*Ht.cm + shrubarea + Fire + DIFN.53,
                  weights = varPower,
                  data= dfABCO[!is.na(dfABCO$DIFN.53),], method = "ML")

AIC(baseline)
AIC(baseline) - AIC(M.DIFN)
AIC(baseline) - AIC(update(M.DIFN, . ~ . - DIFN.53 + DIFN.all))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - DIFN.53 + DIFN.all + DIFN.all*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - shrubarea-shrubarea*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . - shrubarea -shrubarea*Ht.cm + DIFN.53*Ht.cm))
AIC(baseline) - AIC(update(M.DIFN, . ~ . + DIFN.53*Ht.cm))

```

#### it looks like the model without shrubarea and with DIFN.53 is slightly better. Check out its R^2 and coefficients
```{r}
M.DIFN <- update(M.DIFN, . ~ . - shrubarea -shrubarea*Ht.cm + DIFN.53*Ht.cm)
dfABCODIFN <- dfABCO[!is.na(dfABCO$DIFN.53),]
pseudoR2_ABCO(M.DIFN, dfABCODIFN)
pseudoR2_ABCO(baseline, dfABCODIFN)
summary(M.DIFN)
E <- resid(M.DIFN, type="normalized")
coplot(E ~ Ht.cm | Fire, data=dfABCODIFN, ylab="Normalized residuals", col=dfABCODIFN$Fire)

```

## Relationships between shrubarea and DIFN
```{r}
plot(dfABCODIFN$shrubarea,dfABCODIFN$DIFN.53)
library(mgcv)
M1 <- gam(dfABCODIFN$DIFN.53 ~ s(dfABCODIFN$shrubarea, fx=F, k=-1, bs="cr"))
plot(M1, se=TRUE)
plot(dfABCODIFN$DIFN.53,log(dfABCODIFN$shrubarea))
M2 <- gam(dfABCODIFN$DIFN.53 ~ s(log(dfABCODIFN$shrubarea), fx=F, k=-1, bs="cr"))
plot(M2, se=TRUE, add=T)

```

