---
title: "checks"
author: "Carmen"
date: "August 16, 2019"
output: html_document
---

```{r, include = F}
require(readxl)
require(dplyr)
require(tidyverse)
```

# Compare pine seedlings from Aug 15 analysis to those from Aug 19 analysis
```{r}
load("../../compiled_data/fire_footprints/pine_vert.Rdata")
dfnew <- df
remove(df)
load("../../compiled_data/archives/pine_vert_Aug15_2019.Rdata")
dfold <- df
remove(df)
```

## Remove irrelevant variables
```{r}
dfold <- dfold %>% 
  select(-E)
```


## Find Sdlg/year combinations that are different
```{r}
antijoin <- anti_join(dfold, dfnew)
antijoin
```

## Find repeats in old df
```{r}
dfold %>% 
  group_by(Sdlg, Year) %>% 
  filter(n()>1) %>% 
  arrange(Sdlg) %>% 
  dplyr::select(Sdlg, Year, everything())

```

```{r}
boo <- (function(x) x[sapply(x, nrow)>0])(lapply(lapply(antijoin, function(x) tapply(x, antijoin$Sdlg, function(x) x[1]!=x[2])), function(x) subset(antijoin, Sdlg %in% names(which(x)))))
print(boo)
```

```{r}
antijoin %>% 
  select(Sdlg, Years) %>% 
  group_by(Sdlg, Years) %>% 
  filter(n()>1) %>% 
  arrange(Sdlg)
```

```{r}
remove(dfnew, dfold, boo, antijoin)
```

# Look at 2016 seedlings not in both seedling df and shrub df and make sure nothing got missed
```{r}
load("../../compiled_data/fire_footprints/seedlings_cleaned_2016.Rdata")
df_seedlings <- df
remove(df)
load("../../compiled_data/fire_footprints/shrub_master_data_2016.Rdata")
```

## Find mismatching seedlings that are ABCO or PIPO
```{r}
anti_join(df_seedlings, shr_by_sdlg)
mismatch <- anti_join(shr_by_sdlg, df_seedlings)
mismatch
```

```{r}
original_seedlings <- read.csv("~/../Dropbox (Stephens Lab)/SORTIE/Shrubs_Summer16/Completed_Data_and_Photos/Master_Compiled_seedlings.csv")
mismatch <- mismatch %>% select(-Species)
mismatch <- left_join(mismatch, original_seedlings, by = c("Sdlg" = "Seedling."))
mismatch %>% 
  filter(Species %in% c("PIPO", "ABCO")) %>% 
  select(Sdlg, Fire, Patch, DataSheet, Note1, Note2)
```

Seedling 16 was marked as PIPO originally but marked as PILA in subsequent measurements.

```{r}
remove(df_seedlings, mismatch, original_seedlings, shr_by_sdlg)
```

# Randomly select trees and check them from original data sheet to present

## Pines
```{r}
load("../../compiled_data/fire_footprints/pine_vert.Rdata")
```

Why aren't both 2016 data sheets in here??

### Randomly choose a row number
```{r}
sample <- sort(sample(1:251, 10))
df[sample,] %>% select(Sdlg, Fire, FirePatch, Datasheet2016, Light_File, DataSheet2017, notes, Note1, Note2, everything()) %>% View()
```

### Look at those seedlings
```{r}
df[sample,] %>% select(Sdlg, Fire, FirePatch, Date1_2016, Date2_2016, Date2017, Datasheet2016, Light_File, DataSheet2017, notes, Note1, Note2, everything()) %>% View()
```

