---
title: "Clean_shrubs"
author: "Carmen"
date: "May 20, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
require(readxl)
require(dplyr)
require(tidyverse)
```

# Load data
```{r}
shr <- read.csv("~/../Dropbox (Stephens Lab)/SORTIE/Shrubs_Summer16/Completed_Data_and_Photos/Master_Compiled_shrubs.csv")
load("../../compiled_data/fire_footprints/seedlings_cleaned_2016.Rdata")
```

# Basic shrub cleaning

## Rename seedling column
```{r}
shr <- shr %>% 
  rename(Sdlg = Seedling.) 
```
 
## Replace NAs with 0
```{r}
shr <- tbl_df(shr)
shr$Dead.[is.na(shr$Dead.)] <- 0
shr$Cover.cm[is.na(shr$Cover.cm)] <- 0
shr$Ht.cm[is.na(shr$Ht.cm)] <- 0
```

## Take out "pea" shrub cover because it's not a shrub
```{r}
shr <- shr %>% filter(ShrubSpp != "pea")
```

## Take out "conifer" since it's not a shrub
```{r}
summary(as.factor(shr$ShrubSpp))
shr <- shr %>% 
  filter(!(ShrubSpp == "conifer"))
```

## Drop levels
```{r}
shr <- droplevels(shr)
```


# Aggregate shrub data by seedling

## Clean up segment names
```{r}
shr$Seg <- 0
for(i in 1:nrow(shr)){
if(shr$Segment[i] %in% c("0-1 m ","0 -1 m","0-1 m")){
  shr$Seg[i] <- 1
} else if (shr$Segment[i]=="1-2 m "){
  shr$Seg[i] <- 2
} else if (shr$Segment[i] %in% c("2-3 m","2-3m")){
  shr$Seg[i] <- 3
}  else
    shr$Seg[i] <- 999
}
summary(as.factor(shr$Seg))
```

# Merge fire with long shrub data to see how often CEIN was a minor component in AMCR
```{r}
fire <- df %>% 
  dplyr::select(Fire, Sdlg)
shr_fire <- left_join(shr, fire)
```

## Take a look at AMCR decomposed shrub data
```{r}
shr_fire %>% 
  filter(Fire == "AMRC") %>% 
  ggplot()+
  geom_boxplot(aes(x = ShrubSpp, y = Cover.cm))
```


## Cover

### Summarize shrub cover by segment
```{r}
cov <- shr %>%
  group_by(Sdlg, Seg) %>%
  summarise(cov =sum(Cover.cm)) %>%
  spread(Seg, cov) %>% 
  rename(Cov1=`1`) %>%
  rename(Cov2=`2`) %>%
  rename(Cov3=`3`)
```

```{r}
cov <- pivot_longer(cov, cols = c(2:4), names_to = "Seg", values_to = "cov") %>% 
  mutate(Seg = case_when(
    Seg == "Cov1" ~ 1,
    Seg == "Cov2" ~ 2,
    Seg == "Cov3" ~ 3
    
  ))
cov
```

### Replace NA shrub cover with 0
```{r}
cov <- cov %>% 
  replace_na(list(cov = 0))
```

### Find seedlings with 0 cover to inform height
```{r}
cov %>% filter(cov==0) %>% arrange(Sdlg)
```


## Height

### Summarize height by segment 
```{r}
ht <- shr %>%
  dplyr::select(Sdlg, ShrubSpp,Cover.cm, Ht.cm, Seg) %>% 
  group_by(Sdlg, Seg) %>% 
  mutate(ht_wm_seg.cm = weighted.mean(Ht.cm, Cover.cm)) %>% 
  ungroup() %>% 
  dplyr::select(Sdlg, Seg, ht_wm_seg.cm) %>% 
  distinct() 
```

### Replace NA height with 0
```{r}
ht %>% filter(is.na(ht_wm_seg.cm)) %>% arrange(Sdlg)
ht <- ht %>% 
  replace_na(list(ht_wm_seg.cm=0))
```

### Add in zeros where there were no shrubs measured in certain segments and it wasn't captured by the replace_na above
```{r}
cov %>% filter(Sdlg %in% c(290, 294, 38))
```

```{r}
ht %>% 
  group_by(Sdlg) %>% 
  filter(n() != 3)
```

```{r}
new_rows <- cov %>% 
  filter(Sdlg %in% c(290, 294) & cov == 0) %>% 
  rename(ht_wm_seg.cm = cov)
new_rows
```

```{r}
ht <- bind_rows(ht, new_rows)
```



## Combine cover and ht values into one data frame
```{r}
shr_by_seg <- full_join(cov,ht)
shr_by_seg
```

## Shrub species 0-3 m
```{r}
spp3 <- shr %>%
  group_by(Sdlg,ShrubSpp) %>% 
  mutate(cov=sum(Cover.cm)) %>%
  ungroup %>%
  group_by(Sdlg)%>%
  filter(cov == max(cov)) %>%
  dplyr::select(Sdlg,ShrubSpp,cov) %>%
  distinct(.keep_all=TRUE) %>% 
  rename(ShrubSpp03 = ShrubSpp)
spp3
```

```{r}
spp3 %>% 
  group_by(Sdlg) %>% 
  filter(n()>1)
```

##### Fix spp3 data for seedling 20
```{r}
shr %>%
  filter(Sdlg == 20 & ShrubSpp %in% c("PREM", "CECO")) %>%
  dplyr::select(Sdlg, ShrubSpp, Cover.cm, Ht.cm) %>%
  arrange(ShrubSpp) %>% 
  mutate(product = Cover.cm*Ht.cm) %>% 
  group_by(ShrubSpp) %>% 
  mutate(total_product = sum(product)) %>% 
  ungroup() %>% 
  mutate(winner = ifelse(total_product == max(total_product), 1, 0))
```

```{r}
spp3 <- spp3 %>%
  filter(!(Sdlg==20 & ShrubSpp03 =="PREM"))
```

##### Fix spp3 data for seedling 21
```{r}
shr %>%
  filter(Sdlg == 21 & ShrubSpp %in% c("PREM", "CECO")) %>%
  dplyr::select(Sdlg, ShrubSpp, Cover.cm, Ht.cm) %>%
  arrange(ShrubSpp) %>% 
  mutate(product = Cover.cm*Ht.cm) %>% 
  group_by(ShrubSpp) %>% 
  mutate(total_product = sum(product)) %>% 
  ungroup() %>% 
  mutate(winner = ifelse(total_product == max(total_product), 1, 0))
```

```{r}
spp3 <- spp3 %>%
  filter(!(Sdlg==21 & ShrubSpp03 =="PREM"))
```

##### Fix spp3 data for seedling 188
```{r}
shr %>%
  filter(Sdlg == 188 & ShrubSpp %in% c("CONU", "CECO")) %>%
  dplyr::select(Sdlg, ShrubSpp, Cover.cm, Ht.cm) %>%
  arrange(ShrubSpp) %>% 
  mutate(product = Cover.cm*Ht.cm) %>% 
  group_by(ShrubSpp) %>% 
  mutate(total_product = sum(product)) %>% 
  ungroup() %>% 
  mutate(winner = ifelse(total_product == max(total_product), 1, 0))
```

```{r}
spp3 <- spp3 %>%
  filter(!(Sdlg==188 & ShrubSpp03 =="CECO"))
```

##### Fix spp3 data for seedling 190
```{r}
shr %>% 
  filter(Sdlg == 190 & ShrubSpp %in% c("willow", "CECO")) %>% 
  dplyr::select(Sdlg, ShrubSpp, Cover.cm, Ht.cm) %>% 
  arrange(ShrubSpp) %>% 
  mutate(math = Cover.cm*Ht.cm) %>% 
  group_by(ShrubSpp) %>% 
  summarise(sum(math))
```

```{r}
spp3 <- spp3 %>% 
  filter(!(Sdlg==190 & ShrubSpp03 =="CECO"))
```

##### Fix spp3 data for seedling 285
```{r}
shr %>% 
  filter(Sdlg == 285 & ShrubSpp %in% c("ARPA", "CECO")) %>% 
  dplyr::select(Sdlg, ShrubSpp, Dead., Cover.cm, Ht.cm) 
shr %>% 
  filter(Sdlg == 285 & ShrubSpp %in% c("ARPA", "CECO")) %>% 
  dplyr::select(Sdlg, ShrubSpp, Dead., Cover.cm, Ht.cm) %>% 
  arrange(ShrubSpp) %>% 
  mutate(math = Cover.cm*Ht.cm) %>% 
  group_by(ShrubSpp) %>% 
  summarise(sum(math))
```

```{r}
spp3 <- spp3 %>% 
  filter(!(Sdlg==285 & ShrubSpp03 =="CECO"))
```

# Combine shrub data together
```{r}
spp3 <- spp3 %>% 
  ungroup() %>% 
  dplyr::select(Sdlg, ShrubSpp03)
shr_by_sdlg <- full_join(shr_by_seg, spp3, by = c("Sdlg" ="Sdlg"))
shr_by_sdlg
```

## Combine segments into 0-1, 0-2, and0-3 m
```{r}
shr_by_sdlg3 <- shr_by_sdlg %>% 
  summarize(Cov1.3 = sum(cov), Ht1.3 = weighted.mean(ht_wm_seg.cm, cov)) %>% 
  replace_na(list(Ht1.3 = 0))
shr_by_sdlg3
```

```{r}
shr_by_sdlg2 <- shr_by_sdlg %>% 
  filter(Seg %in% c(1:2)) %>% 
  summarize(Cov1.2 = sum(cov), Ht1.2 = weighted.mean(ht_wm_seg.cm, cov)) %>% 
  replace_na(list(Ht1.2 = 0))
shr_by_sdlg2
```

```{r}
shr_by_sdlg1 <- shr_by_sdlg %>% 
  filter(Seg ==1) %>% 
  rename(Cov1 = cov, Ht1 = ht_wm_seg.cm) %>% 
  replace_na(list(Ht1 = 0))
shr_by_sdlg1
```

```{r}
shr_by_sdlg <- full_join(shr_by_sdlg1, shr_by_sdlg2)
shr_by_sdlg <- full_join(shr_by_sdlg, shr_by_sdlg3) %>% 
  select(Sdlg, ShrubSpp03, everything(), -Seg)
shr_by_sdlg 
```


# Save
```{r}
save(shr_by_sdlg, file="../../compiled_data/fire_footprints/shrub_master_data_2016.Rdata")
#write.csv(shr_by_sdlg, file = "../../compiled_data/fire_footprints/shrub_master_data_2016.csv", row.names = F)
```

## This version has 286 rows and 8 columns

# Light files and DIFN

## Merge in DIFN and Light File data from seedling df
```{r}
names(df)
df_light <- df %>%
  dplyr::select(Sdlg, Light_File, DIFN, FirePatch, Species)
shr_by_sdlg <- left_join(shr_by_sdlg, df_light)
shr_by_sdlg
```

## Mark the measurements without seedling as "ref" for reference
```{r}
shr_by_sdlg <- shr_by_sdlg %>%
  mutate(ref = ifelse(Sdlg %in% c("ARPA-1",  "ceco-1",  "CEIN-1", "CEIN-2","chfo-1", "DCECO-1", "DCECO-2", "DCECO-3", "LIDE-1",  "LIDE-2", "LIDE-3", "LIDE-4"), 1,0))
summary(as.factor(shr_by_sdlg$ref))
```

# Make Sdlg all caps
```{r}
shr_by_sdlg <- shr_by_sdlg %>%
  ungroup() %>%
  mutate(Sdlg = toupper(Sdlg))
```

## Add light files for non-seedling shrub locations

```{r}
shr_by_sdlg <- shr_by_sdlg %>%
  mutate(Light_File = ifelse(Sdlg == "LIDE-4", 48, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "LIDE-3", 47, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "LIDE-2", 46, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "LIDE-1", 45, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "CEIN-1", 125, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "CEIN-2", 126, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "CECO-1", 74, Light_File)) %>%
  mutate(Light_File = ifelse(Sdlg == "ARPA-1", 94, Light_File))
```

# Add DIFN degrees from FV2200 output file

## Load in DIFN data using only the first 4 gaps (down to 53 degrees)
```{r}
DIFNl53 <- read.table(file="~/../Dropbox (Stephens Lab)/SORTIE/Shrubs_Summer16/Completed_Data_and_Photos/LAI-2000_data/AllDIFNGaps1-4",header=T,sep="\t")
names(shr_by_sdlg)
DIFNl53 <- distinct(DIFNl53)
```

## Delete repeat row that only has data in one of the rows
```{r}
DIFNl53 <- DIFNl53 %>%
  filter(!(LAI_File ==111 & DIFN==0))
```

## Join DIFN.53 data with shrub data
```{r}
shr_by_sdlg <- left_join(shr_by_sdlg, DIFNl53, by=c("Light_File"="LAI_File")) %>%
  rename(DIFN.all = DIFN.x) %>%
  rename(DIFN.53 =DIFN.y) %>%
  rename(LAI.53 = LAI) %>%
  dplyr::select(-SMP, -TransComp, -Model, -Records, -ScattCorr)
shr_by_sdlg
```

## Check for duplicated seedlings
```{r}
shr_by_sdlg %>%
  group_by(Sdlg) %>%
  summarize(n=n()) %>%
  filter(n !=1)
```

## Find which seedlings have DIFN.all but lack DIFN.53
```{r}
shr_by_sdlg %>%
  filter(!(is.na(DIFN.all) | DIFN.all==0) & is.na(DIFN.53))
```

## Fill in a couple seedlings with funky DIFN situations

Seedlings 3 and 9 have two light files. I average them for 3 and use one for 9 because the other one is funky

```{r}
shr_by_sdlg[shr_by_sdlg$Sdlg==9,"DIFN.53"] <- DIFNl53[DIFNl53$LAI_File==15,"DIFN"]
shr_by_sdlg[shr_by_sdlg$Sdlg==3,"DIFN.53"] <- mean(DIFNl53[DIFNl53$LAI_File%in%c(17,19),"DIFN"])
```

## Check again
```{r}
shr_by_sdlg %>%
  filter(!(is.na(DIFN.all) | DIFN.all==0) & is.na(DIFN.53))
```


# Figure out where the ARVI is
```{r}
shr_by_sdlg %>%
  filter(ShrubSpp03=="ARVI")
```

# Save for DIFN
```{r}
save(shr_by_sdlg, file="../../compiled_data/fire_footprints/shrub_master_data_2016_DIFN.Rdata")
save(shr_by_sdlg, file="../../compiled_data/fire_footprints/shrub_master_data_2016.Rdata")
```

This version has 286 rows and 16 columns

# List seedlings
```{r}
sort(shr_by_sdlg$Sdlg)
```


