---
title: "Clean_vert_growth"
author: "Carmen"
date: "April 16, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(tidyverse)
require(ggplot2)
require(effects)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings2.Rdata")
```

# Take out species that aren't PIPO or ABCO
```{r}
df <- df %>% filter(Species %in% c("PIPO", "ABCO"))
```

# Calculate 2015 height values

## Calculate 2015 and 2016 height by subtracting growth from 2017 height, unless there's no 2017 height value

Calculate 2015 and 2016 height as the height BEFORE 2015 growth
```{r}
df <- df %>% 
  mutate_at(c("Growth_15", "Growth_16","Growth_17"), as.numeric) %>%
  rename(Ht2017.cm.fall = Ht2017.cm) %>% 
  mutate(Ht2015.cm_spring_from2017 = Ht2017.cm.fall-Growth_16-Growth_17-Growth_15) %>% 
  mutate(Ht2016.cm_spring_from2017 = Ht2017.cm.fall-Growth_16-Growth_17) %>% 
  mutate(Ht2017.cm_spring_from2017 = Ht2017.cm.fall-Growth_17)
```

### Check: the following should be TRUE
```{r}
df %>% 
  dplyr::select(Ht2017.cm.fall, Growth_17, Growth_16, Growth_15, Ht2015.cm_spring_from2017) %>% 
  ungroup() %>% 
  mutate(check = Ht2015.cm_spring_from2017 + Growth_15 + Growth_16 + Growth_17 - Ht2017.cm.fall) %>% 
  filter(check > 0.0000001 | check < -0.0000001) %>% 
  nrow() ==0
```

## Fill in 2015 height for seedling 6, which for some reason lost that value somewhere along the way
```{r}
df <- df %>% 
  mutate(Ht2015.meas2016 = ifelse(Sdlg ==6, 9, Ht2015.meas2016))
```

### For those seedlings, find average 2016 growing season apical growth relative to 2015 growth 
```{r}
find_apical_growth <- df %>% 
  filter(!is.na(Ht2015.meas2016)) %>% 
  mutate(apical_growth_2016 = 
           Ht16_meas2016.cm - LastYearGrth.cm - Ht2015.meas2016) %>% 
  dplyr::select(Sdlg, Species, Ht16_meas2016.cm, LastYearGrth.cm, Ht2015.meas2016, apical_growth_2016) %>% 
  mutate(apical_ratio = apical_growth_2016/LastYearGrth.cm) %>% 
  distinct()
find_apical_growth
```

## For trees with measurements in both 2016 and 2017, see how apical growth compares to 2016 growth measured in 2017
```{r}
df_both_yrs <- df %>% 
  filter(!is.na(Growth_16) & !is.na(Ht2015.meas2016)) %>% 
  dplyr::select(Sdlg, Ht2016.cm_spring_from2017, Ht16_meas2016.cm, Ht2016.cm_spring_from2017, LastYearGrth.cm, Ht2015.meas2016, Ht2015.cm_spring_from2017, Growth_16, Date1_2016, Ht2017.cm.fall, Ht2017.cm_spring_from2017) %>% 
  distinct()
df_both_yrs <- left_join(df_both_yrs, find_apical_growth) %>% 
  mutate(Ht16_minus_apical = Ht16_meas2016.cm-apical_growth_2016) %>% 
  dplyr::select(Sdlg, Ht2016.cm_spring_from2017, Ht16_minus_apical, everything())
names(df_both_yrs)
```

### Plot it 
```{r}
ggplot(df_both_yrs, aes(x = apical_growth_2016, Growth_16, col = as.factor(Date1_2016)))+
  geom_point()+
  geom_abline(intercept = 0,slope = 1)
```

### Histogram of differences
```{r}
ggplot(df_both_yrs)+
  geom_histogram(aes(x = Growth_16 - apical_growth_2016))+
  xlim(-20, 20)
```

### Count neg vs pos
```{r}
df_both_yrs %>% 
  mutate(neg_growth = ifelse(Growth_16 - apical_growth_2016 < 0, 1, 0)) %>% 
  group_by(neg_growth) %>% 
  summarize(n())
df_both_yrs %>% 
  summarize(mean(Growth_16 - apical_growth_2016))
```

### Plot 2017 height as measured in 2017 against total height as measured in 2016, and check for a pattern with height
```{r}
ggplot(df_both_yrs, aes(x = Ht2017.cm_spring_from2017, y = Ht16_meas2016.cm, col = as.factor(Date1_2016)))+
  geom_point()+
  geom_abline(intercept = 0, slope = 1)+
  geom_smooth()
```

# Save these data up until now
```{r}
save(df, file = "../../compiled_data/fire_footprints/df_vert_wide.Rdata")
```

## Add 2015 heights measured in 2016 for trees still missing 2015 height

```{r}
df <- df %>% 
  mutate(Ht2015.cm_spring = ifelse(is.na(Ht2015.cm_spring_from2017), Ht2015.meas2016, Ht2015.cm_spring_from2017))
```

### Check
```{r}
df %>% 
  dplyr::select(Ht2017.cm.fall, Growth_17, Growth_16, Growth_15, Ht2015.cm_spring_from2017, Ht2015.meas2016, Ht2015.cm_spring) %>% 
  arrange(desc(Ht2015.meas2016))
```

### Count how many seedlings this applies to
```{r}
df %>% 
  filter(is.na(Ht2015.cm_spring_from2017) & !is.na(Ht2015.meas2016)) %>% 
  group_by(Species) %>% 
  summarize(n())
```

### Find seedlings this didn't work for
```{r}
df %>% 
  filter(is.na(Ht2015.cm_spring)) %>% 
  dplyr::select(Sdlg, FirePatch, LastYearGrth.cm, Ht16_meas2016.cm, Ht2017.cm.fall, Ht2015.meas2016, Growth_17, Growth_16, Growth_15, Ht2015.cm_spring)
```

### Count seedlings without 2017 height or 2015 height measured in 2016
```{r}
df %>% 
  filter(is.na(Ht2017.cm.fall) & is.na(Ht2015.meas2016)) %>% 
  group_by(Species) %>% 
  summarize(n())
```

### List them
```{r}
df %>% 
  filter(is.na(Ht2017.cm.fall) & is.na(Ht2015.meas2016)) %>% 
  select(Sdlg, Species, FirePatch, Datasheet1_2016, Datasheet2_2016, Date1_2016, Date2_2016) %>% 
  arrange(Date1_2016)
```

### For those seedlings, find 2015 growth by creating a regression model

### Find average 2016 growing season apical growth for the seedlings with data on it
```{r}
find_apical_growth <- df %>%
  filter(!is.na(Ht2015.meas2016)) %>%
  mutate(apical_growth_2016 =
           Ht16_meas2016.cm - LastYearGrth.cm - Ht2015.meas2016) %>%
  dplyr::select(Sdlg, Species, Ht16_meas2016.cm, LastYearGrth.cm, Ht2015.meas2016, apical_growth_2016, Date1_2016) %>%
  mutate(apical_ratio = apical_growth_2016/LastYearGrth.cm) %>%
  distinct()
find_apical_growth
```

### Check whether that ratio changes with height or date, excluding the very high values that seem like mistakes

#### Reformat data for LM

```{r}
df <- df %>% 
  mutate(Species = as.factor(Species))
test <- find_apical_growth %>% 
  mutate(Date1_2016 = as.numeric(Date1_2016)) %>% 
  mutate(Species = as.factor(Species))
```

#### Model growth based on species, date, and last year of growth
```{r}
lm <- lm(apical_growth_2016 ~ Date1_2016 + Species*LastYearGrth.cm, data = test)
summary(lm)
drop1(lm, test= "Chisq")
plot(predictorEffect("Species", lm))
plot(predictorEffect("LastYearGrth.cm", lm))
plot(predictorEffect("Date1_2016", lm))
```

### Look at seedlings who have a weirdly high ratio
```{r}
find_apical_growth %>%
  filter(apical_ratio>5)
```

### Find predicted growth for each seedling based on the linear model above
```{r}
df$apical_growth_pred <- predict(lm, newdata = df %>% mutate(Date1_2016 = as.numeric(Date1_2016)))
```

### Then use this value for the seedlings that don't have 2015 height recorded
```{r}
df <- df %>%
   mutate(Ht2015.calc2016 =
           ifelse(is.na(Ht2017.cm.fall) &
                    is.na(Ht2015.meas2016),
                  Ht16_meas2016.cm - LastYearGrth.cm - apical_growth_pred,
                  Ht2015.meas2016)) 

df %>%
  filter(is.na(Ht2017.cm.fall)) %>%
  dplyr::select(Sdlg, Ht16_meas2016.cm, LastYearGrth.cm, Ht2015.meas2016, Ht2015.calc2016)
```

## Assign 2015 spring height as Ht2015 calculated in 2016 if there's no 2017 value
```{r}
df <- df %>% 
  mutate(Ht2015.cm_spring = ifelse(is.na(Ht2015.cm_spring), Ht2015.calc2016, Ht2015.cm_spring))
```

## Find seedlings STILL missing 2015 height
```{r}
df %>% 
  filter(is.na(Ht2015.cm_spring)) %>% 
  dplyr::select(Sdlg, LastYearGrth.cm, Ht16_meas2016.cm, Ht2017.cm.fall, Ht2015.meas2016, Growth_17, Growth_16, Growth_15, Note1_2016, Note2_2016, Notes2017)
```

## Add height for those special cases
```{r}
df <- df %>% 
  mutate(Ht2015.cm_spring = ifelse(Sdlg == 6, Ht2017.cm.fall-Growth_15, Ht2015.cm_spring)) %>% 
  mutate(Ht2015.cm_spring = ifelse(Sdlg == 72, Ht2017.cm.fall-Growth_17 - Growth_16, Ht2015.cm_spring)) %>% 
  mutate(Ht2015.cm_spring = ifelse(Sdlg == 131, Ht2017.cm.fall-Growth_15, Ht2015.cm_spring)) %>% 
  mutate(Ht2015.cm_spring = ifelse(Sdlg == 161, Ht2017.cm.fall-Growth_15 - Growth_16, Ht2015.cm_spring))
```

# Add columns for 2016 and 2017 spring heights
```{r}
df <- df %>% 
  mutate(Ht2017.cm_spring = Ht2017.cm.fall - Growth_17) %>% 
  mutate(Ht2016.cm_spring = Ht2017.cm.fall - Growth_17 - Growth_16)
```

# Look at odd seedlings
```{r}
df %>% 
  filter(Sdlg %in% c(75, 161)) %>% 
  dplyr::select(Sdlg, LastYearGrth.cm, Ht16_meas2016.cm, Ht2017.cm.fall, Ht2015.meas2016, Growth_17, Growth_16, Growth_15, Note1_2016, Note2_2016, Notes2017, Ht2016.cm_spring, Ht2017.cm_spring, Ht2015.cm_spring)
```

## Fix them
```{r}
df <- df %>% 
  mutate(Ht2016.cm_spring = ifelse(Sdlg == 75, Ht2017.cm.fall - Growth_16, Ht2016.cm_spring)) %>% 
  mutate(Ht2015.cm_spring = ifelse(Sdlg == 75, Ht2017.cm.fall - Growth_16 - Growth_15, Ht2015.cm_spring)) %>% 
  mutate(Ht2016.cm_spring = ifelse(Sdlg == 161, Ht2017.cm.fall - Growth_16, Ht2016.cm_spring))
```


# Create data frame with one row per growth measurement
```{r}
df <- tbl_df(df)
df <- df %>% 
  rename(`2017` = Growth_17, `2016` = Growth_16, `2015` = Growth_15)
df <- gather(df, "Year", "VertGrowth_cm",c(`2017`, `2016`, `2015`))
df$Year <- as.factor(df$Year)
```

## Check
```{r}
df %>% dplyr::select(Sdlg, Year, VertGrowth_cm)
summary(df$Year)
```

# Remove bad rows

## Take out 2017 vertical growth measurements for trees that died since last measure
```{r}
summary(as.factor(df$Status))
df <- df %>% 
  filter(!(Status %in% c("DEAD", "dead") & Year == "2017"))
```

## Take out 2016 and 2017 vertical growth measurements for trees that were not found
```{r}
df <- df %>% 
  filter(!(Status %in% c("could not find","not found", "eaten", "herbicide","herbicided") & Year %in% c("2016", "2017")))
```

## Check out NA values and delete years that broke off
```{r}
df %>% 
  filter(is.na(VertGrowth_cm)) %>% 
  dplyr::select(Sdlg, Year, VertGrowth_cm, Status,  Note1_2016, Note2_2016, Notes2017) %>% 
  arrange(Sdlg)
df <- df %>% 
  filter(!(Sdlg %in% c("6","185","131","226","6","74","98", "132") & Year %in% c("2017","2016"))) %>% 
  filter(!(Sdlg %in% c(145, 195,78) & Year == "2016")) %>% 
  filter(!(Sdlg == 75 & Year == 2017))  
```

## Use growth measured in 2016 (which was 2015 growth) if 2015 growth couldn't be measured in 2017
```{r}
df <- df %>% 
  mutate(VertGrowth_cm = ifelse(is.na(VertGrowth_cm) & Year == "2015", LastYearGrth.cm, VertGrowth_cm))
``` 

## Look at odd vertical growth value
```{r}
df %>% 
  filter(Sdlg == "195") %>% 
  dplyr::select(Sdlg, Year, LastYearGrth.cm, VertGrowth_cm, Ht16_meas2016.cm, Ht2017.cm.fall, Ht2015.meas2016)
```

### Replace its 2015 growth with LastYearGrowth measured in 2016
```{r}
df <- df %>% 
  mutate(VertGrowth_cm = ifelse(Sdlg == "195" & Year == "2015", LastYearGrth.cm, VertGrowth_cm))
```

# Make VertGrowth numeric
```{r}
df$VertGrowth_cm <- as.numeric(df$VertGrowth_cm)
```

# Add height column that changes depending on year that represents height BEFORE that year's growth

```{r}
df <- df %>% 
  mutate(Ht_cm = ifelse(Year == "2017", Ht2017.cm_spring, 
                        ifelse(Year == "2016", Ht2016.cm_spring, 
                               ifelse(Year == "2015", Ht2015.cm_spring, 0))))
```

## Find missing heights
```{r}
df %>% 
  filter(is.na(Ht_cm)) %>% 
  dplyr::select(Sdlg, LastYearGrth.cm, Ht2016.cm_spring, Ht2017.cm.fall, Ht2015.meas2016, Year, VertGrowth_cm,  Note1_2016, Note2_2016, Notes2017)
```

# Change years since fire depending on year of growth
```{r}
df %>% 
  group_by(Fire, Year, Years) %>% 
  summarize(n())
```

```{r}
df <- df %>% 
  mutate(FireYear = ifelse(Fire == "AMRC", 2008, ifelse(Fire == "CLVD", 1992, ifelse(Fire == "FRDS", 2004, ifelse(Fire == "STAR", 2001, 1981))))) %>% 
  mutate(Years = as.numeric(paste(Year))-FireYear)
```

## Check
```{r}
df %>% 
  group_by(Fire, FireYear, Year, Years) %>% 
  summarize(n())
```

# Calculate relative vertical growth 
```{r}
df %>% dplyr::select(Sdlg, Year, Ht_cm, VertGrowth_cm) %>% arrange(Sdlg)
df <- df %>% 
  rename(Ht_cm1 = Ht_cm) %>% 
  mutate(Ht_cm2 = Ht_cm1+VertGrowth_cm) %>% 
  #mutate(VertGrowth_Rel = log(Ht_cm2) - log(Ht_cm1))
  mutate(VertGrowth_Rel = (Ht_cm2 - Ht_cm1)/Ht_cm1) %>% 
  mutate(VertGrowth_Rel = log(VertGrowth_Rel))
df %>% dplyr::select(Sdlg, Year, Ht_cm1, Ht_cm2, VertGrowth_cm, VertGrowth_Rel) %>% arrange(Sdlg)
```

## Find NAs
```{r}
df %>% 
  filter(is.na(VertGrowth_cm) | is.na(VertGrowth_Rel))
```


# Consolidate shrub species into fewer categories 
```{r}
df %>% 
  group_by(Species, ShrubSpp03) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) %>% 
  filter(Species == "ABCO") %>% 
  ungroup() %>% 
  mutate(percent = count/sum(count))
```

Pick the species with 5. For ABCO, the winning set is CECO, ARPA, CHSE, CEPR, PREM, and CEIN.

```{r}
df %>% 
  group_by(Species, ShrubSpp03) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) %>% 
  filter(Species == "PIPO") %>% 
  ungroup() %>% 
  mutate(percent = count/sum(count))
```

For PIPO, the winning set is CECO, ARPA, CHFO, CEIN, LIDE

```{r}
df <- df %>% 
  mutate(ShrubSpp03 = ifelse(Species == "ABCO" & !ShrubSpp03 %in% c("CECO", "ARPA", "CHSE", "CEPR", "PREM", "CEIN"), "Other", as.character(ShrubSpp03)))
```

```{r}
df <- df %>% 
  mutate(ShrubSpp03 = ifelse(Species == "PIPO" & !ShrubSpp03 %in% c("CECO", "ARPA", "CHFO", "LIDE", "CEIN"), "Other", as.character(ShrubSpp03)))
```

## Make them factors
```{r}
df <- df %>% 
  mutate(ShrubSpp03 = as.factor(ShrubSpp03))
```

## plot shrub cover growth by shrub species
```{r}
df_shrubs <- df %>% 
  mutate(Years = as.factor(Years)) %>% 
  group_by(ShrubSpp03, Years) %>% 
  summarize(shrub_cover = mean(Cov1.3))
ggplot(df_shrubs)+
  geom_line(aes(x = Years, y = shrub_cover, group=ShrubSpp03, col = ShrubSpp03))
```

## plot shrub cover growth overall
```{r}
df_shrubs <- df %>% 
  group_by(Fire, Cov1.3, Ht1.3) %>% 
  summarise(Years = mean(Years))
ggplot(df_shrubs, aes(x = Years, y = Cov1.3))+
  geom_point()+
  stat_smooth(method = "loess", span = 1)
```

## plot shrub HEIGHT growth overall
```{r}
ggplot(df_shrubs, aes(x = Years, y = Ht1.3))+
  geom_point()+
  stat_smooth(method = "loess", span = 1)
```

# Summarize shrub cover and height
```{r}
perc <- function(x){x/1200}
df %>% 
  summarize(mincov = min(Cov1.3), medcov = median(Cov1.3), meancov = mean(Cov1.3),maxcov = max(Cov1.3)) %>% 
  mutate_all(perc)

df %>% 
  summarize(min(Ht1.3), median(Ht1.3), mean(Ht1.3), max(Ht1.3))
```


# Drop extra levels
```{r}
df <- droplevels(df)
```

# Count by species
```{r}
df %>% 
  dplyr::select(Species, Sdlg) %>% 
  distinct() %>% 
  group_by(Species) %>% 
  summarize(n())
```

# Check for seedlings with too many years or repeat years
```{r}
repeats <- df %>% 
  group_by(Sdlg) %>% 
  summarize(n = n()) %>% 
  filter(n >3)
df_reps <- df %>% 
  filter(Sdlg %in% repeats$Sdlg) 
df %>% 
  group_by(Sdlg, Species, Year) %>% 
  filter(n()>1) %>% 
  arrange(desc(Species))
```

```{r}
boo <- (function(x) x[sapply(x, nrow)>0])(lapply(lapply(df_reps, function(x) tapply(x, df_reps$Sdlg, function(x) x[1]!=x[2])), function(x) subset(df_reps, Sdlg %in% names(which(x)))))
print(boo)
```

# Change species and site class to factors
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg), siteclass= as.factor(siteclass)) 
df <- droplevels(df)
```

# Convert shrub area to sqrt
```{r}
df <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3))
```

# Take out seedling/year combos with 0 growth
```{r}
df <- df %>% 
  filter(VertGrowth_cm!=0)
```

# Save df
```{r}
save(df, file = "../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

# Print final seedlings 
```{r}
sort(df$Sdlg)
```

## By species
```{r}
sort(unique(df$Sdlg[df$Species=="ABCO"]))
sort(unique(df$Sdlg[df$Species=="PIPO"]))

length(unique(df$Sdlg[df$Species=="ABCO"]))
length(unique(df$Sdlg[df$Species=="PIPO"]))
length(df$Sdlg[df$Species=="ABCO"])
length(df$Sdlg[df$Species=="PIPO"])

```

# Print all shrub patches
```{r}
sort(unique(as.factor(df$FirePatch)))
```

