---
title: "Clean 2016 seedling data"
author: "Carmen"
date: "May 20, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
require(readxl)
require(dplyr)
require(tidyverse)
require(lubridate)
```

# Load data
```{r}
sdlt <- read.csv("~/../Dropbox (Stephens Lab)/SORTIE/Shrubs_Summer16/Completed_Data_and_Photos/Master_Compiled_seedlings.csv")
```

# Basic df cleaning

## Rename 2015 height column
```{r}
sdlt <- sdlt %>% 
  dplyr::rename(Ht2015.meas2016 = Ht2015)
```

## Take out blank columns after making sure they're blank
```{r}
names(sdlt)
summary(sdlt[33:45])
sdlt <- sdlt[,c(1:32)]
names(sdlt)
```

## Take out columns related to CADE or "immediately above" data
```{r}
sdlt <- sdlt %>% select(-CADE_Gr_inner, -CADE_Gr_outer, -CADE_dia_at_Gr_cm, -ImmedAboveSpp, -ImmedAboveHt.cm)
```


## Correct typos and create variable for FirePatch
```{r}
sdlt <- sdlt %>% 
  mutate(Fire = as.character(Fire)) %>% 
  mutate(Species = as.character(Species)) %>% 
  mutate(Patch = as.character(Patch)) %>% 
  mutate(Fire = ifelse(Fire == "FRED", "FRDS", Fire)) %>% 
  mutate(Fire = ifelse(Fire == "AMCR", "AMRC", Fire)) %>% 
  mutate(Species = ifelse(Species == "pipo", "PIPO", Species)) %>% 
  mutate(Species = ifelse(Species == "PILA ", "PILA", Species)) %>% 
  mutate(Patch = ifelse(Patch == "s", "S", Patch)) %>% 
  mutate(FirePatch = paste(Fire, Patch, sep = "-"))
```

### Check
```{r}
head(sdlt %>% dplyr::select(Seedling., Fire, Species, Patch, FirePatch))
```

## Take out Seedling 17 because it doesn't have shrub data
```{r}
sdlt <- sdlt %>% 
  filter(Seedling. != "17")
```


## Make a tbl, rename object, and rename Seedling column
```{r}
df <- sdlt
df <- tbl_df(sdlt)
df <- rename(df,Sdlg = Seedling.)
```

# Reshape to include data from both 2016 visits

### Look for trees with 2 different height values
```{r}
df %>% 
  filter(Return==1) %>% 
  filter(Ht.cm>0)
```


## Create separate columns for dates of visit 
```{r}
df <- df %>% 
  mutate(return = as.character(Return))
df <- spread(df, key=Return, value = Date)
df <- df %>% 
  rename(Date1 = `0`, Date2 = `1`) 
```

## Check
```{r}
df %>% 
  select(Sdlg, FirePatch, Date1, Date2) %>% head()
```


## Move over time and personel columns

```{r}
df <- df %>% 
  mutate(Time1 = ifelse(is.na(Date2), paste(Closest_Time), "1:00")) %>% 
  mutate(Time2 = ifelse(is.na(Date1), paste(Closest_Time), "1:00")) %>% 
  mutate(personel2 = ifelse(is.na(Date1), paste(Personel), 0)) %>% 
  mutate(personel1 = ifelse(is.na(Date2), paste(Personel), 0))
           

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Closest_Time, Time1, Time2, Personel, personel1, personel2) %>% head()
```


## Replace NAs in numeric variables with 0s

### Make factors numeric where possible
```{r}
df <- df %>% 
  mutate(Light_File = as.numeric(paste(Light_File)))
```


### Use replace_na function
```{r}
df <- df %>% 
  replace_na(list(Ht.cm=0, 
                  Nearby_Easting = 0, 
                  Nearby_Northing = 0, 
                  Slope.Deg = 0, 
                  Aspect.deg = 0, 
                  BasDia.cm =0, 
                  LastYearGrth.cm=0, 
                  HTLCB=0, 
                  Dia2016=0, 
                  Dia2015=0, 
                  Ht2015=0, 
                  Light_File=0,
                  DIFN = 0))
```

## Replace NA dates with a dummy date
```{r}
df <- df %>% 
  mutate(Date1 = ifelse(is.na(Date1), "9/9/9999", paste(Date1))) %>% 
  mutate(Date2 = ifelse(is.na(Date2), "9/9/9999", paste(Date2))) 

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Time1, Time2) %>% head()
```

## Clean up time columns
```{r}
df <- df %>% 
  mutate(Time1 = ifelse(Time1 == "", "1:00", Time1)) %>% 
  mutate(Time1 = ifelse(Time1 == "1/7/1900 7:00", "1:00", Time1)) %>% 
  mutate(Time1 = ifelse(Time1 == "0.291666667", "1:00", Time1))
df$Time1  
```


## Make date columns into date format and time column into time format
```{r}
df <- df %>% 
  mutate(Date1 = mdy(Date1), Date2 = mdy(Date2)) %>% 
  mutate(Time1 = hm(Time1), Time2 = hm(Time2)) 
df <- df %>%   mutate(Time1 = hms::hms(Time1), Time2 = hms::hms(Time2)) 

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Time1, Time2) %>% head()
```


## Use the real date and time for seedlings with dummy date and time

### Date
```{r}
nrow(df)
df <- df %>% 
  group_by(Sdlg) %>% 
  mutate(Date1 = min(Date1)) %>% 
  mutate(Date2 = min(Date2)) 

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Time1, Time2) %>% head()

```

### Time1
```{r}
max1 <- df %>% 
  filter(Time1 == max(as.numeric(Time1))) %>% 
  dplyr::select(Sdlg, Time1) %>% 
  rename(Time1_final = Time1)


df <- left_join(df, max1) %>% 
  select(-Time1) %>% 
  rename(Time1 = Time1_final)

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Time1, Time2) %>% head()

nrow(df)
```

### Time2
```{r}
max2 <- df %>% 
  filter(Time2 == max(as.numeric(Time2))) %>% 
  dplyr::select(Sdlg, Time2) %>% 
  rename(Time2_final = Time2)

df <- left_join(df, max2) %>% 
  select(-Time2) %>% 
  rename(Time2 = Time2_final)

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Time1, Time2) %>% head()

df %>% 
  select(Sdlg, FirePatch, Date1, Date2, Time1, Time2) %>% tail()

nrow(df)
```

### Delete original time column
```{r}
df <- df %>% 
  select(-Closest_Time)
```


## Merge elevation, lat/long data, photo y/n, slope, etc. by taking the average or max

```{r}
df <- df %>% 
  group_by(Sdlg, FirePatch) %>% 
  mutate(Elevation = mean(Elevation))

df <- df %>% 
  group_by(Sdlg) %>% 
  mutate(Nearby_Easting = max(Nearby_Easting)) %>% 
  mutate(Nearby_Northing = max(Nearby_Northing)) 
```


```{r}
df <- df %>% 
  group_by(Sdlg) %>% 
  mutate(Slope.Deg = max(Slope.Deg)) %>% 
  mutate(Aspect.deg = max(Aspect.deg)) %>% 
  mutate(BasDia.cm = max(BasDia.cm)) %>% 
  mutate(LastYearGrth.cm = max(LastYearGrth.cm)) %>% 
  mutate(HTLCB = max(HTLCB)) %>% 
  mutate(Dia2016 = max(Dia2016)) %>% 
  mutate(Dia2015 = max(Dia2015)) %>% 
  mutate(Ht2015.meas2016 = max(Ht2015.meas2016)) %>% 
  mutate(Light_File = max(Light_File)) %>% 
  mutate(Ht.cm = max(Ht.cm)) %>% 
  mutate(DIFN = max(DIFN))
```

## Move personel info to individual columns
```{r}
df %>% 
  ungroup() %>% 
  filter(Sdlg %in% unlist(df[df$return==1,"Sdlg"]) ) %>% 
  dplyr::select(Personel, personel1, personel2) %>% 
  select(personel1) %>% distinct()

df %>% 
  ungroup() %>% 
  filter(Sdlg %in% unlist(df[df$return==1,"Sdlg"]) ) %>% 
  dplyr::select(Personel, personel1, personel2) %>% 
  select(personel2) %>% distinct()


df <- df %>% 
  mutate(personel2 = ifelse(Sdlg %in% unlist(df[df$return==1,"Sdlg"]), "VF HH", personel2)) %>% 
  select(-Personel)
```


```{r}
df %>% 
  summarize(any(Photo.=="y"))

df <- df %>% 
  mutate(Photo. = as.character(Photo.)) %>% 
  mutate(Photo = ifelse(any(Photo. %in% c("y", "yes")), "1", "0")) %>% 
  dplyr::select(-Photo.)
```


## Check for remaining columns that don't have data synced between the two visits
```{r}
df_reps <- df %>% 
  filter(Sdlg %in% unlist(df[df$return==1,"Sdlg"]) ) 
```

### Check which columns aren't the same for each seedling
```{r}
boo <- (function(x) x[sapply(x, nrow)>0])(lapply(lapply(df_reps, function(x) tapply(x, df_reps$Sdlg, function(x) x[1]!=x[2])), function(x) subset(df_reps, Sdlg %in% names(which(x)))))
print(boo)
```

## Sync notes columns
```{r}
df %>% 
  select(Sdlg, personel1, personel2, return, Note1, Note2) %>% 
  filter(!(Note1=="" & Note2=="")) %>% 
  filter(Sdlg %in% unlist(df[df$return==1,"Sdlg"]))
```

```{r}
df <- df %>% 
  mutate(Note2 = as.character(paste(Note2))) %>% 
  mutate(Note1 = as.character(paste(Note1))) %>% 
  mutate(Note2 = ifelse(Sdlg == 6, "right under log", Note2)) %>% 
  mutate(Note2 = ifelse(Sdlg == 7, "oak greater than 3m tall right above it", Note2)) %>% 
  mutate(Note1 = ifelse(Sdlg == 15, "broken top", Note1)) %>% 
  mutate(Note2 = ifelse(Sdlg == 26, " combine with other seedling 26 row", Note2)) %>% 
  mutate(Note2 = ifelse(Sdlg == 27, " combine with other seedling 27 row", Note2)) 
```


## Make two datasheet columns
```{r}
df <- df %>% rename(DataSheet1 = DataSheet) %>% mutate(DataSheet2 = ifelse(return ==1, paste(DataSheet1), "NA"))
df %>% filter(Sdlg %in% unlist(df[df$return==1,"Sdlg"])) %>% select(Fire, Sdlg, DataSheet1, DataSheet2) 
df <- df %>% 
  mutate(DataSheet2 = ifelse(return==1 & Fire =="CLVD", "Data_Sheet_CLVD_N_JUL_26_2016(1)", DataSheet2)) %>% 
  mutate(DataSheet2 = ifelse(return==1 & Fire == "FRDS", "Data_Sheet_FRDS_MID_JUL_27_2016", DataSheet2))
```


## NOW I can finally delete rows that have return == 1
```{r}
nrow(df)
df <- df %>% 
  filter(return != 1)
```

### Check for duplicate seedling names - this should be TRUE
```{r}
nrow(df) == length(unique(df$Sdlg))
```

# Save data
```{r}
save(df, file = "../../compiled_data/fire_footprints/seedlings_cleaned_2016.Rdata")
```

