---
title: "Fire Footprints Data Compile - 2016 and 2017"
author: "Carmen"
date: "September 20, 2017"
output: 
    html_document:
        toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
require(readxl)
require(dplyr)
require(tidyr)
require(ggplot2)
```

# Read in 2017 data
```{r}
setwd("~/../Dropbox (Stephens Lab)/SORTIE/FireFootprints_2017/data/compiled/")
df17 <- read_excel("Shrubs_Summer17_Compiled_Sep14.xlsx") %>% 
  rename(Ht2017.cm = Ht_cm)
```

# Read in 2016 data
```{r}
setwd("~/Shrubs-Seedlings/compiled_data/fire_footprints/")
load("master_data_2016.Rdata")
df16 <- df %>% 
  rename(Ht16_meas2016.cm = Ht.cm)
remove(df)
```

# Combine

## Make df17 Sdlg uppercase 
```{r}
df17 <- df17 %>% 
  mutate(Sdlg = toupper(Sdlg))
```

## Rename dates columns
```{r}
df16 <- df16 %>% 
  rename(Date1_2016 = Date1) %>% 
  rename(Date2_2016 = Date2) %>% 
  rename(Datasheet1_2016 = DataSheet1) %>% 
  rename(Datasheet2_2016 = DataSheet2)
```

```{r}
df17 <- df17 %>% 
  rename(Date2017 = Date) %>% 
  rename(DataSheet2017 = DataSheet)
```

## Rename notes columns
```{r}
df16 <- df16 %>% 
  rename(Note1_2016 = Note1, Note2_2016 = Note2)
```

```{r}
df17 <- df17 %>% 
  rename(Notes2017 = notes)
```

## Take out non-PIPO/ABCO in 2017 data
```{r}
df17 <- df17 %>% 
  filter(Species %in% c("PIPO", "ABCO"))
```

## Take out Seedling 268 because it doesn't have shrub data
```{r}
df17 <- df17 %>% 
  filter(Sdlg != 268)
```

# TO DO: TRY TO FIGURE OUT WHAT SEEDLING 268 WAS in 2016 DATA

## Full join
```{r}
df <- full_join(df16, df17, by = "Sdlg")
sort(names(df))
```

# Sync up repeat column names

## Check for different species
```{r}
df %>% 
  filter(Species.x != Species.y)
```

## Go with 2017 species when it's listed since one tree's species was corrected
```{r}
df <- df %>% 
  mutate(Species = ifelse(!is.na(Species.y), Species.y, Species.x)) %>% 
  dplyr::select(-Species.x, -Species.y)
```

## Sync fire and fire patch
```{r}
df <- df %>% 
  mutate(FirePatch.y = ifelse(FirePatch.y=="AMCR-NW", "AMRC-NW", FirePatch.y)) %>% 
  mutate(FirePatch.y = ifelse(FirePatch.y=="AMCR-NE", "AMRC-NE", FirePatch.y)) %>% 
  mutate(FirePatch.y = ifelse(FirePatch.y=="AMCR-2", "AMRC-2", FirePatch.y)) %>% 
  mutate(FirePatch.y = ifelse(FirePatch.y=="AMCR-3", "AMRC-3", FirePatch.y)) %>% 
  mutate(FirePatch.y = ifelse(FirePatch.y=="AMCR-A", "AMRC-A", FirePatch.y))
df %>% 
  filter(FirePatch.x != FirePatch.y) %>% 
  dplyr::select(Sdlg, FirePatch.x, FirePatch.y)
df <- df %>% 
  rename(FirePatch = FirePatch.x) %>% 
  dplyr::select(-FirePatch.y)
```

```{r}
summary(as.factor(df17$Fire))
summary(as.factor(df16$Fire))

df %>% 
  filter(Fire.x != Fire.y) %>% 
  dplyr::select(Sdlg, Fire.x, Fire.y)
df <- df %>% 
  rename(Fire = Fire.x) %>% 
  dplyr::select(-Fire.y)
```

## Rename basal diametercolumns
```{r}
df <- df %>% 
  rename(BasDia2016.cm = BasDia.cm) %>% 
  rename(BasDia2017.mm.1 = Dia_1_mm) %>% 
  rename(BasDia2017.mm.2 = Dia_2_mm ) 
```

## Reorder columns
```{r}
df <- df %>% 
  dplyr::select(Sdlg, Species, Fire, FirePatch, Ht16_meas2016.cm, Ht2017.cm, everything())
```

# Find seedlings that are missing from 2017 data AND don't have status recorded already
```{r}
df %>% 
  filter(is.na(Ht2017.cm) & is.na(Status)) %>% 
  dplyr::select(Sdlg, Status, FirePatch, DataSheet2017, Notes2017)
```

## Record why things were left out 
```{r}
df <- df %>% 
  mutate(Status = ifelse(Sdlg %in% c(19,28, 139, 137), "exclude, lone in patch", Status)) %>% 
  mutate(Status = ifelse(FirePatch == "FRDS-JB" | Sdlg %in% c(58,80,91,92,93,100,103,106,107,229,176), "not found",Status)) %>% 
  mutate(Status = ifelse(Sdlg == 7, "exclude, under oak", Status)) %>% 
  mutate(Status = ifelse(Sdlg ==132, "Actually PILA", Status))
```

## Remove seedlings that were excluded for reasons other than dead or couldn't find them
```{r}
df <- df %>% 
  filter(!Status %in% c("exclude, under oak", "exclude, lone in patch"))
summary(as.factor(df$Status))
```

# Create shrubarea variable
```{r}
df <- df %>% 
  mutate(shrubarea3 = Ht1.3 * Cov1.3)
```

# Find seedlings with no 2017 Datasheet recorded
```{r}
df %>% 
  filter(is.na(DataSheet2017)) %>% 
  select(Sdlg, Species, FirePatch, Date2017, Status)
```

# Save
```{r}
save(df, file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings.Rdata")
write.csv(df, file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings.csv", row.names = F)
```

