---
title: "DIFN_analysis_Ecosystems"
date: "Dec 11, 2020"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
require(MuMIn)
set.seed(12346)
```

# Load data
```{r}
setwd("~/../Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file = "../../compiled_data/fire_footprints/DIFN_Ecosystems.Rdata")
```

# Make sure there are no repeats 
```{r}
dfl %>% 
  group_by(DIFN.53, shrubarea3) %>% 
  filter(n()>1) %>% 
  arrange(DIFN.53)
```

# The true repeats are: 
1. Seedlings 80/82
2. Seedlings 91/92/93
3. More

# Make a new data frame with only one row per DIFN measurement
```{r}
dfl_unique <- dfl %>% 
  group_by(DIFN.53, shrubarea3) %>% 
  mutate(Ht_cm_mean = mean(Ht_cm2)) %>% 
  filter(Sdlg == min(Sdlg))
```

# Question 0: Why are there so many more rows of DIFN data than there are seedlings in the final data set?

```{r}
dfl_unique %>% 
  filter(!Species %in% c("ABCO", "PIPO") | ref ==1) %>% 
  group_by(Fire) %>% 
  count()
```

There are about 200 trees in my Chapter 2 analysis. The extra trees here are non-PIPO/ABCO, reference shrub measurements, and trees that subsequently got thrown out of the analysis.

# Question 1: How does DIFN vary with species and along the chronosequence?

## Take out repeats
```{r}
dfl_unique <- dfl %>% 
  filter(Ht.cm == min(Ht.cm))
```

```{r}
dfl_unique <- dfl_unique %>% 
  mutate(ShrubSpp03 = case_when(
    ShrubSpp03 == "ARPA" ~ "Manzanita",
    ShrubSpp03 == "CECO" ~ "Whitethorn",
    ShrubSpp03 == "CEIN" ~ "Deerbrush",
    TRUE ~ "Other")
  )
```

```{r}
ggplot(dfl_unique) + 
  geom_boxplot(aes(x = as.factor(Years), y = DIFN.53)) + 
  geom_point(aes(x = as.factor(Years), y = DIFN.53, col = ShrubSpp03)) + 
  facet_wrap(~ ShrubSpp03)+
  theme(legend.position = "none")+
  xlab("Years since fire")
ggsave("../../results/figures/DIFN/DIFN_reviewer_response.png") 
```

## Model DIFN.53 in relation to shrubarea3, shrub species, height, and years since fire

```{r}
dfl_unique <- dfl_unique %>% 
  mutate(ShrubSpp03 = as.factor(ShrubSpp03))
LM <- lm(DIFN.53 ~ shrubarea3 + ShrubSpp03 + Years + Ht.cm.meanDIFN, data = dfl)
summary(LM)
drop1(LM, test = "Chisq")
```

```{r}
plot(predictorEffect("shrubarea3", LM))
```

```{r}
shr_spp_effects <- predictorEffect("ShrubSpp03", LM)
shr_spp_effects <- as.data.frame(c(shr_spp_effects[6], shr_spp_effects[5], shr_spp_effects[17], shr_spp_effects[18]))
shr_spp_effects <- shr_spp_effects[1:3,]
shr_spp_effects <- shr_spp_effects %>% 
  mutate(ShrubSpp03 = 
           case_when(
             ShrubSpp03 == "CEIN" ~ "Deerbrush",
             ShrubSpp03 == "CECO" ~ "Whitethorn",
             ShrubSpp03 == "ARPA" ~ "Manzanita"
           ))
```

```{r}
ggplot(shr_spp_effects)+
  geom_point(aes(x = ShrubSpp03, y = fit))+
  geom_errorbar(aes(x = ShrubSpp03, ymin = lower, ymax = upper), width = 0.5)+
  ylab("Shrub species effect on diffuse non-interceptance")+
  theme_minimal()+
  xlab("")
ggsave(file = "~/../Shrubs-Seedlings/results/figures/DIFN/DIFN_shr_spp_yrs.png", height = 4)
```

# Question 2: How does DIFN vary along the chronosequence with all species combined?
```{r}
ggplot(dfl_unique) + 
  geom_smooth(aes(x = Years, y = DIFN.53), method = lm) +
  geom_point(aes(x = Years, y = DIFN.53, col = ShrubSpp03)) + 
  xlab("Years since fire")+
  theme(legend.title = element_blank())
ggsave("../../results/figures/DIFN/DIFN_years_lm.png") 
```

# Question 3: Which is a better predictor of growth, DIFN or shrub characteristics? 

Re-run the analyses in my fire foorprints code using DIFN instead of shrub cover and height

## Pines
For pines, that model is: VertGrowth_Rel ~ 
                                            Ht_cm1 + 
                                            Years + 
                                            sqrt_shrubarea3 + 
                                            heatload + 
                                            ShrubSpp03 + 
                                            BasDia2016.cm + 
                                            Year + 
                                            Ht_cm1:sqrt_shrubarea3
                                            
```{r}
load(file="../../compiled_data/fire_footprints/pine_vert_sim.Rdata")
df_pine <- df
```

### Subset data to those than have DIFN AND growth
```{r}
df_pine <- df %>% filter(!is.na(DIFN.53))
```

### Model with DIFN
```{r}
f_DIFN <- VertGrowth_Rel ~ Ht_cm1 + Years + heatload + BasDia2016.cm + Year + DIFN.53 + Ht_cm1:DIFN.53
LM_pine_DIFN.53 <- lme(f_DIFN, data = df_pine, random = ~1|Fire/FirePatch/Sdlg, method = "ML")
summary(LM_pine_DIFN.53)
```

Pseudo R2 of DIFN: 
```{r}
pseudo_R2_DIFN <- r.squaredGLMM(LM_pine_DIFN.53)
pseudo_R2_DIFN
```

### Model with shrub cover and height
```{r}
f_shrub <- VertGrowth_Rel ~ Ht_cm1 + Years + heatload + ShrubSpp03 + BasDia2016.cm + Year + sqrt_shrubarea3 + Ht_cm1:sqrt_shrubarea3
LM_pine_shrubs <- lme(f_shrub, data = df_pine, random = ~1|Fire/FirePatch/Sdlg, method = "ML")
summary(LM_pine_shrubs)
```

Pseudo R2 of shrub model:
```{r}
pseudo_R2_shrubs <- r.squaredGLMM(LM_pine_shrubs)
pseudo_R2_shrubs
pseudo_R2_DIFN
AICcmodavg::AICc(LM_pine_DIFN.53)
AICcmodavg::AICc(LM_pine_shrubs)
```


## Firs
For firs, that model is: VertGrowth_Rel ~   Years + 
                                            heatload + 
                                            incidrad +
                                            Ht_cm1 + 
                                            sqrt_shrubarea3 + 
                                            Slope.Deg. +
                                            Elevation + 
                                            Ht_cm1:sqrt_shrubarea3
                                            
```{r}
load(file="../../compiled_data/fire_footprints/fir_vert.Rdata")
df_fir <- df
```

### Subset data to those than have DIFN AND growth
```{r}
df_fir <- df_fir %>% filter(!is.na(DIFN.53))
```

### Model with DIFN
```{r}
f <- VertGrowth_Rel ~ Years + 
                      heatload + 
                      incidrad +
                      Ht_cm1 + 
                      DIFN.53 + 
                      Slope.Deg +
                      Elevation + 
                      Ht_cm1:DIFN.53
LM_fir_DIFN.53 <- lme(f, data = df_fir, random = ~1|Fire/FirePatch/Sdlg, method = "ML")
summary(LM_fir_DIFN.53)
```

Pseudo R2 of DIFN: 
```{r}
pseudo_R2_DIFN <- r.squaredGLMM(LM_fir_DIFN.53)
pseudo_R2_DIFN
```

### Model with shrub cover and height
```{r}
f <- VertGrowth_Rel ~ Years + 
                      heatload + 
                      incidrad +
                      Ht_cm1 + 
                      sqrt_shrubarea3 + 
                      Slope.Deg +
                      Elevation + 
                      Ht_cm1:sqrt_shrubarea3
LM_fir_shrubs <- lme(f, data = df_fir, random = ~1|Fire/FirePatch/Sdlg, method = "ML")
summary(LM_fir_shrubs)
```

Pseudo R2 of shrub model:
```{r}
pseudo_R2_shrubs <- r.squaredGLMM(LM_fir_shrubs)
pseudo_R2_shrubs
pseudo_R2_DIFN
AICcmodavg::AICc(LM_fir_DIFN.53)
AICcmodavg::AICc(LM_fir_shrubs)
```

