---
title: "DIFN analysis"
author: "Carmen"
date: "May 20, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
set.seed(12346)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_data_2016.Rdata")
names(df)
```

```{r}
summary(df$Fire)
summary(as.factor(df$ref))
summary(as.factor(df$ShrubSpp03))
as.character(df$ShrubSpp03)
class(df$ShrubSpp03)
```

# Add ShrubSpp03 where it's missing for reference measurements
```{r}
df <- df %>% 
  mutate(ShrubSpp03 = as.character(ShrubSpp03)) %>% 
  mutate(ShrubSpp03 = as.factor(paste(ShrubSpp03))) %>% 
  mutate(ShrubSpp03 = ifelse(Sdlg %in% c("ARPA-1"), "ARPA", as.character(ShrubSpp03))) %>% 
  mutate(ShrubSpp03 = ifelse(Sdlg %in% c("ceco-1"), "CECO", ShrubSpp03)) %>% 
  mutate(ShrubSpp03 = ifelse(Sdlg %in% c("CEIN-1", "CEIN-2"), "CEIN", ShrubSpp03)) %>% 
  mutate(ShrubSpp03 = ifelse(Sdlg %in% c("chfo-1"), "CHFO", ShrubSpp03)) %>% 
  mutate(ShrubSpp03 = ifelse(Sdlg %in% c("LIDE-1", "LIDE-2", "LIDE-3", "LIDE-4"), "LIDE", ShrubSpp03)) %>% 
  mutate(ShrubSpp03 = ifelse(ShrubSpp03 == "ARVI", "ARPA", ShrubSpp03)) 
sort(summary(as.factor(df$ShrubSpp03)))
tail(as.factor(df$Sdlg))
```

```{r}
tail(as.factor(df$Sdlg))
df %>% filter(ref == 1)
```


# Narrow down df to just obsv with DIFN and no repeats
```{r}
dfl <- df %>% 
  filter(!is.na(DIFN.53)) %>% 
  dplyr::select(Sdlg, Ht.cm, Fire, FirePatch, Elevation, Slope.Deg, Light_File, DIFN.all, ShrubSpp03, Cov1.3, Ht1.3, ShrG1, Years, LAI, DIFN.53, Ht1, Cov1, ShrubSpp01, Ht1.2, Cov1.2, ref)
```

```{r}
tail(as.factor(df$Sdlg))
dfl %>% filter(ref == 1)
```

# NEED TO: FIND DIFN MEASUREMENTS FOR THE NO-SEEDLING LOCATIONS

# Clean heights

## Add height of 0 for seedling-less measurements
```{r}
dfl <- dfl %>% 
  mutate(Ht.cm = ifelse(ref ==1, Ht.cm ==0, Ht.cm))
summary(dfl$ref)
tail(dfl$Sdlg)
dfl %>% 
  group_by(ref) %>% 
  summarize(n())
```

## Average height for DIFN with multiple seedlings
```{r}
nrow(dfl)
dfl %>% 
  arrange(FirePatch) %>% 
  arrange(Fire) %>% 
  arrange(Cov1.3) %>% 
  arrange(DIFN.all) %>% 
  group_by(Fire, FirePatch, DIFN.all, Cov1.3, Ht1.3) %>% 
  mutate(Ht.cm = mean(Ht.cm)) %>% 
  select(Fire, FirePatch, DIFN.all, Cov1.3, Ht1.3, Sdlg, Ht.cm)
dfl <- dfl %>% 
  arrange(FirePatch) %>% 
  arrange(Fire) %>% 
  arrange(Cov1.3) %>% 
  arrange(DIFN.all) %>% 
  group_by(Fire, FirePatch, DIFN.all, Cov1.3, Ht1.3) %>% 
  mutate(Ht.cm = mean(Ht.cm)) %>% 
  ungroup() %>% 
  select(-Sdlg) %>% 
  distinct()
nrow(dfl)
dfl %>% 
    filter(DIFN.all == 0.047) %>% 
   select(Fire, FirePatch, DIFN.all, Cov1.3, Ht1.3, Ht.cm)
```

## Take out spots with >50 cm tree height
```{r}
dfl <- dfl %>% 
  filter(Ht.cm<50)
```


# NEED TO: CORRECT 2016 HEIGHTS USING 2017 DATA

# NEED TO: CHECK NOTES FROM LIGHT FILES TO SEE IF I TOOK THE LIGHT MEASUREMENTS AT THE HEIGHT OF ONE OR ANOTHER OF THE REPEATED SEEDLINGS


# Group shrub species
```{r}
nrow(dfl)
summary(as.factor(dfl$ShrubSpp03))
dfl <- dfl %>%   
  mutate(ShrubSpp03 = ifelse(!ShrubSpp03 %in% c("CECO", "ARPA", "CHSE", "CHFO", "CEIN", "LIDE"), "Other", as.character(ShrubSpp03))) %>%   mutate(ShrubSpp01 = ifelse(!ShrubSpp01 %in% c("CECO", "ARPA", "LIDE", "CHSE", "CHFO", "CEIN"), "Other", as.character(ShrubSpp01)))
summary(as.factor(dfl$ShrubSpp01))
summary(as.factor(dfl$ShrubSpp03))
dfl <- dfl %>% 
  mutate(ShrubSpp03 = ifelse(ShrubSpp03 %in% c("CEIN", "CECO"), "Ceanothos", ShrubSpp03))
```


# Look at DIFN in comparison to shrub cover, area, and species
```{r}
ggplot(dfl)+
  geom_point(aes(x = Cov1.3, y = DIFN.53))
ggplot(dfl)+
  geom_boxplot(aes(x = ShrubSpp03, y = DIFN.53))
ggplot(dfl)+
  geom_point(aes(x = Years, y = DIFN.53))

hist(dfl$DIFN.53)
hist(log(dfl$DIFN.53))
```

No reason to transform DIFN

# Take out 2 DIFN measurements because they don't make sense
```{r}
dfl <- dfl %>% 
  filter(DIFN.all != 0)
nrow(dfl)
```

# Model DIFN in relation to shrubarea3, shrub species, years since fire

```{r}
dfl$ShrubSpp03 <- as.factor(dfl$ShrubSpp03)

LM <- lm(DIFN.all ~ Years + ShrubSpp03 + Ht1.3 + Cov1.3, data = dfl)
summary(LM)
drop1(LM, test = "Chisq")
```

```{r}
plot(predictorEffect("ShrubSpp03", LM))
plot(predictorEffect("Ht1.3", LM))
plot(predictorEffect("Cov1.3", LM))
plot(predictorEffect("Years", LM))
```

```{r}
plot(LM)
dfl$E <- resid(LM)
ggplot(dfl)+
  geom_boxplot(aes(Fire, E))
```

# To test for the effect of Years since fire when shrubarea and species have been accounted for, model the residuals of that model against Years
```{r}
LM_noyears <- lm(DIFN.all ~ ShrubSpp03 + Ht1.3 + Cov1.3, data = dfl)
dfl$E_noyears <- resid(LM_noyears)
ggplot(dfl, aes(x = as.factor(Years), y = E_noyears))+
  geom_boxplot()
```

```{r}
dfl$Years <- as.numeric(paste(dfl$Years))
summary(dfl$Years)
LM_resid <- lm(E_noyears ~ Years, data = dfl)
summary(LM_resid)
drop1(LM_resid, test= "F")
plot(predictorEffect("Years",LM_resid))
```
