---
title: "DIFN analysis"
author: "Carmen"
date: "May 20, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
set.seed(12346)
```

# Goal

The point of this script is to see if DIFN decreases with time since fire independent of cover and height 

# Load data
```{r}
setwd("~/../Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/shrub_master_data_2016_DIFN.Rdata")
load(file = "../../compiled_data/fire_footprints/seedlings_DIFN_cleaned_2016.Rdata")
```

# Add ShrubSpp03 where it's missing for reference measurements
```{r}
shr_by_sdlg <- shr_by_sdlg %>% 
  mutate(ShrubSpp03 = as.character(ShrubSpp03)) %>% 
  mutate(ShrubSpp03 = as.factor(paste(ShrubSpp03))) %>% 
  mutate(ShrubSpp03 = ifelse(Sdlg %in% c("ARPA-1"), "ARPA", as.character(ShrubSpp03))) %>% 
  mutate(ShrubSpp03 = ifelse(ShrubSpp03 == "ARVI", "ARPA", ShrubSpp03)) 
```

# Add seedling heights
```{r}
df <- left_join(shr_by_sdlg, df)
```

# Throw out seedling without DIFN data
```{r}
df <- df %>% filter(!is.na(DIFN.53))
```

# Look at seedlings with no fire assignment
```{r}
summary(as.factor(df$Fire))
```

```{r}
df <- df %>% 
  mutate(Fire = ifelse(Sdlg == "125F", "FRDS", Fire))
```

```{r}
df %>% 
  select(Sdlg, Date1, Fire, DataSheet1, DataSheet2) %>% 
  head()
```

# Correct data sheet and fire for the rows that are missing it

```{r}
df <- df %>% 
  mutate(DataSheet1 = ifelse(Sdlg %in% c("LIDE-4", "LIDE-3", "LIDE-2", "LIDE-1"),"Data_Sheet_AMRC_NW_JUN_22_1", paste(DataSheet1))) %>% 
  mutate(Fire = ifelse(Sdlg %in% c("LIDE-4", "LIDE-3", "LIDE-2", "LIDE-1"),"AMRC", paste(Fire))) %>% 
  mutate(Fire = ifelse(Sdlg=="CECO-1","AMRC", paste(Fire))) %>% 
  mutate(Fire = ifelse(Sdlg=="ARPA-1","CLVD", paste(Fire))) %>% 
  mutate(DataSheet1 = ifelse(Sdlg=="ARPA-1", "Data_Sheet_CLVD_SE_JUN24_2", paste(DataSheet1))) %>% 
  mutate(DataSheet1 = ifelse(Sdlg %in% c("CEIN-1", "CEIN-2"), "Data_Sheet_PLKN_W_JUL_1", paste(DataSheet1))) %>% 
    mutate(Fire = ifelse(Sdlg %in% c("CEIN-1", "CEIN-2"), "PLKN", paste(Fire))) 
```

# Add years since fire
```{r}
df <- droplevels(df)
summary(as.factor(df$Fire))
df <- df %>% 
  mutate(Years = ifelse(Fire == "AMRC", 8, 0)) %>% 
  mutate(Years = ifelse(Fire == "CLVD", 24, Years)) %>% 
  mutate(Years = ifelse(Fire == "FRDS", 12, Years)) %>% 
  mutate(Years = ifelse(Fire == "PLKN", 43, Years)) %>% 
  mutate(Years = ifelse(Fire == "STAR", 14, Years)) %>% 
  mutate(Years = ifelse(Fire == "WRTS", 35, Years))
summary(as.factor(df$Years))
```

# Narrow down df to just obsv with DIFN and no repeats
```{r}
dfl <- df %>% 
  filter(!is.na(DIFN.53)) %>% 
  dplyr::select(Sdlg, Ht.cm, Fire, FirePatch, Elevation, Slope.Deg, Light_File,ShrubSpp03, Cov1.3, Ht1.3, Years, DIFN.53, ref, Species)
```

# Add shrubarea3 column
```{r}
dfl <- dfl %>% 
  mutate(shrubarea3 = Cov1.3*Ht1.3)
```

# Clean heights

## Load 2017 data with up-to-date heights
```{r}
load("../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
df17 <- df
remove(df)
```

### filter it to only 2016 height data
```{r}
df17 <- df17 %>% 
  filter(Year=="2015") %>% 
  dplyr::select(Sdlg, Ht_cm2)
```

### Join with DIFN data
```{r}
dfl <- left_join(dfl, df17, by = "Sdlg")
```

## Add height of 0 for seedling-less measurements
```{r}
dfl <- dfl %>%
  mutate(Ht.cm = ifelse(ref ==1, Ht.cm ==0, Ht.cm))
```

## Average height for DIFN with multiple seedlings
```{r}
dfl <- dfl %>% 
  group_by(Fire, FirePatch, DIFN.53, Cov1.3, Ht1.3) %>% 
  mutate(Ht.cm.meanDIFN = mean(Ht.cm)) 
```

# Group shrub species into "other" category
```{r}
nrow(dfl)
summary(as.factor(dfl$ShrubSpp03))
dfl <- dfl %>%   
  mutate(ShrubSpp03 = ifelse(!ShrubSpp03 %in% c("CECO", "ARPA", "CHSE", "CHFO", "CEIN", "LIDE"), "Other", as.character(ShrubSpp03))) 

summary(as.factor(dfl$ShrubSpp03))
```

# Look at DIFN.53 in comparison to shrub cover, area, and species
```{r}
ggplot(dfl)+
  geom_point(aes(x = Cov1.3, y = DIFN.53))
ggplot(dfl)+
  geom_boxplot(aes(x = ShrubSpp03, y = DIFN.53))
ggplot(dfl)+
  geom_point(aes(x = Years, y = DIFN.53))
ggplot(dfl)+
  geom_boxplot(aes(x = as.factor(Years), y = DIFN.53))

hist(dfl$DIFN.53)
hist(log(dfl$DIFN.53))
```

No reason to transform DIFN.53

# Take out 2 DIFN.53 measurements = 0 because they don't make sense
```{r}
dfl <- dfl %>% 
  filter(DIFN.53 != 0 & !is.na(shrubarea3))
nrow(dfl)
```

# To test for the effect of Years since fire when shrubarea and species have been accounted for, model the residuals of that model against Years
```{r}
LM_noyears <- lm(DIFN.53 ~ ShrubSpp03 + Ht1.3 + Cov1.3, data = dfl)
dfl$E_noyears <- resid(LM_noyears)
ggplot(dfl, aes(x = as.factor(Years), y = E_noyears))+
  geom_boxplot()
```

```{r}
dfl$Years <- as.numeric(paste(dfl$Years))
summary(dfl$Years)
LM_resid <- lm(E_noyears ~ Years, data = dfl)
summary(LM_resid)
drop1(LM_resid, test= "F")
plot(predictorEffect("Years",LM_resid))
```


# To test for the effect of Years since fire when only shrubarea has been accounted for, model the residuals of that model against Years
```{r}
LM_noyears <- lm(DIFN.53 ~ shrubarea3, data = dfl)
dfl$E_noyears <- resid(LM_noyears)
ggplot(dfl, aes(x = as.factor(Years), y = E_noyears))+
  geom_boxplot()
```

```{r}
dfl$Years <- as.numeric(paste(dfl$Years))
summary(dfl$Years)
LM_resid <- lm(E_noyears ~ Years, data = dfl)
summary(LM_resid)
drop1(LM_resid, test= "F")
plot(predictorEffect("Years",LM_resid))
```

# Save compiled data
```{r}
save(dfl, file = "../../compiled_data/fire_footprints/DIFN_Ecosystems.Rdata")
```

# DO THIS FOR CERTAIN ANALYSES
## Take out spots with >30 cm tree height
```{r}
dfl <- dfl %>% 
  filter(Ht.cm<50)
```

