---
title: "Explore_height_pole_error"
author: "Carmen"
date: "August 16, 2019"
output: html_document
---


```{r, include = F}
require(tidyverse)
require(ggplot2)
```

```{r}
load("../../compiled_data/fire_footprints/df_vert_wide.Rdata")
```


### Count seedlings without 2017 height or 2015 height measured in 2016
```{r}
df %>% 
  filter(is.na(Ht2017.cm.fall) & is.na(Ht2015.meas2016)) %>% 
  group_by(Species) %>% 
  summarize(n())
```

### For those seedlings, find average 2016 growing season apical growth relative to 2015 growth 
```{r}
find_apical_growth <- df %>% 
  filter(!is.na(Ht2015.meas2016)) %>% 
  mutate(apical_growth_2016 = 
           Ht16_meas2016.cm - LastYearGrth.cm - Ht2015.meas2016) %>% 
  dplyr::select(Sdlg, Species, Ht16_meas2016.cm, LastYearGrth.cm, Ht2015.meas2016, apical_growth_2016) %>% 
  mutate(apical_ratio = apical_growth_2016/LastYearGrth.cm) %>% 
  distinct()
find_apical_growth
```

## For trees with measurements in both 2016 and 2017, see how apical growth compares to 2016 growth measured in 2017
```{r}
df_both_yrs <- df %>% 
  filter(!is.na(Growth_16) & !is.na(Ht2015.meas2016)) %>% 
  dplyr::select(Sdlg, Ht2016.cm_spring_from2017, Ht16_meas2016.cm, Ht2016.cm_spring_from2017, LastYearGrth.cm, Ht2015.meas2016, Ht2015.cm_spring_from2017, Growth_16, Date1_2016, Ht2017.cm.fall) %>% 
  distinct()
df_both_yrs <- left_join(df_both_yrs, find_apical_growth) %>% 
  mutate(Ht16_minus_apical = Ht16_meas2016.cm-apical_growth_2016) %>% 
  dplyr::select(Sdlg, Ht2016.cm_spring_from2017, Ht16_minus_apical, everything())
df_both_yrs
```

```{r}
ggplot(df_both_yrs, aes(x = Ht2015.meas2016, y = Ht2015.cm_spring_from2017, col = as.factor(Date1_2016)))+
  geom_point()+
  geom_abline(intercept = 0,slope =1)
```

```{r}
nrow(df_both_yrs)
maybe_bad <- df_both_yrs %>% 
  filter(abs(Ht2015.meas2016 - Ht2015.cm_spring_from2017)>3) %>% 
  mutate(diff_height15 = Ht2015.meas2016 - Ht2015.cm_spring_from2017)
maybe_bad
nrow(maybe_bad)
hist(maybe_bad$diff_height15, breaks = 20)
maybe_bad %>% 
  mutate(neg = ifelse(diff_height15>0, "pos", "neg")) %>% 
  group_by(neg) %>% 
  summarize(n())
```

The fact that about equal numbers of seedlings are wrong in either direction suggests that there wasn't a systemic problem with height pole measurements

