---
title: "fir vertical growth - fire footprints"
author: "Carmen"
date: "November 13, 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
```


```{r set.seed(333)}
knitr::opts_chunk$set(cache = T)
set.seed(333)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

# Filter to just white firs
```{r}
df %>% 
  group_by(Fire,Species) %>% 
  summarize(n())
```

```{r}
df <- df %>% 
  filter(Species == "ABCO" & !df$Fire %in% c("PLKN","FRDS"))
```

## Change species and Sdlg to factor
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg)) 
summary(df$Species)
summary(df$Sdlg)
```

# Count
```{r}
length(unique(df$Sdlg))
nrow(df)
```

# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```

## Take out rows with Inf log(VertGrowth) because growth = 0 cm
```{r}
df <- df %>% 
  filter(is.finite(df$VertGrowth_Rel))
```

# Apply VSURF

## Make df of just the predictor variables I want to include
```{r}
xdf <- df %>% 
  select(Elevation,
         Slope.Deg,
         Aspect.rad.fold.NESW,
         Years,
         Ht_cm1,
         shrubarea1,
         shrubarea2, 
         shrubarea3,
         Cov1, 
         Cov1.2, 
         Cov1.3, 
         Ht1, 
         Ht1.2, 
         Ht1.3,
         ShrubSpp01, 
         ShrubSpp02, 
         ShrubSpp03, 
         BasDia2016.cm, 
         heatload,
         incidrad,
         Year,
         Years) 
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
```

# STOP HERE IF YOU DON'T WANT THINGS TO TAKE FOREVER

## Apply VSURF to vertical growth
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Run 100 random forests on the ones that come up to see whether to delete heatload or incidrad
```{r}
forests100 <- function(vars){
  forest <- randomForest(x = xdf %>% select(heatload, Years, shrubarea2, Ht1.3,incidrad, Ht1.2, BasDia2016.cm), y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
  forest$importance
  imp <- as.data.frame(forest$importance) 
  imp$var <- row.names(imp)
  imp <- imp %>% filter(var %in% vars)
  imp %>% filter(`%IncMSE` == max(imp$`%IncMSE`)) %>% select(var)
}
forests100(c("heatload", "incirad"))
reps <- replicate(100, forests10(c("heatload", "incirad")))
summary(as.factor(unlist(reps)))
```

### Delete incidrad based on the results of the above randomforests
```{r}
xdf <- xdf %>% 
  select(-incidrad)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Delete Aspect since it wasn't in the initial comparison of heatload and incidrad, and then run vsurf again

```{r}
xdf <- xdf %>% 
  select(-Aspect.rad.fold.NESW)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Run a randomforest using those variables
```{r}
forest <- randomForest(x = xdf %>% select(Years, heatload, Ht_cm, shrubarea3, Ht1.3, Cov1.3, Elevation), y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
varImpPlot(forest)
```

Check the correlations of Cov3, Ht1.3, and shrubarea3 to see if one needs to be deleted

```{r}
ggplot(df, aes(x = shrubarea3, y = Cov1.3))+
   geom_point()+
   geom_smooth(method = "loess")
```

```{r}
ggplot(df, aes(x = shrubarea3, y = Ht1.3))+
   geom_point()+
   geom_smooth(method = "loess")
```

Looks like I should delete height, since shrubarea3 has been in the last several models
```{r}
xdf <- xdf %>% 
  select(-Ht1.3)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

# RESUME HERE

## Look at final random forest
```{r}
forest <- randomForest(
  x = xdf %>% 
    select(Years, shrubarea3, heatload, BasDia2016.cm), 
  y = df$VertGrowth_Rel, 
  importance = TRUE, 
  scale = TRUE)
varImpPlot(forest)
```


# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3)) %>% 
  select(Fire, Sdlg, VertGrowth_Rel, Fire, heatload, Years, shrubarea3, sqrt_shrubarea3, log_shrubarea3, Cov1, Elevation, Year, BasDia2016.cm) %>% 
  mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```


## See whether model with shrubarea3 or sqrt_shrubarea3 is better
```{r}
AIC(lme(VertGrowth_Rel ~ heatload+ Years+ shrubarea3*BasDia2016.cm,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  heatload+ Years+ sqrt_shrubarea3*BasDia2016.cm,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
)

AIC(lme(VertGrowth_Rel ~  heatload+ Years+ log_shrubarea3*BasDia2016.cm,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
)
```

Sqrt transformed wins by a tiny amount. Since the difference is so small and pines use untransformed, and beacuse it's better to err on the side of untransformed, I'll use untransformed.

## See whether model with or without interaction is better
```{r}
AIC(lme(VertGrowth_Rel ~ Years + shrubarea3*BasDia2016.cm + heatload, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~ Years + shrubarea3 + BasDia2016.cm + heatload, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

```

With interaction is better.

## Linear model
```{r}
LM <-  lme(VertGrowth_Rel ~Years + shrubarea3*BasDia2016.cm + heatload, 
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

### Test dropping shrubarea3 and diameter individually

#### Diameter
```{r}
LM_noht <- lme(VertGrowth_Rel ~ Years + shrubarea3 + heatload, 
               data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM, LM_noht)
```

#### Shrubarea
```{r}
LM_noshrub <- lme(VertGrowth_Rel ~ Years + BasDia2016.cm + heatload, 
                  data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM, LM_noshrub)
```

Note: anova here is performing likelihood ratio tests

### Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

```{r}
eff_years <- predictorEffect("Years", LM)
plot(eff_years)

eff_shrubarea <- predictorEffect("shrubarea3", LM)
plot(eff_shrubarea)

eff_ht <- predictorEffect("BasDia2016.cm", LM)
plot(eff_ht)

plot(predictorEffect("heatload", LM))
```

# Re-create effects plots using predictorEffects() and ggplot

## Create df for effects and confidence intervals using mean +/- 1 SD
```{r}
quartiles <- c(mean(df$BasDia2016.cm)-sd(df$BasDia2016.cm), mean(df$BasDia2016.cm)+sd(df$BasDia2016.cm))
quartiles
quartiles <- c(mean(LMdf$BasDia2016.cm)-sd(LMdf$BasDia2016.cm), mean(LMdf$BasDia2016.cm)+sd(LMdf$BasDia2016.cm))
quartiles

eff_shrubarea <- predictorEffect("shrubarea3", LM, xlevels = list(BasDia2016.cm = quartiles))
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(BasDia2016.cm = BasDia2016.cm*sd(df$BasDia2016.cm)+mean(df$BasDia2016.cm))
  
effects_df <- effects_df %>% 
  mutate(shrubarea3 = shrubarea3*sd(df$shrubarea3)+mean(df$shrubarea3)) 
```

### test
```{r}
summary(as.factor(effects_df$BasDia2016.cm))
hist(effects_df$shrubarea3)
hist(df$shrubarea3)
```

## Make shrubarea in m instead of cm
```{r}
effects_df <- effects_df %>% 
  mutate(shrubarea3 = shrubarea3/10000)
```

## Round the values of dia
```{r}
effects_df <- effects_df %>% 
  mutate(BasDia2016.cm = round(BasDia2016.cm, digits = 1))
```

## Plot effects of shrubarea
```{r}
ggplot(effects_df)+
  geom_line(aes(x = shrubarea3, y = fit, group = as.factor(BasDia2016.cm)))+
  geom_ribbon(aes(x = shrubarea3, group = as.factor(BasDia2016.cm), fill = as.factor(BasDia2016.cm), ymin = lower, ymax = upper), alpha = .5)+
  theme_bw()+
  xlab(bquote('Shrub competition ' (m^2)))+
  ylab("Relative growth rate")+
  ggtitle("Fir growth in relation to shrub environment within 3 m")+
  geom_rug(data = df, aes(x = shrubarea3/10000, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(-0.1,0.7))+
  labs(fill = "Juvenile tree diameter (cm)")+
  theme(legend.position = c(0.8, 0.15))
ggsave(file = "../../results/figures/FireFootprints/FirVertRel_int.png", width = 6, height = 6, dpi = 400)
```

## Years
```{r}
eff_years <- predictorEffect("Years", LM)
effects_df <- as.data.frame(eff_years)
effects_df <- effects_df %>% 
  mutate(Years = Years*sd(df$Years)+mean(df$Years)) 

ggplot(effects_df)+
  geom_line(aes(x = Years, y = fit))+
  geom_ribbon(aes(x = Years, ymin = lower, ymax = upper), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab("Years since fire")+
  ylab("Predicted relative growth (cm/cm)")+
  ggtitle("Pine growth in relation to time since fire")+
  geom_rug(data = df, aes(x = Years, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(-0.02,0.28))
ggsave(file = "../../results/figures/FireFootprints/PineVertRel_years.png", width = 6, height = 5, dpi = 400)
```

# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea3))
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea3))
```

## Linear model
```{r}
LM <-  lme(VertGrowth_Rel ~ heatload+ Years+ shrubarea3+ BasDia2016.cm,
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

#### Re-run model with non-normalized data
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>% 
  select(Fire, Sdlg, VertGrowth_Rel, shrubarea3, heatload, Year, BasDia2016.cm, Years) %>% 
  mutate(shrubarea3_m = shrubarea3/10000)
LMdf <- droplevels(LMdf)
```

```{r}
LM <-  lme(VertGrowth_Rel ~ heatload+ Years+ shrubarea3_m+ BasDia2016.cm,data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

### Look at individual effects
```{r}
eff_shrubarea <- predictorEffect("shrubarea3_m", LM)
plot(eff_shrubarea)

eff_heat <- predictorEffect("heatload", LM)
plot(eff_heat)

eff_years <- predictorEffect("Years", LM)
plot(eff_years)

plot(predictorEffect("BasDia2016.cm", LM))
```

# Re-create effects plot using predictorEffects() and ggplot
```{r}
effects_df <- as.data.frame(eff_shrubarea)
ggplot(effects_df)+
  geom_line(aes(x = 100*shrubarea3_m/400, y = fit))+
  geom_ribbon(aes(x = 100*shrubarea3_m/400, ymin = lower, ymax = upper), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab(bquote('Shrub cross-sectional area ' (m^2)))+
  ylab("Relative growth rate")+
  ggtitle("Fir growth in relation to shrub environment within 3 m")+
  geom_rug(data = df, aes(x = 100*(shrubarea3/10000)/400, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
    ylim(c(0,0.28))
ggsave(file = "../../results/figures/FireFootprints/FirVertRel.png", width = 6, height = 5, dpi = 400)
```


## Years
```{r}
eff_years <- predictorEffect("Years", LM)
effects_df <- as.data.frame(eff_years)
ggplot(effects_df)+
  geom_line(aes(x = Years, y = fit))+
  geom_ribbon(aes(x = Years, ymin = lower, ymax = upper), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab("Years since fire")+
  ylab("Relative growth rate")+
  ggtitle("Fir growth in relation to time since fire")+
  geom_rug(data = df, aes(x = Years, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(-0.1,0.28))
ggsave(file = "../../results/figures/FireFootprints/FirVertRel_years.png", width = 6, height = 5, dpi = 400)
```


# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea3))
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea3))
```


# Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm, E)
plot(df$Years, E)

ggplot(df, aes(x = shrubarea3, y = E))+
  geom_point()+
  geom_smooth()
```


## Save final data
```{r}
save(df, file ="~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/fir_vert.Rdata")
```
