---
title: "fir vertical growth - fire footprints"
author: "Carmen"
date: "November 13, 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
list.of.packages <- c("nlme", "randomForest", "tree", "knitr", "tidyverse", "ggplot2", "VSURF", "effects", "sjstats", "car")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(effects)
require(sjstats)
require(MuMIn)
require(car)
```


# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

# Load VSURF results
```{r}
load("../../results/VSURF_abco_vars.Rdata")
vars
```

# Before filtering to just ABCO, save min and max shrub cover for growth estiamtes at the end. Use min and max that have BOTH SPECIES
```{r}
df_summary <- df %>% 
  group_by(Species) %>% 
  mutate(shrubarea3 = sqrt(shrubarea3)) %>% 
  summarize(min= min(shrubarea3), max = max(shrubarea3), median = median(shrubarea3))

min <- df_summary %>% 
  filter(min == max(min)) %>% 
  dplyr::select(min) %>% 
  unlist()
min

max <- df_summary %>% 
  filter(max == min(max)) %>% 
  dplyr::select(max) %>% 
  unlist()
max

median <- df %>% 
  mutate(shrubarea3 = sqrt(shrubarea3)) %>%
  summarize(median = median(shrubarea3)) %>% 
  unlist
median


summary <- summary(sqrt(df$shrubarea3))
summary
quart1 <- summary[2]
quart3 <- summary[5]
mean <- mean(sqrt(df$shrubarea3))
```

# Filter to just white firs
```{r}
df %>% 
  group_by(Fire,Species) %>% 
  summarize(n())
```

```{r}
df <- df %>% 
  filter(Species == "ABCO")
```

## Change species and Sdlg to factor
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg))
```

# Convert shrub area to square root
```{r}
df <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3))
```

# Count totals and by fire
```{r}
nrow(df)
```

```{r}
df %>% 
  group_by(Sdlg, Fire) %>%
  dplyr::select(Sdlg, Fire) %>%
  distinct() %>% 
  ungroup() %>% 
  group_by(Fire) %>% 
  summarize(n = n()) %>% 
  mutate(total = sum(n))
```

# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```


# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM

### Normalize function
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

### Filter and normalize
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3)) %>% 
  dplyr::select(Fire, FirePatch, Sdlg, heatload, Years, BasDia2016.cm, shrubarea3, sqrt_shrubarea3, log_shrubarea3, Elevation, Ht_cm1, Slope.Deg, Year, Aspect.rad.fold.NESW, incidrad) %>% 
  mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```

# LME model using VSURF var results

## Convert vars to formula structure
```{r}
vars_one <- paste(vars, collapse = " + ")
vars_one
f <- formula(paste("VertGrowth_Rel ~ ",vars_one))
f
```

## Run lme model
```{r}
LM <- lme(f, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

## Compare to model with FirePatch as random effect
```{r}
LM <- lme(f, data = LMdf, random = ~ 1| Fire/FirePatch/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

# Check collinearity 

<!-- ## Make collinearity plot using Zuur code -->
<!-- ```{r} -->
<!-- source("~/../Documents/HighstatLibV10.R.txt") -->
<!-- z <- cbind(df$VertGrowth_Rel, df$Years, df$heatload, df$Aspect.rad.fold.NESW, df$incidrad, df$Ht_cm1, df$BasDia2016.cm, df$Slope.Deg, df$sqrt_shrubarea3, df$Elevation, df$Year) -->
<!-- colnames(z) <- c("Growth", "Years", "heatload", "Aspect", "incidrad", "Ht_cm1", "BasDia", "Slope", "shrubs", "Elevation", "Year") -->
<!-- pairs(z, lower.panel = panel.smooth2, upper.panel = panel.cor, diag.panel = panel.hist) -->
<!-- ``` -->

## use vif from car pkg because in pines it was the same as the Zuur code
```{r}
vif(LM)
```

# See whether model with or without interaction is better
```{r}
vars_int <- paste(vars_one, "+ Ht_cm1:sqrt_shrubarea3")
f_int <- formula(paste("VertGrowth_Rel ~ ",vars_int))
LM_int <- lme(f_int, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
AICcmodavg::AICc(LM)
AICcmodavg::AICc(LM_int)
if(
  AICcmodavg::AICc(LM) > AICcmodavg::AICc(LM_int)
) print(paste("WITH INTERACTION WINS BY", round(AICcmodavg::AICc(LM)-AICcmodavg::AICc(LM_int), digits = 2), "AIC")) else print(paste("WITHOUT INTERACTION WINS BY", round(AICcmodavg::AICc(LM_int)-AICcmodavg::AICc(LM), digits = 2)))
```

No difference, so I'll leave out the interaction

## Look at summary and drop1 results
```{r}
summary(LM)
```

```{r}
drop <- drop1(LM, test = "Chisq")
drop
drop$var <- row.names(drop)
drop <- tbl_df(drop)
drop %>% 
  arrange( `Pr(>Chi)`) %>% 
  mutate( `Pr(>Chi)` = paste( `Pr(>Chi)`))
```

### Look at individual effects

```{r}
plot(predictorEffect("sqrt_shrubarea3", LM))
```

## Calculate Nakagawa's pseudo R2
```{r}
r.squaredGLMM(LM)
r.squaredGLMM(LM_int)
```

# Re-create effects plots using predictorEffects() and ggplot

## Create df for effects 
```{r}
eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM)
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(sqrt_shrubarea3 = sqrt_shrubarea3*sd(sqrt(df$shrubarea3))+mean(sqrt(df$shrubarea3)) )
```

### test
```{r}
hist(effects_df$sqrt_shrubarea3)
hist(sqrt(df$shrubarea3))
```

# Summarize range of shrub cover and height
```{r}
summary(df$Cov1.3)
summary(df$Ht1.3)
summary(df$shrubarea3)
```

# Summarize cover as percent
```{r}
summary(df$Cov1.3/(300*4))
```

# Find the shrubarea values associated with each quartile, and the mean growth for each of those
```{r}
fit_quart1 <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart1)),]$fit
fit_quart1
fit_quart3 <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart3)),]$fit
fit_median <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - median)),]$fit

fit_min <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - min)),]$fit
fit_max <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - max)),]$fit
fit_mean <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - mean)),]$fit

print(paste("fit_quart1 = ", exp(fit_quart1)))
print(paste("fit_quart3 = ", exp(fit_quart3)))
print(paste("fit_median = ", exp(fit_median)))
print(paste("fit_mean = ", exp(fit_mean)))

print(paste("fit_min = ", exp(fit_min)))
print(paste("fit_max = ", exp(fit_max)))


print(paste("fit_quart1_log = ", fit_quart1))
print(paste("fit_quart3_log = ", fit_quart3))
print(paste("fit_median_log = ", fit_median))
print(paste("fit_min_log = ", fit_min))
print(paste("fit_max_log = ", fit_max))
print(paste("fit_mean_log = ", fit_mean))

gr_minshrub_abco <- exp(fit_min)
gr_maxshrub_abco <- exp(fit_max)
gr_medshrub_abco <- exp(fit_median)
```

## Save them
```{r}
save(gr_minshrub_abco, file = "../../results/data/FireFootprints/gr_minshrub_abco.Rdata")
save(gr_maxshrub_abco, file = "../../results/data/FireFootprints/gr_maxshrub_abco.Rdata")
save(gr_medshrub_abco, file = "../../results/data/FireFootprints/gr_medshrub_abco.Rdata")
```

## Plot effects of shrubarea

### Load colors from PIPO graph
```{r}
load("../../results/figures/FireFootprints/plot_min.Rdata")
load("../../results/figures/FireFootprints/plot_max.Rdata")
```

```{r}
ggplot(effects_df)+
  geom_line(aes(x = sqrt_shrubarea3, y = exp(fit)), col = ("#899DA4"))+
  geom_ribbon(aes(x = sqrt_shrubarea3, ymin = exp(lower), ymax = exp(upper)), alpha = .5, fill = "#8da0cb")+
  theme_bw()+
  xlab(bquote(
    (sqrt(`cover (cm)`%*%`height (cm)`))            
    ))+
  ylab("Relative growth rate")+
  geom_rug(data = df, aes(x = sqrt(shrubarea3), y = exp(VertGrowth_Rel)), alpha = .7, position = "jitter", sides = "b")+
  ylim(c(plot_min, plot_max))+
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    text = element_text(size = 11))
ggsave(file = "../../results/figures/FireFootprints/FirVertRel.png", width = 3, height = 3, dpi = 400)
```



```{r}
ggplot(effects_df)+
  geom_line(aes(x = sqrt_shrubarea3, y = exp(fit)), col = ("#899DA4"))+
  geom_ribbon(aes(x = sqrt_shrubarea3, ymin = exp(lower), ymax = exp(upper)), alpha = .5, fill = "#899DA4")+
  theme_bw()+
  xlab(bquote(
    (sqrt(`cover (cm)`%*%`height (cm)`))            
    ))+
  ylab("Relative growth rate")+
  geom_rug(data = df, aes(x = sqrt(shrubarea3), y = exp(VertGrowth_Rel)), alpha = .7, position = "jitter", sides = "b")+
  ylim(c(plot_min, plot_max))+
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    text = element_text(size = 16))
ggsave(file = "../../results/figures/FireFootprints/FirVertRel.png", width = 3, height = 3, dpi = 400)
```


# Save LM
```{r}
save(LM, file = "../../results/data/FireFootprints/LM_abco.Rdata")
```

# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea3))
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea3))
```

# Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm1, E)
plot(df$Years, E)

ggplot(df, aes(x = shrubarea3, y = E))+
  geom_point()+
  geom_smooth()
```

# Count how many trees have estimated 2015 heights
```{r}
df %>% 
  filter(Year==2015) %>% 
  filter(is.na(Ht2017.cm.fall) & is.na(Ht2015.meas2016)) %>% 
  nrow()
```


## Save final data
```{r}
save(df, file ="../../compiled_data/fire_footprints/fir_vert.Rdata")
```

# Save LM with non-normalized data
```{r}
LMdf_nonnorm <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3)) %>% 
  dplyr::select(Fire, Sdlg, heatload, Years, BasDia2016.cm, shrubarea3, sqrt_shrubarea3, log_shrubarea3, Elevation, Ht_cm1, Slope.Deg, Year, Aspect.rad.fold.NESW, incidrad, VertGrowth_Rel) 
LMdf_nonnorm <- droplevels(LMdf_nonnorm)
```

```{r}
LM_abco_nonnorm <- lme(f, data = LMdf_nonnorm, random = ~ 1| Fire/Sdlg, method = "ML")
```

```{r}
save(LM_abco_nonnorm, file = "../../results/coefficients/LM_abco_nonnorm.Rdata")
```


