---
title: "Fir diameter growth analysis, all years - fire footprints"
author: "Carmen"
date: "November 14, 2017"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(ggplot2)
library(dplyr)
require(nlme)
library(kableExtra)
library(knitr)
library(lme4)
```

# Load data
```{r}
load(file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_dia.Rdata")
```

# Filter out pines and fires with very few firs
```{r}
df_dia <- df_dia[df_dia$Species%in%c("ABCO","ABMA", "PSME") & !df_dia$Fire %in% c("PLKN","FRDS"),]
df_dia$Species <- as.factor(df_dia$Species)
df_dia %>% 
  group_by(Fire) %>% 
  summarise(n())
```

# STATS

## Normalize variables that still need it
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

```{r}
df_dia <- df_dia %>% 
  mutate(Cov1.2_norm = normalize(Cov1.2)) %>% 
  mutate(Cov1.3_norm = normalize(Cov1.3)) %>% 
  mutate(Ht_norm = normalize(Ht_cm)) %>% 
  mutate(Ht_log_norm = normalize(log(Ht_cm)))
```

# Re-analyse and plot using best model that sticks with 0-2 m (AIC model selection deleted)
```{r}
f <- DiaGrowth.mm_rel ~  Cov1.2_norm + Ht_norm 
f_noshrubs <- DiaGrowth.mm_rel ~  Ht_norm
LM2 <- lme(f,data = df_dia, random = ~ 1|Fire, method = "ML")
LM2_noshrubs <- lme(f_noshrubs,data = df_dia, random = ~ 1|Fire, method = "ML")
anova(LM2, LM2_noshrubs)
```


# Look for patterns in residuals of mixed effects model

## In relation to fitted values
```{r}
E <- resid(LM2, type="normalized")
df_dia$resid <- E
plot(LM2, col = df_dia$Fire)
df_dia[fitted.values(LM2)<0,]$resid
plot(df_dia$Ht_norm, E)
```

# What about a variance structure in relation to height?
```{r}
f <- DiaGrowth.mm_rel ~  Cov1.2_norm + Ht_norm 
f_null <- DiaGrowth.mm_rel ~ Ht_norm 
LM3 <- lme(f,
           data = df_dia, 
           random = ~ 1| Fire,  
           method = "REML", 
           weights = varExp(form=~Ht_norm))
LM2 <- lme(f,
           data = df_dia, 
           random = ~ 1| Fire,  
           method = "REML")
anova(LM3, LM2)
```

Looks like a variance structure does not need to be included. Go back above and take it out of the AIC selection.

## Compare to null
# ```{r}
# LM3_null <- lme(f_null,
#            data = df_dia, 
#            random = ~ 1| Fire,  
#            method = "REML", 
#            weights = varExp(form=~Ht_norm))
# anova(LM3, LM3_null)
# summary(LM3)
# ```

Note: the best model from the AIC table is with cover from 0-3  m, and that shows a significant likelihood ratio test. For simplicity, in the poster I include cover from 0-2 m and do not discuss significance

## resids for gls model
```{r}
E <- resid(LM3, type="normalized")
df_dia$resid <- E
plot(LM3, col = df_dia$Fire)
df_dia[fitted.values(LM3)<0,]$resid
plot(df_dia$Ht_norm, E)
```

# Plot once more with the final best model

```{r}
f <- DiaGrowth.mm_rel ~  Cov1.2_norm + Ht_norm 
LM2 <- lme(f,
           data = df_dia, 
           random = ~ 1| Fire,  
           method = "ML")
```

## Make fake data
```{r}
MyData2 <-  rbind(
  expand.grid( 
    Fire = unique(df_dia$Fire),
    Ht_cm = mean(df_dia$Ht_cm),
    Cov1.2 = seq(min(df_dia$Cov1.2), max(df_dia$Cov1.2), length = 20)
    ) 
)
```

## "normalize" MyData
```{r}
MyData2$Ht_norm <- (MyData2$Ht_cm - mean(df_dia$Ht_cm))/sd(df_dia$Ht_cm)
MyData2$Cov1.2_norm <- (MyData2$Cov1.2 - mean(df_dia$Cov1.2))/sd(df_dia$Cov1.2)
```

# Try lmer

# THIS IS THE BEST ONE:
```{r}
LM4 <- lmer(DiaGrowth.mm_rel ~ Cov1.2_norm + Ht_norm + (1|Fire), data=df_dia) 
LM4_null <- lmer(DiaGrowth.mm_rel ~ Ht_norm + (1|Fire), data=df_dia)
anova(LM4, LM4_null)
```

Same results as using nlme

```{r}
MyData2$pred_simple <- predict(LM4, newdata = MyData2)
```



### Summary plot
```{r}
MyData2_summary_simple <- MyData2 %>% 
  group_by(Cov1.2) %>%
  mutate(pred_simple = mean(pred_simple))

figure_summary <- ggplot(MyData2_summary_simple)+
  labs(title = "Fir seedling diameter growth")+
  ylab("relative growth (mm/mm)")+
  xlab("shrub cover 0-2 m from seedling")+
  geom_line(aes(y = pred_simple, x =Cov1.2), col = "#1b3f4c")+
  geom_point(data=df_dia, aes(x = Cov1.2, y = DiaGrowth.mm_rel), col = "#1b3f4c")+
  theme_bw()+
  theme(plot.background = element_rect(fill = '#e9ebe8', colour = '#e9ebe8'),
        panel.background = element_rect(fill="#e9ebe8"),
        text = element_text(size=20),
        axis.title = element_text(colour = "#1b3f4c"),
        plot.title = element_text(colour = "#1b3f4c")
        )+
  scale_y_continuous(limits = c(0,.72)) 
figure_summary
remove(MyData2_summary_simple)
```


# Save figure
```{r}
setwd("C:/Users/Carmen/Documents/Shrubs-Seedlings/results/figures")
pdf("fir_dia.pdf", width = 5.5, height = 7.25, pointsize = 30,useDingbats = F)
figure_summary
dev.off()
```

# Calculate pseudo r squared
```{r}
meanGrth <- mean(df_dia$DiaGrowth.mm_rel)
sse <- (df_dia$DiaGrowth.mm_rel - predict(LM4))^2
sst <- (df_dia$DiaGrowth.mm_rel - meanGrth)^2
R2 <- 1 - (sum(sse)/sum(sst))
R2
```

# Save final data
```{r}
save(df_dia, file = "~/Shrubs-Seedlings/compiled_data/fire_footprints/fir_dia.Rdata")
```

