---
title: "fir vertical growth - fire footprints"
author: "Carmen"
date: "November 13, 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
```


```{r set.seed(333)}
knitr::opts_chunk$set(cache = T)
set.seed(333)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

# Filter to just white firs
```{r}
df %>% 
  group_by(Fire,Species) %>% 
  summarize(n())
```

```{r}
df <- df %>% 
  filter(Species == "ABCO" & !df$Fire %in% c("PLKN","FRDS"))
```

## Change species and Sdlg to factor
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg)) 
summary(df$Species)
summary(df$Sdlg)
```

# Count
```{r}
length(unique(df$Sdlg))
nrow(df)
```

# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```

## Take out rows with Inf log(VertGrowth) because growth = 0 cm
```{r}
df <- df %>% 
  filter(is.finite(df$VertGrowth_Rel))
```

# Apply VSURF

## Make df of just the predictor variables I want to include
```{r}
xdf <- df %>% 
  select(Elevation,
         Slope.Deg,
         heatload,
         incidrad,
         Year,
         Years,
         BasDia2016.cm, 
         Ht_cm1,
         shrubarea1,
         shrubarea2, 
         shrubarea3,
         ShrubSpp01, 
         ShrubSpp02, 
         ShrubSpp03
       ) 
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
```

# STOP HERE IF YOU DON'T WANT THINGS TO TAKE FOREVER

## Apply VSURF to vertical growth
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Run 100 random forests on the ones that come up to see whether to delete heatload or incidrad
```{r}
forests100 <- function(vars){
  forest <- randomForest(x = xdf %>% select(heatload, Years, shrubarea2, incidrad, BasDia2016.cm, Ht_cm1, Elevation, Year), 
                         y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
  forest$importance
  imp <- as.data.frame(forest$importance) 
  imp$var <- row.names(imp)
  imp <- imp %>% filter(var %in% vars)
  imp %>% filter(`%IncMSE` == max(imp$`%IncMSE`)) %>% select(var)
}
forests100(c("heatload", "incirad"))
reps <- replicate(100, forests100(c("heatload", "incirad")))
summary(as.factor(unlist(reps)))
```

The variable with the higher number is better. 

### Delete incidrad based on the results of the above randomforests
```{r}
xdf <- xdf %>% 
  select(-incidrad)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Repeat for shrubarea2 and shrubarea3

## Run 100 random forests on the ones that come up to see whether to delete SHRUBAREA2 or SHRUBAREA3
```{r}
forests100 <- function(vars){
  forest <- randomForest(x = xdf %>% select(heatload, Years, shrubarea2, shrubarea3, BasDia2016.cm, Ht_cm1, Elevation, Year), y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
  forest$importance
  imp <- as.data.frame(forest$importance) 
  imp$var <- row.names(imp)
  imp <- imp %>% filter(var %in% vars)
  imp %>% filter(`%IncMSE` == max(imp$`%IncMSE`)) %>% select(var)
}
forests100(c("shrubarea2", "shrubarea3"))
reps <- replicate(100, forests100(c("shrubarea2", "shrubarea3")))
summary(as.factor(unlist(reps)))
```

This shows that shrubarea2 is better, so take out shrubarea3 and run vsurf yet again


### Delete SHRUBAREA3 and SHRUBAREA1

```{r}
xdf <- xdf %>% 
  select(-shrubarea1, -shrubarea3)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

# RESUME HERE

## Look at final random forest
```{r}
forest <- randomForest(
  x = xdf %>% 
    select(heatload, Years, BasDia2016.cm, shrubarea2, Elevation, Ht_cm1, Slope.Deg, Year), 
  y = df$VertGrowth_Rel, 
  importance = TRUE, 
  scale = TRUE)
varImpPlot(forest)
```

# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea2 = sqrt(shrubarea2)) %>%
  mutate(log_shrubarea2 = log(shrubarea2)) %>% 
  select(Fire, Sdlg, heatload, Years, BasDia2016.cm, shrubarea2, sqrt_shrubarea2, log_shrubarea2, Elevation, Ht_cm1, Slope.Deg, Year) %>% 
  mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```

## See whether model with shrubarea3 or sqrt_shrubarea3 is better
```{r}
AIC(lme(VertGrowth_Rel ~ heatload + Years + BasDia2016.cm + shrubarea2 + Elevation + Ht_cm1 + Slope.Deg + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  heatload + Years + BasDia2016.cm + sqrt_shrubarea2 + Elevation + Ht_cm1 + Slope.Deg + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  heatload + Years + BasDia2016.cm + log_shrubarea2 + Elevation + Ht_cm1 + Slope.Deg + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
)
```

Untransformed wins.

## See whether model with or without interaction is better
```{r}
without <- lme(VertGrowth_Rel ~ 
           heatload + Years + BasDia2016.cm + shrubarea2 + Elevation + Ht_cm1 + Slope.Deg + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")

with <- lme(VertGrowth_Rel ~  
           heatload + Years + BasDia2016.cm + shrubarea2*Ht_cm1 + Elevation + Slope.Deg + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(with, without)
```

No difference, so I'll leave out the interaction

## Linear model
```{r}
LM <-  lme(VertGrowth_Rel ~ heatload + Years + BasDia2016.cm + shrubarea2 + Elevation + Ht_cm1 + Slope.Deg + Year,
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
drop1(LM, test = "Chisq")
```

### Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

```{r}
eff_years <- predictorEffect("Years", LM)
plot(eff_years)

eff_shrubarea <- predictorEffect("shrubarea2", LM)
plot(eff_shrubarea)

plot(predictorEffect("heatload", LM))
```

# Re-create effects plots using predictorEffects() and ggplot

## Create df for effects 
```{r}
eff_shrubarea <- predictorEffect("shrubarea2", LM)
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(shrubarea2 = shrubarea2*sd(df$shrubarea2)+mean(df$shrubarea2)) 
```

### test
```{r}
hist(effects_df$shrubarea2)
hist(df$shrubarea2)
```

## Make shrubarea in m instead of cm
```{r}
effects_df <- effects_df %>% 
  mutate(shrubarea2 = shrubarea2/10000)
```

## Plot effects of shrubarea
```{r}
ggplot(effects_df)+
  geom_line(aes(x = shrubarea2, y = exp(fit)))+
  geom_ribbon(aes(x = shrubarea2, ymin = exp(lower), ymax = exp(upper)), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab(bquote('Shrub competition ' (m^2)))+
  ylab("Relative growth rate")+
  ggtitle("Fir growth in relation to shrub environment within 2 m")+
  geom_rug(data = df, aes(x = shrubarea2/10000, y = exp(VertGrowth_Rel)), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(0.04,0.215))
ggsave(file = "../../results/figures/FireFootprints/FirVertRel_int.png", width = 6, height = 6, dpi = 400)
```

# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea2))
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea2))
```

# Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea2, E)
plot(df$Ht_cm1, E)
plot(df$Years, E)

ggplot(df, aes(x = shrubarea2, y = E))+
  geom_point()+
  geom_smooth()
```


## Save final data
```{r}
save(df, file ="~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/fir_vert.Rdata")
```
