---
title: "Heat Load"
author: "Carmen"
date: "March 28, 2019"
output: 
    html_document:
        toc: TRUE
---

Methods from: McCune, B., Keon, D., 2002. Equations for potential annual direct incident radiation and heat load. J. Veg. Sci. 13, 603-606. https://doi.org/10.1111/j.1654-1103.2002.tb02087.x

```{r, include = F}
require(tidyverse)
```

# Load data
```{r}
load(file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings1.Rdata")
```

# Add elevation and slope where they're missing - look in original data sheets
```{r}
df %>% 
  filter(is.na(Elevation)) %>% 
  arrange(Sdlg)
```

## Create function for adding elevation data to an individual seedling
```{r}
elv <- function(sdlg, elev){
  df <<- df %>% 
    mutate(Elevation = ifelse(Sdlg == sdlg, elev, Elevation))
}
```

## Apply the function using elevation data from the QGIS project file `C:\Users\Carmen\Dropbox (Stephens Lab)\SORTIE\FireFootprints_2017\GIS_maps\Seedling_locations`
```{r}
elv(26, 1642)
elv(27, 1642)
elv(159, 1643)
elv(161, 1676)
elv(162, 1679)
elv(163, 1679)
elv(164, 1679)
elv(165, 1692)
elv(166, 1703)
elv(167, 1703)
elv(168, 1701)
elv(169, 1706)
elv(170, 1644)
elv(171, 1644)
elv(172, 1644)
elv(174, 1645)
elv(176, 1689)
elv(177, 1702)
elv(178, 1697)
elv(179, 1646)
elv(180, 1646)
elv(181, 1642)
elv(220, 1652)
elv(221, 1651)
elv(223, 1663)
elv(224, 1663)
elv(226, 1668)
elv(250, 1685)
elv(251, 1684)
elv(252, 1684)
elv(253, 1681)
elv(254, 1681)
elv(256, 1680)
elv(257, 1680)
elv(258, 1682)
elv(259, 1683)
elv(260, 1683)
elv(262, 1681)
elv(263, 1682)
elv(264, 1684)
elv(265, 1684)
```

## For 222 and 225, and 255, I can't find an avenza point location. So I'll average the elevations of other seedlings with numbers just above and below them

### 222&225 elevation estimation
```{r}
elv_est <- df %>% 
  filter(Sdlg %in% c(220, 221, 223, 224, 226)) %>% 
  dplyr::select(FirePatch, Elevation) %>% 
  distinct() %>% 
  ungroup() %>% 
  summarize(mean(Elevation))
elv_est <- unlist(elv_est)
elv_est
elv(222, elv_est)
elv(225, elv_est)
```

### 255 elevation estimate
```{r}
elv_est <- df %>% 
  filter(Sdlg %in% c(254, 256)) %>% 
  dplyr::select(FirePatch, Elevation) %>% 
  distinct() %>% 
  ungroup() %>% 
  summarize(mean(Elevation))
elv_est <- unlist(elv_est)
elv_est
elv(255, elv_est)
```

### Test
```{r}
df %>% filter(Sdlg %in% c(222, 225, 255)) %>% dplyr::select(Sdlg, FirePatch, Elevation) %>% distinct()
```

## Test for completion
```{r}
df %>% 
  filter(is.na(Elevation)) %>% 
  arrange(Sdlg) %>% 
  dplyr::select(Sdlg, FirePatch, Elevation, Slope.Deg, Aspect.deg) %>% 
  distinct()
```

# Repeat for Aspect
```{r}
df %>% 
  filter(is.na(Aspect.deg)) %>% 
  arrange(Sdlg) %>% 
  dplyr::select(Sdlg, FirePatch, Elevation, Slope.Deg, Aspect.deg) %>% 
  distinct() 
```

## Create function for adding aspect data to an individual seedling
```{r}
asp <- function(sdlg, aspect){
  df <<- df %>% 
    mutate(Aspect.deg = ifelse(Sdlg == sdlg, aspect, Aspect.deg))
}
```

## Apply the function to seedlings with missing aspect data
```{r}
asp(204, 257)
asp(10, 47)
asp(11, 38)
asp(12, 17)
asp(13, 17)
asp(15, 26)
asp(8, 41)
asp(9, 34)
```

## Check
```{r}
df %>% 
  filter(is.na(Aspect.deg)) %>% 
  arrange(Sdlg) %>% 
  dplyr::select(Sdlg, FirePatch, Elevation, Slope.Deg, Aspect.deg) %>% 
  distinct() 
```

# Repeat for Slope
```{r}
df %>% 
  filter(is.na(Slope.Deg)) %>% 
  arrange(Sdlg) %>% 
  dplyr::select(Sdlg, FirePatch, Elevation, Slope.Deg, Aspect.deg) %>% 
  distinct()
```

## Create function for adding slope data to an individual seedling
```{r}
slp <- function(sdlg, slope){
  df <<- df %>% 
    mutate(Slope.Deg = ifelse(Sdlg == sdlg, slope, Slope.Deg))
}
```

## Apply the function to seedlings with missing slope data
```{r}
slp(176, 13)
slp(10, 16)
slp(11, 16)
slp(12, 15)
slp(13, 15)
slp(15, 15)
slp(8, 15)
slp(9, 15)
```

## Check
```{r}
df %>% 
  filter(is.na(Slope.Deg)) %>% 
  arrange(Sdlg) %>% 
  dplyr::select(Sdlg, FirePatch, Elevation, Slope.Deg, Aspect.deg) %>% 
  distinct()
```

# Add latitude

## Compile latitude .csv files exported from QGIS
```{r}
setwd("~/../Dropbox (Stephens Lab)/SORTIE/FireFootprints_2017/Seedling_locations_exported_QGIS/")
files <- list.files()
table_all <- data.frame()
for(i in 1:length(files)){
  table <- read.csv(files[i])
  if("X" %in% names(table)){
    table <- table %>% dplyr::select(X, Y, Name)  
  } else{
    table <- table %>% dplyr::select(Title, Latitude, Longitude) %>% 
      rename(Name = Title, Y = Latitude, X = Longitude)
  }
  table
  
  table$file_lat <- files[i]
  if(nrow(table_all) == 0){
    table_all <- table
  } else(
    table_all <- full_join(table_all, table)
  )
}
table_all %>% arrange(Name)
nrow(table_all)
```

## Join this list with df

### First join rows with only one seedling per row

#### Find those rows
```{r}
table_n1 <- table_all %>% 
  filter(nchar(Name)<4)
```

#### Take out repeats
```{r}
table_n1 %>% 
  group_by(Name) %>% 
  filter(n()>1) %>% 
  arrange(Name)
```

```{r}
table_n1 <- table_n1 %>% 
  filter(!(Name==26 & file_lat == "Freds_JB_A.csv")) %>% 
  filter(!(Name %in% c(76, 77, 80, 81, 83, 88, 89, 94, 97) & file_lat == "AMR - Aerial - 2.csv"))
```


#### Join
```{r}
df <- left_join(df, table_n1, by = c("Sdlg" = "Name"))
```

### Then join rows with more than one seedling per row
```{r}
table_n <- anti_join(table_all, table_n1) %>% arrange(Name)
table_n
```

#### Delete redundant rows
```{r}
dups <- table_n %>% 
  group_by(Name) %>% 
  mutate(n = n()) %>% 
  filter(n>1)
dups
table_n <- table_n %>%
  filter(!(Name %in% dups$Name & file_lat == "AMR - Aerial - 2 - iphone.csv"))
table_n <- table_n %>%
  filter(!(Name %in% dups$Name & file_lat == "AMR - Aerial - NW.csv"))
```

#### Create a function
```{r}
long <- function(sdlgs, name){
  row <- table_n %>% 
    filter(Name == name)
  for(i in 1:length(sdlgs)){
     row$Sdlg = sdlgs[i]
     df[df$Sdlg == sdlgs[i], "X"] <<- row$X
     df[df$Sdlg == sdlgs[i], "Y"] <<- row$Y
     df[df$Sdlg == sdlgs[i], "file_lat"] <<- row$file_lat
  }
  table_n <<- table_n %>% 
    filter(Name != name)
}
```

#### Apply it
```{r}
long(sdlgs = c("116", 119, "120", "121"), name = "116,119-121")
long(sdlgs = c("117", "118"), name = "117-118")
long(sdlgs = c("131", "132"), name = "131-132")
long(c("133","134"), "133-134")
long(c("140","141","142"), "140-142")
long(c("144", "145"), "144-145")
long(c("147","148"), "147-148")
long(c("149","150","151"), "149-151")
long(c("153","154"), "153-154")
long(c("156", 157, 158, 159, 160), "156-160")
long(c("162","163","164"), "162-164")
long(c("170", 171, 172, 173), "170-173")
long(c("179", 180), "179-180, Possible Abcos")
long(c("182","189"), "182 & 189")
long(c("183","184"), "183-184")
long(c("166","167"), "166-7")
long(c("185", 186), "185-186")
long(c("187","188"), "187, 188")
long(c("191","193"), "191 & 193")
long(c("199", "200", "201"), "199-201")
long(c("20", 21, 22, 23, 24, 25), "20-25")
long(c("286", 287, 288, 289), "286-9")
long(c("291", "292"), "291-2")
long(c("12", 13, 14), "12,13,14")
long(c("15"), "15 Remarked")
long(c("277", 278, 276), "277,278,276")
long(c("1",2,3,4),"Seedlings 1,2,3,4")
long(c("227", 228), "227 and 228")
long(c("237", 238), "237 and 238")
long(c("239", 240), "239 and 240")
long(c("242", 243, 244), ("242 - 244"))
long(c("100"), "100 Ceco3")
long(c("202", 203), "202 and 203")
long(c("223", 224), "223 and 224")
long(c("73", 74, 75), "73-75")
long(c("78", 79, 105), "78, 79, and 105")
long(c("85", 86, 87), "85-87")
long(c("90"), "90 Ceco2 ")
long(c("91", 92, 93), "91-93")
long(c("95", 96), "95, 96")
long(c("98","99"), "98-99")
long(c("101", "102"), "101-2")
long(c("251", "252"), "251 and 252")
long(c("256", "257"), "256 and 257")
long(c("264", "265"), "264 and 265")
long(c("268"), "268?")
long(c("35", 36,37,38,39,40), "35 - 40")
long(c("103"), "103 - Return ")
long(c("205", 206), "205,206")
long(c("207", 208), "207,208")
long(c("212", 213), "212,213")
long(c("78", 79), "78, 79")

```

### Check
```{r}
df %>% filter(Sdlg == 227) %>% dplyr::select(Sdlg, X, Y, file_lat)
table_n[1:4,]
```

## Add longitude by hand for the seedlings I couldn't export

### Find all seedlings still needing longitude
```{r}
df %>% 
  filter(is.na(X)) %>% 
  dplyr::select(Sdlg, FirePatch, X, Species) %>% 
  distinct()
```

### Check which patches they're concentrated in
```{r}
df %>% 
  filter(is.na(X)) %>% 
  dplyr::select(FirePatch, X, Species) %>% 
  group_by(FirePatch) %>% 
  summarize(n())
```

### Function for assigning longitude
```{r}
lati <- function(sdlg, longitude, latitude){
  df <<- df %>% 
    mutate(X = ifelse(Sdlg == sdlg, longitude, X)) %>% 
    mutate(Y = ifelse(Sdlg == sdlg, latitude, Y)) 
}
```

```{r}
lati(46, -120.648140, 39.234092)
lati(48, -120.648140,39.234092)
lati(250, -120.635357,39.219178)
lati(253, -120.636404, 39.219158)
lati(254, -120.636404,39.219158)
lati(262, -120.636132,39.219085)
lati(263, -120.635996,39.219089)
```

### For the ones without points, create a function to average the longitudes of the nearby points
```{r}
df <- df %>% ungroup()
latave <- function(sdlg, nearbys){
  meanX <- df %>% filter(Sdlg %in% nearbys) %>% summarise(mean(X))
  meanX <- unlist(meanX)
  meanY <- df %>% filter(Sdlg %in% nearbys) %>% summarise(mean(Y))
  meanY <- unlist(meanY)
  df <<- df %>% 
    mutate(X = ifelse(Sdlg == sdlg, meanX, X))
  df <<- df %>% 
    mutate(Y = ifelse(Sdlg == sdlg, meanY, Y))
}
```

#### Check
```{r}
latave(106, c(105, 108))
latave(107, c(105, 108))
df %>% filter(Sdlg %in% c(105:108)) %>% dplyr::select(Sdlg, X, Y) %>% distinct() %>% arrange(Sdlg) %>% print.data.frame 
```

#### Apply
```{r}
latave(46, c(44,60))
latave(104, c(100, 105))
latave(211, c(212, 210))
latave(55, c(54, 56))
latave(52, c(51, 54))
latave(42, c(41, 43))
latave(216, c(215, 220))
latave(217, c(215, 220))
latave(218, c(215, 220))
latave(219, c(215, 220))
latave(222, c(220, 223))
latave(225, c(224, 226))
latave(230, c(228, 237))
latave(229, c(228, 230))
latave(231, c(228, 237))
latave(232, c(228, 237))
latave(233, c(228, 237))
latave(234, c(228, 237))
latave(235, c(228, 237))
latave(236, c(228, 237))
latave(246, c(245, 244))
latave(247, c(245, 244))
latave(248, c(245, 244))
latave(249, c(245, 244))
latave(255, c(254, 256))
latave(283, c(281, 283))
latave("125w", c(123, 124))
latave(82, c(81, 83))
latave(84, c(83, 85))
latave(106, c(105))
latave(107, c(105))
latave(282, c(281, 283))
latave(135, c(125, 121))
latave(27, c(26))
```

#### Check
```{r}
df %>% 
  filter(is.na(Y)) %>% 
  dplyr::select(Sdlg, FirePatch, X, Y, Species) %>% 
  distinct()
```

# Calculate heat load based on Eqn 3 of McCune, B., Keon, D., 2002. Equations for potential annual direct incident radiation and heat load. J. Veg. Sci. 13, 603-606. https://doi.org/10.1111/j.1654-1103.2002.tb02087.x

Equation: 0.339 + 0.808\*cos(LATITUDE)*cos(SLOPE) - 0.196\*sin(LATITUDE)\*sin(SLOPE) - 0.482\*cos(ASPECT)\*sin(SLOPE)

## Create rows for the sample data from McCune and Keon to check my values against
```{r}
df <- df %>% 
  add_row(Sdlg = "test4030N") %>% 
  add_row(Sdlg = "test4030S") %>% 
  add_row(Sdlg = "test40flat")
df <- df %>% 
  mutate(Aspect.deg = ifelse(Sdlg =="test4030N", 0, Aspect.deg)) %>% 
  mutate(Aspect.deg = ifelse(Sdlg =="test4030S", 180, Aspect.deg)) %>% 
  mutate(Aspect.deg = ifelse(Sdlg =="test40flat", 180, Aspect.deg)) %>% 
  mutate(Slope.Deg = ifelse(Sdlg =="test4030N", 30, Slope.Deg)) %>% 
  mutate(Slope.Deg = ifelse(Sdlg =="test4030S", 30, Slope.Deg)) %>% 
  mutate(Slope.Deg = ifelse(Sdlg =="test40flat", 0, Slope.Deg)) %>% 
  mutate(Y = ifelse(Sdlg %in% c("test4030N", "test4030S", "test40flat"), 40, Y))
  
```


## Convert variables to radians
```{r}
df$Aspect.rad <- (df$Aspect.deg*pi)/180
df$Slope.rad <-  (df$Slope.Deg*pi)/180
df$Lat.rad <-  (df$Y*pi)/180
```

## "Fold" Aspect - once N/S for radiation, and once NE/SW for heat load
```{r}
df$Aspect.rad.fold.NS <- pi - abs(df$Aspect.rad - pi)
df$Aspect.rad.fold.NESW <- abs(pi - abs(df$Aspect.rad - (5*pi)/4))
```

## Apply the equation for radiation
```{r}
df$incidrad <-  0.339 + 0.808*cos(df$Lat.rad)*cos(df$Slope.rad) - 0.196*sin(df$Lat.rad)*sin(df$Slope.rad) - 0.482*cos(df$Aspect.rad.fold.NS)*sin(df$Slope.rad)
```

## APply the equation for heat load
```{r}
df$heatload <-  0.339 + 0.808*cos(df$Lat.rad)*cos(df$Slope.rad) - 0.196*sin(df$Lat.rad)*sin(df$Slope.rad) - 0.482*cos(df$Aspect.rad.fold.NESW)*sin(df$Slope.rad)
```

## Compare test values to McCune and Keon
```{r}
df %>% 
  filter(Sdlg %in% c("test4030N", "test4030S", "test40flat")) %>% 
  dplyr::select(Sdlg, incidrad, heatload)
```

They look good!

## Delete test values
```{r}
df <- df %>% 
  filter(!Sdlg %in% c("test4030N", "test4030S", "test40flat"))
```

## Check for NAs
```{r}
df %>% 
  filter(is.na(heatload)) %>% 
  dplyr::select(Aspect.deg, Aspect.rad, Aspect.rad.fold.NS, Aspect.rad.fold.NESW, Slope.Deg, Slope.rad, Y, Lat.rad, heatload, incidrad)
```


# Save
```{r}
save(df, file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_seedlings2.Rdata")
```

Now I can use incident radiation and heat load in my random forest models