---
title: "Pine volume growth analysis -fire footprints"
author: "Carmen"
date: "November 19, 2017"
output: 
    html_document:
        toc: TRUE
---


```{r, include = F}
library(ggplot2)
library(dplyr)
require(nlme)
library(kableExtra)
library(knitr)
library(lme4)
library(tidyr)
```


# Load final data frame
```{r}
load("~/Shrubs-Seedlings/compiled_data/fire_footprints/pinus_dia.Rdata")
load("~/Shrubs-Seedlings/compiled_data/fire_footprints/pinus_vert.Rdata")
df_vert <- df
remove(df)
```

# Define genus color
```{r}
color <- "#a02313"
```


# Delete rows for vert growth before 2016
```{r}
df_vert <- df_vert %>% 
  filter(Year != "2015")
```

# Restructure vert data so there's one row per seedling
```{r}
df_vert <- df_vert %>% dplyr::select(-VertGrowth_Rel)
df_vert <- spread(df_vert, key = Year, value = VertGrowth) %>% 
  rename(VertGr_2016 = `2016`) %>% 
  rename(VertGr_2017 = `2017`)
```


# Filter to only seedlings with workable data for both diameter and vertical growth
```{r}
df_vert <- df_vert %>% dplyr::select(Sdlg, VertGr_2016, VertGr_2017)
df <- inner_join(df_dia, df_vert)
```


# Clean up extreme and negative values

```{r}
nrow(df)
hist(df$DiaGrowth_rel, breaks = 10)
hist(df$DiaGrowth_rel)
nrow(df)
```


## Take out outliers on both ends, defined as greater than two SD from the mean
```{r}
nrow(df)
hist(df$DiaGrowth, breaks = 20)
mean_minSD <- mean(df$DiaGrowth) - 2*sd(df$DiaGrowth)
mean_minSD
mean_plusSD <- mean(df$DiaGrowth) + 2*sd(df$DiaGrowth)
mean_plusSD
df <- df %>% 
  filter(!(DiaGrowth < mean_minSD |
             DiaGrowth > mean_plusSD))
hist(df$DiaGrowth)
nrow(df)
```



## Adjust final diameters so everything's above 0
```{r}
hist(df$DiaGrowth)
min(df$BasDia2017.mm.ave)
hist(df$BasDia2017.mm.ave)
df <- df %>% 
  mutate(BasDia2017.mm.ave = (BasDia2017.mm.ave - min(df$DiaGrowth)))
min(df$BasDia2017.mm.ave)
hist(df$BasDia2017.mm.ave)
df <- df %>% 
  mutate(DiaGrowth = (DiaGrowth - min(DiaGrowth)))
```


# Calculate relative growth rate

## Calculate pre volume
```{r}
#fix NA vert growth for math purposes
df <- df %>%
  mutate(VertGr_2017 = ifelse(is.na(VertGr_2017), 0, VertGr_2017)) %>% 
   mutate(VertGr_2016 = ifelse(is.na(VertGr_2016), 0, VertGr_2016))

df <- df %>% 
  mutate(vol_pre_cm2 = ((BasDia2016.mm/10)^2) * (Ht_cm - VertGr_2017  - VertGr_2016) * (pi/12))
hist(df$vol_pre_cm2)
```

## Calculate post volume
```{r}
df <- df %>% 
  mutate(vol_post_cm2 = ((BasDia2017.mm.ave/10)^2) * Ht_cm * (pi/12))
hist(df$vol_post_cm2)
```



## Calculate RGR
```{r}
df <- df %>% 
  mutate(RGR = (log(vol_post_cm2)- log(vol_pre_cm2))/2)
hist(df$RGR)
```

### Take a look

```{r}
ggplot(df, aes(x = Ht_cm, y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
ggplot(df, aes(x = log(Ht_cm), y = RGR))+
  geom_point()+
  stat_smooth(method = "gam", formula = y~s(x))
```

# Normalize

```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

```{r}
df <- df %>% 
  mutate(ht_norm = normalize(Ht_cm)) %>% 
  mutate(shrubarea2_sqrt_norm = normalize(sqrt(shrubarea2)) ) %>% 
  mutate(Ht_log_norm = normalize(log(Ht_cm)))
```


```{r}
ggplot(df)+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_pre_cm2)-log(vol_post_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)
ggplot(df, aes(x = log(vol_pre_cm2), y = RGR))+
  geom_point()
```

# Mixed effects with island effect
```{r}
MM1 <- lme(RGR ~Cov1.2_norm + Ht_log_norm, random =~1|Fire, data = df, method = "ML")
MM1_null <- lme(RGR ~ Ht_log_norm, random =~1|Fire, method = "ML",data = df)
anova(MM1, MM1_null)
summary(MM1)
```

### Residuals
```{r}
E <- resid(MM1, type = "normalized")
pred <- predict(MM1)
plot(MM1)
plot(df$shrubarea2_sqrt_norm, E)
plot(df$Ht_log_norm, E)
```

# Plot

## Make fake data
```{r}
MyData <-  rbind(
  expand.grid( 
    Fire = unique(df$Fire),
    Ht_cm = exp(mean(log(df$Ht_cm))),
    Cov1.2 = seq(min(df$Cov1.2), max(df$Cov1.2), length = 20)
    ) 
)
```

## "normalize" MyData
```{r}
MyData$Ht_norm <- (MyData$Ht_cm - mean(df$Ht_cm))/sd(df$Ht_cm)
MyData$Ht_log_norm <- (log(MyData$Ht_cm) - mean(log(df$Ht_cm)))/sd(log(df$Ht_cm))
MyData$Cov1.2_norm <- (MyData$Cov1.2 - mean(df$Cov1.2))/sd(df$Cov1.2)
```


```{r}
MyData$pred <- predict(MM1, newdata = MyData)
```


### Summary plot
```{r}
MyData_summary_simple <- MyData %>% 
  group_by(Cov1.2) %>%
  mutate(pred_simple = mean(pred)) %>% 
  ungroup()

brown <- "#5b3d07"
figure_summary <- ggplot(MyData_summary_simple)+
  labs(title = "Pine seedling volumetric growth")+
  ylab("relative growth rate (RGR)")+
  xlab("shrub cover 0-2 m from seedling")+
  geom_line(aes(y = pred_simple, x =Cov1.2), col = color)+
  #geom_point(data=df, aes(x = Cov1.2, y = RGR), col = color)+
  theme_bw()+
  theme(plot.background = element_rect(fill = '#e9ebe8', colour = '#e9ebe8'),
        panel.background = element_rect(fill="#e9ebe8"),
        text = element_text(size=20, colour = brown),
        axis.text.x = element_text(colour=brown), 
        axis.text.y = element_text(colour=brown),
        panel.border = element_rect(colour = brown)
        )
  #scale_y_continuous(limits = c(0,1.12))
figure_summary
```


# Save figure
```{r}
setwd("C:/Users/Carmen/Documents/Shrubs-Seedlings/results/figures")
pdf("pine_RGR.pdf", width = 5.5, height = 7.25, pointsize = 30,useDingbats = F)
figure_summary
dev.off()
```


# Calculate pseudo r squared
```{r}
meanGrth <- mean(df$RGR)
sse <- (df$RGR - predict(MM1))^2
sst <- (df$RGR - meanGrth)^2
R2 <- 1 - (sum(sse)/sum(sst))
R2
```


# Save final data
```{r}
save(df_dia, file = "~/Shrubs-Seedlings/compiled_data/fire_footprints/pine_vol.Rdata")
```
