---
title: Fir volume growth analysis exploration - fire footprints
author: Carmen
date: November 19, 2017
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(ggplot2)
library(dplyr)
require(nlme)
library(kableExtra)
library(knitr)
library(lme4)
library(tidyr)
library(randomForest)
library(VSURF)
library(effects)
```

# Load data
```{r}
load(file = "~/Shrubs-Seedlings/compiled_data/fire_footprints/pine_vol.Rdata")
```

# Visualize pre and post volumes on a log scale
```{r}
ggplot(df)+
  geom_point(aes(x = log(vol_pre_cm2), y = log(vol_post_cm2)), col = "red")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))+
  geom_abline(slope = 1,intercept = 0)
```

```{r}
ggplot(df)+
  geom_point(aes(x = log(vol_pre_cm2), y = (log(vol_post_cm2)-log(vol_pre_cm2)) ), col = "green")+
  scale_x_continuous(limits = c(-5,7.5))+
  scale_y_continuous(limits = c(-5,7.5))
```

## Plot relative growth rate vs. volume to check for a possible relationship

### least squares
```{r}
ggplot(df, aes(x = log(vol_pre_cm2), y = RGR))+
  geom_point()+
  geom_smooth(method = 'lm')
```

### loess
```{r}
ggplot(df, aes(x = log(vol_pre_cm2), y = RGR))+
  geom_point()+
  geom_smooth()
```

This looks like I should maybe include volume in the model of growth rate or adjust how I analyze growth rates

# Try using VSURF

## Make df of just the predictor variables I want to include
```{r}
xdf <- df %>% 
  select(Fire, 
         #FirePatch,
         Elevation,
         Slope.Deg,
         Aspect.rad.fold.NESW,
         #DIFN.all,
         Years,
         #LAI,
         #DIFN.53,
         tree_Ht_cm,
         Lat.rad,
         shrubarea2, 
         shrubarea3,
         Cov1, 
         Cov1.2, 
         Cov1.3, 
         Ht1, 
         Ht1.2, 
         Ht1.3,
         ShrubSpp01, 
         ShrubSpp02, 
         ShrubSpp03, 
         ShrG1, 
         BasDia2016.cm, 
         vol_pre_cm2, 
         heatload,
         incidrad) 
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
```

## Apply VSURF to RGR
```{r}
vsurf <- VSURF(x = xdf, y = df$RGR, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

# See how that compares to a simple random forest using the same variables
```{r}
forest <- randomForest(x = xdf, y = df$RGR, importance = TRUE, scale = TRUE)
varImpPlot(forest)
```

# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM
```{r}
LMdf <- df %>% 
  select(Fire, RGR, BasDia2016.cm, vol_pre_cm2, Elevation, Ht1.3) %>% 
  mutate_if(is.numeric, normalize)
LMdf <- droplevels(LMdf)
```

## Linear model
```{r}
LM <-  lme(RGR ~ BasDia2016.cm + Elevation + vol_pre_cm2+ Ht1.3 ,data = LMdf, random = ~ 1| Fire, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

### Look at individual effects
```{r}
eff_shrubht<- effect("Ht1.3", LM)
plot(eff_shrubht)

eff_shrubspp <- effect("Elevation", LM)
plot(eff_shrubspp)
```

### Compare to model with no shrub data
```{r}
LMnull <- lme(RGR ~ BasDia2016.cm + Elevation + vol_pre_cm2,data = LMdf, random = ~ 1| Fire, method = "ML")
anova(LM, LMnull)
```

### Residuals
```{r}
E <- resid(LM, type = "normalized")
pred <- predict(LM)
plot(LM)
plot(df$shrubarea2_sqrt_norm, E)
plot(df$tree_Ht_cm, E)
plot(log(df$vol_pre_cm2), E)
```

# Now do a random forest model for the vert:dia growth ratio

## Create growth ratio variable
```{r}
df <- df %>% 
  mutate(grthratio = VertGr_2017/DiaGrowth.mm) %>% 
  mutate(grthratio = ifelse(DiaGrowth.mm == 0, 10, grthratio))
```

**** Bold move: I adjusted the relative growth ratio to 10 for the seedling with 0 diameter growth

## Apply VSURF to grthratio
```{r}
vsurf <- VSURF(x = xdf, y = df$grthratio, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

# See how that compares to a simple random forest using the same variables
```{r}
forest <- randomForest(x = xdf, y = df$grthratio, importance = TRUE, scale = TRUE)
varImpPlot(forest)
```

# **NEXT STEPS: FIX THE SHRUB VARIABLES FOR GROWTH RATIO, REPEAT THAT STEP FOR FIRS, AND DECIDE ON FINAL VARIABLES TO INCLUDE**
