---
title: "Pine diameter growth analysis, all years"
author: "Carmen"
date: "November 14, 2017"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(ggplot2)
library(dplyr)
require(nlme)
library(knitr)
library(kableExtra)
```

# Load data
```{r}
load(file = "~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/master_dia.Rdata")
```

# Filter out firs and fires with very few pines
```{r}
df_dia <- df_dia %>% 
  filter(Species %in% c("PIPO","PILA")) %>% 
  filter(!(Fire == "WRTS"))
df_dia$Species <- as.factor(df_dia$Species)
df_dia %>% 
  group_by(Fire) %>% 
  summarise(n())
```

# STATS

```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

```{r}
df_dia <- df_dia %>% 
  mutate(Cov1.2_norm = normalize(Cov1.2)) %>% 
  mutate(Cov1.3_norm = normalize(Cov1.3)) %>% 
  mutate(Ht_norm = normalize(Ht_cm))
```

## Mixed effects model: How does relative diameter growth relate to shrub area?
```{r}
f <- DiaGrowth.mm_rel ~  shrubarea2_sqrt_norm + Ht_norm + ShrG1 + Ht_norm:shrubarea2_sqrt_norm
f_noshrubs <- DiaGrowth.mm_rel ~  Ht_norm  + ShrG1
LM1 <- lme(f,data = df_dia, random = ~ 1| Fire, method = "ML")
LM1_noshrubs <- lme(f_noshrubs,data = df_dia, random = ~ 1| Fire, method = "ML")
anova(LM1, LM1_noshrubs)
```

# Look for patterns in residuals of mixed effects model

## In relation to fitted values
```{r}
E <- resid(LM1, type="normalized")
df_dia$resid <- E
plot(LM1, col = df_dia$Fire)
df_dia[fitted.values(LM1)<0,]$resid
```

## In relation to height
```{r}
plot(E~df_dia$Ht_norm)
coplot(E ~ Ht_norm | Fire, data=df_dia, ylab="Normalized residuals", col=df_dia$Fire)
```

## In relation to shrub area and fire
```{r}
plot(df_dia$shrubarea2_sqrt_norm, E)
plot(df_dia$Fire, E)
```

# Is the variance different between larger and smaller seedlings? 
```{r}
sd(unlist(df_dia[df_dia$Ht_cm < median(df_dia$Ht_cm),"resid"]))
sd(unlist(df_dia[df_dia$Ht_cm >= median(df_dia$Ht_cm),"resid"]))
sd(df_dia$resid)
(sd(unlist(df_dia[df_dia$Ht_cm < median(df_dia$Ht_cm),"resid"])))/(sd(unlist(df_dia[df_dia$Ht_cm >= median(df_dia$Ht_cm),"resid"])))
```

Not bad.

# Plot 

## Make fake data
```{r}
MyData <-  rbind(
  expand.grid( 
    Fire = unique(df_dia$Fire), 
    shrubarea2_sqrt_norm = seq(min(df_dia$shrubarea2_sqrt_norm), max(df_dia$shrubarea2_sqrt_norm), length = 10),
    Ht_norm = seq(min(df_dia$Ht_norm), max(df_dia$Ht_norm), length = 10),
    Species = c("PIPO","PILA"),
    ShrG1 = unique(df_dia$ShrG1)
    ) 
)
```


## Plot
```{r}
MyData$pred_simple <- predict(LM1, newdata = MyData)
```

### Summary plot
```{r}
MyData_summary_simple <- MyData %>% 
  group_by(shrubarea2_sqrt_norm, ShrG1) %>% 
  mutate(pred_simple = mean(pred_simple))
figure_summary <- ggplot(MyData_summary_simple)+
  labs(title = "Predicted pine seedling diameter growth (lines) \nand actual growth (points)")+
  ylab("Predicted relative seedling growth")+
  xlab("shrub cover * shrub height 0-2 m from seedling")+
  geom_line(aes(y = pred_simple, x = shrubarea2_sqrt_norm, group = ShrG1, col = ShrG1))+
  geom_point(data=df_dia, aes(x = shrubarea2_sqrt_norm, y = DiaGrowth.mm_rel))
figure_summary
```

```{r}
df_dia %>% 
  group_by(ShrG1) %>% 
  summarise(n())
```

# Add date measured in 2016??

```{r}
summary(LM1)
```


# AIC table

## Find all possible variable combinations
```{r}
v1 <- c("shrubarea2_sqrt_norm", "Cov1.2_norm" , "Cov1.3_norm" , "IAG", "ShrG1", "shrubarea2_sqrt_norm:Ht_norm", "Cov1.2_norm:Ht_norm")
combs <- do.call("c", lapply(seq_along(v1), function(i) combn(v1, i, FUN = list)))
list.bad <- list()
for(i in 1:length(combs)){
  if("Cov1.2_norm" %in% combs[i][[1]] & "Cov1.3_norm" %in% combs[i][[1]])
    {
    list.bad <- c(list.bad, i)
    }
  if("shrubarea2_sqrt_norm" %in% combs[i][[1]] & ("Cov1.2_norm" %in% combs[i][[1]] | "Cov1.3_norm" %in% combs[i][[1]]     )){
    list.bad <- c(list.bad, i)
    } 
}
list.bad <- unlist(list.bad)
combs.bad <- combs[list.bad]
combs <- setdiff(combs, combs.bad)
length(combs)
```

## Table
```{r}
table <- data.frame(Model = character(), Formula = character(), AIC = numeric())
for(i in 1:length(combs)){
  vars <- combs[i]
  vars <- paste(vars[[1]], collapse="+")
  f <- formula(paste("(DiaGrowth.mm_rel ~ Ht_norm +", vars,")"))
  M <- lme(f,data = df_dia, random = ~ 1| Fire)
  ModelName <- paste("M", i, sep="_")
  add <- data.frame(c(ModelName),c(vars), c(AIC(M)))
  table <- rbind(table, add)
}
colnames(table) <- c("Model","Variables", "AIC")
table_final <- kable(table %>% arrange(AIC), format = "markdown") 
table_final
```

# Re-analyse and plot using best model from above that sticks with 0-2 m
```{r}
f <- DiaGrowth.mm_rel ~  Cov1.2_norm + Ht_norm 
f_null <- DiaGrowth.mm_rel ~  Ht_norm
LM2 <- lme(f,data = df_dia, random = ~ 1|Fire, method = "ML")
LM2_null <- lme(f_null,data = df_dia, random = ~ 1|Fire, method = "ML")
anova(LM2, LM2_null)
```

## Look at residuals
```{r}
E <- resid(LM2, type="normalized")
df_dia$resid <- E
plot(LM2, col = df_dia$Fire)
plot(df_dia$Ht_norm, E)
```


# THIS IS THE BEST:
```{r}
summary(LM2)
```

# Re-plot

## Make fake data
```{r}
MyData2 <-  rbind(
  expand.grid( 
    Fire = unique(df_dia$Fire),
    Ht_cm = mean(df_dia$Ht_cm),
    Cov1.2 = seq(min(df_dia$Cov1.2), max(df_dia$Cov1.2), length = 20)
    ) 
)
```

## "normalize" MyData
```{r}
MyData2$Ht_norm <- (MyData2$Ht_cm - mean(df_dia$Ht_cm))/sd(df_dia$Ht_cm)
MyData2$Cov1.2_norm <- (MyData2$Cov1.2 - mean(df_dia$Cov1.2))/sd(df_dia$Cov1.2)
```


## Plot
```{r}
MyData2$pred_simple <- predict(LM2, newdata = MyData2)
```

### Summary plot
```{r}
MyData2_summary_simple <- MyData2 %>% 
  group_by(Cov1.2) %>%
  mutate(pred_simple = mean(pred_simple))

figure_summary <- ggplot(MyData2_summary_simple)+
  labs(title = "Pine seedling diameter growth")+
  ylab("relative growth (mm/mm)")+
  xlab("shrub cover 0-2 m from seedling")+
  geom_line(aes(y = pred_simple, x =Cov1.2), col = "#a02313")+
  geom_point(data=df_dia, aes(x = Cov1.2, y = DiaGrowth.mm_rel), col = "#a02313")+
  theme_bw()+
  theme(plot.background = element_rect(fill = '#e9ebe8', colour = '#e9ebe8'),
        panel.background = element_rect(fill="#e9ebe8"),
        text = element_text(size=20),
        axis.title = element_text(colour = "#a02313"),
        plot.title = element_text(colour = "#a02313")
        )+
  scale_y_continuous(limits = c(0,.72))
figure_summary
remove(MyData2_summary_simple)
```


# Save figure
```{r}
setwd("C:/Users/Carmen/Documents/Shrubs-Seedlings/results/figures")
pdf("pine_dia.pdf", width = 5.5, height = 7.25, pointsize = 30,useDingbats = F)
figure_summary
dev.off()
```

# Calculate pseudo r squared
```{r}
meanGrth <- mean(df_dia$DiaGrowth.mm_rel)
sse <- (df_dia$DiaGrowth.mm_rel - predict(LM2))^2
sst <- (df_dia$DiaGrowth.mm_rel - meanGrth)^2
R2 <- 1 - (sum(sse)/sum(sst))
R2
```

# Save final data
```{r}
save(df_dia, file = "~/Shrubs-Seedlings/compiled_data/fire_footprints/pinus_dia.Rdata")
```

