---
title: 'Analysis 6: Pinus with shrubarea vertical analysis - fire footprints'
author: "Carmen"
date: "March 9 + 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
```

# Load data
```{r}
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

## Filter to just PIPO
```{r}
df %>% 
  filter(Species == "PIPO") %>% 
  group_by(Fire) %>% 
  summarize(n()) %>% 
  arrange(`n()`)
df <- df %>% 
  filter(Species == "PIPO")
```

## Change species and Sdlg to factor
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg)) 
summary(df$Species)
summary(df$Sdlg)

```

# Count the seedlings
```{r}
length(unique(df$Sdlg))
nrow(df)
```


# Try using VSURF

## Make df of just the predictor variables I want to include
```{r}
xdf <- df %>% 
  select(Elevation,
         Slope.Deg,
         Aspect.rad.fold.NESW,
         heatload,
         incidrad,
         Year,
         Years,
         BasDia2016.cm, 
         Ht_cm,
         shrubarea1,
         shrubarea2, 
         shrubarea3,
         Cov1, 
         Cov1.2, 
         Cov1.3, 
         Ht1, 
         Ht1.2, 
         Ht1.3,
         ShrubSpp01, 
         ShrubSpp02, 
         ShrubSpp03
         ) 
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
```

# STOP HERE TO SAVE TIME

## Apply VSURF to vertical growth
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Check out shrubarea3/shrubarea2 and ShrubSpp01/03

### See how correlated they are
```{r}
ggplot(df, aes(x = shrubarea3, y = shrubarea2))+
  geom_point()+
  geom_smooth(method = "loess")
```

### Run a random forest
```{r}
forest <- randomForest(x = xdf %>% select(Ht_cm, Years, shrubarea3, shrubarea2, Ht1.3, BasDia2016.cm, ShrubSpp01, Year, ShrubSpp03, incidrad), y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
varImpPlot(forest)
```

### Delete SHRUBAREA3 and SHRUBSPP03 based on the results of the above randomforest
```{r}
xdf <- xdf %>% 
  select(-shrubarea3, -ShrubSpp03)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```


# RESUME HERE

## Look at final random forest
```{r}
forest <- randomForest(
  x = xdf %>% 
    select(Ht_cm, Years, shrubarea2, Ht1.3, BasDia2016.cm, ShrubSpp01, Year), 
  y = df$VertGrowth_Rel, 
  importance = TRUE, 
  scale = TRUE)
varImpPlot(forest)
```


# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea2 = sqrt(shrubarea2)) %>% 
  select(Sdlg, VertGrowth_Rel, Year, Ht_cm, shrubarea2, sqrt_shrubarea2, Ht1.3, BasDia2016.cm, ShrubSpp01, Year) %>% 
 mutate_if(is.numeric, normalize)
LMdf <- droplevels(LMdf)
```

## See whether model with shrubarea3 or sqrt_shrubarea3 is better
```{r}
AIC(lme(VertGrowth_Rel ~ Year + Ht_cm + shrubarea2 + sqrt_shrubarea2 + Ht1.3 + BasDia2016.cm + ShrubSpp01 + Year,
        data = LMdf, random = ~ 1| Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  Year + Ht_cm + sqrt_shrubarea2 + sqrt_shrubarea2 + Ht1.3 + BasDia2016.cm + ShrubSpp01 + Year,
        data = LMdf, random = ~ 1| Sdlg, method = "ML")
)
```

Transformed wins.


## Linear model
```{r}
LM <-  lme(VertGrowth_Rel ~ Year + Ht_cm + sqrt_shrubarea2 + sqrt_shrubarea2 + Ht1.3 + BasDia2016.cm + ShrubSpp01 + Year,data = LMdf, random = ~ 1| Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

### Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

#### Re-run model with non-normalized data
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>% 
  select(Sdlg, VertGrowth_Rel, shrubarea3, BasDia2016.cm, Year, Cov1.2, incidrad, Year) %>% 
  mutate(shrubarea3_m = shrubarea3/10000)
LMdf <- droplevels(LMdf)
```

```{r}

LM <-  lme(VertGrowth_Rel ~ shrubarea3_m + BasDia2016.cm + Cov1.2 + incidrad + Year,data = LMdf, random = ~ 1| Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

```{r}
eff_shrubarea <- predictorEffect("shrubarea3_m", LM)
plot(eff_shrubarea)

eff_year <- predictorEffect("Year", LM)
plot(eff_year)
```

### Re-create effects plot using predictorEffects() and ggplot
```{r}
eff_shrubarea <- predictorEffect("shrubarea3_m", LM)
effects_df <- as.data.frame(eff_shrubarea)
ggplot(effects_df)+
  geom_line(aes(x = shrubarea3_m, y = fit))+
  geom_ribbon(aes(x = shrubarea3_m, ymin = lower, ymax = upper), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab(bquote('Shrub cross-sectional area ' (m^2)))+
  ylab("Predicted relative growth (cm/cm)")+
  ggtitle("Pine growth in relation to shrub environment within 3 m")+
  geom_rug(data = df, aes(x = shrubarea3/10000, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(0,0.25))
ggsave(file = "../../results/figures/FireFootprints/PineVertRel.png")
```


### Compare to model with no shrub data
```{r}
LMnull <- lme(VertGrowth_Rel ~ BasDia2016.cm + incidrad + Year,data = LMdf, random = ~ 1| Sdlg, method = "ML")
anova(LM, LMnull)
```

### Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm, E)
ggplot(df, aes(x = log(df$shrubarea3), y = E))+
  geom_point()+
  geom_smooth()
```


## Save final data
```{r}
save(df, file ="~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/pine_vert.Rdata")
```
