---
title: 'Pinus with shrubarea vertical analysis - fire footprints'
author: "Carmen"
date: "March 9 + 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
require(MuMIn)
require(car)
set.seed(12346)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

# Load VSURF results
```{r}
load("../../results/VSURF_pipo_vars.Rdata")
```

## Filter to just PIPO
```{r}
df %>% 
  filter(Species == "PIPO") %>% 
  group_by(Fire) %>% 
  summarize(n()) %>% 
  arrange(`n()`)
df <- df %>% 
  filter(Species == "PIPO")
```

# Count totals and by fire
```{r}
length(unique(df$Sdlg))
nrow(df)
```

```{r}
df %>% 
  group_by(Sdlg, Fire) %>%
  select(Sdlg, Fire) %>%
  distinct() %>% 
  ungroup() %>% 
  group_by(Fire) %>% 
  summarize(n = n()) %>% 
  mutate(total = sum(n))
```

# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```

# Structure data for LM

## Normalize function
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

## filter and normalize df
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3+1)) %>%
  dplyr::select(Fire, 
         Sdlg, 
         VertGrowth_Rel, 
         Years, 
         Ht_cm1, 
         sqrt_shrubarea3, 
         log_shrubarea3, 
         BasDia2016.cm, 
         incidrad, 
         Year, 
         Elevation, 
         siteclass,
         heatload, 
         ShrubSpp03) %>% 
 mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```

# Model using VSURF var results

## Convert vars to formula structure
```{r}
vars_one <- paste(vars, collapse = " + ")
vars_one
f <- formula(paste("VertGrowth_Rel ~ ",vars_one))
f
```

## Run lme model
```{r}
LM <- lme(f, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
```

## Check collinearity

### Compare Zuur model with vif function from car package
```{r}
source("~/../Documents/HighstatLibV10.R.txt") # from Zuur https://highstat.com/index.php/mixed-effects-models-and-extensions-in-ecology-with-r
corvif(LMdf %>% select(!!vars))
vif(LM)
```

A VIF cutoff value of 5 is often used, and none of my variables have values larger than that, so I think I'm ok.

# See whether model with or without interaction is better
```{r}
vars_int <- paste(vars_one, "+ Ht_cm1:sqrt_shrubarea3")
f_int <- formula(paste("VertGrowth_Rel ~ ",vars_int))
LM_int <- lme(f_int, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
AIC(LM)
AIC(LM_int)
if(
  AIC(LM) > AIC(LM_int)
) print(paste("WITH INTERACTION WINS BY", round(AIC(LM)-AIC(LM_int), digits = 2), "AIC")) else print(paste("WITHOUT INTERACTION WINS BY", round(AIC(LM_int)-AIC(LM), digits = 2)))
```

# See whether residuals are better for log transformed or not
```{r}
plot(LM)
```

```{r}
f_no_log <- formula(paste("exp(VertGrowth_Rel) ~", vars_one))
LM_no_log <-  lme(f_no_log, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
plot(LM_no_log)
```

Log transformed looks better.

# Look at summary and drop1 results
```{r}
summary(LM)
```

```{r}
drop <- drop1(LM, test = "Chisq")
drop$var <- row.names(drop)
drop <- tbl_df(drop)
drop %>% 
  arrange( `Pr(>Chi)`) %>% 
  mutate( `Pr(>Chi)` = paste( `Pr(>Chi)`))
```

## Run anova by hand for the interacting variables since drop1 doesn't do that by default
```{r}
vars_no_shrub <- str_replace(vars_one, "sqrt_shrubarea3 +", "")
vars_no_shrub
f_no_shrub <- formula(paste("VertGrowth_Rel ~ ", vars_no_shrub))
LM_noshrub <- lme(f_no_shrub, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM_int, LM_noshrub)
```

```{r}
vars_no_ht <- vars[vars != "Ht_cm1"]
vars_no_ht_one <- paste(vars_no_ht, collapse = " + ")
f_no_ht <- formula(paste("VertGrowth_Rel ~ ", vars_no_ht_one))
LM_noht <- lme(f_no_ht, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM_int, LM_noht)
```

# Look at collinearity between height and shrub competition
```{r}
cor(df$Ht_cm1, df$sqrt_shrubarea3)
ggplot(df, aes(x = Ht_cm1, y = sqrt_shrubarea3))+
  geom_point()+
  geom_smooth(method = "lm")
```

# Look at residuals

### Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm1, E)
plot(df$Years, E)
plot(df$Ht1.3, E)
plot(df$BasDia2016.cm, E)
plot(df$incidrad, E)
plot(df$Year, E)
ggplot(df, aes(x = log(df$shrubarea3), y = E))+
  geom_point()+
  geom_smooth()
```

# Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

```{r}
plot(predictorEffect("Years", LM))
plot(predictorEffect("sqrt_shrubarea3", LM))
plot(predictorEffect("Ht_cm1", LM))
plot(predictorEffect("Year", LM))
plot(predictorEffect("ShrubSpp03", LM))
```

# Calculate Nakagawa's pseudo R2
```{r}
r.squaredGLMM(LM)
```

# Plot shrub effects with an interaction

```{r}
eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM_int)
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Create df for effects and confidence intervals using mean +/- 1 SD
```{r}
quartiles <- summary(df$Ht_cm1)[c(2,5)]
quartiles <- (quartiles-mean(df$Ht_cm1))/sd(df$Ht_cm1)
eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM_int, xlevels = list(Ht_cm1 = quartiles))
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(sqrt_shrubarea3 = sqrt_shrubarea3*sd(sqrt(df$shrubarea3))+mean(sqrt(df$shrubarea3)) ) %>% 
  mutate(Ht_cm1 = Ht_cm1*sd(df$Ht_cm1)+mean(df$Ht_cm1) ) %>% 
  mutate(Ht_cm1 = as.factor(round(Ht_cm1)) )
head(effects_df)
```

### test
```{r}
hist(effects_df$sqrt_shrubarea3)
hist(sqrt(df$shrubarea3))
summary(as.factor(effects_df$Ht_cm1))
summary(df$Ht_cm1)
```

## Check for whether certain pine tree sizes only occur at certain shrub competition levels
```{r}
summary(df$sqrt_shrubarea3[df$Ht_cm1< mean(df$Ht_cm1)])
summary(df$sqrt_shrubarea3[df$Ht_cm1> mean(df$Ht_cm1)])
```

## Plot effects of shrubarea WITH interaction
```{r}
ggplot(effects_df)+
  geom_line(aes(x = sqrt_shrubarea3, y = exp(fit), group = Ht_cm1, col = Ht_cm1))+
  geom_ribbon(aes(x = sqrt_shrubarea3, ymin = exp(lower), ymax = exp(upper), group = Ht_cm1, fill = Ht_cm1), alpha = .5)+
  theme_bw()+
  #xlab("                                  Shrub competition")+
  xlab(bquote('                           Shrub competition ' ))+
  ylab("Relative growth rate")+
  #ggtitle("Pine growth in relation to shrub environment within 2 m")+
  geom_rug(data = df, aes(x = sqrt(shrubarea3), y = exp(VertGrowth_Rel)), alpha = .5, position = "jitter", sides = "b")+
  #ylim(c(0.04,0.3))+
  labs(fill = "Juvenile tree height (cm)")+
  labs(col = "Juvenile tree height (cm)")+
  theme(legend.position = c(0.62, 0.8))+
  theme(text = element_text(size = 11),
        #panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())
ggsave(file = "../../results/figures/FireFootprints/PineVertRel_int.png", width = 3, height = 3, dpi = 400)
```


# Plot shrub effects WITHOUT interaction

### Re-create effects plots using predictorEffects() and ggplot

```{r}
eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM)
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

### Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(sqrt_shrubarea3 = sqrt_shrubarea3*sd(sqrt(df$shrubarea3))+mean(sqrt(df$shrubarea3)) ) 
```

```{r}
ggplot(effects_df)+
  geom_line(aes(x = sqrt_shrubarea3, y = exp(fit)))+
  geom_ribbon(aes(x = sqrt_shrubarea3, ymin = exp(lower), ymax = exp(upper)), alpha = .2)+
  theme_bw()+
  xlab(bquote('                           Shrub competition ' ))+
  ylab("Relative growth rate")+
  #ggtitle("Pine growth in relation to shrub environment within 2 m")+
  geom_rug(data = df, aes(x = sqrt(shrubarea3), y = exp(VertGrowth_Rel)), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(0.04,0.35))+
  labs(fill = "Juvenile tree height (cm)")+
  labs(col = "Juvenile tree height (cm)")+
  theme(legend.position = c(0.62, 0.8))+
  theme(
        #panel.grid.minor = element_blank(),
        #panel.grid.major = element_blank(),
        text = element_text(size = 11))
ggsave(file = "../../results/figures/FireFootprints/PineVertRel_noint.png", width = 3, height = 3, dpi = 400)
```


# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea3))
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea3))
```

# Summarize range of shrub cover and height
```{r}
summary(df$Cov1.3)
summary(df$Ht1.3)
summary(df$shrubarea3)
```

# Convert cover to Percent
```{r}
summary(df$Cov1.3/(300*4))
```

# Find the shrubarea values associated with each quartile, and the mean growth for each of those
```{r}
summary <- summary(sqrt(df$shrubarea3))
summary
quart1 <- summary[2]
quart3 <- summary[5]
median <- summary[3]
min <- summary[1]
max <- summary[6]
mean <- mean(sqrt(df$shrubarea3))

fit_quart1 <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart1)),]$fit
fit_quart1
fit_quart3 <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart3)),]$fit
fit_median <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - median)),]$fit
fit_min <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - min)),]$fit
fit_max <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - max)),]$fit
fit_mean <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - mean)),]$fit
mean <- mean(df$shrubarea3/10000)
print(paste("fit_quart1 = ", exp(fit_quart1)))
print(paste("fit_quart3 = ", exp(fit_quart3)))
print(paste("fit_median = ", exp(fit_median)))
print(paste("fit_min = ", exp(fit_min)))
print(paste("fit_max = ", exp(fit_max)))


print(paste("fit_quart1_log = ", fit_quart1))
print(paste("fit_quart3_log = ", fit_quart3))
print(paste("fit_median_log = ", fit_median))
print(paste("fit_min_log = ", fit_min))
print(paste("fit_max_log = ", fit_max))
print(paste("fit_mean_log = ", fit_mean))
```

## Repeat for 60% shrub cover and 86 cm ht similar to what's in COllins and ROller
```{r}
collins_shrubarea <- sqrt((.6*1200)*86)
effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart1)),]$fit
```

```{r}
sd <- sd(df$shrubarea3/10000)
```

# Pretty plot of shrub species effects
```{r}
eff_shrubspp <- predictorEffect("ShrubSpp03", LM)
effects_df <- as.data.frame(eff_shrubspp)
head(effects_df)
```

```{r}
ggplot(effects_df)+
  geom_point(aes(x = ShrubSpp03, y = exp(fit)))+
  ylab("Juvenile pine relative growth rate")+
  theme_bw()+
  geom_errorbar(aes(x = ShrubSpp03, ymin=exp(lower), ymax=exp(upper)), width=.1) +
  scale_x_discrete(labels=c("ARPA" = "Arctostaphylos",
                            "CECO" = "Ceanothus\ncordulatus",
                            "CEIN" = "Ceanothus\nintegerrimus",
                            "CHFO" = "Chamaebatia\nfoliolosa",
                            "LIDE" = "Notholithocarpus\ndensiflorus"
                            ))+
  theme(  text = element_text(size = 11),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(face = "bold.italic",
                                   angle=45, 
                                   vjust=1, 
                                   hjust=1))+
  xlab("Dominant shrub species")
  


 ggsave(file = "../../results/figures/FireFootprints/PineVertShrubSpp.png", width = 4, height = 4, dpi = 400)
```

# List the final seedlings used
```{r}
sort(unique(df$Sdlg))
```

# Save final data
```{r}
save(df, file ="../../compiled_data/fire_footprints/pine_vert.Rdata")
```
