---
title: 'Pinus with shrubarea vertical analysis - fire footprints'
author: "Carmen"
date: "March 9 + 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
set.seed(12346)
```

# Load data
```{r}
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

## Filter to just PIPO
```{r}
df %>% 
  filter(Species == "PIPO") %>% 
  group_by(Fire) %>% 
  summarize(n()) %>% 
  arrange(`n()`)
df <- df %>% 
  filter(Species == "PIPO")
```

## Change species and Sdlg to factor
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg)) 
summary(df$Species)
summary(df$Sdlg)

```

# Count the seedlings
```{r}
length(unique(df$Sdlg))
nrow(df)
```

# Try using VSURF

## Make df of just the predictor variables I want to include
```{r}
xdf <- df %>% 
  select(Elevation,
         Slope.Deg,
         Aspect.rad.fold.NESW,
         heatload,
         incidrad,
         Year,
         Years,
         BasDia2016.cm, 
         Ht_cm1,
         shrubarea1,
         shrubarea2, 
         shrubarea3,
         Cov1, 
         Cov1.2, 
         Cov1.3, 
         Ht1, 
         Ht1.2, 
         Ht1.3,
         ShrubSpp01, 
         ShrubSpp02, 
         ShrubSpp03
         ) 
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
```


# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```


# STOP HERE TO SAVE TIME

## Apply VSURF to vertical growth
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Check out shrubarea3/shrubarea2 

### See how correlated they are
```{r}
ggplot(df, aes(x = shrubarea3, y = shrubarea2))+
  geom_point()+
  geom_smooth(method = "loess")
```


## Run 10 random forests on the ones that come up to see whether to delete heatload or incidrad
```{r}
forests100 <- function(vars){
  forest <- randomForest(
    x = xdf %>% select(Years, Ht_cm1, shrubarea3, shrubarea2, Ht1.3, Year, ShrubSpp01), 
    y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
  forest$importance
  imp <- as.data.frame(forest$importance) 
  imp$var <- row.names(imp)
  imp <- imp %>% filter(var %in% vars)
  imp %>% filter(`%IncMSE` == max(imp$`%IncMSE`)) %>% select(var)
}
forests100(c("shrubarea3", "shrubarea2"))
reps <- replicate(100, forests10(c("shrubarea3", "shrubarea2")))
summary(as.factor(unlist(reps)))
```

This shows that shrubarea3 has a higher importance value than shrubarea2 most of the time, so I should delete shrubarea2.

### Delete SHRUBAREA2 based on the results of the above randomforest
```{r}
xdf <- xdf %>% 
  select(-shrubarea2)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Repeat for Ht1.3/Ht1.2/Ht1

### See how correlated they are
```{r}
ggplot(df, aes(x = Ht1.3, y = Ht1))+
  geom_point()+
  geom_smooth(method = "loess")
```

### Run a random forest
```{r}
forest <- randomForest(
  x = xdf %>% select(Ht_cm, Years, shrubarea3, Ht1.3, Ht1.2, BasDia2016.cm, Ht1, Year, incidrad, ShrubSpp03), 
  y = df$VertGrowth_Rel, importance = TRUE, scale = TRUE)
varImpPlot(forest)
```

### Delete ht1.2 and ht1.3 based on the results of the above randomforest
```{r}
xdf <- xdf %>% 
  select(-Ht1.2, -Ht1.3)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Repeat for shrubarea3/shrubarea1

### See how correlated they are
```{r}
ggplot(df, aes(x = shrubarea3, y = shrubarea1))+
  geom_point()+
  geom_smooth(method = "loess")
```

### Just delete shrubarea1 since shrubarea3 has been in the model for multiple steps
```{r}
xdf <- xdf %>% 
  select(-shrubarea1)
```

### Then run VSURF again
```{r}
vsurf <- VSURF(x = xdf, y = df$VertGrowth_Rel, na.action = na.omit)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

# RESUME HERE

## Look at final random forest
```{r}
forest <- randomForest(
  x = xdf %>% 
    select(Ht_cm, Years, shrubarea3, BasDia2016.cm, incidrad, ShrubSpp03, Year), 
  y = df$VertGrowth_Rel, 
  importance = TRUE, 
  scale = TRUE)
varImpPlot(forest)
```


# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3)) %>%
  select(Fire, Sdlg, VertGrowth_Rel, Years, Ht_cm, shrubarea3, sqrt_shrubarea3, log_shrubarea3, BasDia2016.cm, incidrad, ShrubSpp03, Year) %>% 
 mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```

## See whether model with shrubarea3 or sqrt_shrubarea3 is better
```{r}
AIC(lme(VertGrowth_Rel ~ Years + Ht_cm*shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  Years + Ht_cm*sqrt_shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  Years + Ht_cm*log_shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

```

Untransformed wins

## See whether model with or without interaction is better
```{r}
AIC(lme(VertGrowth_Rel ~ Years + Ht_cm*shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  Years + Ht_cm+shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
)
```

With interaction is better.

## Linear model
```{r}
LM <-  lme(VertGrowth_Rel ~ Years + Ht_cm*shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

### Test dropping shrubarea3 and ht individually

#### Height
```{r}
LM_noht <- lme(VertGrowth_Rel ~ Years + shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year, 
               data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM, LM_noht)
```

#### Shrubarea
```{r}
LM_noshrub <- lme(VertGrowth_Rel ~ Years + Ht_cm + BasDia2016.cm + incidrad + ShrubSpp03 + Year, 
                  data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM, LM_noshrub)
```

#### Interaction
```{r}
LM_noshrub <- lme(VertGrowth_Rel ~ Years + Ht_cm + shrubarea3 + BasDia2016.cm + incidrad + ShrubSpp03 + Year, 
                  data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
anova(LM, LM_noshrub)
```


Note: anova here is performing likelihood ratio tests

### Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

```{r}
eff_years <- predictorEffect("Years", LM)
plot(eff_years)

eff_shrubarea <- predictorEffect("shrubarea3", LM)
plot(eff_shrubarea)

eff_ht <- predictorEffect("Ht_cm", LM)
plot(eff_ht)

eff_shrspp <- predictorEffect("ShrubSpp03", LM)
plot(eff_shrspp)

plot(predictorEffect("BasDia2016.cm", LM))

plot(predictorEffect("incidrad", LM))

plot(predictorEffect("Year", LM))
```

# Re-create effects plots using predictorEffects() and ggplot

## Create df for effects and confidence intervals using mean +/- 1 SD
```{r}
quartiles <- c(mean(df$Ht_cm)-sd(df$Ht_cm), mean(df$Ht_cm)+sd(df$Ht_cm))
quartiles
quartiles <- c(mean(LMdf$Ht_cm)-sd(LMdf$Ht_cm), mean(LMdf$Ht_cm)+sd(LMdf$Ht_cm))
quartiles

eff_shrubarea <- predictorEffect("shrubarea3", LM, xlevels = list(Ht_cm = quartiles))
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(Ht_cm = Ht_cm*sd(df$Ht_cm)+mean(df$Ht_cm))
  
effects_df <- effects_df %>% 
  mutate(shrubarea3 = shrubarea3*sd(df$shrubarea3)+mean(df$shrubarea3)) 
```

### test
```{r}
summary(as.factor(effects_df$Ht_cm))
hist(effects_df$shrubarea3)
hist(df$shrubarea3)
```

## Make shrubarea in m instead of cm
```{r}
effects_df <- effects_df %>% 
  mutate(shrubarea3 = shrubarea3/10000)
```

## Round the values of ht
```{r}
effects_df <- effects_df %>% 
  mutate(Ht_cm = round(Ht_cm))
```

## Plot effects of shrubarea
```{r}
ggplot(effects_df)+
  geom_line(aes(x = shrubarea3, y = fit, group = as.factor(Ht_cm)))+
  geom_ribbon(aes(x = shrubarea3, group = as.factor(Ht_cm), fill = as.factor(Ht_cm), ymin = lower, ymax = upper), alpha = .5)+
  theme_bw()+
  xlab(bquote('Shrub competition ' (m^2)))+
  ylab("Relative growth rate")+
  ggtitle("Pine growth in relation to shrub environment within 3 m")+
  geom_rug(data = df, aes(x = shrubarea3/10000, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(-0.1,0.7))+
  labs(fill = "Juvenile tree height (cm)")+
  theme(legend.position = c(0.8, 0.8))
ggsave(file = "../../results/figures/FireFootprints/PineVertRel_int.png", width = 6, height = 6, dpi = 400)
```

## Years
```{r}
eff_years <- predictorEffect("Years", LM)
effects_df <- as.data.frame(eff_years)
effects_df <- effects_df %>% 
  mutate(Years = Years*sd(df$Years)+mean(df$Years)) 

ggplot(effects_df)+
  geom_line(aes(x = Years, y = fit))+
  geom_ribbon(aes(x = Years, ymin = lower, ymax = upper), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab("Years since fire")+
  ylab("Predicted relative growth (cm/cm)")+
  ggtitle("Pine growth in relation to time since fire")+
  geom_rug(data = df, aes(x = Years, y = VertGrowth_Rel), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(-0.02,0.28))
ggsave(file = "../../results/figures/FireFootprints/PineVertRel_years.png", width = 6, height = 5, dpi = 400)
```

# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea3))
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea3))
```

### Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm, E)
plot(df$Years, E)
plot(df$Ht1.3, E)
plot(df$BasDia2016.cm, E)
plot(df$incidrad, E)
plot(df$Year, E)
ggplot(df, aes(x = log(df$shrubarea3), y = E))+
  geom_point()+
  geom_smooth()
```


## Save final data
```{r}
save(df, file ="~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/pine_vert.Rdata")
```
