---
title: 'Pinus with shrubarea vertical analysis - fire footprints'
author: "Carmen"
date: "March 9 + 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
set.seed(12346)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

## Filter to just PIPO
```{r}
df %>% 
  filter(Species == "PIPO") %>% 
  group_by(Fire) %>% 
  summarize(n()) %>% 
  arrange(`n()`)
df <- df %>% 
  filter(Species == "PIPO")
```

## Change species and Sdlg and soil order to factor
```{r}
df <- df %>% 
  mutate(Species = as.factor(Species)) %>% 
  mutate(Sdlg = as.factor(Sdlg)) %>% 
  mutate(soil_order = as.factor(soil_order))
```

# Count the seedlings
```{r}
length(unique(df$Sdlg))
nrow(df)
```

# Try using VSURF

## Make df of just the predictor variables I want to include
```{r}
xdf <- df %>% 
  select(Elevation,
         Slope.Deg,
         heatload,
         incidrad,
         Aspect.rad.fold.NESW,
         siteclass,
         Year,
         Years,
         BasDia2016.cm, 
         Ht_cm1,
         shrubarea3,
         ShrubSpp03) 
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
```

# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```

# Run a random forest on everything
```{r}
forest <- randomForest(
     x = xdf,
     y = df$VertGrowth_Rel, 
     importance = TRUE, 
     scale = TRUE, 
     ntree=5000)
varImpPlot(forest)
```

# STOP HERE TO SAVE TIME

# Apply VSURF to vertical growth
```{r}
vsurf <- VSURF(x = xdf, 
               y = df$VertGrowth_Rel, 
               na.action = na.omit,
               parallel = T)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

## Save the results
```{r}
vars <- names(xdf[vsurf$varselect.pred])
vars
```

# RESUME HERE

## Look at final random forest
```{r}
forest <- randomForest(
  x = xdf %>% 
    select(vars), 
  y = df$VertGrowth_Rel, 
  importance = TRUE, 
  scale = TRUE)
varImpPlot(forest)
```

# Look at a linear model using the variables that were identified by VSURF, plus interactions that seem biologically meaningful

## Structure data for LM
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3+1)) %>%
  select(Fire, 
         Sdlg, 
         VertGrowth_Rel, 
         Years, 
         Ht_cm1, 
         shrubarea3, 
         sqrt_shrubarea3, 
         log_shrubarea3, 
         BasDia2016.cm, 
         incidrad, 
         Year, 
         Elevation, 
         heatload, 
         ShrubSpp03) %>% 
 mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```

```{r}
vars
LM <- lme(VertGrowth_Rel ~  Years +  Ht_cm1 + shrubarea3 + heatload + incidrad + Year + ShrubSpp03,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
```

## See whether model with shrubarea2 or sqrt_shrubarea2 is better
```{r}
AIC(lme(VertGrowth_Rel ~ 
         Years +  Ht_cm1 + shrubarea3 + heatload +  incidrad + Year + ShrubSpp03, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  
          Years +  Ht_cm1 + sqrt_shrubarea3 + heatload +  incidrad + Year + ShrubSpp03, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))

AIC(lme(VertGrowth_Rel ~  
           Years +  Ht_cm1 + log_shrubarea3 + heatload +  incidrad + Year + ShrubSpp03, 
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML"))
```

Square root transformed wins

## See whether model with or without interaction is better
```{r}
without <- lme(VertGrowth_Rel ~ 
           Years +  Ht_cm1 + sqrt_shrubarea3 + heatload +  incidrad + Year+ ShrubSpp03,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
AIC(without)
with <- lme(VertGrowth_Rel ~  
           Years +  Ht_cm1*sqrt_shrubarea3 + heatload +  incidrad + Year+ ShrubSpp03,
        data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
AIC(with)
anova(with, without)
```

Without wins.

## Look at drop1 results
```{r}
LM <-  lme(VertGrowth_Rel ~ 
            Years +  Ht_cm1 + sqrt_shrubarea3 + heatload +  incidrad + Year + ShrubSpp03,
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

No need to include interaction after all

# See whether residuals are better for log transformed or not
```{r}
LM <-  lme(VertGrowth_Rel ~ 
            Years +  Ht_cm1 + sqrt_shrubarea3 + heatload +  incidrad + Year + ShrubSpp03,
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
plot(LM)
```

```{r}
LM <-  lme(exp(VertGrowth_Rel) ~ 
            Years +  Ht_cm1 + sqrt_shrubarea3 + heatload +  incidrad + Year+ ShrubSpp03,
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
plot(LM)
```

Log transformed looks better 

# Re-establish the best LM so far:
```{r}
LM <-  lme(VertGrowth_Rel ~ 
           Years +  Ht_cm1 + sqrt_shrubarea3 + heatload +  incidrad + Year+ ShrubSpp03,
           data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
summary(LM)
drop1(LM, test = "Chisq")
```

Note: Chisq here is performing likelihood ratio tests

# Look at residuals

### Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm1, E)
plot(df$Years, E)
plot(df$Ht1.3, E)
plot(df$BasDia2016.cm, E)
plot(df$incidrad, E)
plot(df$Year, E)
ggplot(df, aes(x = log(df$shrubarea3), y = E))+
  geom_point()+
  geom_smooth()
plot(df$soil_order, E)
```

# Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

```{r}
eff_years <- predictorEffect("Years", LM)
plot(eff_years)

eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM)
plot(eff_shrubarea)

eff_ht <- predictorEffect("Ht_cm1", LM)
plot(eff_ht)

plot(predictorEffect("Year", LM))
```

# Re-create effects plots using predictorEffects() and ggplot

```{r}
eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM)
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(sqrt_shrubarea3 = sqrt_shrubarea3*sd(sqrt(df$shrubarea3))+mean(sqrt(df$shrubarea3)) )
effects_df
```

### test
```{r}
hist(effects_df$sqrt_shrubarea3)
hist(sqrt(df$shrubarea3))
```


## Plot effects of shrubarea
```{r}
ggplot(effects_df)+
  geom_line(aes(x = sqrt_shrubarea3, y = exp(fit)))+
  geom_ribbon(aes(x = sqrt_shrubarea3, ymin = exp(lower), ymax = exp(upper)), alpha = .5, fill = "#fdbb84")+
  theme_bw()+
  xlab(bquote('Shrub competition ' (sqrt(`cover (cm)`%*%`height (cm)`))))+
  ylab("Relative growth rate")+
  #ggtitle("Pine growth in relation to shrub environment within 2 m")+
  geom_rug(data = df, aes(x = sqrt(shrubarea3), y = exp(VertGrowth_Rel)), alpha = .5, position = "jitter", sides = "b")+
  ylim(c(0.04,0.3))+
  labs(fill = "Juvenile tree height (cm)")+
  theme(legend.position = c(0.8, 0.8))
ggsave(file = "../../results/figures/FireFootprints/PineVertRel.png", width = 6, height = 6, dpi = 400)
```

# Look at relationship between years and shrubarea
```{r}
ggplot(df)+
  geom_boxplot(aes(x = Fire, y = shrubarea3))
ggplot(df)+
  geom_boxplot(aes(x = as.factor(Years), y = shrubarea3))
```

# Find the shrubarea values associated with each quartile, and the mean growth for each of those
```{r}
summary <- summary(sqrt(df$shrubarea3))
quart1 <- summary[2]
quart3 <- summary[5]
median <- summary[3]
min <- summary[1]
max <- summary[6]
mean <- mean(sqrt(df$shrubarea3))

fit_quart1 <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart1)),]$fit
fit_quart1
fit_quart3 <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - quart3)),]$fit
fit_median <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - median)),]$fit
fit_min <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - min)),]$fit
fit_max <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - max)),]$fit
fit_mean <- effects_df[which.min(abs(effects_df$sqrt_shrubarea3 - mean)),]$fit
mean <- mean(df$shrubarea3/10000)
print(paste("fit_quart1 = ", exp(fit_quart1)))
print(paste("fit_quart3 = ", exp(fit_quart3)))
print(paste("fit_median = ", exp(fit_median)))
print(paste("fit_min = ", exp(fit_min)))
print(paste("fit_max = ", exp(fit_max)))


print(paste("fit_quart1_log = ", fit_quart1))
print(paste("fit_quart3_log = ", fit_quart3))
print(paste("fit_median_log = ", fit_median))
print(paste("fit_min_log = ", fit_min))
print(paste("fit_max_log = ", fit_max))
print(paste("fit_mean_log = ", fit_mean))

```

```{r}

sd <- sd(df$shrubarea3/10000)
```


## Save final data
```{r}
save(df, file ="~/../Documents/Shrubs-Seedlings/compiled_data/fire_footprints/pine_vert.Rdata")
```
