---
title: "soil_moisture_analysis"
author: "Carmen"
date: "November 10, 2017"
output: 
    html_document:
      toc: TRUE
---


```{r, include = F}
library(dplyr)
library(readxl)
library(ggplot2)
library(nlme)
library(effects)
library(tidyr)
```

# set wd
```{r}
setwd_SE <- function(){setwd("~/../Dropbox (Stephens Lab)/Shrub_experiment/Data/soil_moisture_data/")}
setwd_SE()
```

# Load data
```{r, warning=FALSE}
setwd_SE()
files <- list.files()
files <- files[grep(".xls", files)]
for(i in 1:length(files)){
  df <- read_excel(files[i], sheet = 1, skip = 2, col_names = T, col_types = c("date", "numeric", "numeric", "numeric", "numeric"), range = cell_cols("A:E"))
  assign(files[i], df)
}
```

# Combine data from the same plot
```{r}
D_180 <- full_join(`180_D_20170803.xls`, `180_D_20171102.xls`)
L_380 <- full_join(`380_L_20170802.xls`, `380_L_20171102.xls`)
F_180 <- full_join(`180_F_20170816.xls`, `180_F_20171102.xls`)
F_180 <- full_join(F_180, `180_F_20170808.xls`)
P_180 <- full_join(`180_P_20170803.xls`, `180_P_20171102.xls`)
Q_180_Control <- full_join(`180_Q_1_Control_20170803.xls`, `180_Q_Control_20171102.xls`)
Q_180_Control <- Q_180_Control[,c(1:3)]
colnames(Q_180_Control) <- c("date", "SM_1", "TEMP_1")
Q_180_Removal <- full_join(`180_Q_Removal_20171102.xls`, `180_Q_2_Removal_20170803.xls`)
Q_180_Removal <- Q_180_Removal[,c(1,3:4)]
colnames(Q_180_Removal) <- c("date","SM_2","TEMP_2")
Q_180 <- full_join(Q_180_Control, Q_180_Removal)
Q_180
```

## Look at duplicates
```{r}
D_180[duplicated(D_180),]
L_380[duplicated(L_380),]
summary(Q_180[duplicated(Q_180),])
P_180[duplicated(P_180),]
F_180[duplicated(F_180),]
```


## Rename columns, take out NA rows, and take out dates before insertion or after study period
```{r}
dfs <- c("D_180", "L_380", "P_180","Q_180", "F_180")
for(i in 1:length(dfs)){
  df <- get(dfs[i])
  colnames(df) <- c("date", "SM_1", "TEMP_1","SM_2","TEMP_2")
    df <- df %>% 
      filter(!is.na(date)) %>% 
      filter(date > "2017-06-11" & date < "2018-01-01")
    assign(paste(dfs[i]), df)
    figure <- ggplot(df)+
      geom_line(aes(x = date, y = SM_1), col = 1)+
      ggtitle(dfs[i])+
      geom_line(aes(x = date, y = SM_2), col = 2)
    print(figure)
}
```

## Take a look at temp -- FINE TUNE THIS LATER
```{r}
ggplot(D_180)+
  geom_line(aes(x = date, y = TEMP_1), col = 3)+
  geom_line(aes(x = date, y = TEMP_2), col = 4)
```

```{r}
islands <- c("L_380", "D_180", "F_180","Q_180", "P_180")
df_full <- data.frame()
for(i in 1:length(islands)){
  df <- get(islands[i])
  df <- gather(df, key = "sensor", value = "datum", c(2:5))
  df <- df %>% 
    mutate(type = ifelse(sensor %in% c("SM_1","SM_2"), "SM", "TEMP")) %>% 
    mutate(sensor = ifelse(sensor %in% c("SM_1","TEMP_1"), "1", "2"))
  df <- spread(df, key = "type", value = "datum")
  df$island <- islands[i]
  df <- df %>% 
    select(island, everything())
  df <- df %>% filter(!is.na(SM))
  # Take out transitional days for F where there's weird stuff happening
  if(islands[i]== "F_180") {
    df <- df %>% 
      filter(!(sensor == "1" & date > "2017-08-15" & date < "2017-08-17")) %>% 
      filter(!(sensor == "2" & date > "2017-08-07" & date < "2017-08-09"))
  }
  figure <- ggplot(df)+
    geom_line(aes(x = date, y = SM, group = sensor, col = sensor))+
    ggtitle(islands[i])
  assign(islands[i], df)
  print(figure)
  df_full <- bind_rows(df, df_full)
}
```

# Add treatment data

## Get plot data
```{r}
setwd_SE()
load("../JOINED_DATA/df_vert_final.Rdata")
df_vert <- df
remove(df)
df_removal <- df_vert %>% 
  select(compartment, island, plot, shrubs, shrub_species) %>% 
  distinct() %>% 
  mutate(island = paste(island, compartment, sep ="_"))
```

## Correct based on notebook
```{r}
df_removal <- df_removal %>% filter(island %in% df_full$island) %>% arrange(island)
df_removal
```

F: plots same as sensors
L: plots same as sensors
Q: plots same as sensors
P: need to be switched
D: need to be switched 

```{r}
df_removal <- df_removal %>% 
  mutate(sensor = ifelse(island == "D_180" & shrubs == "removed", 2, 
                         ifelse(island == "D_180" & shrubs == "retained", 1,
                                ifelse(island == "P_180" & shrubs == "removed", 1,
                                       ifelse(island == "P_180" & shrubs == "retained", 2, 
                                              plot)))))
df_removal
```

## Join
```{r}
df_full$sensor <- as.numeric(df_full$sensor)
df_full <- left_join(df_full, df_removal, by = c("island" = "island", "sensor" = "sensor"))

```

## Plot it all together
```{r}
df_full <- df_full %>% 
  filter(date > "2017-06-12")

ggplot(df_full)+
  geom_line(aes(x = date, y = SM, group = interaction(island, plot), col = shrubs))+
  facet_grid(shrub_species + island ~ .)

```

