---
title: "VSURF_fir"
author: "Carmen"
date: "August 20, 2019"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
require(randomForest)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
set.seed(12346)
```

# Load data
```{r}
setwd("~/../Shrubs-Seedlings/code/FireFootprints_analysis/")
load(file="../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

## Filter to just ABCO
```{r}
df %>% 
  filter(Species == "ABCO") %>% 
  group_by(Fire) %>% 
  summarize(n()) %>% 
  arrange(`n()`)
df <- df %>% 
  filter(Species == "ABCO")
```

# Make df of just the predictor variables I want to include in VSURF
```{r}
xdf <- df %>% 
  dplyr::select(Elevation,
         Slope.Deg,
         heatload,
         incidrad,
         #Aspect.rad.fold.NESW,
         siteclass,
         Year,
         Years,
         BasDia2016.cm, 
         Ht_cm1,
         sqrt_shrubarea1,
         sqrt_shrubarea2,
         sqrt_shrubarea3,
         Ht1,
         Ht1.2,
         Ht1.3,
         Cov1,
         Cov1.2,
         Cov1.3,
         Ht1.3,
         ShrubSpp03) 
```

## Find NA values
```{r}
df[unique (unlist (lapply (xdf, function (x) which (is.na (x))))),] %>% 
  select(Sdlg, 
         Elevation,
         Slope.Deg,
         heatload,
         incidrad,
         #Aspect.rad.fold.NESW,
         siteclass,
         Year,
         Years,
         BasDia2016.cm, 
         Ht_cm1,
         sqrt_shrubarea1,
         sqrt_shrubarea2,
         sqrt_shrubarea3,
         Ht1,
         Ht1.2,
         Ht1.3,
         Cov1,
         Cov1.2,
         Cov1.3,
         ShrubSpp03)
```

## Normalize the numeric ones
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
xdf <- xdf %>% mutate_if(is.numeric, normalize)
xdf
```

## Find NA values
```{r}
unlist (lapply (xdf, function (x) which (is.na (x))))
```

## Find -Inf values
```{r}
df %>% arrange(VertGrowth_Rel) %>% select(Sdlg, VertGrowth_cm, VertGrowth_Rel, Ht_cm1, Ht_cm2)
```

# Apply VSURF to vertical growth
```{r}
vsurf <- VSURF(x = xdf,
               y = df$VertGrowth_Rel,
               na.action = na.omit,
               parallel = T)
summary(vsurf)
vsurf$varselect.pred
xdf[vsurf$varselect.pred]
```

```{r}
vars <- names(xdf[vsurf$varselect.pred])
vars
```

## Save the results
```{r}
save(vars, file = "../../results/VSURF_abco_vars.Rdata")
```

## Look at final random forest
```{r}
forest <- randomForest(
  x = xdf %>%
    dplyr::select(vars),
  y = df$VertGrowth_Rel,
  importance = TRUE,
  ntree = 1000,
  scale = TRUE)
varImpPlot(forest)
```


## Look at random forest of variables after interpretation step
```{r}
vars.interp <- names(xdf[vsurf$varselect.interp])
vars.interp
forest <- randomForest(
  x = xdf %>%
    dplyr::select(vars.interp),
  y = df$VertGrowth_Rel,
  importance = TRUE,
  ntree = 1000,
  scale = TRUE)
varImpPlot(forest)
```


## Look at random forest of variables after interpretation step
```{r}
vars.thres <- names(xdf[vsurf$varselect.thres])
vars.thres
forest <- randomForest(
  x = xdf %>%
    dplyr::select(vars.thres),
  y = df$VertGrowth_Rel,
  importance = TRUE,
  ntree = 1000,
  scale = TRUE)
varImpPlot(forest)
```
