---
title: "Diameter ~ Height"
author: "Carmen"
date: "November 1, 2019"
output: 
  html_document:
    toc: TRUE
---

# This code creates linear models for diameter in relation to height for ABCO and PIPO

```{r}
require(tidyverse)
require(readxl)
require(effects)
```

# Load experiment data
```{r}
setwd("~/Shrubs-Seedlings/code/Simulations/")
load(file = "../../compiled_data/shrub_experiment/shrub_removal_cleaned.Rdata")
load( file = "../../compiled_data/dfshrubs.Rdata")
```

# Add shrub data to df
```{r}
dfshrub <- dfshrub %>% 
  ungroup() %>% 
  mutate(compartment = as.factor(compartment), plot = as.factor(plot))
```

```{r}
df <- full_join(df, dfshrub)
```

## If shrubs = removed, make shrub variables = 0
```{r}
df <- df %>% 
  mutate(Cov1.3 = ifelse(shrubs == "removed", 0, Cov1.3)) %>% 
  mutate(Ht1.3 = ifelse(shrubs == "removed", 0, Ht1.3)) %>% 
  mutate(shrubarea = ifelse(shrubs == "removed", 0, shrubarea))
```

# Look
```{r}
names(df)
ggplot(df, aes(x = dia17_mm_Aug, y = dia18_mm))+
  geom_point()+
  geom_smooth()+
  geom_abline(aes(intercept=0,slope=1))+
  theme_bw()
```

# Convert dia to cm
```{r}
df <- df %>% 
  mutate(dia17_cm_Aug = dia17_mm_Aug/10) %>% 
  mutate(dia18_cm = dia18_mm/10)
```


# Calculate relative diameter growth 
```{r}
df <- df %>% 
  mutate(relgrdia = (dia17_cm_Aug-dia18_cm)/dia17_cm_Aug)
```

# Model dia growth in relation to previous dia and height and vertical growth

## both species together
```{r}
lm <- lm(relgrdia ~ dia17_cm_Aug * height_2018*gr_2018_cm + species + shrubarea , data = df)
summary(lm)
```

```{r}
plot(predictorEffect("dia17_cm_Aug", lm))
plot(predictorEffect("height_2018", lm))
plot(predictorEffect("species", lm))
```

## abco

## pipo

# Model dia in relation to ht and competition, species separate

## ABCO
```{r}
lmabco <- lm(dia18_cm ~ sqrt(shrubarea) + height_2018 * dia17_cm_Aug+ gr_2018_cm, data = df %>% filter(species == "ABCO"))
summary(lmabco)
drop1(lmabco, test = "F")
```

```{r}
plot(predictorEffect("dia17_cm_Aug", lmabco))
plot(predictorEffect("height_2018", lmabco))
plot(predictorEffect("shrubarea", lmabco))
```

## PIPO
```{r}
lmpipo <- lm(dia18_cm ~ sqrt(shrubarea) + height_2018 * dia17_cm_Aug, data = df %>% filter(species == "PIPO"))
summary(lmpipo)
drop1(lmabco, test = "F")
```

```{r}
plot(predictorEffect("dia17_cm_Aug", lmpipo))
plot(predictorEffect("height_2018", lmpipo))
plot(predictorEffect("shrubarea", lmpipo))
```

# Save
```{r}
save(lmabco, file ="../../results/coefficients/LM_dia_abco.Rdata")
save(lmpipo, file ="../../results/coefficients/LM_dia_pipo.Rdata")
```

