---
title: "Shrub_growth_analysis_COVER_EXPLORE"
author: "Carmen"
date: "1/30/2020"
output: 
  html_document:
    toc: TRUE
---



# Load data
```{r, include = F}
require(tidyverse)
require(effects)
require(mgcv)
require(gridExtra)
load(file = "~/Shrubs-Seedlings/compiled_data/shrub_growth/My_FireFootprint_data.Rdata")
me <- ht_by_spp
remove(ht_by_spp)
load(file = "~/Shrubs-Seedlings/compiled_data/shrub_growth/all_combined_COVER_LUMPED.Rdata")
```

```{r}
df <- df %>% ungroup()
```


# Model for each shrub species

## CECO
```{r}
dfCECO <- df %>% filter(Shrub_species == "CECO")
```

### Quadratic
```{r}
ggplot(dfCECO, aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(shape = df_source))+
  geom_smooth(method = "lm", formula = y ~ x + poly(x, 3))+
  theme_bw()
```

```{r}
lmCECO <- lm(Shrub_cover ~ poly(years_since_fire, 2), data = dfCECO)
summary(lmCECO)
dfCECO$pred <- predict(lmCECO)
```

### Take out the one data point in the oldest year
```{r}
dfCECO <- dfCECO %>% 
  filter(years_since_fire < 12.5)
```

### GAM
```{r}
gamCECO <- gam(Shrub_cover ~ s(years_since_fire, k = 3, bs = "cr"), data = dfCECO)
summary(gamCECO)
predCECO <- predict.gam(gamCECO, se =T, type = "response")

dfCECO <- dfCECO %>% 
  mutate(predGAM = predCECO$fit) %>% 
  mutate(predGAMseL = predGAM - predCECO$se.fit) %>% 
  mutate(predGAMseU = predGAM + predCECO$se.fit)
```

```{r}
ggplot(dfCECO)+
  geom_point(aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(x = years_since_fire, y = predGAM), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseL), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseU), col = "purple")
```

### Make fake data to make the plot lines smoother
```{r}
x <- seq(min(dfCECO$years_since_fire), max(dfCECO$years_since_fire), .5)
fake <- data.frame(years_since_fire = x)
preddf = predict.gam(gamCECO, newdata = fake, se =T)
fake <- fake %>% 
  mutate(pred = preddf$fit) %>% 
  mutate(predseL = pred - preddf$se.fit) %>% 
  mutate(predseU = pred + preddf$se.fit)
```

### Pretty plot
```{r}
ceco_plot <- ggplot()+
  geom_point(data = dfCECO, aes(x = years_since_fire, y = Shrub_cover), alpha = .3)+
  geom_line(data = fake, aes(x = years_since_fire, y = pred))+
  geom_ribbon(data = fake, aes(x = years_since_fire, ymin = predseL, ymax = predseU), fill = "royal blue",   alpha = .2)+
  theme_minimal()+
  ggtitle("CECO")
ceco_plot
```

## ARPA/ARVI (Manzanita)
```{r}
dfARPA <- df %>% filter(Shrub_species == "ARPA6_ARVI")
```

```{r}
ggplot(dfARPA, aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(shape = df_source))+
  geom_smooth(method = "lm", formula = y ~ x + poly(x, 3))+
  theme_bw()
```

```{r}
lmARPA <- lm(Shrub_cover ~ poly(years_since_fire, 2), data = dfARPA)
summary(lmARPA)
dfARPA$pred <- predict(lmARPA)
```

```{r}
gamARPA <- gam(Shrub_cover ~ s(years_since_fire, k = 3, bs = "cr"), data = dfARPA)
summary(gamARPA)
predARPA <- predict.gam(gamARPA, se =T, type = "response")
dfARPA <- dfARPA %>% 
  mutate(predGAM = predARPA$fit) %>% 
  mutate(predGAMseL = predGAM - predARPA$se.fit) %>% 
  mutate(predGAMseU = predGAM + predARPA$se.fit)
```

```{r}
ggplot(dfARPA)+
  geom_point(aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(x = years_since_fire, y = predGAM), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseL), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseU), col = "purple")
```

### Make fake data to make the plot lines smoother
```{r}
x <- seq(min(dfARPA$years_since_fire), max(dfARPA$years_since_fire), .5)
fake <- data.frame(years_since_fire = x)
preddf = predict.gam(gamARPA, newdata = fake, se =T)
fake <- fake %>% 
  mutate(pred = preddf$fit) %>% 
  mutate(predseL = pred - preddf$se.fit) %>% 
  mutate(predseU = pred + preddf$se.fit)
```

### Pretty plot
```{r}
arpa_plot <- ggplot()+
  geom_point(data = dfARPA, aes(x = years_since_fire, y = Shrub_cover), alpha = .2)+
  geom_line(data = fake, aes(x = years_since_fire, y = pred))+
  geom_ribbon(data = fake, aes(x = years_since_fire, ymin = predseL, ymax = predseU), fill = "purple",   alpha = .3)+
  theme_minimal()+
  ggtitle("ARPA and ARVI")
arpa_plot
```


## CEIN
```{r}
dfCEIN <- df %>% filter(Shrub_species == "CEIN3")
```

### Plot using polynomial formula in ggplot

```{r}
ggplot(dfCEIN, aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(shape = df_source))+
  geom_smooth(method = "lm", formula = y ~ x + poly(x, 3))+
  theme_bw()
```

```{r}
lmCEIN <- lm(Shrub_cover ~ poly(years_since_fire, 2), data = dfCEIN)
summary(lmCEIN)
dfCEIN$pred <- predict(lmCEIN)
```

```{r}
gamCEIN <- gam(Shrub_cover ~ s(years_since_fire, k = 3, bs = "cr"), data = dfCEIN)
summary(gamCEIN)
predCEIN <- predict.gam(gamCEIN, se =T, type = "response")
dfCEIN <- dfCEIN %>% 
  mutate(predGAM = predCEIN$fit) %>% 
  mutate(predGAMseL = predGAM - predCEIN$se.fit) %>% 
  mutate(predGAMseU = predGAM + predCEIN$se.fit)
```

```{r}
ggplot(dfCEIN)+
  geom_point(aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(x = years_since_fire, y = predGAM), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseL), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseU), col = "purple")
```

### Make fake data to make the plot lines smoother
```{r}
x <- seq(min(dfCEIN$years_since_fire), max(dfCEIN$years_since_fire), .5)
fake <- data.frame(years_since_fire = x)
preddf = predict.gam(gamCEIN, newdata = fake, se =T)
fake <- fake %>% 
  mutate(pred = preddf$fit) %>% 
  mutate(predseL = pred - preddf$se.fit) %>% 
  mutate(predseU = pred + preddf$se.fit)
```

### Pretty plot
```{r}
cein_plot <- ggplot()+
  geom_point(data = dfCEIN, aes(x = years_since_fire, y = Shrub_cover), alpha = .2)+
  geom_line(data = fake, aes(x = years_since_fire, y = pred))+
  geom_ribbon(data = fake, aes(x = years_since_fire, ymin = predseL, ymax = predseU), fill = 2,   alpha = .3)+
  theme_minimal()+
  ggtitle("CEIN")
cein_plot  
```


## OTHER

### Filter to just shrub species that are not one of the species above
```{r}
dfOTHER <- df %>% 
   filter(!Shrub_species %in% c("ARPA6_ARVI", "CECO", "CEIN"))
```

```{r}
gamOTHER <- gam(Shrub_cover ~ s(years_since_fire, k = 3, bs = "cr"), data = dfOTHER)
summary(gamOTHER)
predOTHER <- predict.gam(gamOTHER, se =T, type = "response")
dfOTHER <- dfOTHER %>% 
  mutate(predGAM = predOTHER$fit) %>% 
  mutate(predGAMseL = predGAM - predOTHER$se.fit) %>% 
  mutate(predGAMseU = predGAM + predOTHER$se.fit)
```

### Make fake data to make the plot lines smoother
```{r}
x <- seq(min(dfOTHER$years_since_fire), max(dfOTHER$years_since_fire), .5)
fake <- data.frame(years_since_fire = x)
preddf = predict.gam(gamOTHER, newdata = fake, se =T)
fake <- fake %>% 
  mutate(pred = preddf$fit) %>% 
  mutate(predseL = pred - preddf$se.fit) %>% 
  mutate(predseU = pred + preddf$se.fit)
```

### Pretty plot
```{r}
OTHER_plot <- ggplot()+
  geom_point(data = dfOTHER, aes(x = years_since_fire, y = Shrub_cover), alpha = .2)+
  geom_line(data = fake, aes(x = years_since_fire, y = pred))+
  geom_ribbon(data = fake, aes(x = years_since_fire, ymin = predseL, ymax = predseU), fill =11,   alpha = .3)+
  theme_minimal()+
  ggtitle("OTHER")
OTHER_plot
```

# Model for everything together

## GAM
```{r}
gamALL <- gam(Shrub_cover ~ s(years_since_fire, k = 3, bs = "cr"), data = df)
summary(gamALL)
predALL <- predict.gam(gamALL, se =T, type = "response")

df <- df %>% 
  ungroup() %>% 
  mutate(predGAM = predALL$fit) %>% 
  mutate(predGAMseL = predGAM - predALL$se.fit) %>% 
  mutate(predGAMseU = predGAM + predALL$se.fit)
```

```{r}
ggplot(df)+
  geom_point(aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(x = years_since_fire, y = predGAM), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseL), col = "purple")+
  geom_line(aes(x = years_since_fire, y = predGAMseU), col = "purple")
```

### Make fake data to make the plot lines smoother
```{r}
x <- seq(min(df$years_since_fire), max(df$years_since_fire), .5)
fake <- data.frame(years_since_fire = x)
preddf = predict.gam(gamALL, newdata = fake, se =T)
fake <- fake %>% 
  mutate(pred = preddf$fit) %>% 
  mutate(predseL = pred - preddf$se.fit) %>% 
  mutate(predseU = pred + preddf$se.fit)
```

### Pretty plot
```{r}
ALL_plot <- ggplot()+
  geom_point(data = df, aes(x = years_since_fire, y = Shrub_cover), alpha = .3)+
  geom_line(data = fake, aes(x = years_since_fire, y = pred))+
  geom_ribbon(data = fake, aes(x = years_since_fire, ymin = predseL, ymax = predseU), fill = "royal blue",   alpha = .2)+
  theme_minimal()+
  ggtitle("ALL SPECIES")
ALL_plot
```

# Plot the 4 species together
```{r}
grid.arrange(ceco_plot, arpa_plot, cein_plot, OTHER_plot, ALL_plot)
final_plot <- arrangeGrob(ceco_plot, arpa_plot, cein_plot, OTHER_plot, ALL_plot, ncol = 3, nrow = 2)
ggsave("~/Shrubs-Seedlings/results/figures/Shrub_growth_time/gam_4_spp_COVER_LUMPED.png", final_plot, width = 10, height = 6)
```


# Changes in shrub cover for MY fire footprints

## Join ARPA and ARVI
```{r}
me <- me %>% 
  mutate(Shrub_species = as.character(Shrub_species)) %>% 
  mutate(Shrub_species = case_when(
    Shrub_species == "ARPA" ~ "ARPA-ARVI",
    Shrub_species == "ARVI" ~ "ARPA-ARVI",
    TRUE ~ Shrub_species
  ))
```

## Plot summed cover by plot 

### Summarize
```{r}
me_totals <- me %>% 
  group_by(Sdlg, Fire, Year_meas, years_since_fire) %>% 
  summarize(cover_sum = sum(cover), ht_wt_mean_cm = weighted.mean(modal_ht_cm, cover))
```

### Plot
```{r}
ggplot(me_totals, aes(x = years_since_fire, y = cover_sum))+
  geom_point()+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cr"))
```

## Plot summed cover by plot grouped by dominant shrub species

### Summarize
```{r}
spp <- me %>% 
  group_by(Sdlg) %>% 
  filter(cover == max(cover)) %>% 
  dplyr::select(Sdlg, Shrub_species)
```

```{r}
me_totals <- full_join(me_totals, spp)
```

```{r}
me_totals <- me_totals %>% 
  mutate(cover_percent = cover_sum/(300*4)*100)
```

### Remove the one data point for ARPA-ARVI in year 11-12
```{r}
me_totals %>% 
  filter(Shrub_species == "ARPA-ARVI") %>% 
  group_by(years_since_fire) %>% 
  count()
```

```{r}
me_totals <- me_totals %>% 
  filter(!(Shrub_species == "ARPA-ARVI" & years_since_fire == 12))
```

### Plot
```{r}
CEINplot <- ggplot(me_totals %>% filter(Shrub_species == "CEIN"), aes(x = years_since_fire, y = cover_percent))+
  geom_point(alpha = .3)+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3, bs = "cr"), fill = 2, alpha = .2, col = "black")+
  theme_minimal()+
  ggtitle("CEIN")+
  xlim(8, 44)+
  ylim(0, 120)
CEINplot
```

```{r}
CECOplot <- ggplot(me_totals %>% filter(Shrub_species == "CECO"), aes(x = years_since_fire, y = cover_percent))+
  geom_point(alpha = .3)+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3, bs = "cr"), fill = "royal blue", alpha = .2, col = "black")+
  theme_minimal()+
  ggtitle("CECO")+
  xlim(8, 44)+
  ylim(0, 120)

CECOplot
```


```{r}
ARPAplot <- ggplot(me_totals %>% filter(Shrub_species == "ARPA-ARVI"), aes(x = years_since_fire, y = cover_percent))+
  geom_point(alpha = .3)+
  geom_smooth(method = "loess", fill = "purple", alpha = .2, col = "black")+
  theme_minimal()+
  ggtitle("ARPA")+
  xlim(8, 44)+
  ylim(0, 120)

ARPAplot
```

```{r}
OTHERplot <- ggplot(me_totals %>% filter(!Shrub_species %in% c("CECO", "CEIN", "ARPA-ARVI")), 
                   aes(x = years_since_fire, y = cover_percent))+
  geom_point(alpha = .3)+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3, bs = "cr"), fill = 11, alpha = .2, col = "black")+
  theme_minimal()+
  ggtitle("OTHER")+
  xlim(8, 44)+
  ylim(0, 120)

OTHERplot
```


```{r}
ALLplot <- ggplot(me_totals,   aes(x = years_since_fire, y = cover_percent))+
  geom_point(alpha = .3)+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3, bs = "cr"), fill = "grey", alpha = .5, col = "black")+
  theme_minimal()+
  ggtitle("ALL")+
  xlim(8, 44)+
  ylim(0, 120)

ALLplot
```


# Plot the species together
```{r}
grid.arrange(CECOplot, ARPAplot, CEINplot, OTHERplot, ALLplot, ncol = 2)
final_plot <- arrangeGrob(CECOplot, ARPAplot, CEINplot, OTHERplot, ALLplot, ncol = 3, nrow = 2)
ggsave("~/Shrubs-Seedlings/results/figures/Shrub_growth_time/cover_mydata_grid.png", final_plot, width = 10, height = 6)
```
