---
title: "Diameter ~ Height for fire footprints"
author: "Carmen"
date: "November 1, 2019"
output: 
  html_document:
    toc: TRUE
---

# This code creates linear models for diameter in relation to height for ABCO and PIPO based on fire footprint data

```{r}
require(tidyverse)
require(readxl)
require(effects)
require(nlme)
require(MuMIn)
```
#-

# PIPO ONLY

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/Simulations/parameter_dev/")
load(file="../../../compiled_data/fire_footprints/pine_vert.Rdata")
```


#- 
# CLEAN

# Filter to 2017
```{r}
df <- df %>% 
  filter(Year ==2017)
```


# Calculate average 2017 diameter

## Make numeric
```{r}
df <- df %>% 
  mutate(BasDia2017.mm.1 = as.numeric(paste(BasDia2017.mm.1))) %>% 
  mutate(BasDia2017.mm.2 = as.numeric(paste(BasDia2017.mm.2))) 
```

## Look at NA dia
```{r}
df %>% 
  filter(is.na(BasDia2017.mm.1)) %>% 
  dplyr::select(Sdlg, Species, BasDia2016.cm, BasDia2017.mm.1, BasDia2017.mm.2, return, Years, Year)
```

## Take out rows that don't have 2017 data, AKA no 2017 diameter
```{r}
df <- df %>% 
  filter(!is.na(BasDia2017.mm.1))
```

## Calculate average 2017 dia in cm
```{r}
df <- df %>% 
  mutate(BasDia2017.cm = ((BasDia2017.mm.1 + BasDia2017.mm.2)/2)/10)
```

## Check
```{r}
df %>% 
  dplyr::select(BasDia2016.cm, BasDia2017.cm, BasDia2017.mm.1, BasDia2017.mm.2)
```

# Look at ht:dia relationships
```{r}
ggplot(df, aes(x = Ht_cm1, y = BasDia2017.cm, col = Species))+
  geom_point()+
  geom_smooth()
```

# Look at trees whose ht:dia ratio seems off
```{r}
df %>% 
  filter(BasDia2017.cm<2 & Ht_cm1>100) %>% 
  dplyr::select(Sdlg, Ht_cm1, return, Species, BasDia2016.cm, BasDia2017.cm, BasDia2017.mm.1, BasDia2017.mm.2, Year)
```

# Fix -- clearly it's 8.7 cm, not mm
```{r}
df <- df %>% 
  mutate(BasDia2017.cm = ifelse(Sdlg == 290, 8.7, BasDia2017.cm))
```

# Plot range of dia and ht in these data

## diameter
```{r}
ggplot(df)+
  geom_freqpoly(aes(BasDia2017.cm,stat(density)))+
  theme_minimal()+
  xlab("Diameter (cm)")+
  theme(text = element_text(size = 20))
```

## height

```{r}
ggplot(df)+
  geom_freqpoly(aes(Ht_cm1, stat(density)))+
  theme_minimal()+
  xlab("Height (cm)")+
  theme(text = element_text(size = 20))
```

## Ratio
```{r}
ggplot(df)+
  geom_freqpoly(aes(BasDia2017.cm, stat(density)))+
  theme_minimal()+
  xlab("Diameter:Height (cm:cm)")+
  theme(text = element_text(size = 20))
```

# MODEL dia growth in relation to previous dia and height
```{r}
PIPO1 <- lm(BasDia2017.cm ~ Ht_cm1 + sqrt_shrubarea3, data = df)
summary(PIPO1)
```

# Do AICc model selection
```{r}
AICcmodavg::AICc(PIPO1)
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 + sqrt_shrubarea3 + Years, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 + sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 + ShrubSpp03, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 + ShrubSpp03 + VertGrowth_cm, data = df))
```

# Select best model
```{r}
PIPO_final <- lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3, data = df)
summary(PIPO_final)
```

# Plot predictor effects
```{r}
plot(predictorEffect("sqrt_shrubarea3", PIPO_final))
plot(predictorEffect("Ht_cm1", PIPO_final))
```

# Test residual standard error meaning to make sure it's the same as root mean square error = root mean square deviation
```{r}
df$predict_pipo_final <- predict(PIPO_final)
df %>% 
  mutate(residual = BasDia2017.cm - predict_pipo_final) %>% 
  mutate(square_resid = residual^2) %>% 
  summarize(sum_resid = sum(square_resid)) %>% 
  mutate(mean_resid = sum_resid/nrow(df)) %>% 
  mutate(RMSE = sqrt(mean_resid))
```

# Bootstrap 
```{r}
coef_all <- data.frame()
models_all <- list()
model <- PIPO_final
for(i in 1:1000){
  bootdata <- df[sample(nrow(df), nrow(df), replace=TRUE), ]
  bootmodel <- try(update(model, data = bootdata), TRUE)
  if(isTRUE(class(bootmodel)=="try-error")) { next } 
  coef <- as.data.frame(bootmodel$coefficients) %>% 
    mutate(coef = row.names(as.data.frame(bootmodel$coefficients) )) %>% 
    mutate(i = i)
  coef_all <- rbind(coef, coef_all)
  models_all <- c(models_all, list(bootmodel))

}
coef_all <- coef_all %>% 
  rename(value = `bootmodel$coefficients`)
head(coef_all)
head(models_all)
```

# Plot bootstrap results
```{r}
coef_all %>% 
  filter(coef != "(Intercept)") %>% 
  group_by(coef) %>% 
  summarize(mean = mean(value), sd = sd(value)) %>% 
  mutate(coef = fct_reorder(coef, desc(mean))) %>% 
  ggplot()+
  geom_point(aes(x = mean, y = coef))+
  theme_minimal()+
  xlab("Mixed effects model coefficient")+
  ylab(element_blank())+
  geom_errorbarh(aes(y = coef, xmin = mean-sd, xmax = mean+sd), height = .2)+
  geom_vline(aes(xintercept = 0), linetype = 2)+
  xlim(c(-.03, 0.06))+
  scale_y_discrete(labels = c("Juvenile tree height", 
                              "Height/shrub interaction", 
                              "Shrub competition"))+
  theme(panel.grid = element_blank(),
        text = element_text(size = 12))

ggsave(file = "../../../results/figures/Simulations/Pine_dia_coefficients_Supp.png", width = 6, height = 2, dpi = 400)
```

# Save
```{r}
save(PIPO_final, file ="../../../results/coefficients/LM_dia_PIPO_footprints.Rdata")
save(coef_all, file = "../../../results/coefficients/LM_dia_PIPO_coef_all.Rdata")
save(models_all, file = "../../../results/coefficients/LM_dia_PIPO_models_all.Rdata")
```

#-
# ABCO ONLY

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/Simulations/parameter_dev/")
load(file="../../../compiled_data/fire_footprints/fir_vert.Rdata")
```


#- 
# CLEAN

# Filter to 2017
```{r}
df <- df %>% 
  filter(Year ==2017)
```


# Calculate average 2017 diameter

## Make numeric
```{r}
df <- df %>% 
  mutate(BasDia2017.mm.1 = as.numeric(paste(BasDia2017.mm.1))) %>% 
  mutate(BasDia2017.mm.2 = as.numeric(paste(BasDia2017.mm.2))) 
```

## Look at NA dia
```{r}
df %>% 
  filter(is.na(BasDia2017.mm.1)) %>% 
  dplyr::select(Sdlg, Species, BasDia2016.cm, BasDia2017.mm.1, BasDia2017.mm.2, return, Years, Year)
```

## Calculate average 2017 dia in cm
```{r}
df <- df %>% 
  mutate(BasDia2017.cm = ((BasDia2017.mm.1 + BasDia2017.mm.2)/2)/10)
```

## Check
```{r}
df %>% 
  dplyr::select(BasDia2016.cm, BasDia2017.cm, BasDia2017.mm.1, BasDia2017.mm.2)
```

# Look at ht:dia relationships
```{r}
ggplot(df, aes(x = Ht_cm1, y = BasDia2017.cm, col = Species))+
  geom_point()+
  geom_smooth()
```


# PLOT range of dia and ht in these data

## diameter
```{r}
ggplot(df)+
  geom_freqpoly(aes(BasDia2017.cm,stat(density)))+
  theme_minimal()+
  xlab("Diameter (cm)")+
  theme(text = element_text(size = 20))
```

## height

```{r}
ggplot(df)+
  geom_freqpoly(aes(Ht_cm1, stat(density)))+
  theme_minimal()+
  xlab("Height (cm)")+
  theme(text = element_text(size = 20))
```

## Ratio
```{r}
ggplot(df)+
  geom_freqpoly(aes(BasDia2017.cm, stat(density)))+
  theme_minimal()+
  xlab("Diameter:Height (cm:cm)")+
  theme(text = element_text(size = 20))
```


# MODEL dia growth in relation to previous dia and height for abco
```{r}
ABCO1 <- lm(BasDia2017.cm ~ Ht_cm1 + sqrt_shrubarea3, data = df)
summary(ABCO1)
plot(predictorEffect("sqrt_shrubarea3", ABCO1))
```

# Do AIC model selection
```{r}
AICcmodavg::AICc(ABCO1)
AICcmodavg::AICc(lm(BasDia2017.cm ~ sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 + ShrubSpp03, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 + ShrubSpp03*sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 + ShrubSpp03* sqrt_shrubarea3, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 + ShrubSpp03* Ht_cm1, data = df))
AICcmodavg::AICc(lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3 * ShrubSpp03 , data = df))
```

# Select best model
```{r}
ABCO_final <- lm(BasDia2017.cm ~ Ht_cm1 * sqrt_shrubarea3, data = df)
summary(ABCO1)
drop1(ABCO_final, test = "F")
```

NOTE: I didn't select the model with the lowest AIC for the sake of keeping things simple and the same as the PIPO model. 

# Test residual standard error meaning to make sure it's the same as root mean square error = root mean square deviation
```{r}
df$predict_ABCO_final <- predict(ABCO_final)
df %>% 
  mutate(residual = BasDia2017.cm - predict_ABCO_final) %>% 
  mutate(square_resid = residual^2) %>% 
  summarize(sum_resid = sum(square_resid)) %>% 
  mutate(mean_resid = sum_resid/(nrow(df)-4)) %>% 
  mutate(RMSE = sqrt(mean_resid))
summary(ABCO_final)
sigma(ABCO_final)
```

# Plot predictor effects
```{r}
plot(predictorEffect("sqrt_shrubarea3", ABCO_final))
plot(predictorEffect("Ht_cm1", ABCO_final))
```

# Bootstrap 
```{r}
remove(coef_all)
coef_all <- data.frame()
models_all <- list()
model <- ABCO_final
for(i in 1:1000){
  bootdata <- df[sample(nrow(df), nrow(df), replace=TRUE), ]
  bootmodel <- try(update(model, data = bootdata), TRUE)
  if(isTRUE(class(bootmodel)=="try-error")) { next } 
  coef <- as.data.frame(bootmodel$coefficients) %>% 
    mutate(coef = row.names(as.data.frame(bootmodel$coefficients) )) %>% 
    mutate(i = i)
  coef_all <- rbind(coef, coef_all)
  models_all <- c(models_all, list(bootmodel))
}
coef_all <- coef_all %>% 
  rename(value = `bootmodel$coefficients`)
head(coef_all)
head(models_all)
```

# Plot bootstrap results
```{r}
coef_all %>% 
  filter(coef != "(Intercept)") %>% 
  group_by(coef) %>% 
  summarize(mean = mean(value), sd = sd(value)) %>% 
  mutate(coef = fct_reorder(coef, desc(mean))) %>% 
  ggplot()+
  geom_point(aes(x = mean, y = coef))+
  theme_minimal()+
  xlab("Mixed effects model coefficient")+
  ylab(element_blank())+
  geom_errorbarh(aes(y = coef, xmin = mean-sd, xmax = mean+sd), height = .2)+
  geom_vline(aes(xintercept = 0), linetype = 2)+
  xlim(c(-.03, 0.06))+
  scale_y_discrete(labels = c("Juvenile tree height", 
                              "Height/shrub interaction", 
                              "Shrub competition"))+
  theme(panel.grid = element_blank(),
        text = element_text(size = 12))

ggsave(file = "../../../results/figures/Simulations/Abco_dia_coefficients_Supp.png", width = 6, height = 2, dpi = 400)
```


# Save
```{r}
save(ABCO_final, file ="../../../results/coefficients/LM_dia_ABCO_footprints.Rdata")
save(coef_all, file = "../../../results/coefficients/LM_dia_ABCO_coef_all.Rdata")
save(models_all, file = "../../../results/coefficients/LM_dia_ABCO_models_all.Rdata")
```

