---
title: "simulations"
author: "Carmen"
date: "May 16, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
require(MASS)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/Simulations/parameter_dev/")
load(file = "../../../compiled_data/growth_mortality/df_vert.Rdata")
```

#  Make dead a factor
```{r}
df <- df %>% 
  mutate(DEAD_ALIVE = as.factor(DEAD_ALIVE))
```

# Make separate df for each species
```{r}
dfa <- df %>% filter(SPECIES == "ABCO")
dfp <- df %>% filter(SPECIES == "PIPO")
```


#  _
# ABCO SIMULATIONS

## Visually compare distributions
```{r}
means <- dfa %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfa %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(sd = sd(log_gr_rel_ave))
means
sd
```

### Create smooth PDF for live and dead

#### Dead
```{r}
x <- seq(-4, 0, length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

#### Live
```{r}
x <- seq(-4, 0, length=1000)
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

#### Plot PDF with histogram of growth values
```{r}
ggplot()+
  geom_histogram(data = dfa %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfa %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")

```

# Actual simulations for ABCO

## Count abco n in the POPULATION for live and dead
```{r}
n_dead_pop <- unlist(dfa %>% 
                   filter(DEAD_ALIVE=="DEAD") %>% 
                   summarize(n()))
n_dead_pop
```

## Come up with # of live seedlings in simulated population by multiplying the dead ones by the ratio of live:dead from surveys
```{r}
n_alive_pop <- round(n_dead_pop*(1103/25))
```

## Count abco n in MEASURED TREES for live and dead
```{r}
n_alive <- nrow(dfa %>% filter(DEAD_ALIVE == "ALIVE"))
n_dead <- nrow(dfa %>% filter(DEAD_ALIVE == "DEAD"))
```

## Find the sets of live and dead growth rates to draw from
```{r}
gr_set_dead <- dfa %>% 
    filter(DEAD_ALIVE=="DEAD") %>% 
    dplyr::select(log_gr_rel_ave)
gr_set_dead <- gr_set_dead[[1]]

gr_set_alive <- dfa %>% 
  filter(DEAD_ALIVE=="ALIVE") %>%
  dplyr::select(log_gr_rel_ave)
gr_set_alive <- gr_set_alive[[1]]
```

## Loop the simulation 1000 times
```{r}
gr_coef  <- c()
int_coef  <- c()
gr_se <- c()

for(i in 1:1000){
  dead_sample <- sample(gr_set_dead, size = n_dead_pop, replace = TRUE) 
  alive_sample <- sample(gr_set_alive, size = n_alive_pop, replace = TRUE)
  live_df <- data.frame(gr = c(alive_sample))
  live_df$dead <- 0
  dead_df <- data.frame(gr = c(dead_sample))
  dead_df$dead <- 1
  df <- bind_rows(dead_df, live_df)
  glm <- glm(dead ~ gr, family = binomial(link = "logit"), data = df)
  gr_coef <- c(gr_coef , glm$coefficient[2])
  int_coef <- c(int_coef , glm$coefficient[1])
  gr_se <- c(gr_se, summary(glm)$coefficients["gr","Std. Error"])
  all_coefficients <- data.frame(gr_coef, int_coef, gr_se)
}
mean_coef_gr <- mean(gr_coef)
mean_coef_int <- mean(int_coef)
head(all_coefficients)
```

## Save
```{r}
save(mean_coef_gr, file = "../../../results/coefficients/mean_coef_gr_abco.Rdata")
save(mean_coef_int, file = "../../../results/coefficients/mean_coef_int_abco.Rdata")
save(all_coefficients, file = "../../../results/coefficients/gr_mort_all_coefficients_abco.Rdata")
```

# _
# PIPO SIMULATIONS

## Visually compare distributions
```{r}
means <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(sd = sd(log_gr_rel_ave))
means
sd
```

```{r}
x <- seq(-4, 0, length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

```{r}
x <- seq(-4, 0, length=1000)
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

```{r}
ggplot()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")
  
ggsave("../../../results/figures/GrowthMortality/histogram_dead_alive_pipo.png")
```

# Actual simulations for PIPO

## Count pipo n for live and dead
```{r}
n_dead_pop <- unlist(dfp %>% 
                   filter(DEAD_ALIVE=="DEAD") %>% 
                   summarize(n()))
```

```{r}
n_alive_pop <- n_dead_pop*(524/6)
```
    
## Count pipo n in MEASURED TREES for live and dead
```{r}
n_alive <- nrow(dfp %>% filter(DEAD_ALIVE == "ALIVE"))
n_dead <- nrow(dfp %>% filter(DEAD_ALIVE == "DEAD"))
```

## Find the sets of live and dead growth rates to draw from
```{r}
gr_set_dead <- dfp %>% 
    filter(DEAD_ALIVE=="DEAD") %>% 
    dplyr::select(log_gr_rel_ave)
gr_set_dead <- gr_set_dead[[1]]

gr_set_alive <- dfp %>% 
  filter(DEAD_ALIVE=="ALIVE") %>%
  dplyr::select(log_gr_rel_ave)
gr_set_alive <- gr_set_alive[[1]]
```

## Loop it 
```{r}
gr_coef  <- c()
int_coef  <- c()
gr_se <- c()

for(i in 1:1000){
  dead_sample <- sample(gr_set_dead, size = n_dead_pop, replace = TRUE) 
  alive_sample <- sample(gr_set_alive, size = n_alive_pop, replace = TRUE)
  live_df <- data.frame(gr = c(alive_sample))
  live_df$dead <- 0
  dead_df <- data.frame(gr = c(dead_sample))
  dead_df$dead <- 1
  df <- bind_rows(dead_df, live_df)
  glm <- glm(dead ~ gr, family = binomial(link = "logit"), data = df)
  gr_coef <- c(gr_coef , glm$coefficient[2])
  int_coef <- c(int_coef , glm$coefficient[1])
  gr_se <- c(gr_se, summary(glm)$coefficients["gr","Std. Error"])
  all_coefficients <- data.frame(gr_coef, int_coef, gr_se)
}
mean_coef_gr <- mean(gr_coef)
mean_coef_int <- mean(int_coef)
head(all_coefficients)
```

```{r}
se_list <- tbl_df(gr_se) 

for(i in 1:nrow(se_list)){
  se_list$rand_obvs[i] <- rnorm(1, mean(gr_coef), se_list$value[i])
}
  
se_list
```

# Find distribution of the coefficient for growth
```{r}
sd(gr_coef)
hist(rnorm(1000,mean(gr_coef), sd(gr_coef)), breaks = 20)
ggplot()+
  geom_histogram(aes(x = gr_coef), alpha = .2, fill = "green", bins =20)+
  theme_minimal()+
  geom_histogram(aes(x = rnorm(1000,mean(gr_coef), sd(gr_coef))), alpha = .2, fill = "pink", bins =20)+
  geom_histogram(aes(x = rnorm(1000, mean(gr_coef), mean(gr_se))), alpha = .2, fill = "blue", bins =20)+
  geom_histogram(aes(x = se_list$rand_obvs), alpha = .2, fill = "orange", bins =20)
```

THIS HISTOGRAM SHOWS: the coefficient estimate varies with a normal distribution with a standard deviation equal to the standard deviation of the list of coefficients from the simulations, which is also equal to the mean standard error or randomly sampled standard error split out by the glm models.

Therefore, to account for uncertainty in my model, I can take the standard deviation of the list of 1000 coefficients from the simulation runs and draw from that distribution, or just randomly draw from the simulation runs, which would also allow me to link the growth coefficient with the intercept coefficient.

## Save
```{r}
save(mean_coef_gr, file = "../../../results/coefficients/mean_coef_gr_pipo.Rdata")
save(mean_coef_int, file = "../../../results/coefficients/mean_coef_int_pipo.Rdata")
save(all_coefficients, file = "../../../results/coefficients/gr_mort_all_coefficients_pipo.Rdata")
```
