---
title: 'Pinus with shrubarea vertical analysis - fire footprints'
author: "Carmen"
date: "March 9 + 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, include = F}
require(nlme)
require(randomForest)
require(tree)
require(knitr)
require(tidyverse)
require(ggplot2)
require(VSURF)
require(effects)
require(MuMIn)
require(car)
set.seed(12346)
```

# Load data
```{r}
setwd("~/../Shrubs-Seedlings/code/Simulations/parameter_dev/")
load(file="../../../compiled_data/fire_footprints/master_seedlings_vert.Rdata")
```

# Load VSURF results
```{r}
load("../../../results/VSURF_pipo_vars_sim.Rdata")
vars
```

# Filter to just PIPO
```{r}
df %>% 
  filter(Species == "PIPO") %>% 
  group_by(Fire) %>% 
  summarize(n()) %>% 
  arrange(`n()`)
df <- df %>% 
  filter(Species == "PIPO")
```

# Count totals and by fire
```{r}
length(unique(df$Sdlg))
nrow(df)
```

```{r}
df %>% 
  group_by(Sdlg, Fire) %>%
  dplyr::select(Sdlg, Fire) %>%
  distinct() %>% 
  ungroup() %>% 
  group_by(Fire) %>% 
  summarize(n = n()) %>% 
  mutate(total = sum(n))
```

# Find average growth rate
```{r}
mean(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)
sd(df$VertGrowth_Rel)/sqrt(nrow(df))
ggplot(df)+
  geom_boxplot(aes(y = VertGrowth_Rel))
```

# Structure data for LM

## Normalize function
```{r}
normalize <- function(x) {
    return ((x - mean(x)) / sd(x))
  }
```

## filter and normalize df
```{r}
LMdf <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  mutate(log_shrubarea3 = log(shrubarea3+1)) %>%
  dplyr::select(Fire, 
        FirePatch,
         Sdlg, 
         VertGrowth_Rel, 
         Years, 
         Ht_cm1, 
         sqrt_shrubarea3, 
         log_shrubarea3, 
         BasDia2016.cm, 
         incidrad, 
         Year, 
         Elevation, 
         siteclass,
         heatload, 
         ShrubSpp03) %>% 
 mutate_if(is.numeric, normalize) %>% 
  mutate(VertGrowth_Rel = df$VertGrowth_Rel)
LMdf <- droplevels(LMdf)
```

# Model using VSURF var results

## Convert vars to formula structure
```{r}
vars_one <- paste(vars, collapse = " + ")
vars_one
f <- formula(paste("VertGrowth_Rel ~ ",vars_one))
f
```

## Run lme model
```{r}
LM <- lme(f, data = LMdf, random = ~ 1| Fire/FirePatch/Sdlg, method = "ML")
summary(LM)
r.squaredGLMM(LM)
```

# Check collinearity

## Calculate vif 
```{r}
vif(LM)
```

# See whether model with or without HEIGHT interaction is better
```{r}
vars_int <- paste(vars_one, "+ Ht_cm1:sqrt_shrubarea3")
f_int <- formula(paste("VertGrowth_Rel ~ ",vars_int))
LM_int <- lme(f_int, data = LMdf, random = ~ 1| Fire/FirePatch/Sdlg, method = "ML")
AICcmodavg::AICc(LM)
AICcmodavg::AICc(LM_int)
if(
  AICcmodavg::AICc(LM) > AICcmodavg::AICc(LM_int)
) print(paste("WITH INTERACTION WINS BY", round(AICcmodavg::AICc(LM)-AICcmodavg::AICc(LM_int), digits = 2), "AIC")) else print(paste("WITHOUT INTERACTION WINS BY", round(AICcmodavg::AICc(LM_int)-AICcmodavg::AICc(LM), digits = 2)))
```

Do include an interaction because it's better for simulation purposes and AICc are very similar with and without interaction

```{r}
LM <- LM_int
```

## Summary
```{r}
summary(LM)
```


# See whether residuals are better for log transformed or not
```{r}
plot(LM)
```

```{r}
f_no_log <- formula(paste("exp(VertGrowth_Rel) ~", vars_one))
LM_no_log <-  lme(f_no_log, data = LMdf, random = ~ 1| Fire/Sdlg, method = "ML")
plot(LM_no_log)
```

Log transformed looks better.

# Look at collinearity between seedling height and shrub competition
```{r}
cor(df$Ht_cm1, df$sqrt_shrubarea3)
ggplot(df, aes(x = Ht_cm1, y = sqrt_shrubarea3))+
  geom_point()+
  geom_smooth(method = "lm")
```

Not too bad.

# Look at residuals for validation

## Residuals
```{r}
E <- resid(LM, type = "normalized")
df$E <- E
pred <- predict(LM)
plot(LM)
plot(df$shrubarea3, E)
plot(df$Ht_cm1, E)
plot(df$Years, E)
plot(df$Ht1.3, E)
plot(df$BasDia2016.cm, E)
plot(df$incidrad, E)
plot(df$Year, E)
ggplot(df, aes(x = log(df$shrubarea3), y = E))+
  geom_point()+
  geom_smooth()
```

# Look at individual effects

Note: the plots look exactly the same but with different axes if you don't normalize the numeric predictors. So for including plots to visualize in the paper, I just use non-normalized data

```{r}
plot(predictorEffect("Years", LM))
plot(predictorEffect("sqrt_shrubarea3", LM))
plot(predictorEffect("Ht_cm1", LM))
plot(predictorEffect("ShrubSpp03", LM))
```

# Calculate Nakagawa's pseudo R2
```{r}
r.squaredGLMM(LM)
```

# Plot shrub effects without interaction

## Re-create effects plots using predictorEffects() and ggplot

```{r}
eff_shrubarea <- predictorEffect("sqrt_shrubarea3", LM)
effects_df <- as.data.frame(eff_shrubarea)
head(effects_df)
```

## Transform variables back to original units
```{r}
effects_df <- effects_df %>% 
  mutate(sqrt_shrubarea3 = sqrt_shrubarea3*sd(sqrt(df$shrubarea3))+mean(sqrt(df$shrubarea3)) )
```

# Summarize range of shrub cover and height
```{r}
summary(df$Cov1.3)
summary(df$Ht1.3)
summary(df$shrubarea3)
```

# Convert cover to Percent
```{r}
summary(df$Cov1.3/(300*4))
```

# Pretty plot of shrub species effects
```{r}
eff_shrubspp <- predictorEffect("ShrubSpp03", LM)
effects_df <- as.data.frame(eff_shrubspp)
head(effects_df)
```

```{r}
ggplot(effects_df %>% filter(ShrubSpp03 != "Other"  ))+
  geom_point(aes(x = ShrubSpp03, y = exp(fit)))+
  ylim(.1, .25)+
  ylab("Juvenile pine relative growth rate")+
  theme_bw()+
  geom_errorbar(aes(x = ShrubSpp03, ymin=exp(lower), ymax=exp(upper)), width=.1) +
  scale_x_discrete(labels=c("ARPA" = "Arctostaphylos\npatula",
                            "CECO" = "Ceanothus\ncordulatus",
                            "CEIN" = "Ceanothus\nintegerrimus",
                            "CHFO" = "Chamaebatia\nfoliolosa",
                            "LIDE" = "Notholithocarpus\ndensiflorus"
                            ))+
  theme(  text = element_text(size = 16),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(face = "bold.italic",
                                   angle=45, 
                                   vjust=1, 
                                   hjust=1))+
  xlab("Dominant shrub species")

ggsave(file = "../../../results/figures/FireFootprints/PineVertShrubSpp.png", width = 4, height = 4, dpi = 400)
```

# List the final seedlings used
```{r}
sort(unique(df$Sdlg))
```

# Count how many trees have estimated 2015 heights
```{r}
df %>% 
  filter(Year==2015) %>% 
  filter(is.na(Ht2017.cm.fall) & is.na(Ht2015.meas2016)) %>% 
  nrow()
```

# Save final data
```{r}
save(df, file ="../../../compiled_data/fire_footprints/pine_vert_sim.Rdata")
```

# Save LM with non-normalized data
```{r}
LMdf_nonnorm <- df %>% 
  mutate(sqrt_shrubarea3 = sqrt(shrubarea3)) %>%
  dplyr::select(Fire, 
         FirePatch,
         Sdlg, 
         VertGrowth_Rel, 
         Years, 
         Ht_cm1, 
         sqrt_shrubarea3, 
         BasDia2016.cm, 
         incidrad, 
         Year, 
         Elevation, 
         siteclass,
         heatload, 
         ShrubSpp03) 
LMdf_nonnorm <- droplevels(LMdf_nonnorm)
```

```{r}
f_int
```

```{r}
LM_pine_nonnorm_sim <- lme(f_int, data = LMdf_nonnorm, random = ~ 1| Fire/FirePatch/Sdlg, method = "ML")
```

```{r}
plot(predictorEffect("Years", LM_pine_nonnorm_sim))
```

```{r}
summary(LM_pine_nonnorm_sim)
```


# Calculatel residual standard error NOT within-group, since I don't include grouping in the simulation
```{r}
coefpipo <<- LM_pine_nonnorm_sim$coefficients$fixed

LMdf_nonnorm <- LMdf_nonnorm %>% 
  mutate(pred = coefpipo["(Intercept)"] +
             coefpipo["Years"]*Years+
             coefpipo["Ht_cm1"]*Ht_cm1+
             coefpipo["sqrt_shrubarea3"]*sqrt_shrubarea3+
             coefpipo["heatload"]*heatload+
             coefpipo["Elevation"]*Elevation+
             coefpipo["Ht_cm1:sqrt_shrubarea3"]*sqrt_shrubarea3*Ht_cm1) %>%
    mutate(pred = case_when(
      ShrubSpp03 == "CECO" ~ pred + coefpipo["ShrubSpp03CECO"],
      ShrubSpp03 == "CEIN" ~ pred + coefpipo["ShrubSpp03CEIN"],
      ShrubSpp03 == "CHFO" ~ pred + coefpipo["ShrubSpp03CHFO"],
      ShrubSpp03 == "LIDE" ~ pred + coefpipo["ShrubSpp03LIDE"],
      ShrubSpp03 == "Other" ~ pred + coefpipo["ShrubSpp03Other"],
      TRUE ~ as.numeric(pred)) ) %>% 
    mutate(pred = case_when(
      Year == "2016" ~ pred + coefpipo["Year2016"],
      Year == "2017" ~ pred + coefpipo["Year2017"],
      TRUE ~ as.numeric(pred)) )

ggplot(LMdf_nonnorm, aes(x = VertGrowth_Rel, y = pred, col = heatload))+
  geom_point()+
  geom_abline(aes(intercept = 0, slope = 1))+
  theme_minimal()
```

```{r}
k <- length(LM_pine_nonnorm_sim$coefficients$fixed)-1 #Subtract one to ignore intercept
```


```{r}
RMSE_pine_growth <- LMdf_nonnorm %>% 
  mutate(residual = VertGrowth_Rel - pred) %>% 
  mutate(square_resid = residual^2) %>% 
  summarize(sum_resid = sum(square_resid)) %>% 
  mutate(mean_resid = sum_resid/(nrow(LMdf_nonnorm)-(k+1))) %>% 
  mutate(RMSE = sqrt(mean_resid)) %>% 
  dplyr::select(RMSE)
RMSE_pine_growth
sigma(LM_pine_nonnorm_sim)
```

# Save RMSE
```{r}
save(RMSE_pine_growth, file = "../../../results/coefficients/RMSE_pine_growth.Rdata")
```

# Save final model
```{r}
save(LM_pine_nonnorm_sim, file = "../../../results/coefficients/LM_pine_nonnorm_sim.Rdata")
```

# Bootstrap the model, drawing same number of observations each time with replacement
```{r}
coef_all <- data.frame()
model <- lme(f_int, data = LMdf_nonnorm, random = ~ 1| Fire/FirePatch/Sdlg, method = "ML")
for(i in 1:1000){
  bootdata <- LMdf_nonnorm[sample(nrow(LMdf_nonnorm), nrow(LMdf_nonnorm), replace=TRUE), ]
  bootmodel <- try(update(model, data = bootdata), TRUE)
  if(isTRUE(class(bootmodel)=="try-error")) { next } 
  coef <- as.data.frame(bootmodel$coefficients$fixed) %>% 
    mutate(coef = row.names(as.data.frame(bootmodel$coefficients$fixed) )) %>% 
    mutate(i = i)
  coef_all <- rbind(coef, coef_all)
}
coef_all <- coef_all %>% 
  rename(value = `bootmodel$coefficients$fixed`)
head(coef_all)
coef_all %>% 
  group_by(i) %>% 
  count()
hist(coef_all$value[coef_all$coef=="sqrt_shrubarea3"])
```

## Save
```{r}
save(coef_all, file = "../../../results/coefficients/LM_pine_bootstrap_coef.Rdata")
```


# Repeat bootstrapping with standardized coefficients for figure
```{r}
coef_all_norm <- data.frame()
model <- lme(f_int, data = LMdf, random = ~ 1| Fire/FirePatch/Sdlg, method = "ML")
for(i in 1:1000){
  bootdata <- LMdf[sample(nrow(LMdf), nrow(LMdf), replace=TRUE), ]
  bootmodel <- try(update(model, data = bootdata), TRUE)
  if(isTRUE(class(bootmodel)=="try-error")) { next } 
  coef <- as.data.frame(bootmodel$coefficients$fixed) %>% 
    mutate(coef = row.names(as.data.frame(bootmodel$coefficients$fixed) )) %>% 
    mutate(i = i)
  coef_all_norm <- rbind(coef, coef_all_norm)
}
coef_all_norm <- coef_all_norm %>% 
  rename(value = `bootmodel$coefficients$fixed`)
head(coef_all_norm)
coef_all_norm %>% 
  group_by(i) %>% 
  count()
```

```{r}
coef_all_norm %>% 
  filter(coef != "(Intercept)") %>% 
  group_by(coef) %>% 
  summarize(mean = mean(value), sd = sd(value)) %>% 
  mutate(coef = fct_reorder(coef, desc(mean))) %>% 
  ggplot()+
  geom_point(aes(x = mean, y = coef))+
  theme_minimal()+
  xlab("Mixed effects model coefficient")+
  ylab(element_blank())+
  geom_errorbarh(aes(y = coef, xmin = mean-sd, xmax = mean+sd), height = .2)+
  geom_vline(aes(xintercept = 0), linetype = 2)

```

Make sure labels match the figure below:

```{r}
coef_all_norm %>% 
  filter(coef != "(Intercept)") %>% 
  group_by(coef) %>% 
  summarize(mean = mean(value), sd = sd(value)) %>% 
  mutate(coef = fct_reorder(coef, desc(mean))) %>% 
  ggplot()+
  geom_point(aes(x = mean, y = coef))+
  theme_minimal()+
  xlab("Mixed effects model coefficient")+
  ylab(element_blank())+
  geom_errorbarh(aes(y = coef, xmin = mean-sd, xmax = mean+sd), height = .2)+
  geom_vline(aes(xintercept = 0), linetype = 2)+
  scale_y_discrete(labels = c("Basal diameter", 
                              "Shrub species: deerbrush", 
                              "Shrub species: other", 
                              "Shrub species: whitethorn", 
                              "Year 2016", 
                              "Shrub species: bearclover", 
                              "Height/shrub interaction",
                              "Year 2017", 
                              "Heat load", 
                              "Shrub species: tanoak", 
                              "Shrub competition",
                              "Years since fire",  
                              "Juvenile tree height"))+
  theme(panel.grid = element_blank(),
        text = element_text(size = 12))

ggsave(file = "../../../results/figures/FireFootprints/Pine_coefficients_Supp.png", width = 5, height = 5, dpi = 400)
```

