---
title: "Seedling Establishment - Welch Data - Analysis 1: Shrubs"
author: "Carmen"
date: "March 9, 2017"
output: html_document
---

## Analysis question: how does shrub area relate to time since fire, by shrub species?


## 1. Load data
```{r, include=F}
library(RODBC)
library(dplyr)
library(ggplot2)
```
```{r}
conn <- odbcDriverConnect("Driver={Microsoft Access Driver (*.mdb, *.accdb)};DBQ=C:/Users/Carmen/Dropbox (Stephens Lab)/SORTIE/Data_from_others/Hugh/RegenDatabase_1_13_16_updatedDates.accdb")

```

## 2. Combine shrub and plot data and filter for high severity fire
```{r}
plots <- sqlFetch(conn, "Plot_data")
shrubs <- sqlFetch(conn, "shrub_regen")
plots[is.na(plots)] <- 0
plots$ObsYr <- substr(plots$Date,1,4)
plots$ObsYr <- as.numeric(plots$ObsYr)
plots$Years <- as.factor(plots$ObsYr-plots$`Year of Fire`)
shrubplot <- left_join(shrubs, plots,by="Regen_Plot")
shrubplotHS <- shrubplot %>%
  filter(FIRE_SEV>3) 
shrubplotHS <- subset(shrubplotHS,!is.na(shrubplotHS$modal_ht_cm & !is.na(shrubplotHS$Cover)))
shrubplotHS[is.na(shrubplotHS)] <- 0
```

## 3. Calculate shrub area, cover, and height by plot
```{r}
shrubarea_plot <- shrubplotHS %>%
  group_by(Regen_Plot) %>%
  summarize(shrubarea = sum(Cover*modal_ht_cm))
shrubarea_plot <- inner_join(shrubarea_plot, shrubplotHS,by='Regen_Plot')
shrubarea_plot <- shrubarea_plot%>%
  select(Regen_Plot, shrubarea, Years, Fire, FIRE_SEV) %>%
  distinct()

cover_plot <- shrubplotHS %>%
  group_by(Regen_Plot) %>%
  summarize(cover = sum(Cover))
cover_plot <- inner_join(cover_plot, shrubplotHS,by='Regen_Plot')
cover_plot <- cover_plot%>%
  select(Regen_Plot, cover, Years, Fire, FIRE_SEV) %>%
  distinct()  

height_plot <- shrubplotHS %>%
  group_by(Regen_Plot) %>%
  summarize(height = sum(modal_ht_cm))
height_plot <- inner_join(height_plot, shrubplotHS,by='Regen_Plot')
height_plot <- height_plot%>%
  select(Regen_Plot, height, Years, Fire, FIRE_SEV) %>%
  distinct()  

# Check
shrubarea_plot
head(shrubplotHS[shrubplotHS$Regen_Plot=="AMR1300134",])
 45*35+55*.5
```


## 3. Plot shrub characteristics vs year since fire by plot
```{r}
ggplot(data=cover_plot,aes(Years, cover))+
  geom_violin()+
  stat_summary(geom="point",fun.y="mean",size=3,color="red")+
  labs(title="How does shrub *cover* change with time in initial years after fire?")+
  ylab("Shrub cover by plot")+
  theme(text=element_text(size=20),
        title=element_text(size=12))

ggplot(data=height_plot,aes(Years, height))+
  geom_violin()+
  stat_summary(geom="point",fun.y="mean",size=3,color="red")+
  labs(title="How does shrub *height* change with time in initial years after fire?")+
  ylab("Shrub height by plot")+
  theme(text=element_text(size=20),
        title=element_text(size=12))


ggplot(data=shrubarea_plot)+
  geom_violin(aes(Years, shrubarea))+
  stat_summary(geom="point",fun.y="mean",aes(x=Years,y=shrubarea),size=3,color="red")+
  labs(title="How does *shrubarea* change with time in initial years after fire?")+
  ylab("Shrub cover * shrub modal height by plot")+
  theme(text=element_text(size=20),
        title=element_text(size=12))

ggplot(data=shrubarea_plot)+
  geom_boxplot(aes(Years, shrubarea))+
  labs(title="shrubarea by year")

```

## 4. Plot shrub characteristics *by species* by year since fire
```{r}
shrubarea_plot_spp <- shrubplotHS %>%
  mutate(shrubarea=Cover*modal_ht_cm)
shrubarea_plot_spp <- tbl_df(shrubarea_plot_spp)
shrubarea_plot_spp
View(shrubarea_plot_spp[,c("Regen_Plot","Species","Cover","modal_ht_cm","shrubarea")])

fewspp <- names(tail(sort(summary(shrubarea_plot_spp$Species)),n=10))

ggplot(data=shrubarea_plot_spp[shrubarea_plot_spp$Species%in%fewspp,])+
  geom_boxplot(aes(Years, shrubarea, fill=Species))+
  labs(title="shrubarea by year and shrub species")


ggplot(data=shrubarea_plot_spp[shrubarea_plot_spp$Species%in%fewspp,])+
  geom_boxplot(aes(Years, Cover, fill=Species))+
  labs(title="shrub cover by year and shrub species")


ggplot(data=shrubarea_plot_spp[shrubarea_plot_spp$Species%in%fewspp,])+
  geom_boxplot(aes(Years, modal_ht_cm, fill=Species))+
  labs(title="shrub height by year and shrub species")

ggplot()+
  stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="CEIN3",],aes(y=modal_ht_cm, x=as.numeric(Years)),color="darkgreen",size=2)+
   stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="ARVI",],aes(y=modal_ht_cm, x=as.numeric(Years)),color="darkorange",size=2)+
  stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="ARPA6",],aes(y=modal_ht_cm, x=as.numeric(Years)),color="salmon",size=2)+
  labs(title= "Changes in shrub height over time for CEIN, ARVI, and ARPA")+
  xlab("Years")+
  ylab("Shrub height (cm)")

ggplot()+
  stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="CEIN3",],aes(y=Cover, x=as.numeric(Years)),color="darkgreen",size=2)+
   stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="ARVI",],aes(y=Cover, x=as.numeric(Years)),color="darkorange",size=2)+
  stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="ARPA6",],aes(y=Cover, x=as.numeric(Years)),color="salmon",size=2)+
  labs(title= "Changes in shrub cover over time for CEIN, ARVI, and ARPA")+
  xlab("Years")+
  ylab("Shrub cover")


ggplot()+
  stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="CEIN3",],aes(y=shrubarea, x=as.numeric(Years)),color="darkgreen",size=2)+
   stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="ARVI",],aes(y=shrubarea, x=as.numeric(Years)),color="darkorange",size=2)+
  stat_summary(fun.y="mean",geom="line",data=shrubarea_plot_spp[shrubarea_plot_spp$Species=="ARPA6",],aes(y=shrubarea, x=as.numeric(Years)),color="salmon",size=2)+
  labs(title= "Changes in shrubarea time for CEIN, ARVI, and ARPA")+
  xlab("Years")+
  ylab("Shrub height (cm) * shrub cover")
```

## 5. Look at relationships between canopy cover and shrub characteristics 
```{r}

```

