---
title: "Compile and Clean Dendro Data"
author: "Carmen"
date: "December 11, 2018"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
require(tidyverse)
require(dplR)
require(readxl)
```

# Load clean detailed data
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load("../../compiled_data/growth_mortality/df_detailed_clean.Rdata")
df_full <- df
load("../../compiled_data/dendro_joined.Rdata")
load("../../compiled_data/rwl_joined.Rdata")
```

# Dendro data basic clean

## Make series all caps
```{r}
df <- df %>% 
  mutate(series = toupper(series))
```

## Separate out tree number and radius number

### Radius number
```{r}
df$radius <- substr(df$series, nchar(df$series), nchar(df$series)) 
```

### Tree
```{r}
df$tree <- substr(df$series, 1, nchar(df$series)-2) 
```

### Find trees where this technique didn't work and fix them
```{r}
df %>% 
  filter(!radius %in% c(1, 2, 3))
```

#### Radius
```{r}
df <- df %>% 
  mutate(radius = ifelse(
    !radius %in% c(1, 2, 3),
    substr(series, nchar(series)-2, nchar(series)),
    radius
    )) 
df %>% 
  filter(!radius %in% c(1, 2, 3))
```

#### Series
```{r}
df <- df %>% 
  mutate(tree = ifelse(
    !radius %in% c(1, 2, 3),
    substr(series, 1, nchar(series)-4),
    tree
    )) 
df %>% 
  filter(!radius %in% c(1, 2, 3))
head(df)
```

### Fix tree names for 184A they sync up with detailed data later on
```{r}
df %>%
  filter(tree %in% c("184A", "184BA"))
df_full %>% filter(SEEDLING %in% c("184-A", "184B-A"))
```

```{r}
df <- df %>% 
  mutate(tree = ifelse(tree == "184A", "184-A", tree)) %>% 
  mutate(tree = ifelse(tree == "184BA", "184B-A", tree))
```

## Rearrange names
```{r}
df <- df %>% 
  dplyr::select(board, tree, series, radius, everything())
head(df)
```

## Throw out trees from sample board
```{r}
df <- df %>% 
  filter(board != "P1BdS ")
```

## Find trees that don't have the right number of radii
```{r}
df %>% 
  group_by(tree) %>% 
  mutate(count = n()) %>% 
  filter(count != 3) %>% 
  dplyr::select(board, tree, series, radius)
```

## Make seedling color column all lower case 
```{r}
df_full %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR))
```

### Fix duplicates

#### Tree 19

##### Take a look
```{r}
df %>% filter(tree %in% c(19, "19BU"))
```

##### Clearly only the 19BU belongs, so delete the other one and fix the mislabeled radius on the BU board
```{r}
df <- df %>% 
  filter(!(tree == 19 & board == "P1Bd62")) %>% 
  mutate(tree = ifelse(tree %in% c(19, "19BU"), 19, tree))
df %>% filter(tree %in% c(19, "19BU"))
```

## Take out duplicates for trees that have both BU and non-BU records

### List the trees that are on backup boards
```{r}
dfBU <- df[grep("BU", df$series),] %>% 
  arrange(tree)
```

### Move BU data to a separate column
```{r}
df_bu <- df
df <- df %>% 
  mutate(BU = ifelse(tree %in% dfBU$tree, 1, 0)) %>% 
  ungroup() 
BUs <- grep("BU", df$tree, value = T)
df <- df %>% 
  mutate(tree = ifelse(tree %in% BUs, substr(tree, 1, nchar(tree)-2), tree)) %>% 
  arrange(desc(BU))
```

### View BU duplicates and delete non-BU ones
```{r}
dups <- df %>% 
  group_by(tree) %>% 
  mutate(count = n()) %>% 
  filter(count != 3) %>% 
  arrange(tree)
dups
```

### Delete the ones that got replaced by backups 
```{r}
df <- df %>% 
  filter(!(tree %in% dups$tree & BU == 0))
```

## Correct the name of seedling 185B
```{r}
df %>% filter(tree == "185B")
```

Based on notes in the master details excel spreadsheet, 185B-P is on board 45, so I just need to relabel this tree. I can always check the board too, if need be. 

```{r}
df <- df %>% 
  mutate(tree = ifelse(tree == "185B", "185B-P", tree))
```

# Check these changes against the compiled detailed data

## Join dendro data with detailed field data
```{r}
join <- full_join(df, df_full, by = c("tree" = "SEEDLING"))
```

## Take out thrown out seedlings
```{r}
join <- join %>% 
  filter(thrown_out != 1)
```

## Look at seedlings without dendro data
```{r}
join %>%
  filter(is.na(board) | tree %in% (join %>% filter(is.na(board)))$tree) %>% 
  dplyr::select(board, tree, BAS_DIA_2_mm, NOTES, Notes, `Tops cut?`) 
```

# Check for mateless seedlings not marked as such

## Create a column for PAIR, aka dead #
```{r}
join$PAIR <- NA
bs <- grep("B", join$tree, value = T)
bs <- substr(bs, 1, nchar(bs)-1)
join[grep("B", join$tree), "PAIR"] <- bs
head(bs)
join <- join %>% 
  mutate(PAIR = ifelse(is.na(PAIR), tree, PAIR))
```

### Check
```{r}
head(join %>% 
  dplyr::select(tree, PAIR))
```

# Find mateless seedlings not marked as such and not already flagged for needing me to check them
```{r}
join %>% 
  group_by(PAIR) %>% 
  mutate(count = n()) %>% 
  filter(count != 6) %>% 
  arrange(PAIR) %>% 
  filter(mateless == 0) %>% 
  dplyr::select(board, tree, series, radius, BU, PAIR, mateless, thrown_out, exclude_reason) %>% 
  filter(!PAIR %in% c(12, 155, 26, 98, 63, 163, 159, 195, 205))
```

## Fix the funky 183/184/185 ones 
```{r}
join <- join %>% 
  mutate(PAIR = ifelse(tree == "184B-P", "184-P",PAIR))
```

```{r}
join %>% 
  filter(PAIR == "183B-") %>% 
  dplyr::select(board, tree, series, radius, BU, PAIR, mateless, thrown_out, exclude_reason) 
```

```{r}
join <- join %>% 
  mutate(PAIR = ifelse(tree == "185B-P", "185-P",PAIR)) %>% 
  mutate(PAIR = ifelse(tree == "185B-A", "185-A",PAIR)) %>% 
  mutate(PAIR = ifelse(tree == "184B-A", "184-A",PAIR)) %>% 
  mutate(PAIR = ifelse(tree == "183B-A", "183-A",PAIR)) %>% 
  mutate(PAIR = ifelse(tree == "183B-P", "183-P",PAIR)) 
```

The 185B-P correction is based on the "dendro" sheet of the Mort_details... .xls file

# Check again for mateless trees
```{r}
join %>% 
  group_by(PAIR) %>% 
  mutate(count = n()) %>% 
  filter(count != 6) %>% 
  arrange(PAIR) %>% 
  filter(mateless == 0) %>% 
  dplyr::select(board, tree, series, radius, BU, PAIR, mateless, thrown_out, exclude_reason) %>% 
  filter(!PAIR %in% c(12, 155, 26, 98, 63, 163, 159, 195, 205))
```

# See about pairing up mateless seedlings with each other if possible
```{r}
join %>% 
  group_by(PAIR) %>% 
  mutate(count = n()) %>% 
  filter(count != 6) %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  dplyr::select(tree, PAIR, thrown_out, board, DEAD_ALIVE, mateless) %>% 
  distinct() %>% 
  arrange(tree) %>% 
  filter(!is.na(board))
```

What species are these missing seedlings?
```{r}
join %>% 
  filter(tree %in% c("12", "155", "26", "98", "63B", "163B", "159B", "195B", "205B")) %>% 
  dplyr::select(tree, SPECIES)
```

A lot of them are pines, which means I should really try to find them

# Check out how many usable seedlings I have as of now of each species
```{r}
join %>% 
  filter(!is.na(board)) %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  group_by(SPECIES) %>% 
  summarise(n())
```

# Check for seedlings with <3 years worth of growth data
```{r}
nas <- join %>% 
    dplyr::select(5:17) %>% 
    is.na %>% 
    `!` %>% 
    rowSums
```

```{r}
join[nas<3,] %>% 
   dplyr::select(tree, PAIR, 5:17, thrown_out, board, DEAD_ALIVE, mateless, exclude_reason) %>% 
  filter(thrown_out ==0)
```

## Take out 74 because even the backup only has 2 years of growth
```{r}
join %>% 
  filter(PAIR ==74) %>% 
   dplyr::select(5:17, tree, PAIR, thrown_out, board, DEAD_ALIVE, mateless, exclude_reason)
```

```{r}
join <- join %>% 
  mutate(thrown_out = ifelse(tree == 74, 1, thrown_out)) %>% 
  mutate(mateless = ifelse(tree == "74B", 1, mateless)) %>% 
  mutate(thrown_out = ifelse(tree == "74B", "?", thrown_out)) %>% 
  mutate(exclude_reason = ifelse(tree == 74, "<3 yrs growth", exclude_reason))
```

## Check that the live mates with <3 years growth have counterparts with 3+ yrs growth
```{r}
join %>% 
  filter(PAIR %in% c(222, 207, 91)) %>% 
  dplyr::select(5:17, tree, PAIR, thrown_out, board, DEAD_ALIVE, mateless, exclude_reason) %>% 
  arrange(tree)
```

```{r}
join <- join %>% 
  mutate(thrown_out = ifelse(tree %in% c("222B","207B", "91B"), "?", thrown_out)) %>% 
  mutate(exclude_reason = ifelse(tree %in% c("222B","207B", "91B"), "<3 yrs growth", exclude_reason)) %>% 
  mutate(mateless = ifelse(tree %in% c("222","207", "91"), "?", mateless))
```

# Take out thrown out seedlings from rwl

## Make colnames of rwl all caps
```{r}
rwl_bu <- rwl
colnames(rwl) <- toupper(colnames(rwl))
```

```{r}
thrown_out <- join %>% 
  filter(thrown_out != 0 | mateless == 1) %>% 
  dplyr::select(series) %>% 
  distinct()
nrow(thrown_out)
ncol_old <- ncol(rwl)
ncol_old
sum(colnames(rwl) %in% thrown_out$series)
rwl <- rwl[, !colnames(rwl) %in% thrown_out$series]
ncol_old - ncol(rwl) == nrow(thrown_out)
ncol(rwl)
```

## CHeck that seedlings with backups were accounted for correctly
```{r}
sort(names(rwl))
```


THey were NOT. So do NOT use rwl for ring width analysis. Use dffull instead. 

# Calculate means across radii for each tree
```{r}
rwl_bu <- rwl
ncol(rwl)
list <- join %>% filter(series %in% names(rwl)) %>% dplyr::select(tree) %>% distinct()

list <- list[,1]
list <- sort(list)

for(i in 1:length(list)){
  treei <- list[i]
  series <- join %>% filter(tree == treei) %>% dplyr::select(series) %>% distinct()
  series
  rwl_subset <- rwl[,names(rwl)%in% unlist(series)]
  rwl_subset
  rwl[ncol(rwl)+1] <- rowMeans(rwl_subset)
  colnames(rwl)[ncol(rwl)] <- treei
  ncol(rwl)
  rwl <- rwl[,!(names(rwl) %in% c(names(rwl_subset)))]
  ncol(rwl)
}
ncol(rwl)
ncol(rwl)*3
```

## Check on the ones that didn't get means calculated
```{r}
sort(names(rwl[,!(names(rwl) %in% join$tree)]))
```

## Delete those ones - they got removed while filtering join/df_full, eg ones with backups
```{r}
rwl <- rwl[,(names(rwl) %in% join$tree)]
ncol(rwl)
```

# Save df with all columns
```{r}
save(join, file = "../../compiled_data/growth_mortality/dendro_all_vars.Rdata")
```


# Save rwl separately for live, dead for each species
```{r}
summary(as.factor(join$SPECIES))
```

```{r}
head(rwl[1:10])
```

## Separate live from dead
```{r}
ncol(rwl)
rwl_live <- rwl[,grep("B", names(rwl))]
ncol(rwl_live)
rwl_dead <- rwl[,!names(rwl) %in% names(rwl_live)]
ncol(rwl_dead)
ncol(rwl_dead)+ncol(rwl_live) == ncol(rwl)
```

```{r}
ncol(rwl_live)
rwl_live_pipo <- rwl_live[,names(rwl_live) %in% join$tree[join$SPECIES=="PIPO"]]
ncol(rwl_live_pipo)
rwl_live_abco <- rwl_live[,names(rwl_live) %in% join$tree[join$SPECIES=="ABCO"]]
ncol(rwl_live_abco)
```

```{r}
ncol(rwl_dead)
rwl_dead_pipo <- rwl_dead[,names(rwl_dead) %in% join$tree[join$SPECIES=="PIPO"]]
ncol(rwl_dead_pipo)
rwl_dead_abco <- rwl_dead[,names(rwl_dead) %in% join$tree[join$SPECIES=="ABCO"]]
ncol(rwl_dead_abco) + ncol(rwl_dead_pipo) == ncol(rwl_dead)
```


# Save 
```{r}
save(rwl, file = "../../data/GrowthMortality/rwl.Rdata")
save(rwl_live_pipo, file = "../../data/GrowthMortality/live_pipo_rwl.Rdata")
save(rwl_live_abco, file = "../../data/GrowthMortality/live_abco_rwl.Rdata")
write.rwl(rwl_live_pipo, fname = "../../data/GrowthMortality/live_pipo.rwl")
write.rwl(rwl_live_abco, fname = "../../data/GrowthMortality/live_abco.rwl")
```

```{r}
save(rwl, file = "../../data/GrowthMortality/rwl.Rdata")
save(rwl_dead_pipo, file = "../../data/GrowthMortality/dead_pipo_rwl.Rdata")
save(rwl_dead_abco, file = "../../data/GrowthMortality/dead_abco_rwl.Rdata")
write.rwl(rwl_dead_pipo, fname = "../../data/GrowthMortality/dead_pipo.rwl")
write.rwl(rwl_dead_abco, fname = "../../data/GrowthMortality/dead_abco.rwl")
```


# Filter by # years

## PIPO Take out trees with only a few years
```{r}
rwl_live_pipo_t <- rwl_live_pipo
rwl_live_pipo_t[!is.na(rwl_live_pipo_t)] <- 1
rwl_live_pipo_t[is.na(rwl_live_pipo_t)] <- 0

sums <- t(rwl_live_pipo_t) %>% rowSums()
hist(sums)
head(sums)
summary(as.factor(sums))
sums <- sums[sums>6]

rwl_live_pipo_gr6 <- rwl_live_pipo[,names(rwl_live_pipo) %in% names(sums)]
ncol(rwl_live_pipo_gr6)
rwl_live_pipo_gr6 <- rwl_live_pipo[,names(rwl_live_pipo) %in% names(sums)]
rwl_live_pipo_gr6_nochron <- rwl_live_pipo[,!(names(rwl_live_pipo) %in% names(sums))]
ncol(rwl_live_pipo_gr6_nochron)
ncol(rwl_live_pipo_gr6_nochron) + ncol(rwl_live_pipo_gr6) == ncol(rwl_live_pipo)
```


###  Save filtered set
```{r}
write.rwl(rwl_live_pipo_gr6, fname = "../../data/GrowthMortality/live_pipo_gr6.rwl")
write.rwl(rwl_live_pipo_gr6_nochron, fname = "../../data/GrowthMortality/live_pipo_gr6_nochron.rwl")
```


## ABCO Take out trees with only a few years
```{r}
rwl_live_abco_t <- rwl_live_abco
rwl_live_abco_t[!is.na(rwl_live_abco_t)] <- 1
rwl_live_abco_t[is.na(rwl_live_abco_t)] <- 0

sums <- t(rwl_live_abco_t) %>% rowSums()
hist(sums)
head(sums)
summary(as.factor(sums))
sums <- sums[sums>6]

rwl_live_abco_gr6 <- rwl_live_abco[,names(rwl_live_abco) %in% names(sums)]
ncol(rwl_live_abco_gr6)
rwl_live_abco_gr6_nochron <- rwl_live_abco[,!(names(rwl_live_abco) %in% names(sums))]
ncol(rwl_live_abco_gr6_nochron) + ncol(rwl_live_abco_gr6) == ncol(rwl_live_abco)
```


###  Save filtered set
```{r}
write.rwl(rwl_live_abco_gr6, fname = "../../data/GrowthMortality/live_abco_gr6.rwl")
write.rwl(rwl_live_abco_gr6_nochron, fname = "../../data/GrowthMortality/live_abco_gr6_nochron.rwl")
```


## Save filtered set of ABCO and PIPO together
```{r}
rwl_live_both_gr6 <- cbind(rwl_live_abco_gr6, rwl_live_pipo_gr6)
```

### Check
```{r}
rwl_live_both_gr6[1:2]
rwl_live_abco_gr6[1:2]

ncol(rwl_live_both_gr6)
rwl_live_both_gr6[47:48]
ncol(rwl_live_pipo_gr6)
rwl_live_pipo_gr6[13:14]
```

```{r}
write.rwl(rwl_live_both_gr6, fname = "../../data/GrowthMortality/live_both_gr6.rwl")
```

```{r}
rwl_live_both_gr6_nochron <- cbind(rwl_live_abco_gr6_nochron, rwl_live_pipo_gr6_nochron)
```

```{r}
write.rwl(rwl_live_both_gr6_nochron, fname = "../../data/GrowthMortality/live_both_gr6_nochron.rwl")
```

# Filter by size

## Calculate average and median diameter
```{r}
join <- join %>% 
  rowwise() %>% 
  mutate(dia_ave = mean(c(BAS_DIA_1_mm, BAS_DIA_2_mm))) %>% 
  ungroup()
head(join %>% dplyr::select(tree, series, BAS_DIA_1_mm, BAS_DIA_2_mm, dia_ave) )
quarters <- join %>% 
  group_by(SPECIES) %>% 
  summarize(quantile(na.omit(dia_ave), probs = .75))
quarters
```

## PIPO

```{r}
quarter_pipo <- quarters %>% filter(SPECIES == "PIPO")
quarter_pipo
big_pipo <- join %>% filter(SPECIES == "PIPO") %>% filter(dia_ave > unlist(quarter_pipo[2]))
nrow(big_pipo)/nrow(join %>% filter(SPECIES == "PIPO"))
```

### PIPO Take out smallest half of trees
```{r}
rwl_live_pipo_big <- rwl_live_pipo[,names(rwl_live_pipo) %in% big_pipo$tree]
ncol(rwl_live_pipo_big)
rwl_live_pipo_big_nochron <- rwl_live_pipo[,!(names(rwl_live_pipo) %in% big_pipo$tree)]
ncol(rwl_live_pipo_big_nochron)
ncol(rwl_live_pipo_big_nochron) + ncol(rwl_live_pipo_big) == ncol(rwl_live_pipo)
```

### PIPO Save filtered set
```{r}
write.rwl(rwl_live_pipo_big, fname = "../../data/GrowthMortality/live_pipo_big.rwl")
write.rwl(rwl_live_pipo_big_nochron, fname = "../../data/GrowthMortality/live_pipo_big_nochron.rwl")
```

## ABCO

```{r}
quarter_abco <- quarters %>% filter(SPECIES == "ABCO")
quarter_abco
big_abco <- join %>% filter(SPECIES == "ABCO") %>% filter(dia_ave > unlist(quarter_abco[2]))
nrow(big_abco)/nrow(join %>% filter(SPECIES == "ABCO"))
```

### abco Take out smallest half of trees
```{r}
rwl_live_abco_big <- rwl_live_abco[,names(rwl_live_abco) %in% big_abco$tree]
ncol(rwl_live_abco_big)
rwl_live_abco_big_nochron <- rwl_live_abco[,!(names(rwl_live_abco) %in% big_abco$tree)]
ncol(rwl_live_abco_big_nochron)
ncol(rwl_live_abco_big_nochron) + ncol(rwl_live_abco_big) == ncol(rwl_live_abco)
```

### abco Save filtered set
```{r}
write.rwl(rwl_live_abco_big, fname = "../../data/GrowthMortality/live_abco_big.rwl")
write.rwl(rwl_live_abco_big_nochron, fname = "../../data/GrowthMortality/live_abco_big_nochron.rwl")
```


## Save filtered set of ABCO and PIPO together
```{r}
rwl_live_both_big <- cbind(rwl_live_abco_big, rwl_live_pipo_big)
```

```{r}
write.rwl(rwl_live_both_big, fname = "../../data/GrowthMortality/live_both_big.rwl")
```

```{r}
rwl_live_both_big_nochron <- cbind(rwl_live_abco_big_nochron, rwl_live_pipo_big_nochron)
```

### Check
```{r}
rwl_live_both_big_nochron[1:2]
rwl_live_abco_big_nochron[1:2]
```

```{r}
write.rwl(rwl_live_both_big_nochron, fname = "../../data/GrowthMortality/live_both_big_nochron.rwl")
```


# MAJOR THINGS I NEED TO DO:

Seedlings needing checking: 12, 155, 26, 98, 63B, 163B, 159B, 195B, 205B

1. Which 12 is the real 12?
2. Check why Seedling 155 was thrown out - it's on board 32 but no data 
3. Look for seedling 26 on board 39 - it doesn't have measurements in the original board 39 data
2. seedlings 98, 63B, and 163B - they're not listed as having a board, so look for it in the thrown out pile
3. Seedling 159B - not listed as having a board
4. Look for seedling 195B - on board 28 but not in data
5. 205B - missing
4. See which trees do and don't have mates
3. APPLY THESE CHANGES TO THE rwl DATA FRAME
6. ASK JODI WHETHER DIAMETER CHANGES AS TREES AGE AFTER DEATH


# Just for funsies, compare growth in dead trees vs. live trees, roughly

## Tidy growth data
```{r}
join <- gather(join, key = "YEAR", value = "GROWTH_mm", 5:17)
```

## Compare
```{r}
prelim <- join %>% 
  filter(!is.na(GROWTH_mm)) %>% 
  filter(thrown_out == 0) %>% 
  filter(YEAR %in% c(2014, 2015, 2016)) %>% 
  mutate(rel_growth = GROWTH_mm/BAS_DIA_1_mm) %>% 
  group_by(tree, SPECIES, DEAD_ALIVE) %>% 
  summarize(rel_growth = mean(rel_growth), growth_mm = mean(GROWTH_mm), growth_vert = mean(na.omit(as.numeric(paste(LAST_YR_GR_cm))/HEIGHT)))

ggplot(prelim)+
  geom_boxplot(aes(x = interaction(SPECIES, DEAD_ALIVE), y= rel_growth))
anova(lm(rel_growth ~ DEAD_ALIVE, data = prelim %>% filter(SPECIES == "PIPO")))
ggplot(prelim)+
  geom_boxplot(aes(x = interaction(SPECIES, DEAD_ALIVE), y= growth_mm))
ggplot(prelim)+
  geom_boxplot(aes(x = interaction(SPECIES, DEAD_ALIVE), y= growth_vert))

```

# It looks preliminarily like dead seedlings have greater radial growth and lower vertical growth than live seedlings