---
title: "Clean growth ~ mortality data"
author: "Carmen"
date: "April 17, 2019"
output: 
  html_document:
      toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
```

# Load data
```{r}
setwd("~/../../Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../compiled_data/growth_mortality/df_detailed_clean.Rdata")
load("../../results/data/GrowthMortality/died2017_abco.Rdata")
load("../../results/data/GrowthMortality/died2016_abco.Rdata")
load("../../results/data/GrowthMortality/died2015_abco.Rdata")
died2015_abco <- died2015
died2016_abco <- died2016
died2017_abco <- died2017
load("../../results/data/GrowthMortality/died2017_pipo.Rdata")
load("../../results/data/GrowthMortality/died2016_pipo.Rdata")
load("../../results/data/GrowthMortality/died2015_pipo.Rdata")
died2015_pipo <- died2015
died2016_pipo <- died2016
died2017_pipo <- died2017
```

## Create a list of trees to remove that are missing any year of vertical growth
```{r}
df_toss <- df %>% 
  filter(is.na(LAST_YR_GR_cm) | is.na(MINUS_1_GR_cm) | is.na(MINUS_2_GR_cm))
nrow(df_toss)
```

## delete pairs in df_toss from df
```{r}
length(unique(df_toss$PAIR))
length(unique(df$PAIR))

nrow(df)
df <- df %>% 
  filter(!PAIR %in% df_toss$PAIR)
nrow(df)
```


# Check how it looks

## Check for mateless seedlings again
```{r}
df %>% 
  group_by(PAIR) %>% 
  mutate(count = n()) %>% 
  filter(count != 2)
```

## Check out how many usable seedlings I have as of now of each species
```{r}
df %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  group_by(SPECIES) %>% 
  summarise(n())
```

# Calculate average relative growth rate per seedling, using only years based on year of death

## Restructure so there's a column for each year of growth

## Height columns
```{r}
df <- df %>% 
  mutate(ht1 = HEIGHT) %>% 
  mutate(ht2 = HEIGHT - LAST_YR_GR_cm) %>% 
  mutate(ht3 = HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm) %>% 
  mutate(ht4 = HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm - MINUS_2_GR_cm)
```

## Growth per year
```{r}
df <- df %>% 
  mutate(gr1 = (ht1-ht2)/ht2) %>% 
  mutate(gr2 = (ht2-ht3)/ht3) %>% 
  mutate(gr3 = (ht3-ht4)/ht4)
```

## Check years of death from lists of trees that died in different years

### Check that all ABCO are in one of the lists
```{r}
df %>% 
  filter(SPECIES == "ABCO" & !SEEDLING %in% c(died2017_abco, died2016_abco, died2015_abco)) %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  dplyr::select(SEEDLING) %>% 
  distinct() %>% 
  arrange(SEEDLING) 
```

```{r}
df %>% 
  filter(SEEDLING == "98") %>%
  dplyr::select(SEEDLING, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, `Growing?`) 
```

### Add seedling 98 to 2017 list based on photo and physical characteristics
```{r}
died2017_abco <- c(died2017_abco, "98")
```

## Check that all ABCO are in a pair with a year assigned
```{r}
df %>% 
  filter(SPECIES == "ABCO" & !PAIR %in% c(died2017_abco, died2016_abco, died2015_abco)) %>% 
  dplyr::select(SEEDLING)
```
 
## Average growth based on pair's year of death
```{r}
df <- df %>% 
  mutate(death_year = ifelse(PAIR %in% died2017_abco, 2017, 
                             ifelse(PAIR %in% died2016_abco, 2016, 
                             ifelse(PAIR %in% died2015_abco, 2015, NA)))) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2017_abco, (gr1+gr2+gr3)/3, 999)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016_abco & DEAD_ALIVE == "DEAD", (gr1+gr2)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016_abco & DEAD_ALIVE == "ALIVE", (gr2+gr3)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015_abco & DEAD_ALIVE == "DEAD", gr1, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015_abco & DEAD_ALIVE == "ALIVE", gr3, gr_rel_ave))
```

## Repeat for PIPO
```{r}
df <- df %>% 
  mutate(death_year = ifelse(PAIR %in% died2017_pipo, 2017, 
                             ifelse(PAIR %in% died2016_pipo, 2016, 
                             ifelse(PAIR %in% died2015_pipo, 2015, death_year)))) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2017_pipo, (gr1+gr2+gr3)/3, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016_pipo & DEAD_ALIVE == "DEAD", (gr1+gr2)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016_pipo & DEAD_ALIVE == "ALIVE", (gr2+gr3)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015_pipo & DEAD_ALIVE == "DEAD", gr1, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015_pipo & DEAD_ALIVE == "ALIVE", gr3, gr_rel_ave))
```


# Check
```{r}
df %>% filter(gr_rel_ave == 999) %>% nrow() ==0
```

```{r}
df %>% 
  filter(gr_rel_ave == 0) %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, SEEDLING, PAIR, gr1, gr2, gr3, death_year, gr_rel_ave) %>% 
  arrange(PAIR) %>% 
  nrow()==0
```

## Throw out seedling pair that likely died in 2015 AND had 0 growth in 2015
```{r}
df <- df %>% 
  filter(PAIR != 214)
```

## Add columns for log10 and natural log growth
```{r}
df <- df %>% 
  mutate(log10_gr_rel_ave = log10(gr_rel_ave)) %>% 
  mutate(log_gr_rel_ave = log(gr_rel_ave))
summary(df$log_gr_rel_ave)
summary(df$log10_gr_rel_ave)
hist(df$gr_rel_ave)
hist(df$log_gr_rel_ave)
```

## Count dead in each species
```{r}
df %>% 
  group_by(SPECIES, DEAD_ALIVE) %>% 
  summarise(n())
```

## Summarize height by species
```{r}
df %>% 
  ggplot()+
  geom_boxplot(aes(y=ht4, x = SPECIES))
anova(lm(ht4 ~ SPECIES, data = df))
```

# List the final seedlings by species
```{r}
(df %>% filter(SPECIES=="ABCO") %>% select(SEEDLING) %>% arrange(SEEDLING))[[1]]
(df %>% filter(SPECIES=="PIPO") %>% select(SEEDLING) %>% arrange(SEEDLING))[[1]]
```

# Check growth calculations by selecting random seedlings
```{r}
test <- (df %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, SEEDLING, PAIR, HEIGHT, LAST_YR_GR_cm, MINUS_1_GR_cm,MINUS_2_GR_cm, gr1, gr2, gr3, death_year, gr_rel_ave, log_gr_rel_ave) %>% 
  arrange(PAIR))[sample(nrow(df),1),]
test
```


# Save vertical growth df
```{r}
save(df, file = "../../compiled_data/growth_mortality/df_vert.Rdata")
write.csv(df, file = "../../compiled_data/growth_mortality/df_vert.csv", row.names = F)
```
