---
title: "Clean growth ~ mortality data"
author: "Carmen"
date: "April 17, 2019"
output: 
  html_document:
      toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../compiled_data/growth_mortality/df_detailed_clean.Rdata")
load(file = "../../results/data/GrowthMortality/died2017.Rdata")
load(file = "../../results/data/GrowthMortality/died2016.Rdata")
load(file = "../../results/data/GrowthMortality/died2015.Rdata")
```

## Create a list of trees to remove that are missing any year of vertical growth
```{r}
df_toss <- df %>% 
  filter(is.na(LAST_YR_GR_cm) | is.na(MINUS_1_GR_cm) | is.na(MINUS_2_GR_cm))
nrow(df_toss)
```


## delete pairs in df_toss from df
```{r}
length(unique(df_toss$PAIR))
length(unique(df$PAIR))

nrow(df)
df <- df %>% 
  filter(!PAIR %in% df_toss$PAIR)
nrow(df)
```


# Check how it looks

## Check for mateless seedlings again
```{r}
df %>% 
  group_by(PAIR) %>% 
  mutate(count = n()) %>% 
  filter(count != 2)
```

## Check out how many usable seedlings I have as of now of each species
```{r}
df %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  group_by(SPECIES) %>% 
  summarise(n())
```


# Calculate average relative growth rate per seedling, using only years based on year of death


## Restructure so there's a column for each year of growth

## Height columns
```{r}
df <- df %>% 
  mutate(ht1 = HEIGHT) %>% 
  mutate(ht2 = HEIGHT - LAST_YR_GR_cm) %>% 
  mutate(ht3 = HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm) %>% 
  mutate(ht4 = HEIGHT - LAST_YR_GR_cm - MINUS_1_GR_cm - MINUS_2_GR_cm)
```

## Growth per year
```{r}
df <- df %>% 
  mutate(gr1 = (ht1-ht2)/ht2) %>% 
  mutate(gr2 = (ht2-ht3)/ht3) %>% 
  mutate(gr3 = (ht3-ht4)/ht4)
```

## Check years of death from lists of trees that died in different years

### Check that all ABCO are in one of the lists
```{r}
df %>% 
  filter(SPECIES == "ABCO" & !SEEDLING %in% c(died2017, died2016, died2015)) %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  dplyr::select(SEEDLING) %>% 
  distinct() %>% 
  arrange(SEEDLING) 
```

```{r}
df %>% 
  filter(SEEDLING == "98") %>%
  dplyr::select(SEEDLING, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, `Growing?`) 
```

### Add seedling 98 to 2017 list based on photo and physical characteristics
```{r}
died2017 <- c(died2017, "98")
```

## Check that all ABCO are in a pair with a year assigned
```{r}
df %>% 
  filter(SPECIES == "ABCO" & !PAIR %in% c(died2017, died2016, died2015)) %>% 
  dplyr::select(SEEDLING)
```
 
### Delete that one
```{r}
df <- df %>% 
  filter(SEEDLING != "202B")
```

## Average growth based on pair's year of death
```{r}
df <- df %>% 
  mutate(death_year = ifelse(PAIR %in% died2017, 2017, 
                             ifelse(PAIR %in% died2016, 2016, 
                             ifelse(PAIR %in% died2015, 2015, NA)))) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2017, (gr1+gr2+gr3)/3, 999)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016 & DEAD_ALIVE == "DEAD", (gr1+gr2)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016 & DEAD_ALIVE == "ALIVE", (gr2+gr3)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015 & DEAD_ALIVE == "DEAD", gr1, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015 & DEAD_ALIVE == "ALIVE", gr3, gr_rel_ave))
```

## For PIPO, for now assign all trees as having died in 2017
```{r}
df <- df %>% 
  mutate(gr_rel_ave = ifelse(SPECIES == "PIPO", (gr1+gr2+gr3)/3, gr_rel_ave))
```


### Check
```{r}
df %>% 
  filter(SPECIES == "ABCO") %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, SEEDLING, PAIR, gr1, gr2, gr3, death_year, gr_rel_ave) %>% 
  arrange(PAIR)
df %>% 
  filter(SPECIES == "PIPO") %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, SEEDLING, PAIR, gr1, gr2, gr3, death_year, gr_rel_ave) %>% 
  arrange(PAIR)
summary(df$gr_rel_ave)
```

#### Look at seedling with 0 average growth
```{r}
df %>% 
  filter(gr_rel_ave == 0) %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, SEEDLING, PAIR, gr1, gr2, gr3, death_year, gr_rel_ave) %>% 
  arrange(PAIR)
```

## Throw out seedling pair that likely died in 2015 AND had 0 growth in 2015
```{r}
df <- df %>% 
  filter(PAIR != 214)
```


## Add column for log growth
```{r}
df <- df %>% 
  mutate(log_gr_rel_ave = log(gr_rel_ave))
summary(df$log_gr_rel_ave)
hist(df$gr_rel_ave)
hist(df$log_gr_rel_ave)
```

# Save vertical growth df
```{r}
save(df, file = "../../compiled_data/growth_mortality/df_vert.Rdata")
```

# NEXT STEP: MAYBE TRY TO PAIR UP MATELESS SEEDLINGS WITH EACH OTHER SO I DON'T HAVE TO DELETE AS MANY PAIRS