---
title: "Clean_all"
author: "Carmen"
date: "May 30, 2019"
output: 
    html_document:
      toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
df <- read_excel("~/Dropbox (Stephens Lab)/SORTIE/Growth_mortality/data/details/compiled/archives/Mort_details_compiled_June25_2018.xlsx", sheet = 2)
```

# Delete rows with no data
```{r}
df <- df %>% 
  filter(!is.na(SEEDLING))
```

# Fix typos
```{r}
df <- df %>% 
  mutate(DEAD_ALIVE = ifelse(DEAD_ALIVE == "alive", "ALIVE", DEAD_ALIVE)) %>% 
  mutate(DEAD_ALIVE = ifelse(DEAD_ALIVE == "dead", "DEAD", DEAD_ALIVE)) %>% 
  mutate(SPECIES = ifelse(SPECIES == "abco", "ABCO", SPECIES)) %>% 
  mutate(SPECIES = ifelse(SPECIES == "pipo", "PIPO", SPECIES))
```

## Correct tree mislabeled as the wrong species
```{r}
df %>% filter(SEEDLING == 202)
df <- df %>% 
  mutate(SPECIES = ifelse(SEEDLING == 202, "PIPO", SPECIES))
```

# Make numeric
```{r}
df <- df %>% 
  mutate(LAST_YR_GR_cm = as.numeric(LAST_YR_GR_cm)) %>% 
  mutate(MINUS_1_GR_cm = as.numeric(MINUS_1_GR_cm)) %>% 
  mutate(MINUS_2_GR_cm = as.numeric(MINUS_2_GR_cm))
```

```{r}
df <- df %>% 
  mutate(DEAD = ifelse(DEAD_ALIVE == "DEAD", 1, 0))
```

# Fix duplicates
```{r}
df %>% 
  filter(duplicated(SEEDLING))
df %>% 
  filter(SEEDLING %in% c("183", "183B"))
```

## Fix duplicates
```{r}
df <- df %>% 
  mutate(SEEDLING = ifelse(SEEDLING == 183 & SPECIES == "ABCO", "183-A", SEEDLING)) %>% 
  mutate(SEEDLING = ifelse(SEEDLING == 183 & SPECIES == "PIPO", "183-P", SEEDLING)) %>% 
  mutate(SEEDLING = ifelse(SEEDLING == "183B" & SPECIES == "PIPO", "183B-P", SEEDLING)) %>% 
  mutate(SEEDLING = ifelse(SEEDLING == "183B" & SPECIES == "ABCO", "183B-A", SEEDLING)) 
```

# Look at missing growth data
```{r}
df %>% filter(is.na(as.numeric(df$LAST_YR_GR_cm)))
df %>% filter(is.na(as.numeric(df$MINUS_1_GR_cm)))
df %>% filter(is.na(as.numeric(df$MINUS_2_GR_cm)))
```

# Correct mis-recorded growth based on photos
```{r}
df <- df %>% 
  mutate(MINUS_2_GR_cm = ifelse(SEEDLING == 98, 7.1, MINUS_2_GR_cm)) %>% 
  mutate(MINUS_2_GR_cm = ifelse(SEEDLING == 77, 6.4, MINUS_2_GR_cm))
``` 

# Make live/dead status a factor
```{r}
df <- df %>% 
  mutate(DEAD = as.factor(DEAD))
```

# Calculate average diameter
```{r}
df <- df %>% 
  mutate(BAS_DIA_AVE = (BAS_DIA_1_mm+BAS_DIA_2_mm)/2)
```

# Create a column for PAIR, or dead #
```{r}
df <- df %>% 
  mutate(SEEDLING = toupper(SEEDLING))
df$PAIR <- NA
bs <- grep("B", df$SEEDLING, value = T)
bs <- substr(bs, 1, nchar(bs)-1)
df[grep("B", df$SEEDLING), "PAIR"] <- bs
head(bs)
df <- df %>% 
  mutate(PAIR = ifelse(is.na(PAIR), SEEDLING, PAIR))

tail(df %>% 
  dplyr::select(SEEDLING, PAIR))
```

## Fix the funky 183/184/185 ones 
```{r}
df$PAIR[222:235]
df <- df %>% 
  mutate(PAIR = ifelse(SEEDLING == "184B-P", "184-P",PAIR)) %>% 
  mutate(PAIR = ifelse(SEEDLING == "184B-A", "184-A",PAIR)) %>% 
  mutate(PAIR = ifelse(SEEDLING == "183B-P", "183-P",PAIR)) %>% 
  mutate(PAIR = ifelse(SEEDLING == "183B-A", "183-A",PAIR)) %>% 
  mutate(PAIR = ifelse(SEEDLING == "185B-P", "185-P",PAIR)) %>% 
  mutate(PAIR = ifelse(SEEDLING == "185B-A", "185-A",PAIR)) 
```

# Look at the ones missing pairs in df_bu

## Find mateless seedlings
```{r}
df %>% 
  group_by(PAIR) %>% 
  mutate(count = n()) %>% 
  dplyr::select(SEEDLING, PAIR, count) %>% 
  arrange(PAIR) %>% 
  filter(count != 2) 
```

# Delete pairs needing to be thrown out by creating a df with all the bad pairs

## Missing data or pathogen damage
```{r}
summary(as.factor(df$Practice))
df_toss <- df %>% 
  filter(Practice == "1") %>% 
  filter(PATH_DAMAGE == "1")
nrow(df_toss)
```

## seedlings marked "thrown out"
```{r}
nrow(df_toss)
df_toss <- full_join(df_toss, df %>% 
  filter(thrown_out == "1"))
nrow(df_toss)
```


## delete pairs in df_toss from df
```{r}
length(unique(df_toss$PAIR))
length(unique(df$PAIR))

nrow(df)
df <- df %>% 
  filter(!PAIR %in% df_toss$PAIR)
nrow(df)
```

## Add reasons for exclusion in the notes section of df 
```{r}
df <- df %>% 
  mutate(exclude_reason = NA)
```

### Create function for Add Exclusion Reason
```{r}
aer <- function(seedling, reason){
  df <<- df %>% 
    mutate(exclude_reason = ifelse(SEEDLING == seedling, reason, exclude_reason))
}
```

### Apply to seedlings one by one by looking at data
```{r}
aer("65B", "mate excluded")
aer("65", "beetles")
aer("94", "PATH_DAMAGE")
aer("92", "PATH_DAMAGE")
aer("92B", "mate excluded")
aer(76, "beetles")
aer(7, "PATH_DAMAGE")
aer(6, "PATH_DAMAGE")
aer("6B", "mate excluded")
aer("59", "beetles")
aer("59B", "mate excluded")
aer("58", "beetles")
aer(53, "PATH_DAMAGE")
aer(5, "PATH_DAMAGE")
aer(28, "PATH_DAMAGE")
aer("22", "MUNCHED")
aer("194", "TOO YOUNG")
aer("194B", "TOO YOUNG")
aer("189", "PATH_DAMAGE")
aer("188", "PATH_DAMAGE")
aer("187", "TOO YOUNG")
aer("187B", "TOO YOUNG")
aer("185-A", "beetles")
aer("17", "MUNCHED")
aer("162", "MUNCHED")
aer("155", "unkn")
aer("147", "MUNCHED")
aer("149", "beetles")
aer(146, "PATH_DAMAGE")
aer("146B", "mate excluded")
aer(145, "beetles")
aer(136, "beetles")
aer(116, "beetles")
aer("116B", "mate excluded")
aer(114, "pitching")
aer(104, "PATH_DAMAGE")
aer(104, "PATH_DAMAGE")
aer(103, "PATH_DAMAGE")
aer("103B", "PATH_DAMAGE")
```

The following are marked as being thrown out, but DO have data, so could potentially be used as a back-up mate if needed,

So make them thrown_out = ?
```{r}
df <- df %>% 
  mutate(thrown_out = ifelse(SEEDLING %in% c("58B", "53B", "7B", "28B", "22B", "189B", "188B", "185BA", "17B", "162B", "155B", "147B", "149B", "145B", "136B", "114B", "104B", "94B"), "?", thrown_out))
```

### Add column to df for mateless seedlings
```{r}
df <- df %>% 
  mutate(mateless = ifelse(SEEDLING %in% c("7B","76B", "58B", "53B", "5B", "28B", "22B", "189B", "188B", "185BA", "17B", "162B", "155B", "147B", "149B", "145B", "136B", "114B", "104B", "94B"), 1, 0)) %>% 
  arrange(desc(mateless))
```

## See what thrown out seedlings remain without a exclude_reason
```{r}
df %>% 
  filter(thrown_out == 1 & is.na(exclude_reason)) %>% 
  dplyr::select(COMP, DATE, SEEDLING, DEAD_ALIVE, thrown_out, NOTES, Notes, PATH_DAMAGE) %>% 
  arrange(desc(SEEDLING)) 
```

## Correct 184A/184-A/185
```{r}
df <- df %>% 
  mutate(SEEDLING = ifelse(SEEDLING == "184A", "184-A", SEEDLING)) %>% 
  mutate(SEEDLING = ifelse(SEEDLING == "184BA", "184B-A", SEEDLING)) %>% 
  mutate(SEEDLING = ifelse(SEEDLING == "185BA", "185B-A", SEEDLING)) %>% 
  mutate(SEEDLING = ifelse(SEEDLING == "185B", "185B-P", SEEDLING))
```

# Save df
```{r}
save(df, file = "../../compiled_data/growth_mortality/df_detailed_clean.Rdata")
```

