---
title: "Dating dead PIPO from dendro"
author: "Carmen Tubbesing"
date: "6/6/2019"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
require(tidyverse)
require(dplR)
require(readxl)
```

# Read in data

### compiled dendro files
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../data/GrowthMortality/dead_pipo_rwl.Rdata")
rwlp <- rwl_dead_pipo
```

### live detrended data
```{r}
load("../../compiled_data/live_chron_pipo.Rdata")
dfp_live <- dfp
remove(dfp)
```

### detailed dead seedling characteristic data to use for creating a year of death rubric
```{r}
load("../../compiled_data/growth_mortality/dendro_all_vars.Rdata")
```

#### separate detailed data into abco and pipo
```{r}
dfpfull <- join %>% 
  filter(SPECIES == "PIPO")
```

# Clean up Growing column
```{r}
dfpfull_bu <- dfpfull
```

## Standardize the response types
```{r}
dfpfull <- dfpfull_bu
dfpfull <- dfpfull %>% 
  rename(Growing = `Growing?`) %>% 
  mutate(Growing = toupper(Growing)) %>% 
  mutate(Growing = ifelse(Growing == "GROWING", "YES", Growing)) %>% 
  mutate(Growing = ifelse(Growing == "NOT GROWING", "NO", Growing)) %>% 
  mutate(Growing = ifelse(Growing == "EARLY 2017 GROWTH", "YES", Growing)) %>% 
  mutate(Growing = ifelse(Growing == "BEFORE 2017 GROWTH", "NO", Growing)) 

unique(dfpfull$Growing)
```

# Create functions

### Make a function to plot the trend for each tree
```{r}
ringplot <- function(i){
  ggplot(data = dfp %>% filter(tree == i) %>% filter(Year >2008))+
  geom_line(aes(x = Year, y = growth_detrended))+
  geom_point(aes(x = Year, y = growth_detrended))+
  ggtitle(paste(i))+
  scale_x_continuous(breaks = seq(min(dfp$Year), max(dfp$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
}
```

### Make vectors to add to 

"died2017" means that the dead tree's "2016" year matches live trees' "2016" - AKA the tree put on a partial ring in 2017 and then died
```{r}
died2017 <- c()
died2016 <- c()
died2015 <- c()
unkn <- c()
```

### Make functions to add and subtract trees from the vectors
```{r}
add2017 <- function(new){
  died2017 <<- c(died2017, new)
}
```

```{r}
add2016 <- function(new){
  died2016 <<- c(died2016, new)
}
```

```{r}
add2015 <- function(new){
  died2015 <<- c(died2015, new)
}
```

```{r}
addunkn <- function(new){
  unkn <<- c(unkn, new)
  if(new %in% died2017){
    died2017<<- died2017[died2017 != new]
  }
}
```

```{r}
subtract2016 <- function(new){
  if(new %in% died2016){
    died2016<<- died2016[died2016 != new]
  }
}
```

### Make a function to add to a vector of trees whose age assignment I feel confident about BASED SOLELY ON RING WIDTHS
```{r}
sure <- c()
addsure <- function(new){
  sure <<- c(sure, new)
}
```

### Make a function to look at all the physical traits of a seedling
```{r}
look <- function(treenum){
  print(dfpfull %>% 
    dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing, `BLACK DOTS`)  %>% 
    filter(tree == treenum) %>% 
    distinct())
}
```

# Assign green trees as 2017 death

## Find which trees have any mention of green
```{r}
summary(as.factor(dfpfull$NEEDLE_COLOR))
sort(unique(dfpfull[grep("green", dfpfull$NEEDLE_COLOR),]$tree))
```

## Check for green in the photos

### Modify needle color column if I see green in the photo that wasn't recorded
```{r}
```

#### And remove it from the 2016 list
```{r}
```

### Make a vector of trees whose photos had green
```{r}
definitely_green <- c(15, 181, 20, 49, 50)
```


## If there's actually green in the photos, add that seedling to the list of trees that died in 2017 if you haven't already
```{r}
add2017(15)
add2017(181)
add2017(20)
add2017(49)
add2017(50)
```

# Detrend using horizontal line method
```{r}
rwlp_dt <- detrend(rwl = rwlp, method = "Mean")
```

# Reshape data frame to identify trees with the most rings
```{r}
rwlp2 <- rwlp %>% 
  mutate(year = row.names(rwlp))
rwlp2 <- gather(rwlp2, key = "tree", value = "growth", 1:ncol(rwlp)) %>% 
  filter(!is.na(growth))
rwlp2_summary <- rwlp2 %>% 
  group_by(tree) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
rwlp2_summary
```

# Reshape detrended means in order to look at each tree's trend
```{r}
dfp <- as.data.frame(t(rwlp_dt))
dfp$tree <- row.names(dfp)
dfp <- gather(dfp, key = "Year", value = "growth_detrended", c(1:13))
dfp <- dfp %>% 
  filter(!is.na(growth_detrended))
```

```{r}
dfp <- dfp %>% 
  mutate(Year =as.numeric(Year))
```

# Categorize trees based on photos in ppt

## Make a list of all the trees that have no needles in the photos
```{r}
died2015 <- c(213, 74, 65, 150, 204, 48, 190, 196, 70, 177, "185-P")
died2015
```

## Look at their characteristics
```{r}
dfpfull %>% 
    dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing, `BLACK DOTS`)  %>% 
    filter(tree %in% died2015) %>% 
    distinct()
```

## Look at the data of trees that look really intact in photos but don't have green
```{r}
look(1)
look(71)
look(59)
look(26)
look(209)

add2016(26) # based on mostly but not fully intact needles and reddish color, and that it's not growing
add2016(167) # based on high needle retention, reddish color, and not growing

look(172) # likely died early 2017 growth because it's reddish and has 95% needles but not growing
add2017(172)

look(11) 
add2016(11) # not growing and not intact enough to have died after 2017 growth

look(75)
look(160)

look(191)

look(166)
add2016(166) # growing but kinda crappy looking

look(158) 
add2016(158) # growing but kinda crappy looking

look(157) 

look(9)
add2016(9) # kinda crappy but growing

look(8) # mostly intact needles, not gray or tan, but not growing and not just dead

look(19)
add2016(19) # mid growth but not 2017

look(21)
add2016(21) # kinda crappy but growing
```


# Categorize trees based on physical characteristics and photos
```{r}
sort(unique(as.numeric(dfp$tree[!dfp$tree %in% definitely_green])))

look(1)
add2017(1) # super intact 

look(8)
add2016(8)

look(9)
addunkn(9)

look(10)
add2016(10) # growing + droopy and discolored -> didn't die 2017 growing season

look(14)
add2016(14)

look(25)
add2016(25)

look(39)
add2016(39)

look(42)
add2015(42)

look(57)
```

# Add year assignments from powerpoint file

## 2015
```{r}
sort(as.numeric(died2015))
sort(died2015)

died2015 <- c(died2015, 12, "184-P", 186, 164, 99, 67, 107, 124, 125, 154, 156, 57, 106, 110, 121, 182, 123)
sort((died2015))
length(died2015)
```

## 2016
```{r}
sort(as.numeric(died2016))

died2016 <- c(died2016, 60, 61, 63, 64, 72, 109, 115, 118, 159, 122, 142, 144, 165, "183-P", 202, 153, 157, 173, 191, 151, 168, 200, 73, 160, 75, 13, 205, 71)
sort((died2016))
length(died2016)
died2016[duplicated(died2016)]
```

## 2017
```{r}
sort(as.numeric(died2017))

died2017 <- c(died2017, 85, 209, 152)
sort((died2017))
length(died2017)
died2017[duplicated(died2017)]
```

## Check for double counted trees
```{r}
c(died2015, died2016, died2017)[duplicated(c(died2015, died2016, died2017))]
length(c(died2015, died2016, died2017))
```

# Summarize physical characteristics of trees that I think died in each year
```{r}
dfpfull <- tbl_df(dfpfull)
```

## 2017
```{r}
dfpfull %>% 
  dplyr::select(tree, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS)  %>% 
  filter(tree %in% died2017) %>% 
  distinct
```

## 2016
```{r}
dfpfull %>% 
  dplyr::select(tree, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing)  %>% 
  filter(tree %in% died2016) %>% 
  distinct() %>% 
  arrange(PERC_NEEDLES) %>% 
  arrange(Growing)
```

## 2015
```{r}
dfpfull %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing)  %>% 
  filter(tree %in% died2015) %>% 
  distinct() %>% 
  arrange(PERC_NEEDLES) %>% 
  arrange(Growing)
```

# Check that everything is accounted for
```{r}
dfp %>%
  filter(!tree  %in% died2017 & !tree %in% died2016 & !tree %in% died2015) %>% 
  dplyr::select(tree) %>% 
  distinct() %>% 
  arrange(tree)
```

# Count trees in each death year
```{r}
length(died2017)
length(died2016)
length(died2015)
```

# Save lists of trees that died in each year
```{r}
save(died2017, file = "../../results/data/GrowthMortality/died2017_pipo.Rdata")
save(died2016, file = "../../results/data/GrowthMortality/died2016_pipo.Rdata")
save(died2015, file = "../../results/data/GrowthMortality/died2015_pipo.Rdata")
```

# Summarize variables for each year of death

## Percent needles
```{r}
dfpfull %>% 
  filter(tree %in% died2017) %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing) %>% 
  summarize(median_perc_needles = median(na.omit(as.numeric(PERC_NEEDLES))), mean_perc_needles = mean(na.omit(as.numeric(PERC_NEEDLES))))
dfpfull %>% 
  filter(tree %in% died2016) %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing) %>% 
  summarize(median_perc_needles = median(na.omit(as.numeric(PERC_NEEDLES))), mean_perc_needles = mean(na.omit(as.numeric(PERC_NEEDLES))))
dfpfull %>% 
  filter(tree %in% died2015) %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing) %>% 
  summarize(median_perc_needles = median(na.omit(as.numeric(PERC_NEEDLES))), mean_perc_needles = mean(na.omit(as.numeric(PERC_NEEDLES))))
```

## Color
```{r}
dfpfull %>% 
  filter(tree %in% died2017) %>%
  dplyr::select(tree, NEEDLE_COLOR) %>% 
  distinct() %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR)) %>% 
  group_by(NEEDLE_COLOR) %>% 
  summarize(n())
dfpfull %>% 
  filter(tree %in% died2016) %>%
  dplyr::select(tree, NEEDLE_COLOR) %>% 
  distinct() %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR)) %>% 
  group_by(NEEDLE_COLOR) %>% 
  summarize(n())
dfpfull %>% 
  filter(tree %in% died2015) %>%
  dplyr::select(tree, NEEDLE_COLOR) %>% 
  distinct() %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR)) %>% 
  group_by(NEEDLE_COLOR) %>% 
  summarize(n())
```

## Bark
```{r}
dfpfull %>% 
  filter(tree %in% died2017) %>%
  dplyr::select(tree, BARK) %>% 
  distinct() %>% 
  mutate(BARK = tolower(BARK)) %>% 
  group_by(BARK) %>% 
  summarize(n())
dfpfull %>% 
  filter(tree %in% died2016) %>%
  dplyr::select(tree, BARK) %>% 
  distinct() %>% 
  mutate(BARK = tolower(BARK)) %>% 
  group_by(BARK) %>% 
  summarize(n())
dfpfull %>% 
  filter(tree %in% died2015) %>%
  dplyr::select(tree, BARK) %>% 
  distinct() %>% 
  mutate(BARK = tolower(BARK)) %>% 
  group_by(BARK) %>% 
  summarize(n())
```

# Re-assign death year 
```{r}
dfptest <- dfp %>% 
  filter(tree %in% c(died2017, died2016, died2015)) %>% 
  mutate(Year = ifelse(tree %in% died2016, Year-1, Year)) %>% 
  mutate(Year = ifelse(tree %in% died2015, Year-2, Year))
head(dfptest)
```

# Plot with standard errors*2
```{r}
dfptest <- dfptest %>% 
  filter(Year > 2008) %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
```

```{r}
ggplot(data = dfptest, aes(x = Year, y = mean_gr_year))+
  geom_point(col = "red")+
  geom_line(col = "red")+
  geom_errorbar(aes(x = Year,ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = Year, y = min), col = "red", linetype = "dashed")+
  geom_line(aes(x = Year, y = max), col = "red", linetype = "dashed")+
  ggtitle("Dead PIPO detrended means +/- 2 SE")+
  ylab("Mean detrended ring widths") +
  scale_x_continuous(breaks = seq(min(dfp$Year), max(dfp$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
ggsave("../../results/figures/GrowthMortality/dead_chron_pipo.jpg", width = 5, height = 5)
```

# Save df version of the detrended dendro data
```{r}
save(dfp, file = "../../compiled_data/pine_rings_detrended_dead.Rdata")
```


# To Do
1. Look at tree 172 to see if it has a micro ring - I think it died early 2017 growing season
