---
title: "date_dead_dendro"
author: "Carmen"
date: "May 22, 2019"
output: 
  html_document:
    toc: TRUE
---

Process:

1.  Compare individual ABCO series to the summarized series from live ABCO
2.  Make lists of trees that look like they pretty clearly died in 2017, 2016, 2015, or unknown
3.  Plot summarized series for the subset of trees I think I can date and compare it to the live ABCO chronology
4.  Summarize physical characteristics from the trees by estimated age of death
5.  Try to use those physical characteristics to age the remainder of the trees by creating a "year of death rubric"


```{r, include = F}
require(tidyverse)
require(dplR)
require(readxl)
```

# Read in data

### compiled dendro files
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../data/GrowthMortality/dead_pipo_rwl.Rdata")
rwlp <- rwl_dead_pipo
load(file = "../../data/GrowthMortality/dead_abco_rwl.Rdata")
rwla <- rwl_dead_abco
```

### live detrended data
```{r}
load("../../compiled_data/live_chron_abco.Rdata")
dfa_live <- dfa
remove(dfa)
load("../../compiled_data/live_chron_pipo.Rdata")
dfp_live <- dfp
remove(dfp)
```

### detailed dead seedling characteristic data to use for creating a year of death rubric
```{r}
load("../../compiled_data/growth_mortality/dendro_all_vars.Rdata")
```

#### separate detailed data into abco and pipo
```{r}
dfafull <- join %>% 
  filter(SPECIES == "ABCO")
dfpfull <- join %>% 
  filter(SPECIES == "PIPO")
```


# Clean up Growing column
```{r}
dfafull_bu <- dfafull
```

## Standardize the response types
```{r}
dfafull <- dfafull_bu
dfafull <- dfafull %>% 
  rename(Growing = `Growing?`) %>% 
  mutate(Growing = toupper(Growing)) %>% 
  mutate(Growing = ifelse(Growing == "GROWING", "YES", Growing)) %>% 
  mutate(Growing = ifelse(Growing == "NOT GROWING", "NO", Growing)) %>% 
  mutate(Growing = ifelse(Growing == "EARLY 2017 GROWTH", "YES", Growing)) %>% 
  mutate(Growing = ifelse(Growing == "BEFORE 2017 GROWTH", "NO", Growing)) 

unique(dfafull$Growing)
```

## Correct tree that says it's not growing but really looks like it's growing from the photo
```{r}
dfafull <- dfafull %>%
  mutate(Growing = ifelse(tree == 139, "YES", Growing))
```


# Create functions

### Make a function to plot the trend for each tree
```{r}
ringplot <- function(i){
  ggplot(data = dfa %>% filter(tree == i) %>% filter(Year >2008))+
  geom_line(aes(x = Year, y = growth_detrended))+
  geom_point(aes(x = Year, y = growth_detrended))+
  ggtitle(paste(i))+
  scale_x_continuous(breaks = seq(min(dfa$Year), max(dfa$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
}
```

### Make vectors to add to 

"died2017" means that the dead tree's "2016" year matches live trees' "2016" - AKA the tree put on a partial ring in 2017 and then died
```{r}
died2017 <- c()
died2016 <- c()
died2015 <- c()
unkn <- c()
```

### Make functions to add and subtract trees from the vectors
```{r}
add2017 <- function(new){
  died2017 <<- c(died2017, new)
}
```

```{r}
add2016 <- function(new){
  died2016 <<- c(died2016, new)
}
```

```{r}
add2015 <- function(new){
  died2015 <<- c(died2015, new)
}
```

```{r}
addunkn <- function(new){
  unkn <<- c(unkn, new)
  if(new %in% died2017){
    died2017<<- died2017[died2017 != new]
  }
}
```

```{r}
subtract2016 <- function(new){
  if(new %in% died2016){
    died2016<<- died2016[died2016 != new]
  }
}
```

### Make a function to add to a vector of trees whose age assignment I feel confident about BASED SOLELY ON RING WIDTHS
```{r}
sure <- c()
addsure <- function(new){
  sure <<- c(sure, new)
}
```

### Make a function to look at all the physical traits of a seedling
```{r}
look <- function(treenum){
  print(dfafull %>% 
    dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing)  %>% 
    filter(tree == treenum) %>% 
    distinct())
}
```

# Assign green trees as 2017 death

## Find which trees have any mention of green
```{r}
summary(as.factor(dfafull$NEEDLE_COLOR))
sort(unique(dfafull[grep("green", dfafull$NEEDLE_COLOR),]$tree))
```

## Check for green in the photos

### Modify needle color column if I see green in the photo that wasn't recorded
```{r}
dfafull <- dfafull %>% 
  mutate(NEEDLE_COLOR = ifelse(tree == 215, "red to light green", NEEDLE_COLOR))
```

#### And remove it from the 2016 list
```{r}
subtract2016(215)
```

### Make a vector of trees whose photos had green
```{r}
definitely_green <- c(201, 100, 30, 127, 16, 211, 32, 52, 55, 169, 215, 208)
```

## If there's actually green in the photos, add that seedling to the list of trees that died in 2017 if you haven't already
```{r}
add2017(201)
add2017(55)
add2017(100)
add2017(30)
add2017(127)
add2017(16)
add2017(211)
add2017(32)
add2017(52)
add2017(169)
add2017(215)
add2017(208)
```

# Detrend using horizontal line method

## PIPO
```{r}
rwlp_dt <- detrend(rwl = rwlp, method = "Mean")
```

## ABCO
```{r}
rwla_dt <- detrend(rwl = rwla, method = "Mean")
```

# Reshape data frame to identify trees with the most rings
```{r}
rwla2 <- rwla %>% 
  mutate(year = row.names(rwla))
rwla2 <- gather(rwla2, key = "tree", value = "growth", 1:ncol(rwla)) %>% 
  filter(!is.na(growth))
rwla2_summary <- rwla2 %>% 
  group_by(tree) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
rwla2_summary
```

# Reshape detrended means in order to look at each tree's trend

### PIPO
```{r}
dfp <- as.data.frame(t(rwlp_dt))
dfp$tree <- row.names(dfp)
dfp <- gather(dfp, key = "Year", value = "growth_detrended", c(1:13))
dfp <- dfp %>% 
  filter(!is.na(growth_detrended))
```

### ABCO
```{r}
dfa <- as.data.frame(t(rwla_dt))
dfa$tree <- row.names(dfa)
dfa <- gather(dfa, key = "Year", value = "growth_detrended", c(1:13))
dfa <- dfa %>% 
  filter(!is.na(growth_detrended))
```

```{r}
dfa <- dfa %>% 
  mutate(Year =as.numeric(Year))
```

# List trees in descening age order in order to start with longest series

Exclude trees with green foliage
```{r}
rwla2_summary_unkn <- rwla2_summary %>% 
  filter(!tree %in% definitely_green)
rwla2_summary_unkn 
```

# Categorize each tree based on best guess death year for trees 6+ y/o

```{r}
ringplot(41)
add2017(41)
addsure(41)

ringplot(129)
addunkn(129)

ringplot(161)
add2017(161) 
addsure(161)

ringplot(51)
add2017(51) # based largely on color and the fact that it looks like it actually is growing

ringplot(68)
look(68)
add2016(68) # based largely on rings, the fact that it's red/brown 

ringplot(78)
add2016(78)
addsure(78)

ringplot(102)
look(102) 
add2017(102) # based largely on photo

ringplot(112)
look(112)
add2017(112) # based largely on the fact that it's growing and looks so red in photo

ringplot(131)
add2016(131)
addsure(131)

ringplot(139)
add2017(139) # Growing and looks like it may have hints of green
```

```{r}
rwla2_summary_unkn[11:20,]
```

```{r}
ringplot("184-A")
addunkn("184-A")

ringplot(195)
look(195)
add2016(195) # based on physical characteristics

ringplot(203)
add2016(203)
addsure(203)

ringplot(206)
addunkn(206)

ringplot(234)
add2016(234)
addsure(234)

ringplot(62)
add2017(62) # based on characteristics

ringplot(77)
look(77)
add2016(77) # based on a combination of rings and characteristics

ringplot(96)
add2016(96) # based largely on needle retention in PHOTO

ringplot(128)
look(128)
add2015(128) # based on a combination of rings and characteristics
```

```{r}
rwla2_summary_unkn[21:30,]
```

```{r}
ringplot(132)
add2017(132)
addsure(132)

ringplot(18)
add2017(18) # based largely on needle retention

ringplot("183-A")
look("183-A")
add2017("183-A") # based on a combination of rings and characteristics

ringplot(197)
add2015(197) # based on physical characteristics

ringplot(198)
add2017(198) # based largely on photos

ringplot(207)
add2017(207) # based in part on color and abundance of foliage

ringplot(214)
add2015(214) # based in part on color and abundance of foliage

ringplot(216)
look(216)
add2016(216) # based mostly on the physical characteristics and the fact that it wasn't growing

ringplot(224)
add2016(224) # hard choice. Lots of needles but ring pattern matches 2016 more closely

ringplot(235)
add2016(235) # Not growing, missing some needles, and not bright red or green
```

```{r}
rwla2_summary_unkn[31:40,]
```

```{r}
ringplot(80)
look(80)
add2015(80) # based on characteristics

ringplot(117)
add2016(117)
addsure(117)

ringplot(119)
add2016(119)
addsure(119)

ringplot(120)
look(120)
add2015(120) # based largely on characteristics

ringplot(130)
addunkn(130)

ringplot(137)
look(137)
add2017(137) # based largely on photo and needles and the fact that it was growing

ringplot(141)
addunkn(141)

ringplot(192)
look(192)
add2017(192) 

ringplot(223)
add2017(223) # based on photo and the fact that it was growing
```

```{r}
rwla2_summary_unkn[41:50,]
```

```{r}
ringplot(229)
addunkn(229) # maybe 2014

ringplot(230)
add2016(230) # because of needle retention, the fact that it's not growing, and photo

ringplot(29)
add2016(29) # based largely on the fact that it's not growing and doesn't look freshly dead

ringplot(31)
add2015(31) # based largely on needle color and quantity

ringplot(33)
add2016(33) # kind of a guess. Based largely on photo and needle retention and the fact that it's not growing

ringplot(66)
add2015(66)
addsure(66)

ringplot(93)
look(93)
add2017(93)
```

## Look at the trees I've added to each list
```{r}
died2017
died2016
died2015
unkn
sure
length(sure)
```

### See what years the "sure" seedlings died in
```{r}
died2016[died2016 %in% sure]
died2017[died2017 %in% sure]
died2015[died2015 %in% sure]
```

# Summarize physical characteristics of trees that I think died in each year
```{r}
dfafull <- tbl_df(dfafull)
```

## 2017
```{r}
dfafull %>% 
  dplyr::select(tree, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS)  %>% 
  filter(tree %in% died2017) %>% 
  distinct
```

## 2016
```{r}
dfafull %>% 
  dplyr::select(tree, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing)  %>% 
  filter(tree %in% died2016) %>% 
  distinct() %>% 
  arrange(PERC_NEEDLES) %>% 
  arrange(Growing)
```

### Look at trees with 90+% needles, red, and not growing, and see if I can date them as 2016 or 2017
```{r}
red90notgrowing <- dfafull %>% 
  filter(as.numeric(PERC_NEEDLES)>=90 & Growing == "NO" & !tree %in% c(died2017, died2016, died2015) & NEEDLE_COLOR == "red") %>%
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing)  %>% 
  distinct() %>% 
  arrange(tree) %>% 
  arrange(PERC_NEEDLES)
red90notgrowing
```

#### Add to 2017 young trees that match the characteristics of other dead 2017 trees based on data and photos
```{r}
add2017(4)
add2017(232)
add2017(193)
add2017("184-A")
add2017(47)
```

## 2015
```{r}
dfafull %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing)  %>% 
  filter(tree %in% died2015) %>% 
  distinct() %>% 
  arrange(PERC_NEEDLES) %>% 
  arrange(Growing)
```

### Look if it makes sense to say all trees with <=50% needles are 2015, and then do that
```{r}
needles50 <- dfafull %>%  
  filter(PERC_NEEDLES <=50 & !tree %in% died2017 & !tree %in% died2016  & !tree %in% died2015 ) %>% 
  dplyr::select(tree, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS)  %>% 
  arrange(PERC_NEEDLES) %>% 
  distinct() 
needles50
```

```{r}
died2015 <- c(died2015, needles50$tree)
died2015
```

# Look at trees that haven't yet been categorized
```{r}
dfa <- tbl_df(dfa)
dfa %>%
  filter(!tree  %in% died2017 & !tree %in% died2016 & !tree %in% died2015) %>% 
  dplyr::select(tree) %>% 
  distinct() %>% 
  arrange()
length(unique(dfa$tree))
```

## Go through each one and assign a year to it

### 222
```{r}
ringplot(222)
look(222)
add2017(222)
```

### 101
```{r}
look(101)
add2017(101)
```

### 111
```{r}
look(111)
ringplot(111)
add2016(111)
```

### 199
```{r}
look(199)
ringplot(199)
add2016(199)
```

### 212
```{r}
look(212)
ringplot(212)
add2017(212)
```

### 225
```{r}
look(225)
ringplot(225)
add2016(225)
```

### 210
```{r}
look(210)
ringplot(210)
add2016(210)
```

### 233
```{r}
look(233)
ringplot(233)
add2017(233)
```

### 221
```{r}
look(221)
ringplot(221)
add2016(221)
```

### 56
```{r}
look(56)
ringplot(56)
add2016(56)
```

### 218
```{r}
look(218)
ringplot(218)
add2016(218)
```

### 219
```{r}
look(219)
ringplot(219)
add2016(219)
```

### 220
```{r}
look(220)
ringplot(220)
add2017(220)
```

### 226
```{r}
look(226)
ringplot(226)
add2016(226)
```

### 227
```{r}
look(227)
ringplot(227)
add2017(227) # looks like it has some green
```

### 36
```{r}
look(36)
ringplot(36)
add2016(36)
```

### 91
```{r}
look(91)
ringplot(91)
add2016(91)
```

### 97
```{r}
look(97)
ringplot(97)
add2017(97)
```

# Check that everything is accounted for
```{r}
dfa %>%
  filter(!tree  %in% died2017 & !tree %in% died2016 & !tree %in% died2015) %>% 
  dplyr::select(tree) %>% 
  distinct() %>% 
  arrange(tree)
```

# Count trees in each death year
```{r}
length(died2017)
length(died2016)
length(died2015)
```

# Save lists of trees that died in each year
```{r}
save(died2017, file = "../../results/data/GrowthMortality/died2017_abco.Rdata")
save(died2016, file = "../../results/data/GrowthMortality/died2016_abco.Rdata")
save(died2015, file = "../../results/data/GrowthMortality/died2015_abco.Rdata")
```


## Find duplicates
```{r}
length(unique(c(died2017, died2016, died2015)))
length(unique(dfa$tree))
```


# Summarize variables for each year of death

## Percent needles
```{r}
dfafull %>% 
  filter(tree %in% died2017) %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing) %>% 
  summarize(median_perc_needles = median(na.omit(as.numeric(PERC_NEEDLES))), mean_perc_needles = mean(na.omit(as.numeric(PERC_NEEDLES))))
dfafull %>% 
  filter(tree %in% died2016) %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing) %>% 
  summarize(median_perc_needles = median(na.omit(as.numeric(PERC_NEEDLES))), mean_perc_needles = mean(na.omit(as.numeric(PERC_NEEDLES))))
dfafull %>% 
  filter(tree %in% died2015) %>% 
  dplyr::select(tree, HEIGHT, PERC_NEEDLES, NEEDLE_COLOR, BARK, BUDS, Growing) %>% 
  summarize(median_perc_needles = median(na.omit(as.numeric(PERC_NEEDLES))), mean_perc_needles = mean(na.omit(as.numeric(PERC_NEEDLES))))
```

## Color
```{r}
dfafull %>% 
  filter(tree %in% died2017) %>%
  dplyr::select(tree, NEEDLE_COLOR) %>% 
  distinct() %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR)) %>% 
  group_by(NEEDLE_COLOR) %>% 
  summarize(n())
dfafull %>% 
  filter(tree %in% died2016) %>%
  dplyr::select(tree, NEEDLE_COLOR) %>% 
  distinct() %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR)) %>% 
  group_by(NEEDLE_COLOR) %>% 
  summarize(n())
dfafull %>% 
  filter(tree %in% died2015) %>%
  dplyr::select(tree, NEEDLE_COLOR) %>% 
  distinct() %>% 
  mutate(NEEDLE_COLOR = tolower(NEEDLE_COLOR)) %>% 
  group_by(NEEDLE_COLOR) %>% 
  summarize(n())
```

## Bark
```{r}
dfafull %>% 
  filter(tree %in% died2017) %>%
  dplyr::select(tree, BARK) %>% 
  distinct() %>% 
  mutate(BARK = tolower(BARK)) %>% 
  group_by(BARK) %>% 
  summarize(n())
dfafull %>% 
  filter(tree %in% died2016) %>%
  dplyr::select(tree, BARK) %>% 
  distinct() %>% 
  mutate(BARK = tolower(BARK)) %>% 
  group_by(BARK) %>% 
  summarize(n())
dfafull %>% 
  filter(tree %in% died2015) %>%
  dplyr::select(tree, BARK) %>% 
  distinct() %>% 
  mutate(BARK = tolower(BARK)) %>% 
  group_by(BARK) %>% 
  summarize(n())
```

# Assign time of death in column of dfafull
```{r}
dfafull <- dfafull %>% 
  mutate(time_of_death = ifelse(tree %in% died2017 & Growing == "YES", "2017 growing season", 0))
```

```{r}
dfafull <- dfafull %>% 
  mutate(time_of_death = ifelse(tree %in% c(203), "between 2016 and 2017 growth", time_of_death))
```

## Assign outer ring column 
**Note: this was not measured for ring width.** 
**Also note: can be different from year of death if the tree died before the ring was put on**

<still need to do this>

# Re-assign death year for best guess of skeleton plot matching, then try plotting master chronology to compare to live tree master chronology

```{r}
dfatest <- dfa %>% 
  filter(tree %in% c(died2017, died2016, died2015)) %>% 
  mutate(Year = ifelse(tree %in% died2016, Year-1, Year)) %>% 
  mutate(Year = ifelse(tree %in% died2015, Year-2, Year))
head(dfatest)
```

# Plot with standard errors*2

```{r}
dfatest <- dfatest %>% 
  filter(Year > 2008) %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
```

```{r}
ggplot(data = dfatest, aes(x = Year, y = mean_gr_year))+
  geom_point(col = "red")+
  geom_line(col = "red")+
  geom_errorbar(aes(x = Year,ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = Year, y = min), col = "red", linetype = "dashed")+
  geom_line(aes(x = Year, y = max), col = "red", linetype = "dashed")+
  ggtitle("Dead ABCO detrended means +/- 2 SE")+
  ylab("Mean detrended ring widths") +
  scale_x_continuous(breaks = seq(min(dfa$Year), max(dfa$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
ggsave("../../results/figures/GrowthMortality/dead_chron_abco.jpg", width = 5, height = 5)
```

# Check the shape of 2015 trees since I didn't have many to base the physical critera on

```{r}
dfatest2015 <- dfatest %>% 
  filter(tree %in% died2015) %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
```

```{r}
ggplot(data = dfatest2015, aes(x = Year, y = mean_gr_year))+
  geom_point(col = "red")+
  geom_line(col = "red")+
  geom_errorbar(aes(x = Year,ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = Year, y = min), col = "red", linetype = "dashed")+
  geom_line(aes(x = Year, y = max), col = "red", linetype = "dashed")+
  ggtitle("Dead ABCO detrended means +/- 2 SE")+
  ylab("Mean detrended ring widths") +
  scale_x_continuous(breaks = seq(min(dfa$Year), max(dfa$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

# Save df version of the detrended dendro data
```{r}
save(dfa, file = "../../compiled_data/fir_rings_detrended_dead.Rdata")
```

# Summarize detrended dead growth to compare to detrended live growth
```{r}
dfa %>% 
  filter(Year %in% (c(2014, 2015, 2016))) %>% 
  ggplot()+
  geom_boxplot(aes(y = growth_detrended))
dfa %>% 
  filter(Year %in% (c(2014, 2015, 2016))) %>% 
  ungroup() %>% 
  summarize(mean(growth_detrended))
```

```{r}
dfp %>% 
  filter(Year %in% (c(2014, 2015, 2016))) %>% 
  ggplot()+
  geom_boxplot(aes(y = growth_detrended))
dfp %>% 
  filter(Year %in% (c(2014, 2015, 2016))) %>% 
  ungroup() %>% 
  summarize(mean(growth_detrended))
```



