---
title: "Dendro_analysis"
author: "Carmen Tubbesing"
date: "5/18/2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
require(MASS)
```

```{r}
load("../../data/GrowthMortality/rwl.Rdata")
load("../../data/GrowthMortality/dendro_all_vars.Rdata")
load(file = "../../results/data/GrowthMortality/died2017.Rdata")
load(file = "../../results/data/GrowthMortality/died2016.Rdata")
load(file = "../../results/data/GrowthMortality/died2015.Rdata")
```

# Reconfigure rwl to ease analysis
```{r}
head(rwl[300:321])
rwl <- t(rwl)
head(rwl)
names <- row.names(rwl)
```

```{r}
df <- tbl_df(rwl)
df <- df %>% 
  mutate(tree = names)
head(df)
```

# Join reconfigured rwl data with dendro_all_vars data
```{r}
join <- join %>% 
  mutate(dia_mm = (BAS_DIA_1_mm + BAS_DIA_2_mm)/2) %>% 
  dplyr::select(tree, PAIR, SPECIES, DEAD_ALIVE, dia_mm, HEIGHT) %>% 
  distinct() %>% 
  arrange(tree)
head(join)
df <- left_join(df, join) 
```

# Make death binary
```{r}
df <- df %>% 
  mutate(dead = ifelse(DEAD_ALIVE == "DEAD", 1, 0))
```

# FIRS

# Narrow down to just firs
```{r}
dfa <- df %>% 
  filter(SPECIES == "ABCO") %>% 
  distinct()
dfa
```


# Find relative growth for last 7 years
```{r}
dfa <- dfa %>% 
  rename(gr1 = `2016`, 
         gr2 = `2015`, 
         gr3 = `2014`, 
         gr4 = `2013`, 
         gr5 = `2012`,
         gr6 = `2011`,
         gr7 = `2010`) %>% 
  rename(dia1 = dia_mm) %>% 
  mutate(dia2 = dia1 - gr1*2) %>%
  mutate(dia3 = dia1 - gr1*2 - gr2*2) %>% 
  mutate(dia4 = dia1 - gr1*2 - gr2*2 - gr3*2) %>% 
  mutate(dia5 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2) %>% 
  mutate(dia6 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2 - gr5*2) %>% 
  mutate(dia7 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2 - gr5*2 - gr6*2) %>% 
  mutate(dia8 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2 - gr5*2 - gr6*2 - gr7*2) %>% 
  mutate(gr1 = (dia1 - dia2)/dia2) %>% 
  mutate(gr2 = (dia2 - dia3)/dia3) %>% 
  mutate(gr3 = (dia3 - dia4)/dia4) %>% 
  mutate(gr4 = (dia4 - dia5)/dia5) %>% 
  mutate(gr5 = (dia5 - dia6)/dia6) %>% 
  mutate(gr6 = (dia6 - dia7)/dia7) %>% 
  mutate(gr7 = (dia7 - dia8)/dia8)
```


# Figure out how many years to go back by seeing how far the growth records go back
```{r}
dfa %>% 
  filter(is.na(gr4)) %>% 
  dplyr::select(tree, DEAD_ALIVE) %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(n())
```

Since there are several trees withonly 3 years of growth, I'm going to stick with going 3 years back.
# Average relative growth for different years depending on when the tree died

## Check years of death from lists of trees that died in different years

### Check that all ABCO are in one of the lists
```{r}
dfa %>% 
  filter(!PAIR %in% c(died2017, died2016, died2015)) %>% 
  dplyr::select(tree) %>% 
  distinct() %>% 
  arrange(tree) 
```

### Delete that one
```{r}
dfa <- dfa %>% 
  filter(!PAIR %in% c(98, 202))
```

## Average growth based on pair's year of death
```{r}
dfa <- dfa %>% 
  mutate(death_year = ifelse(PAIR %in% died2017, 2017, 
                             ifelse(PAIR %in% died2016, 2016, 
                             ifelse(PAIR %in% died2015, 2015, NA)))) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2017, (gr1+gr2+gr3)/3, 999)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016 & DEAD_ALIVE == "DEAD", (gr1+gr2+gr3)/3, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2016 & DEAD_ALIVE == "ALIVE", (gr2+gr3+gr4)/3, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015 & DEAD_ALIVE == "DEAD", (gr1+gr2+gr3)/3, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(PAIR %in% died2015 & DEAD_ALIVE == "ALIVE", (gr3+gr4+gr5)/3, gr_rel_ave))
```

### If there are fewer than 3 years of growth for live trees starting at the year its pair died, bump both # years down
```{r}
dfa %>% 
  filter(is.na(gr_rel_ave)) %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, tree, PAIR, gr1, gr2, gr3, gr4, gr5, death_year, gr_rel_ave) %>% 
  arrange(PAIR)
```

```{r}
dfa <- dfa %>% 
  mutate(gr_rel_ave = ifelse(tree == "141B", (gr3+gr4)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "141", (gr1+gr2)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "230B", (gr2+gr3)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "230", (gr1+gr2)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "33B", (gr2+gr3)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "33", (gr1+gr2)/2, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "228B", gr3, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "228", gr1, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "229", gr1, gr_rel_ave)) %>% 
  mutate(gr_rel_ave = ifelse(tree == "229B", gr3, gr_rel_ave)) 
  
dfa %>% 
  filter(PAIR == 33) %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, tree, PAIR, gr1, gr2, gr3, gr4, gr5, death_year, gr_rel_ave) 
```

#### Check for seedling with 0 average growth
```{r}
dfa %>% 
  filter(gr_rel_ave == 0) %>% 
  dplyr::select(SPECIES, DEAD_ALIVE, tree, PAIR, gr1, gr2, gr3, death_year, gr_rel_ave) %>% 
  arrange(PAIR)
```

## Add column for log growth
```{r}
dfa <- dfa %>% 
  mutate(log_gr_rel_ave = log(gr_rel_ave))
summary(dfa$log_gr_rel_ave)
hist(dfa$gr_rel_ave)
hist(dfa$log_gr_rel_ave)
```

# Plot growth by year
```{r}
df2 <- gather(dfa, key = "year", value = "gr_rel", c(gr1, gr2, gr3, gr4)) %>% 
  mutate(year = ifelse(year == "gr1", "outer ring", year)) %>% 
  mutate(year = ifelse(year == "gr2", "outer ring - 1", year)) %>% 
  mutate(year = ifelse(year == "gr3", "outer ring - 2", year)) %>% 
  mutate(year = ifelse(year == "gr4", "outer ring - 3", year)) %>% 
  mutate(year = as.factor(year))

df2$year <-  factor(df2$year, levels = rev(levels(df2$year)))

df2 %>% dplyr::select(year, gr_rel, tree, SPECIES, DEAD_ALIVE)
ggplot(df2)+
  geom_boxplot(aes(x = year, y = log(gr_rel), fill = DEAD_ALIVE))+
  facet_wrap(~SPECIES)+
  ylim(-3, .5)
```

# Plot relative growth comparison between live and dead
```{r}
ggplot(dfa)+
  geom_boxplot(aes(x = DEAD_ALIVE, y = gr_rel_ave))+
  geom_point(aes(x = DEAD_ALIVE, y = mean(gr_rel_ave)), col = "red")
```

```{r}
ggplot(dfa)+
  geom_boxplot(aes(x = DEAD_ALIVE, y = log_gr_rel_ave))+
  geom_point(aes(x = DEAD_ALIVE, y = mean(log_gr_rel_ave)), col = "red")
```

# Test with glm
```{r}
cor(dfa$gr_rel_ave, dfa$dead)
cor(log(dfa$gr_rel_ave), dfa$dead)
summary(glm(dead ~ log_gr_rel_ave, family = binomial(link = "logit"), data = dfa))
```

# Visually compare distributions
```{r}
dfa <- dfa %>% ungroup()
means <- dfa %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfa %>% 
  group_by(DEAD_ALIVE) %>% 
 summarize(sd = sd(log_gr_rel_ave))
means
sd
```

```{r}
x <- seq(min(dfa$log_gr_rel_ave), max(dfa$log_gr_rel_ave), length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

```{r}
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

```{r}
ggplot()+
  geom_histogram(data = dfa %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 25, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfa %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 25, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")
  
ggsave("../../results/figures/GrowthMortality/histogram_dead_alive_abco_dendro.png")

```


# PINES

# Narrow down to just pines
```{r}
dfp <- df %>% 
  filter(SPECIES == "PIPO") %>% 
  distinct()
dfp
```

# Find average relative growth for last 3 years - WHICH 3 YEARS DEPENDS ON WHETHER IT'S DEAD OR ALIVE
```{r}
dfp <- dfp %>% 
  mutate(dead = ifelse(DEAD_ALIVE == "DEAD", 1, 0))

dfp <- dfp %>% 
  dplyr::select(`2013`, `2014`, `2015`, `2016`, tree, SPECIES, DEAD_ALIVE, dia_mm, dead) %>%
  rename(gr1 = `2016`, gr2 = `2015`, gr3 = `2014`, gr4 = `2013`) %>% 
  rename(dia1 = dia_mm) %>% 
  mutate(dia2 = dia1 - gr1*2) %>%
  mutate(dia3 = dia1 - gr1*2 - gr2*2) %>% 
  mutate(dia4 = dia1 - gr1*2 - gr2*2 - gr3*2) %>% 
  mutate(dia5 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2) %>% 
  mutate(gr_rel_1 = (dia1 - dia2)/dia1) %>% 
  mutate(gr_rel_2 = (dia2 - dia3)/dia2) %>% 
  mutate(gr_rel_3 = (dia3 - dia4)/dia3) %>% 
  mutate(gr_rel_4 = (dia4 - dia5)/dia4) 
dfp <- dfp %>% 
  rowwise() %>% 
  mutate(gr_rel_ave = ifelse(dead == 1, 
                             mean(c(gr_rel_1, gr_rel_2), na.rm=T),
                             mean(c(gr_rel_3, gr_rel_4), na.rm=T)))

summary(dfp$gr_rel_ave)
hist(dfp$gr_rel_ave)
```

## Throw out trees with too few years to calculate growth
```{r}
dfp <- dfp %>% 
  filter(!is.na(gr_rel_ave))
```

```{r}
ggplot(dfp)+
  geom_boxplot(aes(x = DEAD_ALIVE, y = gr_rel_ave))
```

# Test with glm
```{r}
dfp <- dfp %>% 
  mutate(dead = ifelse(DEAD_ALIVE == "DEAD", 1, 0))
cor(dfp$gr_rel_ave, dfp$dead)
cor(log(dfp$gr_rel_ave), dfp$dead)
summary(glm(dead ~ log(gr_rel_ave), family = binomial(link = "logit"), data = dfp))
dfp$log_gr_rel_ave <- log(dfp$gr_rel_ave)
```

# Visually compare distributions
```{r}
dfp <- dfp %>% ungroup()
means <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
 summarize(sd = sd(log_gr_rel_ave))
means
sd
```

```{r}
x <- seq(min(dfp$log_gr_rel_ave), max(dfp$log_gr_rel_ave), length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

```{r}
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

```{r}
ggplot()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")
  
ggsave("../../results/figures/GrowthMortality/histogram_dead_alive_pipo_dendro.png")
```
