---
title: "Dendro_analysis"
author: "Carmen Tubbesing"
date: "5/18/2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
require(MASS)
```

```{r}
load("../../data/GrowthMortality/rwl.Rdata")
load("../../data/GrowthMortality/dendro_all_vars.Rdata")
```

# Reconfigure rwl to ease analysis
```{r}
head(rwl[300:321])
rwl <- t(rwl)
head(rwl)
names <- row.names(rwl)
```

```{r}
df <- tbl_df(rwl)
df <- df %>% 
  mutate(tree = names)
head(df)
```

```{r}
join <- join %>% 
  mutate(dia_mm = (BAS_DIA_1_mm + BAS_DIA_2_mm)/2) %>% 
  dplyr::select(tree, SPECIES, DEAD_ALIVE, dia_mm)
df <- left_join(df, join) 
summary(as.factor(df$SPECIES))
summary(as.factor(df$DEAD_ALIVE))
summary(df$dia_mm)
```


# Plot growth by year
```{r}
df2 <- gather(df, key = "year", value = "gr_rel", c(`2016`, `2015`, `2014`, `2013`)) %>% 
  mutate(year = ifelse(year == "2016", "outer ring", year)) %>% 
  mutate(year = ifelse(year == "2015", "outer ring - 1", year)) %>% 
  mutate(year = ifelse(year == "2014", "outer ring - 2", year)) %>% 
  mutate(year = ifelse(year == "2013", "outer ring - 3", year)) %>% 
  mutate(year = as.factor(year))

df2$year <-  factor(df2$year, levels = rev(levels(df2$year)))

df2 %>% dplyr::select(year, gr_rel, tree, SPECIES)
ggplot(df2)+
  geom_boxplot(aes(x = year, y = log(gr_rel), fill = SPECIES))+
  facet_wrap(~SPECIES)+
  ylim(-3, .5)
```


# FIRS

# Narrow down to just firs
```{r}
dfa <- df %>% 
  filter(SPECIES == "ABCO") %>% 
  distinct()
dfa
```

# Find average relative growth for last 3 years
```{r}
dfa <- dfa %>% 
  mutate(dead = ifelse(DEAD_ALIVE == "DEAD", 1, 0))

dfa <- dfa %>% 
  dplyr::select(`2012`, `2013`, `2014`, `2015`, `2016`, tree, SPECIES, DEAD_ALIVE, dia_mm, dead) %>%
  rename(gr1 = `2016`, gr2 = `2015`, gr3 = `2014`, gr4 = `2013`, gr5 = `2012`) %>% 
  rename(dia1 = dia_mm) %>% 
  mutate(dia2 = dia1 - gr1*2) %>%
  mutate(dia3 = dia1 - gr1*2 - gr2*2) %>% 
  mutate(dia4 = dia1 - gr1*2 - gr2*2 - gr3*2) %>% 
  mutate(dia5 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2) %>% 
  mutate(dia6 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2 - gr5*2) %>% 
  mutate(gr_rel_1 = (dia1 - dia2)/dia2) %>% 
  mutate(gr_rel_2 = (dia2 - dia3)/dia3) %>% 
  mutate(gr_rel_3 = (dia3 - dia4)/dia4) %>% 
  mutate(gr_rel_4 = (dia4 - dia5)/dia5) %>% 
  mutate(gr_rel_5 = (dia5 - dia6)/dia6)
dfa <- dfa %>% 
  rowwise() %>% 
  mutate(gr_rel_ave = ifelse(dead == 1, 
                             mean(c(gr_rel_1, gr_rel_2, gr_rel_3), na.rm=T),
                             mean(c(gr_rel_2, gr_rel_3, gr_rel_4), na.rm=T)))

summary(dfa$gr_rel_ave)
hist(dfa$gr_rel_ave)
```

# Plot growth by year
```{r}
df2 <- gather(dfa, key = "year", value = "gr_rel", c(gr1, gr2, gr3, gr4)) %>% 
  mutate(year = ifelse(year == "gr1", "outer ring", year)) %>% 
  mutate(year = ifelse(year == "gr2", "outer ring - 1", year)) %>% 
  mutate(year = ifelse(year == "gr3", "outer ring - 2", year)) %>% 
  mutate(year = ifelse(year == "gr4", "outer ring - 3", year)) %>% 
  mutate(year = as.factor(year))

df2$year <-  factor(df2$year, levels = rev(levels(df2$year)))

df2 %>% dplyr::select(year, gr_rel, tree, SPECIES, DEAD_ALIVE)
ggplot(df2)+
  geom_boxplot(aes(x = year, y = log(gr_rel), fill = DEAD_ALIVE))+
  facet_wrap(~SPECIES)+
  ylim(-3, .5)
```



## Throw out trees with too few years to calculate growth
```{r}
dfa <- dfa %>% 
  filter(!is.na(gr_rel_ave))
```


```{r}
ggplot(dfa)+
  geom_boxplot(aes(x = DEAD_ALIVE, y = gr_rel_ave))
```

# Test with glm
```{r}
cor(dfa$gr_rel_ave, dfa$dead)
cor(log(dfa$gr_rel_ave), dfa$dead)
summary(glm(dead ~ log(gr_rel_ave), family = binomial(link = "logit"), data = dfa))
dfa$log_gr_rel_ave <- log(dfa$gr_rel_ave)
```

# Visually compare distributions
```{r}
dfa <- dfa %>% ungroup()
means <- dfa %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfa %>% 
  group_by(DEAD_ALIVE) %>% 
 summarize(sd = sd(log_gr_rel_ave))
means
sd
```

```{r}
x <- seq(min(dfa$log_gr_rel_ave), max(dfa$log_gr_rel_ave), length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

```{r}
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

```{r}
ggplot()+
  geom_histogram(data = dfa %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfa %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")
  
ggsave("../../results/figures/GrowthMortality/histogram_dead_alive_abco_dendro.png")

```


# PINES

# Narrow down to just pines
```{r}
dfp <- df %>% 
  filter(SPECIES == "PIPO") %>% 
  distinct()
dfp
```

# Find average relative growth for last 3 years - WHICH 3 YEARS DEPENDS ON WHETHER IT'S DEAD OR ALIVE
```{r}
dfp <- dfp %>% 
  mutate(dead = ifelse(DEAD_ALIVE == "DEAD", 1, 0))

dfp <- dfp %>% 
  dplyr::select(`2013`, `2014`, `2015`, `2016`, tree, SPECIES, DEAD_ALIVE, dia_mm, dead) %>%
  rename(gr1 = `2016`, gr2 = `2015`, gr3 = `2014`, gr4 = `2013`) %>% 
  rename(dia1 = dia_mm) %>% 
  mutate(dia2 = dia1 - gr1*2) %>%
  mutate(dia3 = dia1 - gr1*2 - gr2*2) %>% 
  mutate(dia4 = dia1 - gr1*2 - gr2*2 - gr3*2) %>% 
  mutate(dia5 = dia1 - gr1*2 - gr2*2 - gr3*2 - gr4*2) %>% 
  mutate(gr_rel_1 = (dia1 - dia2)/dia1) %>% 
  mutate(gr_rel_2 = (dia2 - dia3)/dia2) %>% 
  mutate(gr_rel_3 = (dia3 - dia4)/dia3) %>% 
  mutate(gr_rel_4 = (dia4 - dia5)/dia4) 
dfp <- dfp %>% 
  rowwise() %>% 
  mutate(gr_rel_ave = ifelse(dead == 1, 
                             mean(c(gr_rel_1, gr_rel_2), na.rm=T),
                             mean(c(gr_rel_3, gr_rel_4), na.rm=T)))

summary(dfp$gr_rel_ave)
hist(dfp$gr_rel_ave)
```

## Throw out trees with too few years to calculate growth
```{r}
dfp <- dfp %>% 
  filter(!is.na(gr_rel_ave))
```

```{r}
ggplot(dfp)+
  geom_boxplot(aes(x = DEAD_ALIVE, y = gr_rel_ave))
```

# Test with glm
```{r}
dfp <- dfp %>% 
  mutate(dead = ifelse(DEAD_ALIVE == "DEAD", 1, 0))
cor(dfp$gr_rel_ave, dfp$dead)
cor(log(dfp$gr_rel_ave), dfp$dead)
summary(glm(dead ~ log(gr_rel_ave), family = binomial(link = "logit"), data = dfp))
dfp$log_gr_rel_ave <- log(dfp$gr_rel_ave)
```

# Visually compare distributions
```{r}
dfp <- dfp %>% ungroup()
means <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
 summarize(sd = sd(log_gr_rel_ave))
means
sd
```

```{r}
x <- seq(min(dfp$log_gr_rel_ave), max(dfp$log_gr_rel_ave), length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

```{r}
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

```{r}
ggplot()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")
  
ggsave("../../results/figures/GrowthMortality/histogram_dead_alive_pipo_dendro.png")
```
