---
title: "Seedling dendro inventory"
author: "Carmen"
date: "June 25, 2018"
output: 
    html_document:
      toc: TRUE
---

```{r, include =F}
library(readxl)
library(ggplot2)
library(dplyr)
```

# Read in data
```{r}
file <- "~/../Dropbox (Stephens Lab)/SORTIE/Growth_mortality/data/details/compiled/Mort_details_compiled_June25_2018.xlsx"
df <- read_excel(file, sheet = 2)
boards <- read_excel(file, sheet = 3)
```

# Clean

## Take out NA rows
```{r}
df <- df %>% filter(!is.na(SEEDLING))
```

## Make everything uppercase
```{r}
df <- df %>% 
  mutate(SPECIES = toupper(SPECIES))
```

```{r}
df <- df %>% 
  mutate(SEEDLING = toupper(SEEDLING))
boards <- boards %>% 
  mutate(SEEDLING = toupper(SEEDLING))
```

```{r}
df <- df %>% 
  mutate(DEAD_ALIVE = toupper(DEAD_ALIVE))
```

### Correct -A and -P
```{r}
setdiff(boards$SEEDLING, df$SEEDLING)
setdiff(df$SEEDLING, boards$SEEDLING)
```


# Combine the two sheets

## Rename boards notes
```{r}
boards <- boards %>% 
  rename(Mount_notes = Notes)
```

## Sync up seedling names in the two sheets
```{r}
nrow(df)
df <- full_join(df, boards)
nrow(df)
```


# Find boards with fewer than 5 seedlings
```{r}
df %>% 
  group_by(Board) %>% 
  summarise(n = n()) %>% 
  filter(n != 5) 
```



# If two rows for one seedling, remove the one that's not backup

## Create duplicates list
```{r}
dups <- df %>% 
  group_by(SEEDLING) %>% 
  filter(n()>1) %>% 
  select(SEEDLING, Backup) %>% 
  arrange(SEEDLING)
dups
```

## Take out seedling that's duplicated but one isn't backup
```{r}
dups <- dups %>% filter(!is.na(SEEDLING))
dups_nobackup <- dups %>% 
  group_by(SEEDLING) %>% 
  tally(Backup) %>% 
  filter(n==0)
dups_nobackup
dups <- dups %>% filter(!SEEDLING %in% dups_nobackup$SEEDLING)
```

## Remove dups rows that aren't backups from df
```{r}
df <- df %>% 
  filter(!(SEEDLING %in% dups$SEEDLING & Backup==0)) 
nrow(df)
```

# See whether all seedlings that weren't thrown out have been mounted
```{r}
summary(as.factor(df$thrown_out))
summary(as.factor(df$Board))
df %>% 
  filter(thrown_out==0) %>% 
  filter(is.na(Board))
```

# Compare # of seedlings of each species that weren't thrown out and have been mounted

## All
```{r}
df %>% 
  filter(thrown_out==0) %>% 
  filter(!is.na(Board)) %>% 
  group_by(SPECIES) %>% 
  summarise(n())
```

## Dead
```{r}
df %>% 
  filter(thrown_out==0) %>% 
  filter(!is.na(Board)) %>% 
  filter(DEAD_ALIVE == "DEAD") %>% 
  group_by(SPECIES) %>% 
  summarise(n())
```

# Look at dia distributions
```{r}
ggplot(df %>% filter(SPECIES == "ABCO"))+
  geom_histogram(aes(x = BAS_DIA_1_mm), bins = 40)
ggplot(df)+
  geom_histogram(aes(x = BAS_DIA_1_mm, group = SPECIES, fill = SPECIES), bins = 40, position = "dodge2")
```

# Find seedlings that may not have been read yet whose mates are thrown out
```{r}
df %>% 
  filter(thrown_out == 1) %>% 
  filter(Board >65)
```

# Find seedlings that weren't thrown out but maybe should be
```{r}
df %>% 
  filter(thrown_out=="0" & PATH_DAMAGE != "0") %>% 
  select(SEEDLING, SPECIES, Board, PATH_DAMAGE, Notes, Mount_notes)
```
