---
title: "live chronology"
author: "Carmen"
date: "May 16, 2019"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
require(tidyverse)
require(dplR)
require(readxl)
```

# Read in and compile dendro files
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../data/GrowthMortality/live_pipo_rwl.Rdata")
rwlp <- rwl_live_pipo
load(file = "../../data/GrowthMortality/live_abco_rwl.Rdata")
rwla <- rwl_live_abco
load(file = "../../data/GrowthMortality/dead_pipo_rwl.Rdata")
load(file = "../../data/GrowthMortality/dead_abco_rwl.Rdata")
```

## Add dead ABCO with green foliage to the live chronology, since we know they died in 2017
```{r}
rwla_dead_green <- rwl_dead_abco[,colnames(rwl_dead_abco) %in% c(201, 100, 30, 127, 16, 211, 32, 52, 55, 169, 215)]
rwla_dead_green
```

```{r}
#rwla <- cbind(rwla, rwla_dead_green)
```

# Plot data before detrending

## Gather/restructure

### PIPO
```{r}
rwlp2 <- rwlp %>% 
  mutate(year = row.names(rwlp))
rwlp2 <- gather(rwlp2, key = "tree", value = "growth", 1:ncol(rwlp)) %>% 
  filter(!is.na(growth))
head(rwlp2)
```

```{r}
rwlp2_summary <- rwlp2 %>% 
  group_by(tree) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
rwlp2_summary
```

### ABCO
```{r}
rwla2 <- rwla %>% 
  mutate(year = row.names(rwla))
rwla2 <- gather(rwla2, key = "tree", value = "growth", 1:ncol(rwla)) %>% 
  filter(!is.na(growth))
head(rwla2)
```

```{r}
rwla2_summary <- rwla2 %>% 
  group_by(tree) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
rwla2_summary
```

## Plot by # of rings in the series

### PIPO

#### Pines with 8 years:
```{r}
ggplot(rwlp2 %>% filter(tree %in%  rwlp2_summary$tree[1:6])) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

2 pos, 4 neg

#### Pines with 7 years:
```{r}
ggplot(rwlp2 %>% filter(tree %in%  unlist(rwlp2_summary[rwlp2_summary$n == 7,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

5 pos, 3 neg

#### Pines with 6 years:
```{r}
ggplot(rwlp2 %>% filter(tree %in%  unlist(rwlp2_summary[rwlp2_summary$n == 6,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

5 pos, 8 neg, 2 zero

**Of the pines with 6, 7, or 8 rings, 12 have positive slopes, 15 have negative slopes, and 2 have slopes very near zero. In other words, nearly half have positive or zero slopes, so applying a negative exponential curve doesn't make much sense. I'm going to apply a horizontal line curve instead.**


###  ABCO

#### Firs with 10+ years:
```{r}
ggplot(rwla2 %>% filter(tree %in%  rwla2_summary$tree[1:7])) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

4 pos, 3 neg

#### Firs with 9 years:
```{r}
ggplot(rwla2 %>% filter(tree %in%  unlist(rwla2_summary[rwla2_summary$n == 9,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

4 pos, 5 neg

#### Firs with 8 years:
```{r}
ggplot(rwla2 %>% filter(tree %in%  unlist(rwla2_summary[rwla2_summary$n == 8,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

5 pos, 3 neg

**Of the firs with 8-13 rings, 13 have positive slopes and 11 have negative slopes. In other words, over half have positive slopes, so applying a negative exponential curve doesn't make much sense. I'm going to apply a horizontal line curve instead.**

# Detrend

## PIPO
```{r}
rwlp_dt <- detrend(rwl = rwlp, method = "Mean")
```

## ABCO
```{r}
rwla_dt <- detrend(rwl = rwla, method = "Mean")
```

# Build mean chronology

## PIPO
```{r}
rwlp_chron <- chron(rwlp_dt, prefix = "CAM")
rwlp_chron
plot(rwlp_chron, main ="PIPO")
```


```{r}
ggplot(rwlp_chron %>% 
           mutate(year = row.names(rwlp_chron)) %>% 
         ungroup() 
       )+
  geom_point(aes(x = year, y = CAMstd))+
  geom_line(aes(x = year, y = CAMstd, group = 1))+
  ggtitle("PIPO")
```


## ABCO
```{r}
rwla_chron <- chron(rwla_dt, prefix = "CAM")
rwla_chron
plot(rwla_chron, main = "ABCO")
```

```{r}
ggplot(rwla_chron %>% 
           mutate(year = row.names(rwla_chron)) %>% 
         ungroup() 
       )+
  geom_point(aes(x = year, y = CAMstd))+
  geom_line(aes(x = year, y = CAMstd, group = 1))+
  ggtitle("ABCO")
```

# Plot detrended seedlings

## Reshape detrended means with error bars

### PIPO
```{r}
dfp <- as.data.frame(t(rwlp_dt))
dfp$tree <- row.names(dfp)
dfp <- gather(dfp, key = "Year", value = "growth_detrended", c(1:13))
dfp <- dfp %>% 
  filter(!is.na(growth_detrended))
```

#### Take out trees with very few rings

##### Find tallies of lengths of each tree's series
```{r}
dfp %>% 
  group_by(tree) %>% 
  summarize(n_rings = n()) %>% 
  group_by(n_rings) %>% 
  summarize(n_n = n()) %>% 
  arrange(n_rings)
```

So, 21 trees have only 4 rings

<!-- ##### Take out trees with 3 or fewer rings -->
<!-- ```{r} -->
<!-- nrow(dfp) -->
<!-- dfp <- dfp %>%  -->
<!--   group_by(tree) %>%  -->
<!--   filter(n()>3) -->
<!-- nrow(dfp) -->
<!-- ``` -->


### ABCO
```{r}
dfa <- as.data.frame(t(rwla_dt))
dfa$tree <- row.names(dfa)
dfa <- gather(dfa, key = "Year", value = "growth_detrended", c(1:13))
dfa <- dfa %>% 
  filter(!is.na(growth_detrended))
```

#### Take out trees with very few rings

##### Find tallies of lengths of each tree's series
```{r}
dfa %>% 
  group_by(tree) %>% 
  summarize(n_rings = n()) %>% 
  group_by(n_rings) %>% 
  summarize(n_n = n()) %>% 
  arrange(n_rings)
```

So, 7 trees have only 4 rings

##### Take out trees with 4 or fewer rings
# ```{r}
# nrow(dfa)
# dfa <- dfa %>% 
#   group_by(tree) %>% 
#   filter(n()>4)
# nrow(dfa)
# ```

## Plot

### PIPO

#### Calculate standard errors
```{r}
dfpstd <- dfp %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
dfpstd
```

#### Plot
```{r}
dfpstd <- dfpstd %>% 
  ungroup() %>% 
  mutate(Year = as.numeric(Year))
ggplot(dfpstd  %>% filter(Year >2010)) + 
  geom_point(aes(x = Year, y = mean_gr_year), col = "red")+
  geom_line(aes(x = Year, y = mean_gr_year), col = "red")+
  geom_errorbar(aes(x = Year,ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = Year, y = min), col = "red", linetype = "dashed")+
  geom_line(aes(x = Year, y = max), col = "red", linetype = "dashed")+
  ggtitle("Live PIPO detrended means +/- 2 SE")+
   ylab("Mean detrended ring widths") +
  scale_x_continuous(breaks = seq(min(dfpstd$Year), max(dfpstd$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
ggsave("../../results/figures/GrowthMortality/live_chron_pipo_nogreendead.jpg", width = 5, height = 5)
```

### ABCO

#### Calculate standard errors
```{r}
dfastd <- dfa %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
```

#### Plot
```{r, warning = F}
dfastd <- dfastd %>% 
  ungroup() %>% 
  mutate(Year = as.numeric(Year))
ggplot(dfastd %>% filter(Year >2008)) + 
  geom_point(aes(x = Year, y = mean_gr_year), col = "red")+
  geom_line(aes(x = Year, y = mean_gr_year), col = "red")+
  geom_errorbar(aes(x = Year,ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = Year, y = min), col = "red", linetype = "dashed")+
  geom_line(aes(x = Year, y = max), col = "red", linetype = "dashed")+
  ggtitle("Live ABCO detrended means +/- 2 SE")+
  ylab("Mean detrended ring widths") +
  scale_x_continuous(breaks = seq(min(dfastd$Year), max(dfastd$Year), by = 1))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
ggsave("../../results/figures/GrowthMortality/live_chron_abco_nogreendead.jpg", width = 5, height = 5)
```

# Summarize detrended live growth to compare to detrended dead growth
```{r}
dfa %>% 
  ungroup() %>% 
  filter(Year %in% (c(2013, 2014, 2015))) %>% 
  ggplot()+
  geom_boxplot(aes(y = growth_detrended))
dfa %>% 
  filter(Year %in% (c(2013, 2014, 2015))) %>% 
  ungroup() %>% 
  summarize(mean(growth_detrended))

```

```{r}
dfp %>% 
  ungroup() %>% 
  filter(Year %in% (c(2013, 2014, 2015))) %>% 
  ggplot()+
  geom_boxplot(aes(y = growth_detrended))
dfp %>% 
  filter(Year %in% (c(2013, 2014, 2015))) %>% 
  ungroup() %>% 
  summarize(mean(growth_detrended))

```

# Save detrended data
```{r}
save(dfa, file = "../../compiled_data/live_chron_abco.Rdata")
save(dfp, file = "../../compiled_data/live_chron_pipo.Rdata")
```

