---
title: "live chrnology"
author: "Carmen"
date: "May 16, 2019"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
require(tidyverse)
require(dplR)
require(readxl)
```

# Read in and compile dendro files
```{r}
load(file = "../../data/GrowthMortality/live_pipo_rwl.Rdata")
rwlp <- rwl_live_pipo
load(file = "../../data/GrowthMortality/live_abco_rwl.Rdata")
rwla <- rwl_live_abco
```

# Plot data before detrending

## Gather/restructure

### PIPO
```{r}
rwlp2 <- rwlp %>% 
  mutate(year = row.names(rwlp))
rwlp2 <- gather(rwlp2, key = "tree", value = "growth", 1:ncol(rwlp)) %>% 
  filter(!is.na(growth))
head(rwlp2)
```

```{r}
rwlp3 <- rwlp2 %>% 
  group_by(year) %>% 
  mutate(means = mean(na.omit(growth))) %>% 
  mutate(n = n()) %>% 
  ungroup()
```

```{r}
rwlp3_summary <- rwlp3 %>% 
  group_by(tree) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
rwlp3_summary
```

### ABCO
```{r}
rwla2 <- rwla %>% 
  mutate(year = row.names(rwla))
rwla2 <- gather(rwla2, key = "tree", value = "growth", 1:ncol(rwla)) %>% 
  filter(!is.na(growth))
head(rwla2)
```

```{r}
rwla3 <- rwla2 %>% 
  group_by(year) %>% 
  mutate(means = mean(na.omit(growth))) %>% 
  mutate(n = n()) %>% 
  ungroup()
```

```{r}
rwla3_summary <- rwla3 %>% 
  group_by(tree) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
rwla3_summary
```

## Plot by # of rings in the series

### PIPO

#### Pines with 8 years:
```{r}
ggplot(rwlp3 %>% filter(tree %in%  rwlp3_summary$tree[1:6])) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

2 pos, 4 neg

#### Pines with 7 years:
```{r}
ggplot(rwlp3 %>% filter(tree %in%  unlist(rwlp3_summary[rwlp3_summary$n == 7,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

5 pos, 3 neg

#### Pines with 6 years:
```{r}
ggplot(rwlp3 %>% filter(tree %in%  unlist(rwlp3_summary[rwlp3_summary$n == 6,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

5 pos, 8 neg, 2 zero

**Of the pines with 6, 7, or 8 rings, 12 have positive slopes, 15 have negative slopes, and 2 have slopes very near zero. In other words, nearly half have positive or zero slopes, so applying a negative exponential curve doesn't make much sense. I'm going to apply a horizontal line curve instead.**


###  ABCO

#### Firs with 10+ years:
```{r}
ggplot(rwla3 %>% filter(tree %in%  rwla3_summary$tree[1:7])) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

4 pos, 3 neg

#### Firs with 9 years:
```{r}
ggplot(rwla3 %>% filter(tree %in%  unlist(rwla3_summary[rwla3_summary$n == 9,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

4 pos, 5 neg

#### Firs with 8 years:
```{r}
ggplot(rwla3 %>% filter(tree %in%  unlist(rwla3_summary[rwla3_summary$n == 8,"tree"]))) + 
  stat_smooth(aes(x = year, y = growth, group = tree, col = tree), method = "lm", fill = "white",se = F)+
  geom_line(aes(x = year, y = growth, group = tree, col = tree), size = 1, linetype = "dashed")
```

5 pos, 3 neg

**Of the firs with 8-13 rings, 13 have positive slopes and 11 have negative slopes. In other words, over half have positive slopes, so applying a negative exponential curve doesn't make much sense. I'm going to apply a horizontal line curve instead.**

# Detrend

## PIPO
```{r}
rwlp_dt <- detrend(rwl = rwlp, method = "Mean")
```

## ABCO
```{r}
rwla_dt <- detrend(rwl = rwla, method = "Mean")
```

# Build mean chronology

## PIPO
```{r}
rwlp_chron <- chron(rwlp_dt, prefix = "CAM")
rwlp_chron
plot(rwlp_chron, main ="PIPO")
```


```{r}
ggplot(rwlp_chron %>% 
           mutate(year = row.names(rwlp_chron)) %>% 
         ungroup() 
       )+
  geom_point(aes(x = year, y = CAMstd))+
  geom_line(aes(x = year, y = CAMstd, group = 1))+
  ggtitle("PIPO")
```


## ABCO
```{r}
rwla_chron <- chron(rwla_dt, prefix = "CAM")
rwla_chron
plot(rwla_chron, main = "ABCO")
```

```{r}
ggplot(rwla_chron %>% 
           mutate(year = row.names(rwla_chron)) %>% 
         ungroup() 
       )+
  geom_point(aes(x = year, y = CAMstd))+
  geom_line(aes(x = year, y = CAMstd, group = 1))+
  ggtitle("ABCO")
```

# Plot detrended seedlings

## Reshape detrended means with error bars

### PIPO
```{r}
dfp <- as.data.frame(t(rwlp_dt))
dfp$tree <- row.names(dfp)
dfp <- gather(dfp, key = "Year", value = "growth_detrended", c(1:13))
dfp <- dfp %>% 
  filter(!is.na(growth_detrended))
```

#### Take out trees with very few rings

##### Find tallies of lengths of each tree's series
```{r}
dfp %>% 
  group_by(tree) %>% 
  summarize(n_rings = n()) %>% 
  group_by(n_rings) %>% 
  summarize(n_n = n()) %>% 
  arrange(n_rings)
```

So, 21 trees have only 4 rings

##### Take out trees with 3 or fewer rings
```{r}
nrow(dfp)
dfp <- dfp %>% 
  group_by(tree) %>% 
  filter(n()>3)
nrow(dfp)
```


### ABCO
```{r}
dfa <- as.data.frame(t(rwla_dt))
dfa$tree <- row.names(dfa)
dfa <- gather(dfa, key = "Year", value = "growth_detrended", c(1:13))
dfa <- dfa %>% 
  filter(!is.na(growth_detrended))
```

#### Take out trees with very few rings

##### Find tallies of lengths of each tree's series
```{r}
dfa %>% 
  group_by(tree) %>% 
  summarize(n_rings = n()) %>% 
  group_by(n_rings) %>% 
  summarize(n_n = n()) %>% 
  arrange(n_rings)
```

So, 7 trees have only 4 rings

##### Take out trees with 4 or fewer rings
```{r}
nrow(dfa)
dfa <- dfa %>% 
  group_by(tree) %>% 
  filter(n()>4)
nrow(dfa)
```

## Plot

### PIPO

#### Calculate standard errors
```{r}
dfpstd <- dfp %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
dfpstd
```

#### Plot
```{r}
ggplot(dfpstd) + 
  geom_point(aes(x = as.numeric(Year), y = mean_gr_year), col = "red")+
  geom_line(aes(x = as.numeric(Year), y = mean_gr_year), col = "red")+
  geom_errorbar(aes(x = as.numeric(Year),ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = as.numeric(Year), y = min), col = "gray")+
  geom_line(aes(x = as.numeric(Year), y = max), col = "gray")+
  ggtitle("PIPO standardized means +/- 2 SE")
```

### ABCO

#### Calculate standard errors
```{r}
dfastd <- dfa %>% 
  group_by(Year) %>% 
  mutate(mean_gr_year = mean(growth_detrended)) %>% 
  mutate(n_year = n()) %>% 
  mutate(stderror = sd(growth_detrended)/sqrt(n_year)) %>% 
  mutate(min = mean_gr_year-2*stderror) %>% 
  mutate(max = mean_gr_year+2*stderror)
```

#### Plot
```{r}
ggplot(dfastd) + 
  geom_point(aes(x = as.numeric(Year), y = mean_gr_year), col = "red")+
  geom_line(aes(x = as.numeric(Year), y = mean_gr_year), col = "red")+
  geom_errorbar(aes(x = as.numeric(Year),ymin = min, ymax = max), width = .2, col = "gray")+
  geom_line(aes(x = as.numeric(Year), y = min), col = "gray")+
  geom_line(aes(x = as.numeric(Year), y = max), col = "gray")+
  ggtitle("ABCO standardized means +/- 2 SE")
```
