---
title: "Plot simulation results"
author: "Carmen"
date: "September 4, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include=F}
require(tidyverse)
```

# Define ymax to be used throughout both abco and pipo graphs

## Set y limits for graphs later
```{r}
ymax_set <- 10
```

# Load data

## Growth results from fire footprints
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../results/data/FireFootprints/gr_minshrub_abco.Rdata")
load(file = "../../results/data/FireFootprints/gr_maxshrub_abco.Rdata")
load(file = "../../results/data/FireFootprints/gr_medshrub_abco.Rdata")
```


# ABCO

## Load simulation results
```{r}
load(file = "../../results/coefficients/mean_coef_gr_abco.Rdata")
load(file = "../../results/coefficients/mean_coef_int_abco.Rdata")

load(file = "../../results/coefficients/coef_gr_abco.Rdata")
load(file = "../../results/coefficients/coef_int_abco.Rdata")
```

## Plot the mean coefficient and CI

### Calculate probabilities
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
dfgraph <- as.data.frame(log_gr_rel)
dfgraph$prob <- exp(mean_coef_int + mean_coef_gr*log_gr_rel)/(1+exp(mean_coef_int + mean_coef_gr*log_gr_rel))
summary(dfgraph$prob)
```

### Calculate confidence intervals to plot
```{r}
CI_low_int <- sort(coefficient_int)[25]
CI_high_int <- sort(coefficient_int)[975]
CI_low_gr <- sort(coefficient_gr)[25]
CI_high_gr <- sort(coefficient_gr)[975]
```

#### lower
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_lower <- as.data.frame(log_gr_rel)
dfgraph_lower$prob <- exp(mean_coef_int + CI_low_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_low_gr*log_gr_rel))
summary(dfgraph_lower$prob)
```

#### upper
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_upper <- as.data.frame(log_gr_rel)
dfgraph_upper$prob <- exp(mean_coef_int + CI_high_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_high_gr*log_gr_rel))
summary(dfgraph_upper$prob)
```

### Adjust probabilities to be yearly, since I found dead trees from 3 years 
```{r}
head(dfgraph)
dfgraph <- dfgraph %>% 
  mutate(prob = prob/3)
head(dfgraph_lower)
dfgraph_lower <- dfgraph_lower %>% 
  mutate(prob = prob/3)
dfgraph_upper <- dfgraph_upper %>% 
  mutate(prob = prob/3)
```

```{r}
ggplot()+
  geom_line(data = dfgraph, aes(x = exp(log_gr_rel), y = prob*100))+
  geom_line(data = dfgraph_lower, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  geom_line(data = dfgraph_upper, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  theme_bw()+
  #xlim(0.03, .27)+
  ylim(0, ymax_set)+
   theme(text = element_text(size = 12, colour = "black"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(colour = "black", size = 10))+
  xlab(bquote('Annual relative growth defined as '  ~ frac((height[2] - height[1]),height[1])))+
  ylab("Annual % mortality probability")

ggsave("../../results/figures/GrowthMortality/predict_mortality_abco_Feb25.pdf", height = 4, width = 5, dpi = 600)
```

# Calculate the min and max and median mortality probability captured by the fire footprint growth data

## Create a function for calculating mort probability
```{r}
calc_prob_mort <- function(log_growth){
  prob_3_yrs <- (exp(mean_coef_int + mean_coef_gr*log_growth)/(1+exp(mean_coef_int + mean_coef_gr* log_growth)))
  return(prob_3_yrs/3)
}
```

## Check it
```{r}
test_log_growth <- head(dfgraph)[1,1]
test_log_growth
calc_prob_mort(test_log_growth) == head(dfgraph)[1,2]
calc_prob_mort(log(.05))
```

```{r}
prob_median <-calc_prob_mort(log(gr_medshrub_abco))
prob_median
```

## Use it
```{r}
prob_min <- calc_prob_mort(log(gr_minshrub_abco))
prob_max <- calc_prob_mort(log(gr_maxshrub_abco))

prob_min
prob_max

(prob_max-prob_min)/prob_min
```

## Compile the results
```{r}
abco_growth_pred <- cbind(prob_min, prob_max, prob_median)

abco_growth_pred <- as.data.frame(t(abco_growth_pred)) %>% 
  rename(ABCO_perc_mort = V1) %>% 
  mutate(metric = names(as.data.frame(abco_growth_pred))) %>% 
  mutate(ABCO_perc_mort = ABCO_perc_mort*100) %>% 
  dplyr::select(metric, ABCO_perc_mort)
abco_growth_pred
```

## Compile the results
```{r}
abco_growth_pred <- cbind(prob_min, prob_max, prob_median)
abco_growth_pred <- as.data.frame(t(abco_growth_pred)) 
row.names(abco_growth_pred) <- c("min", "max", "median")
abco_growth_pred
abco_growth_pred <- abco_growth_pred %>% 
  mutate(shrubs = row.names(abco_growth_pred)) %>%  
   mutate(growth = c(gr_minshrub_abco, gr_maxshrub_abco, gr_medshrub_abco)) %>% 
  rename(Mort = "V1") %>% 
  mutate(Species = "ABCO") %>% 
  mutate(Perc_mort = Mort*100) %>% 
  dplyr::select(shrubs, Species, growth, Mort, Perc_mort)
abco_growth_pred
```

# PIPO

## Load simulation results
```{r}
load(file = "../../results/coefficients/mean_coef_gr_pipo.Rdata")
load(file = "../../results/coefficients/mean_coef_int_pipo.Rdata")

load(file = "../../results/coefficients/coef_gr_pipo.Rdata")
load(file = "../../results/coefficients/coef_int_pipo.Rdata")
```

## Plot the mean coefficient and CI

### Calculate probabilities
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
dfgraph <- as.data.frame(log_gr_rel)
dfgraph$prob <- exp(mean_coef_int + mean_coef_gr*log_gr_rel)/(1+exp(mean_coef_int + mean_coef_gr*log_gr_rel))
```

### CI
```{r}
CI_low_int <- sort(coefficient_int)[25]
CI_high_int <- sort(coefficient_int)[975]
CI_low_gr <- sort(coefficient_gr)[25]
CI_high_gr <- sort(coefficient_gr)[975]
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_lower <- as.data.frame(log_gr_rel)
dfgraph_lower$prob <- exp(mean_coef_int + CI_low_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_low_gr*log_gr_rel))
summary(dfgraph_lower$prob)
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_upper <- as.data.frame(log_gr_rel)
dfgraph_upper$prob <- exp(mean_coef_int + CI_high_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_high_gr*log_gr_rel))
summary(dfgraph_upper$prob)
```

### Adjust probabilities to be yearly, since I found dead trees from 3 years 
```{r}
head(dfgraph)
dfgraph <- dfgraph %>%
  mutate(prob = prob/3)
head(dfgraph_lower)
dfgraph_lower <- dfgraph_lower %>%
  mutate(prob = prob/3)
dfgraph_upper <- dfgraph_upper %>%
  mutate(prob = prob/3)
```

## Plot it

```{r}
pineplot <- ggplot()+
  geom_line(data = dfgraph, aes(x = exp(log_gr_rel), y = prob*100))+
  geom_line(data = dfgraph_lower, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  geom_line(data = dfgraph_upper, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  theme_bw()+
  ylim(0, ymax_set)+
  #xlim(0.03, .27)+
  ylab("Annual % mortality probability")+
  xlab(bquote('Annual relative growth defined as '  ~ frac((height[2] - height[1]),height[1])))+
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 12, color = "black"),
                axis.text = element_text(colour = "black", size = 10))
ggsave("../../results/figures/GrowthMortality/predict_mortality_pipo_Feb25.pdf", height = 4, width = 5, dpi = 600)
```


## Calculate the range of mortality probabilities captured by the fire footprint growth data

### Load growth results from fire footprings
```{r}
load(file = "../../results/data/FireFootprints/gr_minshrub_pipo.Rdata")
load(file = "../../results/data/FireFootprints/gr_maxshrub_pipo.Rdata")
load(file = "../../results/data/FireFootprints/gr_medshrub_pipo.Rdata")
```

### Create a function for calculating mort probability
```{r}
calc_prob_mort <- function(log_growth){
  prob_3_yrs <- (exp(mean_coef_int + mean_coef_gr*log_growth)/(1+exp(mean_coef_int + mean_coef_gr* log_growth)))
  return(prob_3_yrs/3)
}
```

### Check it
```{r}
test_log_growth <- head(dfgraph)[1,1]
test_log_growth
calc_prob_mort(test_log_growth) == head(dfgraph)[1,2]
```

### Use it
```{r}
prob_min <- calc_prob_mort(log(gr_minshrub_pipo))
prob_max <- calc_prob_mort(log(gr_maxshrub_pipo))  
prob_min
prob_max
prob_min-prob_max
(prob_max-prob_min)/prob_min
```

```{r}
prob_median <-calc_prob_mort(log(gr_medshrub_pipo))
prob_median
```

### Compile the results
```{r}
pipo_growth_pred <- cbind(prob_min, prob_max, prob_median)
pipo_growth_pred <- as.data.frame(t(pipo_growth_pred)) 
row.names(pipo_growth_pred) <- c("min", "max", "median")
pipo_growth_pred
pipo_growth_pred <- pipo_growth_pred %>% 
  mutate(shrubs = row.names(pipo_growth_pred)) %>%  
  mutate(growth = c(gr_minshrub_pipo, gr_maxshrub_pipo, gr_medshrub_pipo)) %>% 
  rename(Mort = "V1") %>% 
  mutate(Species = "PIPO") %>% 
  mutate(Perc_mort = Mort*100) %>% 
  dplyr::select(shrubs, Species, growth, Mort, Perc_mort)
pipo_growth_pred
```

### Find how large of a change that is from smallest to largest
```{r}
pipo_growth_pred[2,4]/pipo_growth_pred[1,4]
```

# All pred results for both species in one table
```{r}
full_join(abco_growth_pred, pipo_growth_pred)
```

