---
title: "simulations_hybrid"
author: "Carmen Tubbesing"
date: "6/12/2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
require(MASS)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/GrowthMortality_Analysis/")
load(file = "../../compiled_data/growth_mortality/df_vert.Rdata")
load("../../results/data/GrowthMortality/def_green_pipo.Rdata")
```

#  Make dead a factor
```{r}
df <- df %>% 
  mutate(DEAD_ALIVE = as.factor(DEAD_ALIVE))
```

# Make separate df for each species
```{r}
dfa <- df %>% filter(SPECIES == "ABCO")
dfp <- df %>% filter(SPECIES == "PIPO")
```

# Define ymax to be used throughout both abco and pipo graphs
```{r}
ymax <- 15
```

# _
# PIPO SIMULATIONS

## Visually compare distributions
```{r}
means <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(mean = mean(log_gr_rel_ave))
sd <- dfp %>% 
  group_by(DEAD_ALIVE) %>% 
  summarize(sd = sd(log_gr_rel_ave))
means
sd
```

```{r}
x <- seq(-4, 0, length=1000)
y <- dnorm(x, mean=means[[2,2]], sd=sd[[2,2]])
pdf_dead <- as.data.frame(cbind(x, y))
```

```{r}
x <- seq(-4, 0, length=1000)
y <- dnorm(x, mean=means[[1,2]], sd=sd[[1,2]])
pdf_live <- as.data.frame(cbind(x, y))
```

```{r}
ggplot()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="DEAD"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "red", fill = "red")+
  theme_bw()+
  geom_histogram(data = dfp %>% filter(DEAD_ALIVE=="ALIVE"), aes(log_gr_rel_ave, stat(density)), bins = 20, alpha = .2, col = "dark green", fill = "dark green")+
  geom_vline(data = means, aes(xintercept = mean, col = DEAD_ALIVE))+
  scale_color_manual(values = c("dark green", "red"))+
  geom_line(data = pdf_dead, aes(x = x, y = y), col = "red")+
  geom_line(data = pdf_live, aes(x = x, y = y), col = "dark green")
  
ggsave("../../results/figures/GrowthMortality/histogram_dead_alive_pipo.png")
```

# Actual simulations for PIPO

## Count pipo n for live and dead
```{r}
n_dead_pop <- unlist(dfp %>% 
                   filter(DEAD_ALIVE=="DEAD") %>% 
                   summarize(n()))
```

```{r}
n_alive_pop <- n_dead_pop*(524/6)
```

# Hybrid approach = randomly assign year of death to each seedling to within 1 year of estimated year of death - AKA 2015 can be either 2015 or 2016, 2016 can be any year, and 2017 can be 2016 or 2017
- EXCEPT trees that have green on them, which I know died in 2017

## Loop it 
```{r}
coefficient_gr <- c()
coefficient_int <- c()
for(i in 1:1000){
  ## Randomly assign year of death for the trees I'm unsure about 
  dfp_new <- dfp %>% 
    group_by(PAIR) %>% 
    mutate(death_year_temp = ifelse(PAIR %in% definitely_green, 2017, 0)) %>% 
    mutate(death_year_temp = ifelse(death_year == 2017 & (!PAIR %in% definitely_green), 
                                    sample(c(2016, 2017), 1, replace = T), death_year_temp)) %>% 
    mutate(death_year_temp = ifelse(death_year == 2016, 
                                    sample(c(2015, 2016, 2017), 1, replace = T), death_year_temp)) %>% 
    mutate(death_year_temp = ifelse(death_year == 2015, 
                                    sample(c(2015, 2016), 1, replace = T), death_year_temp)) %>% 
    ungroup()
  
  dfp_new2 <- dfp_new %>%
    rowwise() %>% 
    mutate(gr_rel_ave = ifelse(death_year_temp == "2017",
                                 mean(c(gr1, gr2, gr3), na.rm=T), 999)) %>%
    mutate(gr_rel_ave = ifelse(death_year_temp == "2016" & DEAD_ALIVE == "ALIVE",
                                 mean(c(gr2, gr3), na.rm=T), gr_rel_ave)) %>% 
    mutate(gr_rel_ave = ifelse(death_year_temp == "2016" & DEAD_ALIVE == "DEAD",
                                 mean(c(gr1, gr2), na.rm=T), gr_rel_ave)) %>% 
    mutate(gr_rel_ave = ifelse(death_year_temp == "2015" & DEAD_ALIVE == "ALIVE",
                                 gr3, gr_rel_ave)) %>% 
    mutate(gr_rel_ave = ifelse(death_year_temp == "2015" & DEAD_ALIVE == "DEAD",
                                 gr1, gr_rel_ave)) %>% 
    mutate(log_gr_rel_ave = log(gr_rel_ave)) 
  
  gr_set_dead <- dfp_new2 %>% 
      filter(DEAD_ALIVE=="DEAD") %>% 
      dplyr::select(log_gr_rel_ave)
  gr_set_dead <- gr_set_dead[[1]]
  
  gr_set_alive <- dfp_new2 %>% 
    filter(DEAD_ALIVE=="ALIVE") %>%
    dplyr::select(log_gr_rel_ave)
  gr_set_alive <- gr_set_alive[[1]]
  
  dead_sample <- sample(gr_set_dead, size = n_dead_pop, replace = TRUE) 
  alive_sample <- sample(gr_set_alive, size = n_alive_pop, replace = TRUE)
  
  live_df <- data.frame(gr = c(alive_sample))
  live_df$dead <- 0
  
  dead_df <- data.frame(gr = c(dead_sample))
  dead_df$dead <- 1
  
  df <- bind_rows(dead_df, live_df)
  
  glm <- glm(dead ~ gr, family = binomial(link = "logit"), data = df)
  
  coefficient_gr <- c(coefficient_gr, glm$coefficient[2])
  coefficient_int <- c(coefficient_int, glm$coefficient[1])
}
mean_coef_gr <- mean(coefficient_gr)
mean_coef_int <- mean(coefficient_int)
mean_coef_gr
mean_coef_int

  
  # Check
  dfp_new2 %>% arrange(PAIR) %>% dplyr::select(SEEDLING, DEAD_ALIVE, gr1, gr2, gr3, death_year, death_year_temp, gr_rel_ave, log_gr_rel_ave)
  ###########
  

```

## Plot the mean coefficient and CI

### Calculate probabilities
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
dfgraph <- as.data.frame(log_gr_rel)
dfgraph$prob <- exp(mean_coef_int + mean_coef_gr*log_gr_rel)/(1+exp(mean_coef_int + mean_coef_gr*log_gr_rel))
summary(dfgraph$prob)
```

### CI
```{r}
CI_low_int <- sort(coefficient_int)[25]
CI_high_int <- sort(coefficient_int)[975]
CI_low_gr <- sort(coefficient_gr)[25]
CI_high_gr <- sort(coefficient_gr)[975]
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_lower <- as.data.frame(log_gr_rel)
dfgraph_lower$prob <- exp(mean_coef_int + CI_low_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_low_gr*log_gr_rel))
summary(dfgraph_lower$prob)
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_upper <- as.data.frame(log_gr_rel)
dfgraph_upper$prob <- exp(mean_coef_int + CI_high_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_high_gr*log_gr_rel))
summary(dfgraph_upper$prob)
```

### Adjust probabilities to be yearly, since I found dead trees from 3 years 
```{r}
head(dfgraph)
dfgraph <- dfgraph %>% 
  mutate(prob = prob/3)
head(dfgraph_lower)
dfgraph_lower <- dfgraph_lower %>% 
  mutate(prob = prob/3)
dfgraph_upper <- dfgraph_upper %>% 
  mutate(prob = prob/3)
```

# Plot it
```{r}
ggplot()+
  geom_line(data = dfgraph, aes(x = exp(log_gr_rel), y = prob*100))+
  geom_line(data = dfgraph_lower, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  geom_line(data = dfgraph_upper, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  theme_bw()+
  xlim(0.03, .25)+
  ylim(0, ymax)+
  ylab("Annual % mortality probability")+
  xlab("Annual relative growth defined as\n(height_2 - height_1)/\nheight_1")+
  geom_vline(aes(xintercept = exp(-1.70428)), col = "red")+
  geom_text(aes(x= exp(-1.70428), label="\nminimum fire footprint shrub competition", y = ymax*.6), colour="red", angle=90) +
  geom_vline(aes(xintercept = exp(-2.5877)), col = "red")+
  geom_text(aes(x= exp(-2.5877), label="\nmaximum fire footprint shrub competition", y = ymax*.6), colour="red", angle=90) +
  geom_vline(aes(xintercept = exp(-1.955775)), col = "red")+
  geom_text(aes(x= exp(-1.955775), label="\nmedian fire footprint shrub competition", y = ymax*.6), colour="red", angle=90) +
  ggtitle("HYBRID SIMULATIONS Pine mortality probability in relation to growth\nbased on simulation results, with 95% confidence intervals")
ggsave("../../results/figures/GrowthMortality/predict_mortality_pipo_hybrid.png", height = 5, width = 6)
```

# Calculate the range of mortality probabilities captured by the fire footprint growth data
```{r}
prob_min <- exp(mean_coef_int + mean_coef_gr*(-1.70428)/(1+exp(mean_coef_int + mean_coef_gr*(-1.70428))))
prob_max <- exp(mean_coef_int + mean_coef_gr*(-2.5877)/(1+exp(mean_coef_int + mean_coef_gr*(-2.5877))))
prob_min
prob_max
prob_min-prob_max
(prob_max-prob_min)/prob_min
```

```{r}
prob_median <- exp(mean_coef_int + mean_coef_gr*(-1.955775)/(1+exp(mean_coef_int + mean_coef_gr*(-1.955775))))
prob_quart1 <- exp(mean_coef_int + mean_coef_gr*(-1.8115886)/(1+exp(mean_coef_int + mean_coef_gr*(-1.8115886))))
prob_quart3 <- exp(mean_coef_int + mean_coef_gr*(-2.10016998)/(1+exp(mean_coef_int + mean_coef_gr*(-2.10016998))))
prob_median
prob_quart1
prob_quart3
prob_median - prob_quart1
(prob_quart3 - prob_median)/prob_median
```


# Next steps: go back and look at how decayed the dead seedlings I found in the surveys were to compare mortality rate by year to the dead seedlings I measured?