---
title: "simulations_hybrid_plot"
author: "Carmen"
date: "September 4, 2019"
output: 
    html_document:
        toc: TRUE
---

```{r, include = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(effects)
```

# Set upper y limit for graphs 
```{r}
ymax_set <- 10
```


# PIPO

## Load simulation results
```{r}
load(file = "../../results/coefficients/mean_coef_gr_pipo_hybrid.Rdata")
load(file = "../../results/coefficients/mean_coef_int_pipo_hybrid.Rdata")

load(file = "../../results/coefficients/coef_gr_pipo_hybrid.Rdata")
load(file = "../../results/coefficients/coef_int_pipo_hybrid.Rdata")
```


## Plot the mean coefficient and CI

### Calculate probabilities
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
dfgraph <- as.data.frame(log_gr_rel)
dfgraph$prob <- exp(mean_coef_int + mean_coef_gr*log_gr_rel)/(1+exp(mean_coef_int + mean_coef_gr*log_gr_rel))
summary(dfgraph$prob)
```

### CI
```{r}
CI_low_int <- sort(coefficient_int)[25]
CI_high_int <- sort(coefficient_int)[975]
CI_low_gr <- sort(coefficient_gr)[25]
CI_high_gr <- sort(coefficient_gr)[975]
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_lower <- as.data.frame(log_gr_rel)
dfgraph_lower$prob <- exp(mean_coef_int + CI_low_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_low_gr*log_gr_rel))
summary(dfgraph_lower$prob)
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_upper <- as.data.frame(log_gr_rel)
dfgraph_upper$prob <- exp(mean_coef_int + CI_high_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_high_gr*log_gr_rel))
summary(dfgraph_upper$prob)
```

### Adjust probabilities to be yearly, since I found dead trees from 3 years 
```{r}
head(dfgraph)
dfgraph <- dfgraph %>% 
  mutate(prob = prob/3)
head(dfgraph_lower)
dfgraph_lower <- dfgraph_lower %>% 
  mutate(prob = prob/3)
dfgraph_upper <- dfgraph_upper %>% 
  mutate(prob = prob/3)
```

# Plot it
```{r}
ggplot()+
  geom_line(data = dfgraph, aes(x = exp(log_gr_rel), y = prob*100))+
  geom_line(data = dfgraph_lower, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  geom_line(data = dfgraph_upper, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  theme_bw()+
  theme(text = element_text(size = 11),
        panel.grid.minor = element_blank())+
  #xlim(0.03, .25)+
  ylim(0, ymax_set)+
  ylab("Annual % mortality probability")+
  xlab(bquote('Annual relative growth defined as '  ~ frac((height[2] - height[1]),height[1])))
ggsave("../../results/figures/GrowthMortality/predict_mortality_pipo_hybrid.png", height = 4, width = 5)
```

# Calculate the range of mortality probabilities captured by the fire footprint growth data
```{r}
prob_min <- exp(mean_coef_int + mean_coef_gr*(-1.70428)/(1+exp(mean_coef_int + mean_coef_gr*(-1.70428))))
prob_max <- exp(mean_coef_int + mean_coef_gr*(-2.5877)/(1+exp(mean_coef_int + mean_coef_gr*(-2.5877))))
prob_min
prob_max
prob_min-prob_max
(prob_max-prob_min)/prob_min
```

```{r}
prob_median <- exp(mean_coef_int + mean_coef_gr*(-1.955775)/(1+exp(mean_coef_int + mean_coef_gr*(-1.955775))))
prob_quart1 <- exp(mean_coef_int + mean_coef_gr*(-1.8115886)/(1+exp(mean_coef_int + mean_coef_gr*(-1.8115886))))
prob_quart3 <- exp(mean_coef_int + mean_coef_gr*(-2.10016998)/(1+exp(mean_coef_int + mean_coef_gr*(-2.10016998))))
prob_median
prob_quart1
prob_quart3
prob_median - prob_quart1
(prob_quart3 - prob_median)/prob_median
```

# __

# ABCO

## Load simulation results
```{r}
load(file = "../../results/coefficients/mean_coef_gr_abco_hybrid.Rdata")
load(file = "../../results/coefficients/mean_coef_int_abco_hybrid.Rdata")

load(file = "../../results/coefficients/coef_gr_abco_hybrid.Rdata")
load(file = "../../results/coefficients/coef_int_abco_hybrid.Rdata")
```


## Plot the mean coefficient and CI

### Calculate probabilities
```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
dfgraph <- as.data.frame(log_gr_rel)
dfgraph$prob <- exp(mean_coef_int + mean_coef_gr*log_gr_rel)/(1+exp(mean_coef_int + mean_coef_gr*log_gr_rel))
summary(dfgraph$prob)
```

### CI
```{r}
CI_low_int <- sort(coefficient_int)[25]
CI_high_int <- sort(coefficient_int)[975]
CI_low_gr <- sort(coefficient_gr)[25]
CI_high_gr <- sort(coefficient_gr)[975]
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_lower <- as.data.frame(log_gr_rel)
dfgraph_lower$prob <- exp(mean_coef_int + CI_low_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_low_gr*log_gr_rel))
summary(dfgraph_lower$prob)
```

```{r}
log_gr_rel <- seq(log(.03), log(.3), length=1000)
head(log_gr_rel)
dfgraph_upper <- as.data.frame(log_gr_rel)
dfgraph_upper$prob <- exp(mean_coef_int + CI_high_gr*log_gr_rel)/(1+exp(mean_coef_int + CI_high_gr*log_gr_rel))
summary(dfgraph_upper$prob)
```

### Adjust probabilities to be yearly, since I found dead trees from 3 years 
```{r}
head(dfgraph)
dfgraph <- dfgraph %>% 
  mutate(prob = prob/3)
head(dfgraph_lower)
dfgraph_lower <- dfgraph_lower %>% 
  mutate(prob = prob/3)
dfgraph_upper <- dfgraph_upper %>% 
  mutate(prob = prob/3)
```

# Plot it
```{r}
ggplot()+
  geom_line(data = dfgraph, aes(x = exp(log_gr_rel), y = prob*100))+
  geom_line(data = dfgraph_lower, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  geom_line(data = dfgraph_upper, aes(x = exp(log_gr_rel), y = prob*100), linetype = "dashed")+
  theme_bw()+
  #xlim(0.03, .25)+
  ylim(0, ymax_set)+
  theme(text = element_text(size = 11),
      panel.grid.minor = element_blank())+
  ylab("Annual % mortality probability")+
    xlab(bquote('Annual relative growth defined as '  ~ frac((height[2] - height[1]),height[1])))
ggsave("../../results/figures/GrowthMortality/predict_mortality_abco_hybrid.png", height = 4, width = 5)
```

# Calculate the range of mortality probabilities captured by the fire footprint growth data
```{r}
prob_min <- exp(mean_coef_int + mean_coef_gr*(-1.70428)/(1+exp(mean_coef_int + mean_coef_gr*(-1.70428))))
prob_max <- exp(mean_coef_int + mean_coef_gr*(-2.5877)/(1+exp(mean_coef_int + mean_coef_gr*(-2.5877))))
prob_min
prob_max
prob_min-prob_max
(prob_max-prob_min)/prob_min
```

```{r}
prob_median <- exp(mean_coef_int + mean_coef_gr*(-1.955775)/(1+exp(mean_coef_int + mean_coef_gr*(-1.955775))))
prob_quart1 <- exp(mean_coef_int + mean_coef_gr*(-1.8115886)/(1+exp(mean_coef_int + mean_coef_gr*(-1.8115886))))
prob_quart3 <- exp(mean_coef_int + mean_coef_gr*(-2.10016998)/(1+exp(mean_coef_int + mean_coef_gr*(-2.10016998))))
prob_median
prob_quart1
prob_quart3
prob_median - prob_quart1
(prob_quart3 - prob_median)/prob_median
```

