---
title: "Clean_shrubs"
author: "Carmen"
date: "May 20, 2019"
output: 
    html_document:
        toc: TRUE
        toc_depth: 1
---

```{r, include = F}
require(readxl)
require(dplyr)
require(tidyverse)
```

# Load data
```{r}
setwd("~/Shrubs-Seedlings/code/FireFootprints_analysis/")
shr <- read.csv("~/../Dropbox (Stephens Lab)/SORTIE/Shrubs_Summer16/Completed_Data_and_Photos/Master_Compiled_shrubs.csv")
load("../../compiled_data/fire_footprints/seedlings_cleaned_2016.Rdata")
```

# Basic shrub cleaning

## Rename seedling column
```{r}
shr <- shr %>% 
  rename(Sdlg = Seedling.) 
```
 
## Replace NAs with 0
```{r}
shr <- tbl_df(shr)
shr$Dead.[is.na(shr$Dead.)] <- 0
shr$Cover.cm[is.na(shr$Cover.cm)] <- 0
shr$Ht.cm[is.na(shr$Ht.cm)] <- 0
```

## Take out "pea" shrub cover because it's not a shrub
```{r}
shr <- shr %>% filter(ShrubSpp != "pea")
```

## Take out "conifer" since it's not a shrub
```{r}
summary(as.factor(shr$ShrubSpp))
shr <- shr %>% 
  filter(!(ShrubSpp == "conifer"))
```

## Drop levels
```{r}
shr <- droplevels(shr)
```

# Clean up segment names
```{r}
shr$Seg <- 0
for(i in 1:nrow(shr)){
if(shr$Segment[i] %in% c("0-1 m ","0 -1 m","0-1 m")){
  shr$Seg[i] <- 1
} else if (shr$Segment[i]=="1-2 m "){
  shr$Seg[i] <- 2
} else if (shr$Segment[i] %in% c("2-3 m","2-3m")){
  shr$Seg[i] <- 3
}  else
    shr$Seg[i] <- 999
}
summary(as.factor(shr$Seg))
```

# Calculate height shrub species rather than across species

## Weighted mean by seedling + shrub species
```{r}
ht_by_spp <- shr %>%
  group_by(Sdlg, ShrubSpp) %>% 
  mutate(ht_wm.cm.spp = weighted.mean(Ht.cm, Cover.cm)) %>% 
  arrange(Sdlg, ShrubSpp)
ht_by_spp
```

## Get one value per seedling + shrub species
```{r}
ht_by_spp <- ht_by_spp %>% 
  dplyr::select(Sdlg, ShrubSpp, ht_wm.cm.spp) %>% 
  distinct() %>% 
  ungroup() %>% 
  arrange(Sdlg, ShrubSpp)
ht_by_spp
```

# Make Sdlg all caps
```{r}
ht_by_spp <- ht_by_spp %>% 
  ungroup() %>% 
  mutate(Sdlg = toupper(Sdlg))
```

## Check for duplicated seedlings
```{r}
ht_by_spp %>% 
  group_by(Sdlg, ShrubSpp) %>% 
  summarize(n=n()) %>% 
  filter(n !=1)
```

# Add Fire names
```{r}
fires <- df %>% 
  dplyr::select(Fire, Sdlg)
ht_by_spp <- left_join(ht_by_spp, fires)
```


# Save
```{r}
save(ht_by_spp, file="../../compiled_data/shrub_growth/My_FireFootprint_data.Rdata")
```

