---
title: "Clean Welch data"
author: "Carmen"
date: "January 3, 2020"
output: 
  html_document:
    toc: TRUE
---

```{r, include = F}
setwd("~/../Shrubs-Seedlings/code/Shrub_growth_time/")
require(tidyverse)
require(lubridate)
require(readxl)
```

# Load Welch data
```{r}
plot <- read_excel("../../data/Data_for_shrub_growth/Welch_data/plot_data_RegenDatabase_1_13_16_updatedDates.xlsx") 
seed_tree <- read_excel("../../data/Data_for_shrub_growth/Welch_data/seed_tree_RegenDatabase_1_13_16_updatedDates.xlsx") 
shrub <- read_excel("../../data/Data_for_shrub_growth/Welch_data/shrub_regen_RegenDatabase_1_13_16_updatedDates.xlsx") 
```

# Load cleaned Young data
```{r}
load("~/../Shrubs-Seedlings/compiled_data/shrub_growth/Young_df.Rdata")
young <- df
remove(df)
```

# Take out unnecessary columns

## Plot
```{r}
plot <- plot %>% 
  dplyr::select(ID, Regen_Plot, Fire, Date, `Year of Fire`, FIRE_SEV)
```

## shrub
```{r}
shrub <- shrub %>% 
  dplyr::select(-comments, Cover, -ID, -modal_ht_ft, -OldPLOTID)
```

# Take out shrub columns with no shrub ht data
```{r}
shrub <- shrub %>% filter(!is.na(modal_ht_cm))
```

# Join shrub data and plot data
```{r}
df <- left_join(shrub, plot, by = "Regen_Plot")
```

# Create columns for year measured, years since fire, and source
```{r}
head(df)
df <- df %>% 
  mutate(Year_meas = lubridate::year(Date)) %>% 
  mutate(years_since_fire = Year_meas - `Year of Fire`) %>% 
   rename(Shrub_species = Species) %>% 
  dplyr::select(years_since_fire, Year_meas, `Year of Fire`, Shrub_species, everything())
head(df)
```

# Filter plots by fire severity
```{r}
df <- df %>% 
  filter(FIRE_SEV > 3)
```

# Reshape seed tree table to join with df
```{r}
seed_tree <- seed_tree %>% 
  group_by(Regen_Plot) %>% 
  filter(Dist_m == min(Dist_m)) %>% 
  dplyr::select(Regen_Plot, Dist_m) %>% 
  rename(seed_tree_dist_m = Dist_m) %>% 
  distinct()
```

# Join seed tree table with df
```{r}
df <- left_join(df, seed_tree)
```

# If seed tree distance is NA, replace it with 999 based on March 27 email with Derek Young
```{r}
df <- df %>% 
  mutate(seed_tree_dist_m = ifelse(is.na(seed_tree_dist_m), 999, seed_tree_dist_m))
```

# Filter df to plots with >20 m distance to seed tree
```{r}
df <- df %>% 
  filter(seed_tree_dist_m>20)
```

# Filter to shrub species of interest
```{r}
summary(as.factor(df$Shrub_species))
df <- df %>% 
  filter(Shrub_species %in% c("ARNE", "ARPA", "ARPA6", "ARVI", "CECO", "CEIN","CEIN3", "CEPR", "CHFO", "CHSE11", "PREM", "QUVA", "RIBES", "RICE", "RIRO", "SYMO"))
summary(as.factor(df$Shrub_species))
```

# Fix species name typos
```{r}
df <- df %>% 
  mutate(Shrub_species = 
           case_when(Shrub_species == "ARPA" ~ "ARPA6",
                     Shrub_species == "CEIN" ~ "CEIN3",
                    Shrub_species == "RICE" ~ "RIBES",
                     Shrub_species == "RIRO" ~ "RIBES",
                     TRUE ~ Shrub_species))
```

# Check for NA modal shrub ht
```{r}
df %>% filter(is.na(modal_ht_cm))
```

# Plot HEIGHT for each species with more than five years
```{r}
years <- df %>% 
  group_by(Shrub_species) %>% 
  select(Shrub_species, years_since_fire) %>% 
  distinct() %>% 
  summarize(years = n()) %>% 
  filter(years>5) %>% 
  dplyr::select(Shrub_species)
years
```

```{r}
ggplot(df %>% 
         filter(Shrub_species %in% years$Shrub_species) %>%
         filter(years_since_fire<24), 
       aes(x = years_since_fire, y = modal_ht_cm, col = Shrub_species))+
  geom_point()+
  geom_smooth(method = "loess", span = 1)

```

# Plot COVER for each species with more than five years
```{r}
ggplot(df %>% 
         filter(Shrub_species %in% years$Shrub_species) %>%
         filter(years_since_fire<24), 
       aes(x = years_since_fire, y = Cover, col = Shrub_species))+
  geom_point()+
  geom_smooth(method = "loess", span = 1)
```

# Look at column names
```{r}
names(df)
```


# Save
```{r}
save(df, file = "~/../Shrubs-Seedlings/compiled_data/shrub_growth/Welch_df.Rdata")
```
