---
title: "combine shrub data with species lumped"
author: "Carmen"
date: "January 21, 2020"
output: 
  html_document:
    toc: TRUE
---

```{r, include=FALSE}
require(tidyverse)
```

# Load data

## Data from others
```{r}
load("~/Shrubs-Seedlings/compiled_data/shrub_growth/Latimer_df.Rdata")
latimer <- df
remove(df)
load("~/Shrubs-Seedlings/compiled_data/shrub_growth/Young_df.Rdata")
young <- df
remove(df)
load("~/Shrubs-Seedlings/compiled_data/shrub_growth/Welch_df.Rdata")
welch <- df
remove(df)
```

# Sync column names

## Latimer
```{r}
latimer <- latimer %>% 
  mutate(Year_meas = 2018) %>% 
  dplyr::select(Fire, FireSev, years_since_fire, Year_meas, Shrub_species, shrub_ht, Shrub_cover, PlotID) %>% 
  rename(modal_ht_cm = shrub_ht)
```

## Welch
```{r}
welch <- welch %>% 
  rename(fire_year = `Year of Fire`, FireSev = FIRE_SEV, Shrub_cover = Cover, PlotID = Regen_Plot) %>% 
  dplyr::select(-seed_tree_dist_m)
```

## Young
```{r}
young <- young %>% 
  rename(Shrub_species = dominant_shrub_1) %>% 
  rename(modal_ht_cm = dominant_shrub_ht_cm, FireSev = FIRE_SEV, Shrub_cover = Shrub_cover_allspp) %>% 
  dplyr::select(-Date)
```

# Restructure data to combine cover for all shrub species in a given plot

## Young

These data are already structured the way I want, as this shows:
```{r}
young %>% 
  group_by(Regen_Plot) %>% 
  filter(n()>1)
head(young)
```

## Welch
```{r}
welch %>% 
  group_by(PlotID) %>% 
  mutate(Shrub_cover = sum(Shrub_cover)) %>% 
  arrange(PlotID) 
```

```{r}
welch <- welch %>% 
  group_by(PlotID) %>% 
  filter(Shrub_cover == max(Shrub_cover)) 
```

## latimer
```{r}
latimer %>% 
  group_by(PlotID) %>% 
  mutate(Shrub_cover = sum(Shrub_cover)) %>% 
  arrange(PlotID) %>% 
  filter(n()>1) %>% 
  dplyr::select(PlotID, everything())
```

```{r}
latimer <- latimer %>% 
  group_by(PlotID) %>% 
  filter(Shrub_cover == max(Shrub_cover)) 
```

# Combine

## Add column for df_source
```{r}
latimer$df_source = "latimer"
welch$df_source = "welch"
young$df_source = "young"
```

## Welch and Latimer data
```{r}
df <- full_join(welch, latimer)
```

## Check
```{r}
nrow(latimer) + nrow(welch) == nrow(df)
```

```{r}
df %>% 
  group_by(df_source, PlotID) %>% 
  count() %>% 
  group_by(df_source) %>% 
  count()
```

## Add young data
```{r}
df <- full_join(df, young)
```


# Sync shrub species names

```{r}
summary(as.factor(df$Shrub_species))
```

```{r}
df <- df %>% 
  mutate(Shrub_species = 
           case_when(Shrub_species == "ARCPAT" ~ "ARPA6",
                     Shrub_species == "ARCVIS" ~ "ARVI",
                     Shrub_species == "CEACOR" ~ "CECO",
                     Shrub_species == "CEAPRO" ~ "CEPR",
                     Shrub_species == "CEAINT" ~ "CEIN3",
                     Shrub_species == "CEINT" ~ "CEIN3",
                     Shrub_species == "CHAFOL" ~ "CHFO",
                     Shrub_species == "CHRSEM" ~ "CHSE11",
                     Shrub_species == "QUEVAC" ~ "QUVA",
                     Shrub_species == "SYMMOL" ~ "SYMO",
                     Shrub_species == "RIBCER" ~ "RIBES",
                     Shrub_species == "RIRO" ~ "RIBES",
                     TRUE ~ Shrub_species))
```

```{r}
summary(as.factor(df$Shrub_species))
```

# Clean

## Take out the oldest fire, Cottonwood, which Derek mentioned is at the sagebrush ecotone so not a good comparison and which I didn't use in the height analysis
```{r}
df <- df %>% 
  filter(years_since_fire<24)
```

## Take out rows with NA shrub cover
```{r}
df %>% filter(is.na(Shrub_cover))
```

```{r}
df <- df %>% 
  filter(!is.na(Shrub_cover))
```

# Plot for all species together
```{r}
ggplot(df, aes(x = years_since_fire, y = Shrub_cover))+
  geom_point()+
  geom_smooth(method = "loess", span = 1)
```


# Plot just for ARVI and ARPA
```{r, warning = F}
ggplot(df %>% filter(Shrub_species %in% c("ARVI", "ARPA6")), aes(x = years_since_fire, y = Shrub_cover, col = Shrub_species))+
  geom_point()+
    geom_smooth(method = "loess")
```

# Lump ARVI and ARPA
```{r}
df <- df %>% 
  mutate(Shrub_species = 
           case_when(Shrub_species == "ARPA6" ~ "ARPA6_ARVI",
                     Shrub_species == "ARVI" ~ "ARPA6_ARVI",
                     TRUE ~ Shrub_species))
```

# Plot for manzanita, whitethorn, and deerbrush
```{r}
ggplot(df %>% 
         filter(Shrub_species %in% c("ARPA6_ARVI", "CEIN3", "CECO")),
       aes(x = years_since_fire, y = Shrub_cover, col = Shrub_species))+
  geom_point(aes(shape = df_source))+
  geom_smooth(method = "loess", span = 1)+
  theme_bw()
```


# Plot faceted
```{r}
df_plot <- df %>% 
  filter(Shrub_species %in% c("ARPA6_ARVI", "CECO", "CEIN3", "CHSE11"))
unique(df_plot$Shrub_species)
df_plot <- droplevels(df_plot)
```

```{r}
ggplot(df_plot, aes(x = years_since_fire, y = Shrub_cover))+
  geom_point(aes(shape = df_source, col = Fire))+
  geom_smooth(method = "gam")+
  facet_wrap(~Shrub_species, ncol = 2)+
  theme_bw()+
  ylim(0, 150)
```


# Save combined data
```{r}
save(df, file = "~/Shrubs-Seedlings/compiled_data/shrub_growth/all_combined_COVER_LUMPED.Rdata")
```

# Look at trends in my data
```{r}

```

