---
title: "Shrub_growth_analysis_COVER"
author: "Carmen"
date: "January 16, 2020"
output: 
  html_document:
    toc: TRUE
---

# Load data
```{r, include = F}
require(tidyverse)
require(effects)
require(mgcv)
require(gridExtra)
load(file = "~/Shrubs-Seedlings/compiled_data/shrub_growth/My_FireFootprint_data.Rdata")
me <- ht_by_spp
remove(ht_by_spp)
load(file = "~/Shrubs-Seedlings/compiled_data/shrub_growth/all_combined_COVER_LUMPED.Rdata")
```

```{r}
df <- df %>% ungroup()
```

# Changes in shrub cover for MY fire footprints

## Join ARPA and ARVI
```{r}
me <- me %>% 
  mutate(Shrub_species = as.character(Shrub_species)) %>% 
  mutate(Shrub_species = case_when(
    Shrub_species == "ARPA" ~ "ARPA-ARVI",
    Shrub_species == "ARVI" ~ "ARPA-ARVI",
    TRUE ~ Shrub_species
  ))
```

## Plot summed cover by plot 

### Summarize
```{r}
me_totals <- me %>% 
  group_by(Sdlg, Fire, Year_meas, years_since_fire) %>% 
  summarize(cover_sum = sum(cover), ht_wt_mean_cm = weighted.mean(modal_ht_cm, cover))
```

### Plot
```{r}
ggplot(me_totals, aes(x = years_since_fire, y = cover_sum))+
  geom_point()+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cr"))
```

### Clean
```{r}
spp <- me %>% 
  group_by(Sdlg) %>% 
  filter(cover == max(cover)) %>% 
  dplyr::select(Sdlg, Shrub_species)
```

```{r}
me_totals <- full_join(me_totals, spp)
```

```{r}
me_totals <- me_totals %>% 
  mutate(cover_percent = cover_sum/(300*4)*100)
```

### Remove the one data point for ARPA-ARVI in year 11-12
```{r}
me_totals %>% 
  filter(Shrub_species == "ARPA-ARVI") %>% 
  group_by(years_since_fire) %>% 
  count()
```

```{r}
me_totals <- me_totals %>% 
  filter(!(Shrub_species == "ARPA-ARVI" & years_since_fire == 12))
```

## Lump shrub species into the categories used in the simulation
```{r}
me_totals <- me_totals %>% 
  mutate(Shrub_species = case_when(
    !Shrub_species %in% c("ARPA-ARVI", "CECO", "CEIN", "CHSE", "LIDE") ~ "Other",
    TRUE ~ Shrub_species
  ))
```

# LINEAR Model of shrub cover increase across my data 

```{r}
lmALL.ME <- lm(cover_sum ~ years_since_fire * Shrub_species + Shrub_species *ht_wt_mean_cm, data = me_totals)
summary(lmALL.ME)
drop1(lmALL.ME, test = "F")
```

### Look at most extreme residual and take it out 
```{r}
me_totals[resid(lmALL.ME)< (-1000),]
me_totals <- me_totals %>% 
  filter(!Sdlg ==294)
```

### Compare AIC for some models
```{r}
AICcmodavg::AICc(lmALL.ME)
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire * Shrub_species +ht_wt_mean_cm, data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire * Shrub_species + Shrub_species*ht_wt_mean_cm, data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire + Shrub_species + ht_wt_mean_cm, data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire * Shrub_species * ht_wt_mean_cm, data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire  + Shrub_species*ht_wt_mean_cm, data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire * Shrub_species + Shrub_species *sqrt(ht_wt_mean_cm), data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire *sqrt(ht_wt_mean_cm), data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire *ht_wt_mean_cm, data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire + sqrt(ht_wt_mean_cm), data = me_totals))
AICcmodavg::AICc(lm(cover_sum ~ years_since_fire + ht_wt_mean_cm, data = me_totals))
```


### Rerun model using the one with the lowest AICc

```{r}
lmALL.ME <- lm(cover_sum ~ years_since_fire * Shrub_species + Shrub_species *sqrt(ht_wt_mean_cm), data = me_totals)
summary(lmALL.ME)
drop1(lmALL.ME, test = "F")
```

### Plot
```{r}
ggplot(me_totals)+
  geom_point(aes(x = years_since_fire, y = cover_sum))+
  geom_point(aes(x = years_since_fire, y = predict(lmALL.ME)), col = "red")
```

```{r}
ggplot(me_totals, aes(x = years_since_fire, y = resid(lmALL.ME)))+
  geom_point()+
  geom_smooth(span = .5)
hist(resid(lmALL.ME), breaks = 20)
```

## Rename columns to match those in the simulations
```{r}
me_totals <- me_totals %>% 
  rename(Years = years_since_fire, ShrubSpp03 = Shrub_species, Ht1.3 = ht_wt_mean_cm)
```

```{r}
me_totals <- me_totals %>% 
  mutate(ShrubSpp03 = ifelse(ShrubSpp03 == "ARPA-ARVI", "ARPA", ShrubSpp03))
```


### Rerun model with new column names

 years_since_fire * Shrub_species + Shrub_species *sqrt(ht_wt_mean_cm)

```{r}
lmALL.ME <- lm(cover_sum ~ Years*ShrubSpp03 + ShrubSpp03*sqrt(Ht1.3), data = me_totals)
summary(lmALL.ME)
drop1(lmALL.ME, test = "F")
```


```{r}
plot(predictorEffect("Years", lmALL.ME))
plot(predictorEffect("Ht1.3", lmALL.ME))
```

# Save final model
```{r}
save(lmALL.ME, file = "~/Shrubs-Seedlings/results/coefficients/LM_shrubcover.Rdata")
```

